### 2020-01-10 work.Rmd
starting another session

```{r}
library(tidyverse)
library(mirt)
R.utils::sourceDirectory("R")
```

```{r}
pars_1_dif <-
    read_csv("data-parameters/kopf2015-8.csv") %>%
    arrange(desc(easy)) %>%
    mutate(
        item = row_number(),
        nuisance_disc = c(0, 0, 0, 0.2, 0, 0, 0, 0)
    )

sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 10000, n_foc = 10000, mean_foc = -1)

dif_constrained_baseline_one_at_a_time(data = sim_1_dif$data, groups = sim_1_dif$groups)
```

```{r}
pars_3_dif <-
    read_csv("data-parameters/kopf2015-8.csv") %>%
    arrange(desc(easy)) %>%
    mutate(
        item = row_number(),
        nuisance_disc = c(0, 0, 0, 0.5, 0.5, 0.5, 0, 0)
    )
```

### 2020-01-10 R/dif.R

@@ -1,0 +1,6 @@
+dif_constrained_baseline_one_at_a_time <- function(data, groups){
+    tibble(
+        item = 1:ncol(data)
+    ) %>%
+        mutate(p = item %>% map_dbl(~ constrained_baseline_test_flex_items(data, groups, flex_items = .)))
+}

### 2020-01-10 R/dif_helpers_level_1.R

@@ -1,0 +1,6 @@
+constrained_baseline_test_flex_items <- function(data, groups, flex_items){
+    quiet(anova(
+        mod_constrained_baseline(data = sim_1_dif$data, groups = sim_1_dif$groups),
+        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = flex_items)
+    ))$p[2]
+}

### 2020-01-10 R/dif_helpers_level_2.R

@@ -1,0 +1,21 @@
+mod_constrained_baseline <- function(data, groups){
+    multipleGroup(
+        data,
+        1,
+        itemtype = "2PL",
+        groups,
+        invariance = c("free_means", "free_var", "intercepts", "slopes"),
+        technical = list(NCYCLES = 100)
+    )
+}
+
+mod_flexible <- function(data, groups, flex_items){
+    multipleGroup(
+        data,
+        1,
+        itemtype = "2PL",
+        groups,
+        invariance = c("free_means", "free_var", paste0("Item_", (1:ncol(data))[-flex_items])),
+        technical = list(NCYCLES = 100)
+    )
+}

### 2020-01-10 work.R

@@ -26,6 +26,7 @@ coef_1f(mod1f)
 
 # not sure why all of the action happens in item 1 here
 # perhaps evidence that modeling multidimensionally is hard
+# and maybe something to do with variance of that second dimension being 0 but us not inputting that
 mod2f <- multipleGroup(sim$data, 2, itemtype = "2PL", sim$groups, invariance = c("free_means","slopes", "intercepts"), technical = list(NCYCLES = 50))
 coef(mod2f)
 

### 2020-01-10 work.Rmd
dif_intuitive works great for plotting logit graphs

@@
         item = row_number(),
-        nuisance_disc = c(0, 0, 0, 0.2, 0, 0, 0, 0)
+        nuisance_disc = c(0, 0, 0, 0.5, 0, 0, 0, 0)
     )
 
-sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 10000, n_foc = 10000, mean_foc = -1)
+sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 50000, n_foc = 50000, mean_foc = -1)
+```
 
+```{r}
 dif_constrained_baseline_one_at_a_time(data = sim_1_dif$data, groups = sim_1_dif$groups)
+```
+
+```{r}
+mod <- 
+    multipleGroup(
+        sim_1_dif$data,
+        1,
+        itemtype = "Rasch",
+        sim_1_dif$groups,
+        invariance = c("free_var"),
+        technical = list(NCYCLES = 100),
+        SE = TRUE
+    )
+
+mod %>% 
+    mod_to_draws_df() %>% 
+    draws_df_to_logit_plot()
 ```

### 2020-01-10 R/dif_helpers_level_2.R

@@ -2,7 +2,7 @@ mod_constrained_baseline <- function(data, groups){
     multipleGroup(
         data,
         1,
-        itemtype = "2PL",
+        itemtype = "Rasch",
         groups,
         invariance = c("free_means", "free_var", "intercepts", "slopes"),
         technical = list(NCYCLES = 100)
@@ -13,7 +13,7 @@ mod_flexible <- function(data, groups, flex_items){
     multipleGroup(
         data,
         1,
-        itemtype = "2PL",
+        itemtype = "Rasch",
         groups,
         invariance = c("free_means", "free_var", paste0("Item_", (1:ncol(data))[-flex_items])),
         technical = list(NCYCLES = 100)

### 2020-01-10 R/dif_intuitive.R

@@ -1,0 +1,30 @@
+mod_to_draws_df <- function(mod){
+    par_draws <- MASS::mvrnorm(n = 10000, mu = extract.mirt(mod, "parvec"), Sigma = extract.mirt(mod, "vcov"))
+
+    par_draws <- par_draws[ , str_detect(colnames(par_draws), "d")]
+
+    draws_df <- tibble(run = 1:nrow(par_draws))
+
+    stopifnot(ncol(par_draws) %% 2 == 0)
+
+    n_items <- ncol(par_draws) / 2
+
+    for (i in 1:n_items) {
+        draws_df[[paste0("item", i)]] <- par_draws[ , i] - par_draws[ , i + n_items]
+    }
+
+    draws_df
+}
+
+draws_df_to_logit_plot <- function(draws_df){
+    draws_df %>%
+        select(-run) %>%
+        gather(var, val) %>%
+        ggplot(aes(x = val, y = var)) +
+        ggridges::geom_density_ridges() +
+        labs(
+            x = "Reference group logits - Focal group logits",
+            y = "",
+            subtitle = "Identify by fixing both means to 0 and let difference show up in item easiness"
+        )
+}

### 2020-01-11 work.Rmd
before fully cleaning to be careful

@@
         item = row_number(),
-        nuisance_disc = c(0, 0, 0, 0.5, 0, 0, 0, 0)
+        nuisance_disc = c(0, 0, 0, 0.3, 0, 0, 0.3, 0)
     )
@@
 ```{r}
-dif_constrained_baseline_one_at_a_time(data = sim_1_dif$data, groups = sim_1_dif$groups)
+get_next_status <- function(status, data, groups){
+    the_flex <- which(status$status == "flex")
+
+    status %>% 
+        filter(status == "unknown") %>% 
+        mutate(
+            p = item %>% 
+                map_dbl(~ test_with_known_flex(data, groups, flex_known = the_flex, flex_test = .))
+        ) %>% 
+        select(-status)
+}
+```
+
+```{r}
+all_other <- dif_all_other(data = sim_1_dif$data, groups = sim_1_dif$groups)
+
+```
+
+```{r}
+dif_iterative_backward(data, groups){
+
+    dif_cb <- dif_all_other(data, groups)
+
+    if (all(dif_cb$p > 0.05)){
+        return(
+            dif_cb %>% 
+                mutate(
+                    status = ifelse(item == which.min(p), "flex", "unknown")
+                )
+        )
+    }
+
+    status <- 
+        dif_cb %>% 
+        mutate(
+            status = ifelse(item == which.min(p), "flex", "unknown")
+        )
+
+    while(TRUE){
+        new_status <- get_next_status(status, sim_1_dif$data, sim_1_dif$groups)
+
+        status <- 
+            status %>% 
+            select(-p) %>% 
+            left_join(new_status, by = "item")
+
+        if(all(status$p > 0.05, na.rm = TRUE)){break}
+
+        status$status[which.min(status$p)] <- "flex"
+    }
+
+    status
+}
+
 ```

### 2020-01-11 R/dif.R

@@ -1,6 +1,0 @@
-dif_constrained_baseline_one_at_a_time <- function(data, groups){
-    tibble(
-        item = 1:ncol(data)
-    ) %>%
-        mutate(p = item %>% map_dbl(~ constrained_baseline_test_flex_items(data, groups, flex_items = .)))
-}

### 2020-01-11 R/dif_all_other.R

@@ -1,0 +1,9 @@
+dif_all_other <- function(data, groups){
+    tibble(
+        item = 1:ncol(data)
+    ) %>%
+        mutate(
+            p = item %>% map_dbl(~ constrained_baseline_test_flex_items(data, groups, flex_items = .)),
+            status = ifelse(p > 0.05, "flex", "unknown")
+        )
+}

### 2020-01-11 R/dif_helpers_level_1.R

@@ -1,6 +1,0 @@
-constrained_baseline_test_flex_items <- function(data, groups, flex_items){
-    quiet(anova(
-        mod_constrained_baseline(data = sim_1_dif$data, groups = sim_1_dif$groups),
-        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = flex_items)
-    ))$p[2]
-}

### 2020-01-11 R/dif_iterative_backward.R

@@ -1,0 +1,46 @@
+get_next_status <- function(status, data, groups){
+    the_flex <- which(status$status == "flex")
+
+    status %>%
+        filter(status == "unknown") %>%
+        mutate(
+            p = item %>%
+                map_dbl(~ test_with_known_flex(data, groups, flex_known = the_flex, flex_test = .))
+        ) %>%
+        select(-status)
+}
+
+dif_iterative_backward <- function(data, groups){
+
+    dif_cb <- dif_all_other(data, groups)
+
+    if (all(dif_cb$p > 0.05)){
+        return(
+            dif_cb %>%
+                mutate(
+                    status = ifelse(item == which.min(p), "flex", "unknown")
+                )
+        )
+    }
+
+    status <-
+        dif_cb %>%
+        mutate(
+            status = ifelse(item == which.min(p), "flex", "unknown")
+        )
+
+    while(TRUE){
+        new_status <- get_next_status(status, sim_1_dif$data, sim_1_dif$groups)
+
+        status <-
+            status %>%
+            select(-p) %>%
+            left_join(new_status, by = "item")
+
+        if(all(status$p > 0.05, na.rm = TRUE)){break}
+
+        status$status[which.min(status$p)] <- "flex"
+    }
+
+    status
+}

### 2020-01-11 R/dif_zelpers_level_1.R

@@ -1,0 +1,13 @@
+constrained_baseline_test_flex_items <- function(data, groups, flex_items){
+    quiet(anova(
+        mod_constrained_baseline(data = sim_1_dif$data, groups = sim_1_dif$groups),
+        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = flex_items)
+    ))$p[2]
+}
+
+test_with_known_flex <- function(data, groups, flex_known, flex_test){
+    quiet(anova(
+        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = flex_known),
+        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = c(flex_known, flex_test))
+    ))$p[2]
+}

### 2020-01-11 work.Rmd
getting functions written is good!

--- 2020-01-10 work.Rmd
@@
         item = row_number(),
-        nuisance_disc = c(0, 0, 0, 0.2, 0, 0, 0, 0)
+        nuisance_disc = c(0, 0, 0, 0.1, 0.1, 0, 0, 0)
     )
 
-sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 10000, n_foc = 10000, mean_foc = -1)
-
-dif_constrained_baseline_one_at_a_time(data = sim_1_dif$data, groups = sim_1_dif$groups)
+sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 50000, n_foc = 50000, mean_foc = -1)
 ```
@@
 ```{r}
-pars_3_dif <-
-    read_csv("data-parameters/kopf2015-8.csv") %>%
-    arrange(desc(easy)) %>%
-    mutate(
-        item = row_number(),
-        nuisance_disc = c(0, 0, 0, 0.5, 0.5, 0.5, 0, 0)
-    )
+all_other <- dif_all_other(sim_1_dif$data, sim_1_dif$groups)
+
+iterative_backward <- dif_iterative_backward(sim_1_dif$data, sim_1_dif$groups, all_other)
 ```

### 2020-01-11 R/dif_all_other.R

@@ -4,6 +4,6 @@ dif_all_other <- function(data, groups){
     ) %>%
         mutate(
             p = item %>% map_dbl(~ constrained_baseline_test_flex_items(data, groups, flex_items = .)),
-            status = ifelse(p > 0.05, "flex", "unknown")
+            status = ifelse(p > 0.05, "unknown", "flex")
         )
 }

### 2020-01-11 R/dif_iterative_backward.R

@@ -10,13 +10,11 @@ get_next_status <- function(status, data, groups){
         select(-status)
 }
 
-dif_iterative_backward <- function(data, groups){
+dif_iterative_backward <- function(data, groups, all_other){
 
-    dif_cb <- dif_all_other(data, groups)
-
-    if (all(dif_cb$p > 0.05)){
+    if (all(all_other$p > 0.05)){
         return(
-            dif_cb %>%
+            all_other %>%
                 mutate(
                     status = ifelse(item == which.min(p), "flex", "unknown")
                 )
@@ -24,7 +22,7 @@ dif_iterative_backward <- function(data, groups){
     }
 
     status <-
-        dif_cb %>%
+        all_other %>%
         mutate(
             status = ifelse(item == which.min(p), "flex", "unknown")
         )

### 2020-01-13 paper/paper.Rmd
work.Rmd works great for summarizing the all other method!

---
output:
  bookdown::pdf_document2:
    includes:
    keep_tex: no
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 2
header-includes: 
- \usepackage{float}
- \usepackage{todonotes}
- \usepackage{caption}
- \usepackage[autostyle, english = american]{csquotes}

fontsize: 11pt
link-citations: yes
spacing: double

title: "Measurement invariance"

subtitle: |
    Ben Stenhaug & Ben Domingue  
    Stanford University

abstract: |
  <!-- https://tinyurl.com/ybremelq -->

bibliography: paper.bib
---

<!-- setup -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    fig.align='center',
    echo=FALSE,
    warning=TRUE,
    fig.pos='H'
)

library(tidyverse)
library(knitr)
library(kableExtra)

theme_set(theme_bw())
```

```{r child = "paper1.Rmd"}
```

# References

### 2020-01-13 paper/paper1.Rmd
work.Rmd works great for summarizing the all other method!

# Hello

but there is no guarantee that it will work for the small sample sizes with which psychometricians often work [@mcdonald1995goodness].

### 2020-01-13 work.Rmd
work.Rmd works great for summarizing the all other method!

@@
 library(mirt)
+set.seed(123)
 R.utils::sourceDirectory("R")
@@
 ```{r}
-pars_1_dif <-
-    read_csv("data-parameters/kopf2015-8.csv") %>%
-    arrange(desc(easy)) %>%
-    mutate(
-        item = row_number(),
-        nuisance_disc = c(0, 0, 0, 0.1, 0.1, 0, 0, 0)
-    )
-
-sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 50000, n_foc = 50000, mean_foc = -1)
+create_pars_1_biased_item <- function(){
+    tibble(
+        item = 1:8,
+        easy = rnorm(8, 0, 1),
+        target_disc = 1,
+        nuisance_disc = c(rep(0.5, 1), rep(0, 7))
+    ) %>%
+    arrange(desc(nuisance_disc), easy)
+}
 ```
@@
 ```{r}
-all_other <- dif_all_other(sim_1_dif$data, sim_1_dif$groups)
+out <- 
+    tibble(
+        pars = rerun(1, create_pars_1_biased_item())
+    ) %>% 
+    mutate(
+        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
+        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
+    ) %>% 
+    mutate(
+        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
+        all_other_mod = 
+            map2(
+                sim, 
+                all_other_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex")
+                )
+            ),
+        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
+    )
 
-iterative_backward <- dif_iterative_backward(sim_1_dif$data, sim_1_dif$groups, all_other)
+out$intuitive_mod[[1]] %>% 
+    mod_intuitive_to_draws_df() %>% 
+    draws_df_to_logit_plot() +
+    geom_point(data = out$all_other_final_dif[[1]], aes(x = dif, y = item), color = "red") +
+    geom_vline(xintercept = filter(out$all_other_final_dif[[1]], anchor)$dif[1], linetype = "dashed", color = "red")
 ```

### 2020-01-13 R-work/multi.R

@@ -1,0 +1,49 @@
+# this look at the idea that maybe i could just fit multi model which
+# is actually pretty hard because of rotation etc unless you pin a few
+# things down - prolly reason its not in the literature
+
+# pars <- tibble(
+#     easy = rnorm(50, 0, 1),
+#     target_disc  = rnorm(50, 1.5, 0.5),
+#     nuisance_disc  = rnorm(50, 1.5, 0.5),
+# )
+
+pars <-
+    read_csv("data-parameters/kopf2015-8.csv") %>%
+    arrange(desc(easy)) %>%
+    mutate(item = row_number())
+
+n_ref <- 200000
+n_foc <- 200000
+
+theta_ref <-
+    tibble(
+        target = rnorm(n_ref, 0, 1),
+        nuisance = rnorm(n_ref, 0, 1)
+    )
+
+theta_foc <-
+    tibble(
+        target = rnorm(n_foc, 0, 1),
+        nuisance = rnorm(n_foc, 0, 1)
+    )
+
+groups <- groups <- c(rep("a_ref", n_ref), rep("b_foc", n_foc))
+
+data <-
+    simdata(
+        a = pars %>% select(ends_with("disc")) %>% as.matrix(),
+        d = pars$easy,
+        itemtype = "2PL",
+        Theta = bind_rows(theta_ref, theta_foc) %>% as.matrix()
+    )
+
+vals <- mirt(data, 2, "2PL", pars = "values")
+
+vals$value[which(vals$name == "a1")][-(4:8)] <- pars$target_disc[-(4:8)]
+
+vals$est[which(vals$name == "a1")][-(4:8)] <- FALSE
+
+mod <- mirt(data, 2, "2PL", pars = vals)
+
+coef(mod, simplify = TRUE)

### 2020-01-13 R/dif_intuitive.R

@@ -1,5 +1,9 @@
-mod_to_draws_df <- function(mod){
-    par_draws <- MASS::mvrnorm(n = 10000, mu = extract.mirt(mod, "parvec"), Sigma = extract.mirt(mod, "vcov"))
+fit_mod_intuitive <- function(data, groups){
+    multipleGroup(data, 1, itemtype = "Rasch", groups, invariance = "free_var", SE = TRUE)
+}
+
+mod_intuitive_to_draws_df <- function(mod_intuitive){
+    par_draws <- MASS::mvrnorm(n = 10000, mu = extract.mirt(mod_intuitive, "parvec"), Sigma = extract.mirt(mod_intuitive, "vcov"))
 
     par_draws <- par_draws[ , str_detect(colnames(par_draws), "d")]
 
@@ -10,7 +14,7 @@ mod_to_draws_df <- function(mod){
     n_items <- ncol(par_draws) / 2
 
     for (i in 1:n_items) {
-        draws_df[[paste0("item", i)]] <- par_draws[ , i] - par_draws[ , i + n_items]
+        draws_df[[paste0("item", i)]] <- par_draws[ , i + n_items] - par_draws[ , i]
     }
 
     draws_df
@@ -23,7 +27,7 @@ draws_df_to_logit_plot <- function(draws_df){
         ggplot(aes(x = val, y = var)) +
         ggridges::geom_density_ridges() +
         labs(
-            x = "Reference group logits - Focal group logits",
+            x = "Focal group logits - Reference group logits",
             y = "",
             subtitle = "Identify by fixing both means to 0 and let difference show up in item easiness"
         )

### 2020-01-13 R/dif_zelpers_level_1.R

@@ -1,13 +1,13 @@
 constrained_baseline_test_flex_items <- function(data, groups, flex_items){
     quiet(anova(
-        mod_constrained_baseline(data = sim_1_dif$data, groups = sim_1_dif$groups),
-        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = flex_items)
+        mod_constrained_baseline(data, groups),
+        mod_flexible(data, groups, flex_items = flex_items)
     ))$p[2]
 }
 
 test_with_known_flex <- function(data, groups, flex_known, flex_test){
     quiet(anova(
-        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = flex_known),
-        mod_flexible(data = sim_1_dif$data, groups = sim_1_dif$groups, flex_items = c(flex_known, flex_test))
+        mod_flexible(data, groups, flex_items = flex_known),
+        mod_flexible(data, groups, flex_items = c(flex_known, flex_test))
     ))$p[2]
 }

### 2020-01-13 R/dif_zelpers_level_3.R

@@ -1,0 +1,10 @@
+mod_to_final_dif <- function(all_other_mod){
+    the_coef <- coef_1f(all_other_mod)
+
+    the_coef$items %>%
+        mutate(
+            item = paste0("item", 1:8),
+            dif = b_foc_easy - a_ref_easy + the_coef$ability$b_foc[1]
+        ) %>%
+        select(item, anchor, dif)
+}

### 2020-01-13 data-documentation/data-documentation.R

@@ -1,8 +1,30 @@
-pars <-
-    read_csv("data-parameters/kopf2015-8.csv") %>%
-    arrange(desc(easy)) %>%
-    mutate(item = row_number())
+library(tidyverse)
+library(mirt)
+set.seed(123)
+R.utils::sourceDirectory("R")
 
-sim <- sim_bias(pars = pars, n_ref = 100000, n_foc = 100000, mean_foc = -1)
+# simulate ----------------------------------------------------------------
+pars <-
+    tibble(
+        item = 1:8,
+        easy = rnorm(8, 0, 1),
+        target_disc = 1,
+        nuisance_disc = c(rep(0.5, 1), rep(0, 7)) # 1 + 7 = 8
+    ) %>%
+    arrange(desc(nuisance_disc), easy)
+
+sim <- sim_bias(pars = pars, n_ref = 10000, n_foc = 10000, mean_foc = -1)
 
 sim %>% write_rds("data-documentation/sim.rds")
+
+# dif ---------------------------------------------------------------------
+
+# intuitive
+mod <- multipleGroup(sim$data, 1, itemtype = "Rasch", sim$groups, invariance = "free_var", SE = TRUE)
+mod %>% mod_to_draws_df() %>% draws_df_to_logit_plot()
+
+# all other
+all_other <- dif_all_other(sim$data, sim$groups)
+
+# iterative backward
+iterative_backward <- dif_iterative_backward(sim$data, sim$groups, all_other)

### 2020-01-13 scratch.R

@@ -1,3 +1,0 @@
-# tried to fit 2-dimensional but it wasn't working
-
-mod <- multipleGroup(out, 2, groups, invariance = c("free_means", "free_var","slopes", "intercepts"), technical = list(NCYCLES = 50))

### 2020-01-14 work.Rmd
end of the day and running overnight simulation

--- 2020-01-11 work.Rmd
@@
+- implement gini!!
+
+- implement the forward SA
+
 ```{r}
@@
 library(mirt)
+library(cluster)
+set.seed(123)
 R.utils::sourceDirectory("R")
@@
 
+# 1
+
 ```{r}
-pars_1_dif <-
-    read_csv("data-parameters/kopf2015-8.csv") %>%
-    arrange(desc(easy)) %>%
+out_1_biased_item <- 
+    tibble(
+        pars = rerun(100, create_pars_1_biased_item()) # swap in pars function here
+    ) %>% 
+    # sim and fit intuitive
     mutate(
-        item = row_number(),
-        nuisance_disc = c(0, 0, 0, 0.1, 0.1, 0, 0, 0)
+        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
+        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
+    ) %>% 
+    # dif: all other
+    mutate(
+        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
+        all_other_mod = 
+            map2(
+                sim, 
+                all_other_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
                 )
-
-sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 50000, n_foc = 50000, mean_foc = -1)
+            ),
+        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
+    ) %>% 
+    # dif: iterative backward
+    mutate(
+        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
+        it_back_mod = 
+            map2(
+                sim, 
+                it_back_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
+    ) %>% 
+    # dif: tibs cluster
+    mutate(
+        cluster = 
+            intuitive_mod %>% 
+            map(mod_to_final_dif) %>% 
+            map(~ mutate(., cluster = irt_cluster(dif)))
+    ) %>% 
+    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>% 
+    mutate(cluster_status = map2(cluster, cluster_stats, left_join, by = "cluster")) %>% 
+    mutate(
+        cluster_mod = 
+            map2(
+                sim, 
+                cluster_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        cluster_final_dif = cluster_mod %>% map(mod_to_final_dif)
+    )
 ```
 
+# 3
+
 ```{r}
-all_other <- dif_all_other(sim_1_dif$data, sim_1_dif$groups)
+out_3_biased_items <- 
+    tibble(
+        pars = rerun(100, create_pars_3_biased_items()) # swap in pars function here
+    ) %>% 
+    # sim and fit intuitive
+    mutate(
+        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
+        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
+    ) %>% 
+    # dif: all other
+    mutate(
+        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
+        all_other_mod = 
+            map2(
+                sim, 
+                all_other_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
+    ) %>% 
+    # dif: iterative backward
+    mutate(
+        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
+        it_back_mod = 
+            map2(
+                sim, 
+                it_back_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
+    ) %>% 
+    # dif: tibs cluster
+    mutate(
+        cluster = 
+            intuitive_mod %>% 
+            map(mod_to_final_dif) %>% 
+            map(~ mutate(., cluster = irt_cluster(dif)))
+    ) %>% 
+    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>% 
+    mutate(cluster_status = map2(cluster, cluster_stats, left_join, by = "cluster")) %>% 
+    mutate(
+        cluster_mod = 
+            map2(
+                sim, 
+                cluster_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        cluster_final_dif = cluster_mod %>% map(mod_to_final_dif)
+    )
+```
 
-iterative_backward <- dif_iterative_backward(sim_1_dif$data, sim_1_dif$groups, all_other)
+# 4
+
+```{r}
+out_4_biased_items_balanced <- 
+    tibble(
+        pars = rerun(100, create_pars_4_biased_items_balanced()) # swap in pars function here
+    ) %>% 
+    # sim and fit intuitive
+    mutate(
+        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
+        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
+    ) %>% 
+    # dif: all other
+    mutate(
+        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
+        all_other_mod = 
+            map2(
+                sim, 
+                all_other_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
+    ) %>% 
+    # dif: iterative backward
+    mutate(
+        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
+        it_back_mod = 
+            map2(
+                sim, 
+                it_back_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
+    ) %>% 
+    # dif: tibs cluster
+    mutate(
+        cluster = 
+            intuitive_mod %>% 
+            map(mod_to_final_dif) %>% 
+            map(~ mutate(., cluster = irt_cluster(dif)))
+    ) %>% 
+    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>% 
+    mutate(cluster_status = map2(cluster, cluster_stats, left_join, by = "cluster")) %>% 
+    mutate(
+        cluster_mod = 
+            map2(
+                sim, 
+                cluster_status, 
+                ~ mod_flexible(
+                    .x$data, 
+                    .x$groups, 
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        cluster_final_dif = cluster_mod %>% map(mod_to_final_dif)
+    )
 ```

### 2020-01-14 R-work/work.R

@@ -1,3 +1,28 @@
+# can i add some sort of sampling error somewhere
+
+out$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    View()
+
+# was playing with clustering for each draw..
+
+mod_intuitive <- out$intuitive_mod[[1]]
+
+
+
+x <- mod %>% mod_to_draws_df(10)
+
+hi <-
+    x %>%
+    gather(item, easy, -run) %>%
+    arrange(run) %>%
+    split(.$run) %>%
+    map(~ irt_cluster(.$easy))
+
+vec <- y$val[1:8]
+
+#
+
 library(tidyverse)
 library(mirt)
 

### 2020-01-14 R/clustering.R

@@ -1,0 +1,32 @@
+# https://blog.dominodatalab.com/clustering-in-r/
+# https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92
+# https://github.com/pimentel/cluster/blob/master/R/clusGap.R
+
+pam1 <- function(x,k) list(cluster = pam(x, k, cluster.only=TRUE))
+
+irt_cluster <- function(vec){
+
+    theGap <- clusGap(matrix(vec), FUNcluster = pam1, K.max = 4)
+
+    n_clust <-
+        maxSE(
+            f = theGap$Tab[ , "gap"],
+            SE.f = theGap$Tab[ , "SE.sim"],
+            method = "Tibs2001SEmax",
+            SE.factor = 1
+        )
+
+    pam1(matrix(vec), n_clust)$cluster
+
+}
+
+get_cluster_stats <- function(cluster){
+    cluster %>%
+        group_by(cluster) %>%
+        summarize(
+            n = n(),
+            sd = sd(dif)
+        ) %>%
+        arrange(desc(n), sd) %>%
+        mutate(status = ifelse(row_number() == 1, "unknown", "flex"))
+}

### 2020-01-14 R/create_pars.R

@@ -1,0 +1,29 @@
+create_pars_1_biased_item <- function(){
+    tibble(
+        item = 1:8,
+        easy = rnorm(8, 0, 1),
+        target_disc = 1,
+        nuisance_disc = c(rep(0.5, 1), rep(0, 7))
+    ) %>%
+        arrange(desc(nuisance_disc), easy)
+}
+
+create_pars_3_biased_items <- function(){
+    tibble(
+        item = 1:8,
+        easy = rnorm(8, 0, 1),
+        target_disc = 1,
+        nuisance_disc = c(rep(0.5, 3), rep(0, 5))
+    ) %>%
+        arrange(desc(nuisance_disc), easy)
+}
+
+create_pars_4_biased_items_balanced <- function(){
+    tibble(
+        item = 1:8,
+        easy = rnorm(8, 0, 1),
+        target_disc = 1,
+        nuisance_disc = c(rep(0.5, 2), rep(-0.5, 2), rep(0, 4))
+    ) %>%
+        arrange(desc(nuisance_disc), easy)
+}

### 2020-01-14 R/dif_intuitive.R

@@ -2,8 +2,8 @@ fit_mod_intuitive <- function(data, groups){
     multipleGroup(data, 1, itemtype = "Rasch", groups, invariance = "free_var", SE = TRUE)
 }
 
-mod_intuitive_to_draws_df <- function(mod_intuitive){
-    par_draws <- MASS::mvrnorm(n = 10000, mu = extract.mirt(mod_intuitive, "parvec"), Sigma = extract.mirt(mod_intuitive, "vcov"))
+mod_intuitive_to_draws_df <- function(mod){
+    par_draws <- MASS::mvrnorm(n = 10000, mu = extract.mirt(mod, "parvec"), Sigma = extract.mirt(mod, "vcov"))
 
     par_draws <- par_draws[ , str_detect(colnames(par_draws), "d")]
 

### 2020-01-14 R/dif_zelpers_level_2.R

@@ -9,13 +9,14 @@ mod_constrained_baseline <- function(data, groups){
     )
 }
 
-mod_flexible <- function(data, groups, flex_items){
+mod_flexible <- function(data, groups, flex_items, se = FALSE){
     multipleGroup(
         data,
         1,
         itemtype = "Rasch",
         groups,
         invariance = c("free_means", "free_var", paste0("Item_", (1:ncol(data))[-flex_items])),
-        technical = list(NCYCLES = 100)
+        technical = list(NCYCLES = 100),
+        SE = se
     )
 }

### 2020-01-14 R/dif_zelpers_level_3.R

@@ -1,10 +1,0 @@
-mod_to_final_dif <- function(all_other_mod){
-    the_coef <- coef_1f(all_other_mod)
-
-    the_coef$items %>%
-        mutate(
-            item = paste0("item", 1:8),
-            dif = b_foc_easy - a_ref_easy + the_coef$ability$b_foc[1]
-        ) %>%
-        select(item, anchor, dif)
-}

### 2020-01-14 R/mod_zelpers.R

@@ -1,0 +1,41 @@
+mod_to_final_dif <- function(all_other_mod){
+    the_coef <- coef_1f(all_other_mod)
+
+    the_coef$items %>%
+        mutate(
+            item = paste0("item", 1:8),
+            dif = b_foc_easy - a_ref_easy + the_coef$ability$b_foc[1]
+        ) %>%
+        select(item, anchor, dif)
+}
+
+# this works for a general model where some are fixed (as opposed to mod intuitive where all flex)
+mod_to_draws_df <- function(mod, n){
+    par_draws <- MASS::mvrnorm(n, mu = extract.mirt(mod, "parvec"), Sigma = extract.mirt(mod, "vcov"))
+
+    par_draws <-
+        par_draws[ , str_detect(colnames(par_draws), "d")] %>%
+        as_tibble() %>%
+        mutate(id = row_number()) %>%
+        gather(var, val, -id) %>%
+        mutate(var = str_remove(var, "d\\.")) %>%
+        separate(var, c("one", "two"), "\\.", fill = "right") %>%
+        gather(var, item, -val, -id) %>%
+        na.omit() %>%
+        select(-var) %>%
+        mutate(item = as.numeric(item)) %>%
+        spread(item, val) %>%
+        select(-id)
+
+    draws_df <- tibble(run = 1:nrow(par_draws))
+
+    stopifnot(ncol(par_draws) %% 2 == 0)
+
+    n_items <- ncol(par_draws) / 2
+
+    for (i in 1:n_items) {
+        draws_df[[paste0("item", i)]] <- par_draws[[i + n_items]] - par_draws[[i]]
+    }
+
+    draws_df
+}

### 2020-01-14 work_graph.R

@@ -1,0 +1,7 @@
+out$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    draws_df_to_logit_plot() +
+    geom_point(data = out$all_other_final_dif[[1]], aes(x = dif, y = item), color = "red", shape = 1) +
+    geom_vline(xintercept = filter(out$all_other_final_dif[[1]], anchor)$dif[1], linetype = "dashed", color = "red") +
+    geom_point(data = out$it_back_final_dif[[1]], aes(x = dif, y = item), color = "blue", shape = 3) +
+    geom_vline(xintercept = filter(out$it_back_final_dif[[1]], anchor)$dif[1], linetype = "dotted", color = "blue")

### 2020-01-16 paper/paper.Rmd
great day of progress

@@
     number_sections: yes
-    toc: yes
+    toc: no
     toc_depth: 2
@@
 - \usepackage{float}
-- \usepackage{todonotes}
-- \usepackage{caption}
 - \usepackage[autostyle, english = american]{csquotes}
+- \setlength{\parindent}{4em}
+- \setlength{\parskip}{4em}
+- \renewcommand{\baselinestretch}{1.5}
 
@@
 link-citations: yes
-spacing: double
+spacing: single
 
@@
 
+# Intro
+
 ```{r child = "paper1.Rmd"}

### 2020-01-16 paper/paper1.Rmd
great day of progress

@@
-# Hello
+# Intro
 
-but there is no guarantee that it will work for the small sample sizes with which psychometricians often work [@mcdonald1995goodness].
+Inspired by @camilli1992conceptual, we view differential item functioning (DIF) as bias towards a group of students that manifests when we impose an item response model with too few ability dimensions.
+From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item.
+For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability but performance also depends on their verbal ability.
+In this case, math ability is the "target ability" and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test will result in students with low verbal ability being given a score systematically lower than their true math ability; this is DIF.
+
+Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item.
+For example, @kopf2015framework simulate students as belonging to either the reference or the focal group.
+The item difficulties for the reference group, ${b_j}^{\text{ref}}$, are fixed.
+The item difficulties for the focal group are ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits.
+Student ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(0,1)$ for students in the focal group.
+The item responses are generated according to the Rasch model.
+The Rasch model specifies that the probability of student $i$ responding correctly to item $j$ is
+$$
+P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
+$$
+where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.
+
+We find it more informative to describe simulation conditions in terms of single dimensional items and multidimensional abilities.
+For example, the @kopf2015framework simulation study can be translated from the DIF-as-item-property view to the DIF-as-student-property view.
+Item difficulty is set to what was previously ${b_j}^{\text{ref}}$ for all students.
+Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
+The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(0,1)$ for students in the focal group.
+The nuisance ability is set to $\eta_i^{\text{ref}} = 0$ for the reference group and $\eta_i^{\text{ref}} = -1$ for the focal group.
+We now need to use a 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF and $a_{2j} = 0$ otherwise.
+Again, 0.6 is the magnitude of DIF in logits.
+According to the two-dimensional 2PL model, the probability student $i$ responding correctly to item $j$ is
+
+$$
+P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
+$$
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students.
+
+If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model?
+@camilli1992conceptual tested this idea with the goal of a "more satisfying description of the secondary abilities." They found that the rotational indetermincy of item response models is difficult to overcome and concluded that "a priori knowledge of the true factor structure" is necessary.
+It's hard to imagine how one would have such knowledge.
+Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias that manifests in the item parameters.
+
+Psychometricians have long been in search of the perfect DIF detection method.
+@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method where each item is independently tested for DIF using all other items as anchors.
+More specifically, testing is done using a nested model comparison between the baseline model where all item parameters are fixed across groups and the flexible model where the parameters of the tested item are freed across groups.
+For example, @edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
+One might wonder how we can be sure its those 10 items and not the other 11 items with DIF.
+Implicit in the use of AOAA is the assumption.
+Most applications of DIF and many DIF simulation studies do not make explicit the assumptions of the DIF methods.
+STroble 2018 lists some of these assumptions.
+For AOAA the assumption is that.
+It's easy to show that AOAA is imperfect: Imagine a test with 3 items.
+Each item will test positive for DIF.
+a phenomenon that went indirectly address in the literature until andrich at all named it artificial DIF.
+
+We advocate for methods that present information to the analyst where the analyst makes decisions as opposed to methods which automate the detection and subsequently the removal of items with DIF as is frequently the case (e.g.
+in PISA).
+The most obvious method is to at first punt on disentangling between ability and bias and instead set both group means to zero and estimate the item parameters separately for each group.
+The difference in item parameters, then, includes both the difference in ability and bias.
+To our knowledge, this method has not been suggested in the literature with the exception of pohl who uses it in a pedagogical example and tashjkdf who uses it pre to post.
+For example, consider a didactic scenario in which an eight item exam where the focal group and reference group.
+Where the difference in ability between groups is one logit.
+And additionally, the first two items have a one logit bias against the students who.
+We fit a model then fix the means.
+We then sample as leah describes (look at how she describes it) then we get this plot.
+It cannot be said strongly enough that this plot contains all of the information about the combination of ability and bias that is available in the data.
+It is then up to the analyst and/or method to choos what's known anchor point, which is essentially the difference in abilities with remaining differences being attributed to bias.

### 2020-01-16 paper/paper.Rmd
starting to look like a paper

--- 2020-01-13 paper/paper.Rmd
@@
     includes:
-    keep_tex: no
+    keep_tex: yes
     latex_engine: xelatex
@@
 - \usepackage{float}
-- \usepackage{todonotes}
-- \usepackage{caption}
 - \usepackage[autostyle, english = american]{csquotes}
-
+- \setlength{\parindent}{4em}
+- \setlength{\parskip}{2em}
+- \usepackage{setspace}\doublespacing
 fontsize: 11pt
 link-citations: yes
-spacing: double
-
 title: "Measurement invariance"
-
 subtitle: |
@@
     Stanford University
-
 abstract: |
+  ABSTRACT TO DO
   <!-- https://tinyurl.com/ybremelq -->
-
 bibliography: paper.bib
@@
 
-<!-- setup -->
+```{r setup, include = FALSE}
+# 
+# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
 
-```{r setup, include = FALSE}
 knitr::opts_chunk$set(
@@
     warning=TRUE,
-    fig.pos='H'
+    fig.pos='H',
+    fig.retina = 5
 )
@@
 library(tidyverse)
+library(mirt)
 library(knitr)
 library(kableExtra)
+set.seed(123)
+R.utils::sourceDirectory(here::here("R"))
 
@@
 
-```{r child = "paper1.Rmd"}
+\clearpage
+
+# Introduction
+
+```{r child = "paper_1_intro.Rmd"}
+```
+
+# DIF methods
+
+```{r child = "paper_2_dif_methods.Rmd"}
 ```
@@
 # References
+
+<!-- weird formatting, solution is to use papaja (look at paper_apa.Rmd) -->

### 2020-01-16 paper/paper.tex
starting to look like a paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Measurement invariance},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{float}
\usepackage[autostyle, english = american]{csquotes}
\setlength{\parindent}{4em}
\setlength{\parskip}{2em}
\usepackage{setspace}\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Measurement invariance}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Ben Stenhaug \& Ben Domingue\\
Stanford University}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle
\begin{abstract}
ABSTRACT TO DO
\end{abstract}

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\clearpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we view differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed. From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is a property of the student, not the item. For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability. In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies the DIF.

Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group. They fix the item difficulties for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study. They set item difficulties for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits. They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group and generate item responses according to the Rasch model. The Rasch model specifies that the probability of student \(i\) responding correctly to item \(j\) is
\begin{align}
    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
\end{align}
where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.

We find it more informative to describe simulation conditions in terms of single-dimensional items and multidimensional abilities. For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item difficulty is set to what was previously \({b_j}^{\text{ref}}\) for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group. We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF and \(a_{2j} = 0\) otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional 2PL model, the probability student \(i\) responding correctly to item \(j\) is
\begin{align}
    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
\end{align}
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students in the same group.

If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model? Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}, 144) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities.} He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary. It's hard to imagine how one would have such knowledge. Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.

\hypertarget{dif-methods}{%
\section{DIF methods}\label{dif-methods}}

Psychometricians have long been in search of the perfect DIF detection method. Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method, where each item is tested one at a time for DIF using all other items as anchors. More specifically, testing is done using a nested model comparison between the baseline model, where all item parameters are fixed across groups, and the flexible model, where the parameters of the tested item are freed across groups. Anchor items have fixed parameters across groups and consequently are used to estimate the ability difference between groups. For example, Edelen et al. (\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF. How can we be sure its those 10 items and not the other 11 items with DIF? Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible. Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method Strobl et al. (\protect\hyperlink{ref-strobl2018anchor}{2018}). In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.

Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}). A simple thought experiment illustrates how AOAA fails: Imagine a test with three items where the first item has DIF, and the other two do not. Using AOAA, all items test positive for DIF. The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated. This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until (Andrich and Hagquist \protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}

One way to attempt to counter artificial DIF is with purification with iterative anchor selection. For example, (Drasgow \protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then continued iteratively to test the remaining items for DIF until no more items tested positively. (Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free. Iterative-backward-AOAA might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage? With a sufficient sample size, this will necessarily be the case. Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. Woods suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items display DIF themselves.

We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge, has not previously been proposed. AOAA-OAT starts with AOAA, and the single item exhibiting the most DIF is removed from the anchor set. Each item in the anchor set is then tested for DIF again (with items outside of the anchor set allowed to have free parameters across groups in both the baseline and flexible model). This process continues until no new items display DIF. Applying AOAA-OAT to our thought experiment demonstrates its logic. The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF. In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion. AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.

DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human. In some cases, it might make sense to present information to the analyst in a way that empowers them to make decisions. Our preferred technique, which we refer to as the \enquote{equal means logit graph} (EMLG), is to set both groups means to 0 so that differences in performance to manifest in the item parameters. Relatedly, Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) fit a model with both groups means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

We use a simple simulation to demonstrate EMLG. 10,000 reference group students and 10,000 focal group students take an eight-item test. Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group. The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise. We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data. We estimate the item parameter covariance matrix using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}). Multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of the difference in difficulty, \({b_j}^{\text{foc}} - {b_j}^{\text{ref}}\), for each item. Figure \ref{fig:emlg} shows these distributions.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emlg-1}

}

\caption{Equal means logit graph (EMLG)}\label{fig:emlg}
\end{figure}

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-ackerman1992didactic}{}%
Ackerman, Terry A. 1992. ``A Didactic Explanation of Item Bias, Item Impact, and Item Validity from a Multidimensional Perspective.'' \emph{Journal of Educational Measurement} 29 (1): 67--91.

\leavevmode\hypertarget{ref-andrich2012real}{}%
Andrich, David, and Curt Hagquist. 2012. ``Real and Artificial Differential Item Functioning.'' \emph{Journal of Educational and Behavioral Statistics} 37 (3): 387--416.

\leavevmode\hypertarget{ref-camilli1992conceptual}{}%
Camilli, Gregory. 1992. ``A Conceptual Analysis of Differential Item Functioning in Terms of a Multidimensional Item Response Model.'' \emph{Applied Psychological Measurement} 16 (2): 129--47.

\leavevmode\hypertarget{ref-chalmers2018numerical}{}%
Chalmers, R Philip. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.

\leavevmode\hypertarget{ref-drasgow1987study}{}%
Drasgow, Fritz. 1987. ``Study of the Measurement Bias of Two Standardized Psychological Tests.'' \emph{Journal of Applied Psychology} 72 (1): 19.

\leavevmode\hypertarget{ref-edelen2006identification}{}%
Edelen, Maria Orlando, David Thissen, Jeanne A Teresi, Marjorie Kleinman, and Katja Ocepek-Welikson. 2006. ``Identification of Differential Item Functioning Using Item Response Theory and the Likelihood-Based Model Comparison Approach: Application to the Mini-Mental State Examination.'' \emph{Medical Care}, S134--S142.

\leavevmode\hypertarget{ref-kopf2015framework}{}%
Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.

\leavevmode\hypertarget{ref-meade2012solving}{}%
Meade, Adam W, and Natalie A Wright. 2012. ``Solving the Measurement Invariance Anchor Item Problem in Item Response Theory.'' \emph{Journal of Applied Psychology} 97 (5): 1016.

\leavevmode\hypertarget{ref-pohl2017cluster}{}%
Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``CluStEr-baSEd anCHor Item Identification and Selection.'' NEPS Working Paper.

\leavevmode\hypertarget{ref-stark2006detecting}{}%
Stark, Stephen, Oleksandr S Chernyshenko, and Fritz Drasgow. 2006. ``Detecting Differential Item Functioning with Confirmatory Factor Analysis and Item Response Theory: Toward a Unified Strategy.'' \emph{Journal of Applied Psychology} 91 (6): 1292.

\leavevmode\hypertarget{ref-strobl2018anchor}{}%
Strobl, Carolin, Julia Kopf, Raphael Hartmann, and Achim Zeileis. 2018. ``Anchor Point Selection: An Approach for Anchoring Without Anchor Items.'' Working Papers in Economics; Statistics.

\leavevmode\hypertarget{ref-talbot2013taking}{}%
Talbot III, Robert M. 2013. ``Taking an Item-Level Approach to Measuring Change with the Force and Motion Conceptual Evaluation: An Application of Item Response Theory.'' \emph{School Science and Mathematics} 113 (7): 356--65.

\leavevmode\hypertarget{ref-yang2012characterizing}{}%
Yang, Ji Seung, Mark Hansen, and Li Cai. 2012. ``Characterizing Sources of Uncertainty in Item Response Theory Scale Scores.'' \emph{Educational and Psychological Measurement} 72 (2): 264--90.

\end{document}

### 2020-01-16 paper/paper_1_intro.Rmd
starting to look like a paper

Inspired by @camilli1992conceptual, we view differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed. From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item. For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability. In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies the DIF.

Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, @kopf2015framework simulate students as belonging to either a reference or focal group. They fix the item difficulties for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study. They set item difficulties for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits. They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group and generate item responses according to the Rasch model. The Rasch model specifies that the probability of student $i$ responding correctly to item $j$ is
\begin{align}
    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
\end{align}
where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.

We find it more informative to describe simulation conditions in terms of single-dimensional items and multidimensional abilities. For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item difficulty is set to what was previously ${b_j}^{\text{ref}}$ for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group. We now need to use a 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF and $a_{2j} = 0$ otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional 2PL model, the probability student $i$ responding correctly to item $j$ is
\begin{align}
    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
\end{align}
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students in the same group.

If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model? @camilli1992conceptual [p. 144] tested this idea with the goal of a "more satisfying description of the secondary abilities." He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that "a priori knowledge of the true factor structure" is necessary. It's hard to imagine how one would have such knowledge. Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.

### 2020-01-16 paper/paper_2_dif_methods.Rmd
starting to look like a paper

--- 2020-01-16 paper/paper.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-  11pt,
-]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Measurement invariance},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-\usepackage{float}
-\usepackage[autostyle, english = american]{csquotes}
-\setlength{\parindent}{4em}
-\setlength{\parskip}{2em}
-\usepackage{setspace}\doublespacing
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\title{Measurement invariance}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Ben Stenhaug \& Ben Domingue\\
-Stanford University}
-\author{}
-\date{\vspace{-2.5em}}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-ABSTRACT TO DO
-\end{abstract}
-
-{
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-\clearpage
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we view differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
-From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is a property of the student, not the item.
-For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies the DIF.
-
-Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item.
-For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item difficulties for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
-They set item difficulties for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group and generate item responses according to the Rasch model.
-The Rasch model specifies that the probability of student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
-\end{align}
-where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
-
-We find it more informative to describe simulation conditions in terms of single-dimensional items and multidimensional abilities.
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item difficulty is set to what was previously \({b_j}^{\text{ref}}\) for all students.
-Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF and \(a_{2j} = 0\) otherwise.
-Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional 2PL model, the probability student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
-\end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students in the same group.
-
-If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model?
-Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}, 144) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities.} He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary.
-It's hard to imagine how one would have such knowledge.
-Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
-
-\hypertarget{dif-methods}{%
-\section{DIF methods}\label{dif-methods}}
-
 Psychometricians have long been in search of the perfect DIF detection method.
-Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method, where each item is tested one at a time for DIF using all other items as anchors.
+@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method, where each item is tested one at a time for DIF using all other items as anchors.
 More specifically, testing is done using a nested model comparison between the baseline model, where all item parameters are fixed across groups, and the flexible model, where the parameters of the tested item are freed across groups.
 Anchor items have fixed parameters across groups and consequently are used to estimate the ability difference between groups.
-For example, Edelen et al.
-(\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
+For example, @edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
 How can we be sure its those 10 items and not the other 11 items with DIF?
 Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}).
+Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method @strobl2018anchor.
 In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
 
-Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
+Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies [@stark2006detecting].
 A simple thought experiment illustrates how AOAA fails: Imagine a test with three items where the first item has DIF, and the other two do not.
@@
 The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until (Andrich and Hagquist \protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until [@andrich2012real] coined the term "artificial DIF."
 
 One way to attempt to counter artificial DIF is with purification with iterative anchor selection.
-For example, (Drasgow \protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then continued iteratively to test the remaining items for DIF until no more items tested positively.
-(Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
+For example, [@drasgow1987study] started with AOAA, removed items displaying DIF from the anchor set, then continued iteratively to test the remaining items for DIF until no more items tested positively.
+[@kopf2015framework] named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
 Iterative-backward-AOAA might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
 With a sufficient sample size, this will necessarily be the case.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
+@kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
 Woods suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
@@
 AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
+<!-- the two kopf 2015 papers love the iterative forward SA. I bet AOAA-OAT is as good or better while fitting fewer models -->
 
@@
 In some cases, it might make sense to present information to the analyst in a way that empowers them to make decisions.
-Our preferred technique, which we refer to as the \enquote{equal means logit graph} (EMLG), is to set both groups means to 0 so that differences in performance to manifest in the item parameters.
-Relatedly, Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) fit a model with both groups means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
+Our preferred technique, which we refer to as the "equal means logit graph" (EMLG), is to set both groups means to 0 so that differences in performance to manifest in the item parameters.
+Relatedly, @pohl2017cluster fit a model with both groups means set to 0 in a pedagogical example, and @talbot2013taking fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
 
@@
 10,000 reference group students and 10,000 focal group students take an eight-item test.
-Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
+Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
+Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group.
+The slope on target ability is set to $a_{1j} = 1$ for all items, and the slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise.
 We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data.
-We estimate the item parameter covariance matrix using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of the difference in difficulty, \({b_j}^{\text{foc}} - {b_j}^{\text{ref}}\), for each item.
+We estimate the item parameter covariance matrix using Oakes' identity [@chalmers2018numerical].
+Multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of the difference in difficulty, ${b_j}^{\text{foc}} - {b_j}^{\text{ref}}$, for each item.
 Figure \ref{fig:emlg} shows these distributions.
 
-\begin{figure}[H]
+```{r emlg, fig.cap = 'Equal means logit graph (EMLG)', out.width="70%", warning = FALSE, message = FALSE}
+pars <-
+    tibble(
+        item = 1:8,
+        easy = rnorm(8, 0, 1),
+        target_disc = 1,
+        nuisance_disc = c(rep(0, 5), rep(0.5, 3)) # 1 + 7 = 8
+    )
 
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emlg-1}
+sim <- sim_bias(pars = pars, n_ref = 10000, n_foc = 10000, mean_foc = -1)
 
-}
+multipleGroup(
+    sim$data, 1, itemtype = "Rasch", sim$groups, 
+    invariance = "free_var", SE = TRUE, verbose = FALSE
+) %>% 
+    mod_intuitive_to_draws_df() %>% 
+    mutate_at(vars(starts_with("item")), function(x) -x) %>% 
+    draws_df_to_logit_plot() +
+    labs(
+        x = latex2exp::TeX("$b_j^{foc} - b_j^{ref}$ (group means fixed to 0)")
+    )
 
-\caption{Equal means logit graph (EMLG)}\label{fig:emlg}
-\end{figure}
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ackerman1992didactic}{}%
-Ackerman, Terry A. 1992. ``A Didactic Explanation of Item Bias, Item Impact, and Item Validity from a Multidimensional Perspective.'' \emph{Journal of Educational Measurement} 29 (1): 67--91.
-
-\leavevmode\hypertarget{ref-andrich2012real}{}%
-Andrich, David, and Curt Hagquist. 2012. ``Real and Artificial Differential Item Functioning.'' \emph{Journal of Educational and Behavioral Statistics} 37 (3): 387--416.
-
-\leavevmode\hypertarget{ref-camilli1992conceptual}{}%
-Camilli, Gregory. 1992. ``A Conceptual Analysis of Differential Item Functioning in Terms of a Multidimensional Item Response Model.'' \emph{Applied Psychological Measurement} 16 (2): 129--47.
-
-\leavevmode\hypertarget{ref-chalmers2018numerical}{}%
-Chalmers, R Philip. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
-
-\leavevmode\hypertarget{ref-drasgow1987study}{}%
-Drasgow, Fritz. 1987. ``Study of the Measurement Bias of Two Standardized Psychological Tests.'' \emph{Journal of Applied Psychology} 72 (1): 19.
-
-\leavevmode\hypertarget{ref-edelen2006identification}{}%
-Edelen, Maria Orlando, David Thissen, Jeanne A Teresi, Marjorie Kleinman, and Katja Ocepek-Welikson.
-2006.
-``Identification of Differential Item Functioning Using Item Response Theory and the Likelihood-Based Model Comparison Approach: Application to the Mini-Mental State Examination.'' \emph{Medical Care}, S134--S142.
-
-\leavevmode\hypertarget{ref-kopf2015framework}{}%
-Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.
-
-\leavevmode\hypertarget{ref-meade2012solving}{}%
-Meade, Adam W, and Natalie A Wright. 2012. ``Solving the Measurement Invariance Anchor Item Problem in Item Response Theory.'' \emph{Journal of Applied Psychology} 97 (5): 1016.
-
-\leavevmode\hypertarget{ref-pohl2017cluster}{}%
-Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``CluStEr-baSEd anCHor Item Identification and Selection.'' NEPS Working Paper.
-
-\leavevmode\hypertarget{ref-stark2006detecting}{}%
-Stark, Stephen, Oleksandr S Chernyshenko, and Fritz Drasgow.
-2006.
-``Detecting Differential Item Functioning with Confirmatory Factor Analysis and Item Response Theory: Toward a Unified Strategy.'' \emph{Journal of Applied Psychology} 91 (6): 1292.
-
-\leavevmode\hypertarget{ref-strobl2018anchor}{}%
-Strobl, Carolin, Julia Kopf, Raphael Hartmann, and Achim Zeileis. 2018. ``Anchor Point Selection: An Approach for Anchoring Without Anchor Items.'' Working Papers in Economics; Statistics.
-
-\leavevmode\hypertarget{ref-talbot2013taking}{}%
-Talbot III, Robert M.
-2013.
-``Taking an Item-Level Approach to Measuring Change with the Force and Motion Conceptual Evaluation: An Application of Item Response Theory.'' \emph{School Science and Mathematics} 113 (7): 356--65.
-
-\leavevmode\hypertarget{ref-yang2012characterizing}{}%
-Yang, Ji Seung, Mark Hansen, and Li Cai. 2012. ``Characterizing Sources of Uncertainty in Item Response Theory Scale Scores.'' \emph{Educational and Psychological Measurement} 72 (2): 264--90.
-
-\end{document}
+# labeling good enough for now maybe check here to fiddle more:
+# https://stackoverflow.com/questions/8190087/italic-greek-letters-latex-style-math-in-plot-titles
+```

### 2020-01-16 R/dif_intuitive.R

@@ -26,9 +26,5 @@ draws_df_to_logit_plot <- function(draws_df){
         gather(var, val) %>%
         ggplot(aes(x = val, y = var)) +
         ggridges::geom_density_ridges() +
-        labs(
-            x = "Focal group logits - Reference group logits",
-            y = "",
-            subtitle = "Identify by fixing both means to 0 and let difference show up in item easiness"
-        )
+        labs(x = "", y = "")
 }

### 2020-01-16 data-documentation/data-documentation.R

@@ -9,7 +9,7 @@ pars <-
         item = 1:8,
         easy = rnorm(8, 0, 1),
         target_disc = 1,
-        nuisance_disc = c(rep(0.5, 1), rep(0, 7)) # 1 + 7 = 8
+        nuisance_disc = c(rep(0.5, 3), rep(0, 5)) # 1 + 7 = 8
     ) %>%
     arrange(desc(nuisance_disc), easy)
 
@@ -21,7 +21,7 @@ sim %>% write_rds("data-documentation/sim.rds")
 
 # intuitive
 mod <- multipleGroup(sim$data, 1, itemtype = "Rasch", sim$groups, invariance = "free_var", SE = TRUE)
-mod %>% mod_to_draws_df() %>% draws_df_to_logit_plot()
+mod %>% mod_intuitive_to_draws_df() %>% draws_df_to_logit_plot()
 
 # all other
 all_other <- dif_all_other(sim$data, sim$groups)

### 2020-01-22 paper/paper.Rmd
after adding AOAA table and before thinking through a Chalmers-like statistic

@@
 library(mirt)
+library(cluster)
 library(knitr)
@@
 
-<!-- weird formatting, solution is to use papaja (look at paper_apa.Rmd) -->
+<!-- weird formatting, solution is to use papaja (look at paper_apa.Rmd) or this might also be helpful https://stackoverflow.com/questions/4073674/how-to-disable-indentation-in-particular-section-in-latex-->

### 2020-01-22 paper/paper.tex
after adding AOAA table and before thinking through a Chalmers-like statistic

@@
 Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
-A simple thought experiment illustrates how AOAA fails: Imagine a test with three items where the first item has DIF, and the other two do not.
+A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not.
 Using AOAA, all items test positive for DIF.
@@
 (Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
-Iterative-backward-AOAA might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
+We find this name confusing and instead refer to this method as all-others-as-anchors-all-significiant (AOAA-AS) where appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF are removed from the anchor set.
+AOAA-AS might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
 With a sufficient sample size, this will necessarily be the case.
@@
 
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge, has not previously been proposed.
+We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAAT), which, to our knowledge (and surprise), has not previously been proposed.
 AOAA-OAT starts with AOAA, and the single item exhibiting the most DIF is removed from the anchor set.
@@
 This process continues until no new items display DIF.
+AOAA-OAAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAAT takes the more conservative approach of removing only the item exhibiting the most DIF.
 Applying AOAA-OAT to our thought experiment demonstrates its logic.
@@
 AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
+Table X summarizes.
+
+\begin{center}
+ \begin{tabular}{||c c c c c||} 
+ \hline
+ Name & Abbreviation & Description & Assumptions & Literature \\ [0.5ex] 
+ \hline\hline
+ All-others-as-anchors & AOAA & 87837 & 787 & 1\\ 
+ \hline
+ All-others-as-anchors-all-significant & AOAA-AS & 78 & 5415 & 1\\
+ \hline
+ All-others-as-anchors-one-at-a-time & AOAA-OAAT & 778 & 7507 & 1\\
+ \hline
+ 4 & 545 & 18744 & 7560 & 1\\
+ \hline
+ 5 & 88 & 788 & 6344 & 1\\ [1ex] 
+ \hline
+ \end{tabular}
+\end{center}
 
@@
 \end{figure}
+
+It cannot be stated strongly enough that Figure \ref{fig:emlg} contains all possible information about the difference in group performance.
+The challenge, then, is to identify the difference in group ability.
+The analyst might assume that, because there are five items showing a difference of about 1 logit and only three items showing a difference of 1.5 logits, the group ability difference is 1 logit and items 6, 7, and 8 contain approximately 0.5 logit bias against the focal group.
+The implicit assumption they are making is that it is the minority of items that contain DIF.
+
+DIF methods use an algorith with its own set of assumptions to identify the group ability difference.
+
+The beauty of the EMLG graph is it makes clear all information instead of allowing the analyst the false sense of security of running an algorithm and wholeheartedlyr believing the output.
+
+We know that this test contains DIF.
+Which items depends on ability difference which we can be sure of.
+We can, however, quantify the amount of DIF on this test.
+We can take the SD or we can assume between and calculate chalmers statistic for each.
+We might not know for sure the ability difference but we can quantify the potential for DIF.
+For each we calculate the SD across item parameters.
 

### 2020-01-22 paper/paper_2_dif_methods.Rmd
after adding AOAA table and before thinking through a Chalmers-like statistic

@@
 Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies [@stark2006detecting].
-A simple thought experiment illustrates how AOAA fails: Imagine a test with three items where the first item has DIF, and the other two do not.
+A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not.
 Using AOAA, all items test positive for DIF.
 The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until [@andrich2012real] coined the term "artificial DIF."
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF."
 
@@
 [@kopf2015framework] named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
-Iterative-backward-AOAA might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
+We find this name confusing and instead refer to this method as all-others-as-anchors-all-significiant (AOAA-AS) where appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF are removed from the anchor set.
+AOAA-AS might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
 With a sufficient sample size, this will necessarily be the case.
@@
 AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
+Table X summarizes.
 <!-- the two kopf 2015 papers love the iterative forward SA. I bet AOAA-OAT is as good or better while fitting fewer models -->
+
+\begin{center}
+ \begin{tabular}{||c c c c c||} 
+ \hline
+ Name & Abbreviation & Description & Assumptions & Literature \\ [0.5ex] 
+ \hline\hline
+ All-others-as-anchors & AOAA & 87837 & 787 & 1\\ 
+ \hline
+ All-others-as-anchors-all-significant & AOAA-AS & 78 & 5415 & 1\\
+ \hline
+ All-others-as-anchors-one-at-a-time & AOAA-OAAT & 778 & 7507 & 1\\
+ \hline
+ 4 & 545 & 18744 & 7560 & 1\\
+ \hline
+ 5 & 88 & 788 & 6344 & 1\\ [1ex] 
+ \hline
+ \end{tabular}
+\end{center}
 
@@
 ```{r emlg, fig.cap = 'Equal means logit graph (EMLG)', out.width="70%", warning = FALSE, message = FALSE}
-pars <-
-    tibble(
-        item = 1:8,
-        easy = rnorm(8, 0, 1),
-        target_disc = 1,
-        nuisance_disc = c(rep(0, 5), rep(0.5, 3)) # 1 + 7 = 8
-    )
+source("paper_2_sim.R")
 
-sim <- sim_bias(pars = pars, n_ref = 10000, n_foc = 10000, mean_foc = -1)
-
-multipleGroup(
-    sim$data, 1, itemtype = "Rasch", sim$groups, 
-    invariance = "free_var", SE = TRUE, verbose = FALSE
-) %>% 
+out_3_biased_items$intuitive_mod[[1]] %>% 
     mod_intuitive_to_draws_df() %>% 
@@
 ```
+
+It cannot be stated strongly enough that Figure \ref{fig:emlg} contains all possible information about the difference in group performance.
+The challenge, then, is to identify the difference in group ability.
+The analyst might assume that, because there are five items showing a difference of about 1 logit and only three items showing a difference of 1.5 logits, the group ability difference is 1 logit and items 6, 7, and 8 contain approximately 0.5 logit bias against the focal group.
+The implicit assumption they are making is that it is the minority of items that contain DIF.
+
+DIF methods use an algorith with its own set of assumptions to identify the group ability difference.
+
+The beauty of the EMLG graph is it makes clear all information instead of allowing the analyst the false sense of security of running an algorithm and wholeheartedlyr believing the output.
+
+We know that this test contains DIF.
+Which items depends on ability difference which we can be sure of.
+We can, however, quantify the amount of DIF on this test.
+We can take the SD or we can assume between and calculate chalmers statistic for each.
+We might not know for sure the ability difference but we can quantify the potential for DIF.
+For each we calculate the SD across item parameters.

### 2020-01-22 R/dif_intuitive.R

@@ -1,5 +1,5 @@
 fit_mod_intuitive <- function(data, groups){
-    multipleGroup(data, 1, itemtype = "Rasch", groups, invariance = "free_var", SE = TRUE)
+    multipleGroup(data, 1, itemtype = "Rasch", groups, invariance = "free_var", SE = TRUE, verbose = FALSE)
 }
 
 mod_intuitive_to_draws_df <- function(mod){

### 2020-01-22 R/dif_zelpers_level_2.R

@@ -5,7 +5,8 @@ mod_constrained_baseline <- function(data, groups){
         itemtype = "Rasch",
         groups,
         invariance = c("free_means", "free_var", "intercepts", "slopes"),
-        technical = list(NCYCLES = 100)
+        technical = list(NCYCLES = 100),
+        verbose = FALSE
     )
 }
 
@@ -17,6 +18,7 @@ mod_flexible <- function(data, groups, flex_items, se = FALSE){
         groups,
         invariance = c("free_means", "free_var", paste0("Item_", (1:ncol(data))[-flex_items])),
         technical = list(NCYCLES = 100),
-        SE = se
+        SE = se,
+        verbose = FALSE
     )
 }

### 2020-01-22 gini_l1.R

@@ -1,0 +1,32 @@
+a <- c(0, 0, 1, 1)
+b <- c(0, 0, 0, 0)
+
+a <- c(0, 0, 0, 1, 1)
+b <- c(0, 0, 0, 0, 0)
+
+a <- c(0.1, 0.2, 0.3, 1, 1)
+b <- c(0, 0, 0, 0, 0)
+
+library(tidyverse)
+library(DescTools)
+
+tibble(
+    p = seq(0, 1, 0.05)
+) %>%
+    mutate(
+        gini = p %>% map_dbl(~ Gini(abs(a - b - .))),
+        #llog = p %>% map_dbl(~ mean(log(abs(a - b - .)))),
+        #l12 = p %>% map_dbl(~ mean(abs((a - b - .))^1/2)),
+        l1 = p %>% map_dbl(~ mean(abs((a - b - .))^1)),
+        #l2 = p %>% map_dbl(~ mean(abs((a - b - .))^2)),
+        #l3 = p %>% map_dbl(~ mean(abs((a - b - .))^3))
+    ) %>%
+    gather(var, val, -p) %>%
+    ggplot(aes(x = p, y = val, color = var)) +
+    geom_path()
+
+mean((values - 0.5)^2)
+
+Gini(values - 0)
+
+Gini

### 2020-01-22 paper/paper_2_sim.R

@@ -1,0 +1,64 @@
+out_3_biased_items <-
+    tibble(
+        pars = rerun(1, create_pars_3_biased_items())
+    ) %>%
+    # sim and fit intuitive
+    mutate(
+        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
+        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
+    ) %>%
+    # dif: all other
+    mutate(
+        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
+        all_other_mod =
+            map2(
+                sim,
+                all_other_status,
+                ~ mod_flexible(
+                    .x$data,
+                    .x$groups,
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
+    ) %>%
+    # dif: iterative backward
+    mutate(
+        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
+        it_back_mod =
+            map2(
+                sim,
+                it_back_status,
+                ~ mod_flexible(
+                    .x$data,
+                    .x$groups,
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
+    ) %>%
+    # dif: tibs cluster
+    mutate(
+        cluster =
+            intuitive_mod %>%
+            map(mod_to_final_dif) %>%
+            map(~ mutate(., cluster = irt_cluster(dif)))
+    ) %>%
+    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>%
+    mutate(cluster_status = map2(cluster, cluster_stats, left_join, by = "cluster")) %>%
+    mutate(
+        cluster_mod =
+            map2(
+                sim,
+                cluster_status,
+                ~ mod_flexible(
+                    .x$data,
+                    .x$groups,
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        cluster_final_dif = cluster_mod %>% map(mod_to_final_dif)
+    )

### 2020-02-24 R-work-in-progress/work-rough.Rmd
its working and just doing gini stuff now

--- 2020-01-14 work.Rmd
@@
     tibble(
-        pars = rerun(100, create_pars_3_biased_items()) # swap in pars function here
+        pars = rerun(1, create_pars_3_biased_items())
     ) %>% 
@@
 ```
+
+```{r}
+save.image("work_run_1.Rdata")
+```

### 2020-02-24 R-work-in-progress/work.Rmd
its working and just doing gini stuff now

```{r}
library(tidyverse)
library(mirt)
R.utils::sourceDirectory("R")
```

```{r}
pars_1_dif <-
    read_csv("data-parameters/kopf2015-8.csv") %>%
    arrange(desc(easy)) %>%
    mutate(
        item = row_number(),
        nuisance_disc = c(0, 0, 0, 0.1, 0.1, 0, 0, 0)
    )

sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 50000, n_foc = 50000, mean_foc = -1)

sim <- read_rds("data-documentation/sim.rds")

fit_mod_intuitive(sim$data, sim$groups) %>% 
    mod_intuitive_to_draws_df() %>% 
    draws_df_to_logit_plot()
```

```{r}
all_other <- dif_all_other(sim_1_dif$data, sim_1_dif$groups)

iterative_backward <- dif_iterative_backward(sim_1_dif$data, sim_1_dif$groups, all_other)

all_other

mod_to_draws_df(all_other)

out_3_biased_items$intuitive_mod[[1]] %>%
    mod_intuitive_to_draws_df() %>%
    draws_df_to_logit_plot() +
    geom_point(data = out_3_biased_items$all_other_final_dif[[1]], aes(x = dif, y = item), color = "red", shape = 1) +
    geom_vline(xintercept = filter(out_3_biased_items$all_other_final_dif[[1]], anchor)$dif[1], linetype = "dashed", color = "red") +
    geom_point(data = out_3_biased_items$it_back_final_dif[[1]], aes(x = dif, y = item), color = "blue", shape = 3) +
    geom_vline(xintercept = filter(out_3_biased_items$it_back_final_dif[[1]], anchor)$dif[1], linetype = "dotted", color = "blue")
```

### 2020-02-24 paper/TEST.Rmd
its working and just doing gini stuff now

---
output:
  bookdown::pdf_document2:
    includes:
    keep_tex: no
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 2
header-includes: 
- \usepackage{float}
- \usepackage[autostyle, english = american]{csquotes}
- \setlength{\parindent}{4em}
- \setlength{\parskip}{2em}
- \usepackage{setspace}\doublespacing
- \usepackage{amsmath}
- \DeclareMathOperator*{\argmax}{arg\,max}
fontsize: 11pt
link-citations: yes
title: "Measurement invariance"
subtitle: |
    Ben Stenhaug & Ben Domingue  
    Stanford University
abstract: |
  ABSTRACT TO DO
  <!-- https://tinyurl.com/ybremelq -->
bibliography: paper.bib
---

```{r setup, include = FALSE}
# 
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

knitr::opts_chunk$set(
    fig.align='center',
    echo=FALSE,
    warning=TRUE,
    fig.pos='H',
    fig.retina = 5
)

library(tidyverse)
library(ggtext)
library(here)
library(mirt)
library(cluster)
library(DescTools)
library(knitr)
library(kableExtra)
set.seed(123)
R.utils::sourceDirectory(here("R"))

theme_set(theme_bw())
```

The Gini index is maximized to select the anchor point. The intuition is that we want to find the anchor point such that a small minority of items have bias while the vast majority do not. Denote the easiness difference between the two groups when both group means are fixed to zero as $d_j = {b_j}^\text{ref} - {b_j}^\text{foc}$, a function that calculates the Gini index from a vector of non-negative elements as $G(\mathbf{x})$, and the mean of the focal group ability when the reference group ability is fixed to zero (i.e. the anchor point) as $\Theta$.

$$
\argmax_\Theta G(\mathbf{d} + \Theta)
$$

### 2020-02-24 paper/paper_1_intro.Rmd
its working and just doing gini stuff now

@@
 From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item.
-For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies the DIF.
+For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but their performance also depends on their verbal ability.
+In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.
 
@@
 \end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students in the same group.
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students of the same group.
 

### 2020-02-24 paper/paper_2_dif_methods.Rmd
its working and just doing gini stuff now

@@
 Psychometricians have long been in search of the perfect DIF detection method.
-@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method, where each item is tested one at a time for DIF using all other items as anchors.
+In the following three sections, we summarize DIF methods and propose new DIF methods.
+We use a simple simulation to demonstrate DIF methods: 10,000 reference group students and 10,000 focal group students take an eight-item test.
+Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
+Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group.
+The slope on target ability is set to $a_{1j} = 1$ for all items, and the slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise.
+
+## All others as anchors
+
+@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method in which each item is tested one at a time for DIF using all other items as anchors.
 More specifically, testing is done using a nested model comparison between the baseline model, where all item parameters are fixed across groups, and the flexible model, where the parameters of the tested item are freed across groups.
@@
 @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-Woods suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
+@woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
 It's unclear if one should proceed if those four items display DIF themselves.
@@
 AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
-Table X summarizes.
+Table X summarizes these methods.
+In wrapping one's mind around these methods it's useful to remember that AOAA is not an iterative procedure, while the methods with a dash, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the most conservative.
+
+ALSO MAKE A TABLE OF SIMULATION RESULTS
+
 <!-- the two kopf 2015 papers love the iterative forward SA. I bet AOAA-OAT is as good or better while fitting fewer models -->
@@
 
+## Presenting information to the analyst
+
 DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human.
-In some cases, it might make sense to present information to the analyst in a way that empowers them to make decisions.
-Our preferred technique, which we refer to as the "equal means logit graph" (EMLG), is to set both groups means to 0 so that differences in performance to manifest in the item parameters.
+We refer to DIF methods that do not use an analyst's judgement as "algorithmic DIF methods." On the other hand, it might make sense to present information to the analyst in a way that empowers them to make decisions in some cases.
+We propose a technique, the "equal means logit graph" (EMLG), where both group means are to 0 so that differences in performance manifest exclusively in the item parameters.
 Relatedly, @pohl2017cluster fit a model with both groups means set to 0 in a pedagogical example, and @talbot2013taking fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
 
-We use a simple simulation to demonstrate EMLG.
-10,000 reference group students and 10,000 focal group students take an eight-item test.
-Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group.
-The slope on target ability is set to $a_{1j} = 1$ for all items, and the slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise.
 We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data.
@@
 Multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of the difference in difficulty, ${b_j}^{\text{foc}} - {b_j}^{\text{ref}}$, for each item.
-Figure \ref{fig:emlg} shows these distributions.
+Figure \ref{fig:emlg} shows these distributions in the EMLG.
 
 ```{r emlg, fig.cap = 'Equal means logit graph (EMLG)', out.width="70%", warning = FALSE, message = FALSE}
-source("paper_2_sim.R")
+# source(here("paper/paper_2_sim.R"))
+# out_3_biased_items %>% write_rds(here("paper", "paper_2_out_3_biased_items.rds"))
+out_3_biased_items <- read_rds(here("paper", "paper_2_out_3_biased_items.rds"))
 
@@
     mod_intuitive_to_draws_df() %>%
-    mutate_at(vars(starts_with("item")), function(x) -x) %>% 
     draws_df_to_logit_plot() +
     labs(
-        x = latex2exp::TeX("$b_j^{foc} - b_j^{ref}$ (group means fixed to 0)")
+        x = latex2exp::TeX("$b_j^{ref} - b_j^{foc}$ (group means fixed to 0)")
     )
@@
 
-It cannot be stated strongly enough that Figure \ref{fig:emlg} contains all possible information about the difference in group performance.
+It cannot be stated strongly enough that the EMLG contains all possible information about the difference in group performance.
 The challenge, then, is to identify the difference in group ability.
-The analyst might assume that, because there are five items showing a difference of about 1 logit and only three items showing a difference of 1.5 logits, the group ability difference is 1 logit and items 6, 7, and 8 contain approximately 0.5 logit bias against the focal group.
-The implicit assumption they are making is that it is the minority of items that contain DIF.
+The analyst might assume that — because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and should be used as anchor items.
+At this point, the analyst re-fits the model with the anchor items having fixed parameters across groups so that they are used to estimate the difference in group ability.
+The parameters of the other items are free to vary across groups, and, indeed, the new model finds that the item easiness difference, ${b_j}^{\text{ref}} - {b_j}^{\text{foc}}$, is about 0.5 logits for each of these items.
 
-DIF methods use an algorith with its own set of assumptions to identify the group ability difference.
+One way to show this:
 
-The beauty of the EMLG graph is it makes clear all information instead of allowing the analyst the false sense of security of running an algorithm and wholeheartedlyr believing the output.
+```{r who, fig.cap = 'fix this', out.width="70%"}
+out_3_biased_items$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    draws_df_to_logit_plot() +
+    geom_vline(xintercept = filter(out_3_biased_items$AOAA_OAT_final_dif[[1]], anchor)$difference_in_easy[1], linetype = "dashed", color = "#009E73") +
+    geom_point(
+        data = out_3_biased_items$AOAA_OAT_final_dif[[1]],
+        aes(x = difference_in_easy, y = item),
+        color = "#009E73",
+        size = 2
+    ) +
+    labs(
+        title =
+            "<span style='color:#009E73'> Analyst decision based on EMLG <br>
+            Anchor on items 1-5 and freely estimate items 6-8 </span>",
+        x = latex2exp::TeX("$b_j^{ref} - b_j^{foc} - \\mu_j^{foc}$")
+    ) +
+    theme(
+        plot.title = element_markdown(lineheight = 1.1)
+    )
+```
 
-We know that this test contains DIF.
-Which items depends on ability difference which we can be sure of.
-We can, however, quantify the amount of DIF on this test.
-We can take the SD or we can assume between and calculate chalmers statistic for each.
-We might not know for sure the ability difference but we can quantify the potential for DIF.
-For each we calculate the SD across item parameters.
+Another way to show this:
+
+```{r why, fig.cap = 'fix this', out.width="70%"}
+out_3_biased_items$AOAA_OAT_final_dif[[1]] %>%
+    select(item, anchor, a_ref_easy, b_foc_easy) %>%
+    gather(var, val, -item, -anchor) %>%
+    mutate(var = case_when(anchor ~ "anchor", var == "a_ref_easy" ~ "ref_easy", var == "b_foc_easy" ~ "foc_easy")) %>%
+    ggplot(aes(x = item, y = val, color = var)) +
+    geom_point() +
+    coord_flip() +
+    scale_color_manual(values = c("black", "blue", "red"))
+```
+
+One of our key concerns with algorithmic DIF methods is that they can lull the analyst into a false sense of security.
+Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
+The EMLG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst is may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even in cases where algorithmic DIF methods are used, the analyst should use the EMLG to give them a sense of their item response data.
+
+## Clustering
+
+@bechger2015statistical proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue of item response models is to arbitrarily set an item easiness parameter to 0.
+For each group, the relative difficulties for each pair of items are stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$.
+The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in difficulty between items" [p.
+323].
+
+The general idea of identifying clusters of items is interesting.
+However, their approach is needlessly complicated and they did not give specific suggestions on how to move from $\Delta \mathbf{R}$ to a set of anchor items.
+@pohl2017cluster extended their work by proposing an approach for identifying anchor items.
+$\Delta \mathbf{R}$ is skew-symmetric and of rank 1 so that all information is contained in a single row or column.
+Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
+They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
+However, in their simulation study, they find that BIC identifies only a single cluster, so they use all items as anchors.
+
+We propose a new cluster-based approach which we refer to as equal means clustering (EMC).
+Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector of differences in item easiness, ${b_j}^\text{ref} - {b_j}^\text{foc}$, when both groups have their means fixed to 0 as is the case in the EMLG.
+Instead of choosing k with BIC, we use the gap statistic method reccommended by @hastie2001estimating.
+The largest cluster is chosen as the anchor cluster.
+If there is a tie for the largest cluster, the cluster with the lowest standard deviation of item easiness differences is selected.
+
+Using this method, we find that _________.
+
+```{r}
+out_3_biased_items$cluster_status[[1]] %>% 
+    select(item, a_ref_easy, b_foc_easy, difference_in_easy, cluster) %>% 
+    mutate(anchor = ifelse(cluster == 1, TRUE, FALSE))
+```
+
+## Anchor points
+
+All of the previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
+The difference in group abilities is then estimated using exclusively the anchor items.
+An alternative strategy is to select an anchor point, which essentially specifies the difference in group abilities directly.
+Anchor point methods have the advantage of not assuming that any particular item is DIF-free and therefore, all items can be tested for DIF.
+
+### Gini index
+
+@strobl2018anchor suggest using the Gini index -@gini1912variabilita for selecting anchor points.
+The Gini index is typically used to measure the inequality of wealth distribution in a country.
+South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution @chitiga2015income.
+In general, the Gini index "takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little" [@strobl2018anchor, p.
+7].
+
+The Gini index is maximized to select the anchor point.
+The intuition is that we want to find the anchor point such that a small minority of items have bias while the vast majority do not.
+Denote the easiness difference between the two groups when both group means are fixed to zero as $d_j = {b_j}^\text{ref} - {b_j}^\text{foc}$, a function that calculates the Gini index from a vector of non-negative elements as $G(\mathbf{x})$, and the mean of the focal group ability when the reference group ability is fixed to zero (i.e.
+the anchor point) as $\Theta$.
+The anchor point is set to
+$$
+\argmax_\Theta G(\mathbf{d} + \Theta)
+$$
+where $\Theta \in (-\infty, \infty)$ and $\Theta$ is added to each element of $\mathbf{d}$.
+
+```{r}
+emlg <- 
+    coef_1f(out_3_biased_items$intuitive_mod[[1]])$items %>% 
+    select(item, ref_easy = a_ref_easy, foc_easy = b_foc_easy) %>% 
+    mutate(difference_in_easy = ref_easy - foc_easy)
+
+gini_results <- emlg_to_gini_l1(emlg, lo = -2, hi = 0)
+```
+
+Figure \ref{fig:workplease} shows these distributions in the EMLG.
+
+```{r workplease, fig.cap = 'always need a capt', out.width="70%"}
+gini_results$grid %>% 
+    ggplot(aes(x = anchor_point, y = gini)) +
+    geom_point()
+```
+
+Strobl used the Gini coefficient
+
+Also the method where we say okay we can always add 1 to group ability and subtract 1 from group difference so we need a way to select from all of these solutions which are equivalent from a likelihood perspective.
+An obvious solution is to min total between curves
+
+- min between curves perform similarly to minimizing l1 norm in Rasch case but has clearer solution for 2PL and on
+
+```{r workplease2, fig.cap = 'always need a capt', out.width="70%"}
+# min_bc <-
+#     out_3_biased_items$intuitive_mod[[1]] %>%
+#     min_between_curves(lo = 0.95, hi = 1.1, by = 0.001)
+
+# min_bc %>% write_rds(here("paper", "paper_2_min_bc.rds"))
+
+min_bc <- read_rds(here("paper", "paper_2_min_bc.rds"))
+
+min_bc$graph
+```

### 2020-02-24 paper/paper_3_uncertainty.Rmd
its working and just doing gini stuff now

--- 2020-01-16 paper/paper.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-  11pt,
-]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Measurement invariance},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-\usepackage{float}
-\usepackage[autostyle, english = american]{csquotes}
-\setlength{\parindent}{4em}
-\setlength{\parskip}{2em}
-\usepackage{setspace}\doublespacing
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
+We know that this test contains DIF.
+Which items depends on ability difference which we can be sure of.
+We can, however, quantify the amount of DIF on this test.
+We can take the SD or we can assume between and calculate chalmers statistic for each.
+We might not know for sure the ability difference but we can quantify the potential for DIF.
+For each we calculate the SD across item parameters.
 
-\title{Measurement invariance}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Ben Stenhaug \& Ben Domingue\\
-Stanford University}
-\author{}
-\date{\vspace{-2.5em}}
+Something with DTF?
 
-\begin{document}
-\maketitle
-\begin{abstract}
-ABSTRACT TO DO
-\end{abstract}
+Do a simulation where the second dimension varies - a more real testing ground
 
-{
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-\clearpage
+Example with real data perhaps the Mike Frank data
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we view differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
-From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is a property of the student, not the item.
-For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies the DIF.
-
-Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item.
-For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item difficulties for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
-They set item difficulties for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group and generate item responses according to the Rasch model.
-The Rasch model specifies that the probability of student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
-\end{align}
-where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
-
-We find it more informative to describe simulation conditions in terms of single-dimensional items and multidimensional abilities.
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item difficulty is set to what was previously \({b_j}^{\text{ref}}\) for all students.
-Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF and \(a_{2j} = 0\) otherwise.
-Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional 2PL model, the probability student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
-\end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students in the same group.
-
-If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model?
-Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}, 144) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities.} He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary.
-It's hard to imagine how one would have such knowledge.
-Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
-
-\hypertarget{dif-methods}{%
-\section{DIF methods}\label{dif-methods}}
-
-Psychometricians have long been in search of the perfect DIF detection method.
-Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method, where each item is tested one at a time for DIF using all other items as anchors.
-More specifically, testing is done using a nested model comparison between the baseline model, where all item parameters are fixed across groups, and the flexible model, where the parameters of the tested item are freed across groups.
-Anchor items have fixed parameters across groups and consequently are used to estimate the ability difference between groups.
-For example, Edelen et al.
-(\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can we be sure its those 10 items and not the other 11 items with DIF?
-Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}).
-In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
-
-Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
-A simple thought experiment illustrates how AOAA fails: Imagine a test with three items where the first item has DIF, and the other two do not.
-Using AOAA, all items test positive for DIF.
-The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until (Andrich and Hagquist \protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}
-
-One way to attempt to counter artificial DIF is with purification with iterative anchor selection.
-For example, (Drasgow \protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then continued iteratively to test the remaining items for DIF until no more items tested positively.
-(Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
-Iterative-backward-AOAA might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
-With a sufficient sample size, this will necessarily be the case.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-Woods suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items display DIF themselves.
-
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge, has not previously been proposed.
-AOAA-OAT starts with AOAA, and the single item exhibiting the most DIF is removed from the anchor set.
-Each item in the anchor set is then tested for DIF again (with items outside of the anchor set allowed to have free parameters across groups in both the baseline and flexible model).
-This process continues until no new items display DIF.
-Applying AOAA-OAT to our thought experiment demonstrates its logic.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
-In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
-
-DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human.
-In some cases, it might make sense to present information to the analyst in a way that empowers them to make decisions.
-Our preferred technique, which we refer to as the \enquote{equal means logit graph} (EMLG), is to set both groups means to 0 so that differences in performance to manifest in the item parameters.
-Relatedly, Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) fit a model with both groups means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
-
-We use a simple simulation to demonstrate EMLG.
-10,000 reference group students and 10,000 focal group students take an eight-item test.
-Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
-We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data.
-We estimate the item parameter covariance matrix using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of the difference in difficulty, \({b_j}^{\text{foc}} - {b_j}^{\text{ref}}\), for each item.
-Figure \ref{fig:emlg} shows these distributions.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emlg-1}
-
-}
-
-\caption{Equal means logit graph (EMLG)}\label{fig:emlg}
-\end{figure}
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ackerman1992didactic}{}%
-Ackerman, Terry A. 1992. ``A Didactic Explanation of Item Bias, Item Impact, and Item Validity from a Multidimensional Perspective.'' \emph{Journal of Educational Measurement} 29 (1): 67--91.
-
-\leavevmode\hypertarget{ref-andrich2012real}{}%
-Andrich, David, and Curt Hagquist. 2012. ``Real and Artificial Differential Item Functioning.'' \emph{Journal of Educational and Behavioral Statistics} 37 (3): 387--416.
-
-\leavevmode\hypertarget{ref-camilli1992conceptual}{}%
-Camilli, Gregory. 1992. ``A Conceptual Analysis of Differential Item Functioning in Terms of a Multidimensional Item Response Model.'' \emph{Applied Psychological Measurement} 16 (2): 129--47.
-
-\leavevmode\hypertarget{ref-chalmers2018numerical}{}%
-Chalmers, R Philip. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
-
-\leavevmode\hypertarget{ref-drasgow1987study}{}%
-Drasgow, Fritz. 1987. ``Study of the Measurement Bias of Two Standardized Psychological Tests.'' \emph{Journal of Applied Psychology} 72 (1): 19.
-
-\leavevmode\hypertarget{ref-edelen2006identification}{}%
-Edelen, Maria Orlando, David Thissen, Jeanne A Teresi, Marjorie Kleinman, and Katja Ocepek-Welikson.
-2006.
-``Identification of Differential Item Functioning Using Item Response Theory and the Likelihood-Based Model Comparison Approach: Application to the Mini-Mental State Examination.'' \emph{Medical Care}, S134--S142.
-
-\leavevmode\hypertarget{ref-kopf2015framework}{}%
-Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.
-
-\leavevmode\hypertarget{ref-meade2012solving}{}%
-Meade, Adam W, and Natalie A Wright. 2012. ``Solving the Measurement Invariance Anchor Item Problem in Item Response Theory.'' \emph{Journal of Applied Psychology} 97 (5): 1016.
-
-\leavevmode\hypertarget{ref-pohl2017cluster}{}%
-Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``CluStEr-baSEd anCHor Item Identification and Selection.'' NEPS Working Paper.
-
-\leavevmode\hypertarget{ref-stark2006detecting}{}%
-Stark, Stephen, Oleksandr S Chernyshenko, and Fritz Drasgow.
-2006.
-``Detecting Differential Item Functioning with Confirmatory Factor Analysis and Item Response Theory: Toward a Unified Strategy.'' \emph{Journal of Applied Psychology} 91 (6): 1292.
-
-\leavevmode\hypertarget{ref-strobl2018anchor}{}%
-Strobl, Carolin, Julia Kopf, Raphael Hartmann, and Achim Zeileis. 2018. ``Anchor Point Selection: An Approach for Anchoring Without Anchor Items.'' Working Papers in Economics; Statistics.
-
-\leavevmode\hypertarget{ref-talbot2013taking}{}%
-Talbot III, Robert M.
-2013.
-``Taking an Item-Level Approach to Measuring Change with the Force and Motion Conceptual Evaluation: An Application of Item Response Theory.'' \emph{School Science and Mathematics} 113 (7): 356--65.
-
-\leavevmode\hypertarget{ref-yang2012characterizing}{}%
-Yang, Ji Seung, Mark Hansen, and Li Cai. 2012. ``Characterizing Sources of Uncertainty in Item Response Theory Scale Scores.'' \emph{Educational and Psychological Measurement} 72 (2): 264--90.
-
-\end{document}
+Summer learning loss as applied DIF?

### 2020-02-24 R-documentation/doc_min_between_curves.R

@@ -1,0 +1,228 @@
+# This is all about thinking through how the difference between curves integration
+# depends on which group I integrate based on.
+# Below I first show that this matters in a simple example
+# Then I walk through integration based on reference group, then focal group, and
+# finally based on combining the two populations (which is what I end up using)
+
+# FOR THE 2PL:
+# useful to think in terms of that grid and then find which minimizes
+# the difference in curves where the two parameters are how
+# much i add or shift ability and then how much i divide each by
+
+# VERIFY THAT WHICH GROUP I INTEGRATE BASED ON MATTERS --------------------
+# imagine a single item
+# there is no bias
+# but ref_mu = 0 and foc_mu = -1
+
+Th <- seq(-6, 6, 0.01)
+
+df <-
+    tibble(
+        theta = Th
+    ) %>%
+    mutate(
+        ref = boot::inv.logit(theta + 0),
+        foc = boot::inv.logit(theta + -1),
+        bet_curv = abs(ref - foc)
+    ) %>%
+    mutate(
+        wt_ref = dnorm(Th, 0, 1) / sum(dnorm(Th, 0, 1)),
+        wt_foc = dnorm(Th, -1, 1) / sum(dnorm(Th, -1, 1))
+    )
+
+df %>%
+    select(theta, ref, foc) %>%
+    gather(var, val, -theta) %>%
+    ggplot(aes(x  = theta, y = val, color = var)) +
+    geom_point()
+
+df %>%
+    ggplot(aes(x = theta, y = bet_curv)) +
+    geom_point()
+
+df %>%
+    select(theta, starts_with("wt")) %>%
+    gather(var, val, -theta) %>%
+    ggplot(aes(x = theta, y = val, color = var)) +
+    geom_point()
+
+weighted.mean(df$bet_curv, df$wt_ref)
+
+weighted.mean(df$bet_curv, df$wt_foc)
+
+# FIX PARAMETERS ----------------------------------------------------------
+intuitive_mod <- out_3_biased_items$intuitive_mod[[1]]
+lo <- 0.9
+hi <- 1.1
+by <- 0.01
+
+# REFERENCE GROUP FOCUS ---------------------------------------------------
+REF_min_between_curves <- function(intuitive_mod, lo, hi, by){
+
+    n_items <- length(intuitive_mod@Data$K)
+
+    Theta <- seq(-6, 6, length.out = 61)
+
+    mu <- 0 # ref group never changes
[…]

### 2020-02-24 R-work-complete/2PL.R

@@ -1,0 +1,65 @@
+# my whole goal here is to think about what happens when we move from Rasch to the 2PL
+# in particular with the cluster based approach how will I cluster on both of these things
+# what I confirm below is that just like for Rasch fixing means across groups is enough,
+# for the 2PL, fixing means and variance across groups is enough
+# the way to think about it is the standard for the Rasch or 2PL is to fix one group
+# to N(0, 1) and then have the other group N(?, ?) and all of the item parameters unknown
+# I basically need to fix one thing for the Rasch - in the paper I argue the first ?
+# and two things for the 2PL - both of the ? ?
+# the end of page 335 of the bechger piece is useful for thinking
+# and confirms like we think of differences in difficulty we can think of
+# ratios of discriminations
+# two ideas for the cluster scenario
+# 1. just choose which items to anchor for difficulty and discrimination separately
+# 2. do some cluster on the matrix of differences in difficulty and ratio of discrimination
+# but that kind of seems like it puts us in some weird space
+# SEE doc_min_between_curves for some thinking on doing this with the integration thing
+# (basically first intuition is to make a grid of group means and variance slopes to divide by
+# and then calculate at each of them)
+
+pars <-
+    tibble(
+        item = 1:8,
+    ) %>%
+    mutate(
+        easy = 0,
+        disc = 1
+    )
+
+theta_ref <- tibble(target = rnorm(100000, 0, 1))
+theta_foc <- tibble(target = rnorm(100000, 0, 2))
+
+hi <- simdata(
+    a = pars %>% select(disc) %>% as.matrix(),
+    d = pars$easy,
+    itemtype = "2PL",
+    Theta = bind_rows(theta_ref, theta_foc) %>% as.matrix(),
+    returnList = TRUE
+)
+
+mod <-
+    multipleGroup(
+        hi$data,
+        1,
+        itemtype = "2PL",
+        group = c(rep("aref", 100000), rep("bfoc", 100000)),
+        technical = list(NCYCLES = 500),
+        verbose = TRUE
+    )
+
+coef(mod)
+
+mod <-
+    multipleGroup(
+        hi$data,
+        1,
+        itemtype = "2PL",
+        group = c(rep("ref", 100000), rep("foc", 100000)),
+        technical = list(NCYCLES = 500),
+        verbose = TRUE,
+        invariance = c("intercepts","slopes", "free_var")
+    )
+
+coef(mod)
+
+# to get ref down we have sqrt(5.47 * 1.15^4)

### 2020-02-24 R-work-complete/DTF_chalmers_uses_uniform_weights - supplement.R

@@ -1,0 +1,133 @@
+# this is the full one with the lapply and all
+# in the other non-supplement i just run for one value
+# to figure out what's going on
+
+data <- out_3_biased_items$sim[[1]]$data
+groups <- out_3_biased_items$sim[[1]]$groups
+
+mod <-
+    multipleGroup(
+        data,
+        1,
+        itemtype = "Rasch",
+        group = groups,
+        invariance = c("Item_1", "free_means", "free_var")
+    )
+
+coef_1f(mod)
+
+debug(DTF)
+undebug(DTF)
+
+DTF(mod)
+
+# Parameters --------------------------------------------------------------
+npts <- 1000
+theta_lim <- c(-6, 6)
+
+type <- "score"
+integration <- "quad"
+plot <- "none"
+Theta_nodes <- NULL
+impute <- FALSE
+
+# Main --------------------------------------------------------------------
+theta <- matrix(seq(theta_lim[1L], theta_lim[2L], length.out = npts))
+Theta <- thetaComb(theta, mod@Model$nfact)
+theta <- matrix(seq(theta_lim[1L], theta_lim[2L], length.out = npts))
+Theta <- thetaComb(theta, mod@Model$nfact)
+max_score <- sum(mod@Data$mins + mod@Data$K - 1L)
+list_scores <- lapply(1L, fn, omod = mod, impute = FALSE,
+                        Theta_nodes = Theta_nodes, imputenums = NULL, max_score = max_score,
+                        Theta = Theta, plot = plot, integration = integration,
+                        theta_lim = theta_lim, type = type)
+
+# Depth 1 -----------------------------------------------------------------
+fn <- function(x, omod, impute, imputenums, Theta, max_score,
+               Theta_nodes = NULL, plot, integration, theta_lim, type,
+               pre.ev, shortpars, longpars) {
+
+    mod <- omod
+    if (impute) {
+        mod <- imputePars2(MGmod = mod, pre.ev = pre.ev,
+                           imputenums = imputenums, shortpars = shortpars,
+                           longpars = longpars)
+    }
+    if (!is.null(Theta_nodes)) {
+        if (type == "score") {
+            T1 <- expected.test(mod, Theta_nodes, group = 1L,
+                                mins = FALSE)
+            T2 <- expected.test(mod, Theta_nodes, group = 2L,
[…]

### 2020-02-24 R-work-complete/DTF_chalmers_uses_uniform_weights.R

@@ -1,0 +1,106 @@
+# this is me diving into the chalmers DTF function
+# and finding out that it just weights each points the
+# same within the integration bands
+# seems like a big problem with their work
+
+data <- out_3_biased_items$sim[[1]]$data
+groups <- out_3_biased_items$sim[[1]]$groups
+
+mod <-
+    multipleGroup(
+        data,
+        1,
+        itemtype = "Rasch",
+        group = groups,
+        invariance = c("Item_1", "free_means", "free_var")
+    )
+
+coef_1f(mod)
+
+debug(DTF)
+undebug(DTF)
+
+DTF(mod)
+
+# Parameters --------------------------------------------------------------
+npts <- 13
+theta_lim <- c(-6, 6)
+
+type <- "score"
+integration <- "quad"
+plot <- "none"
+Theta_nodes <- NULL
+impute <- FALSE
+
+# Main --------------------------------------------------------------------
+theta <- matrix(seq(theta_lim[1L], theta_lim[2L], length.out = npts))
+Theta <- thetaComb(theta, mod@Model$nfact)
+theta <- matrix(seq(theta_lim[1L], theta_lim[2L], length.out = npts))
+Theta <- thetaComb(theta, mod@Model$nfact)
+max_score <- sum(mod@Data$mins + mod@Data$K - 1L)
+
+calc_DTFs(mod = mod, Theta = Theta, plot = plot, max_score = max_score, type = type)
+
+# Depth 2 -----------------------------------------------------------------
+calc_DTFs <- function (mod, Theta, plot, max_score, type) {
+    Ds <- matrix(NA, nrow(Theta), 2L)
+    colnames(Ds) <- c("score", "null")
+    if ("score" %in% type) {
+        T1 <- expected.test(mod, Theta, group = 1L, mins = FALSE)
+        T2 <- expected.test(mod, Theta, group = 2L, mins = FALSE)
+        Ds[, 1L] <- T1 - T2
+    }
+    if (plot != "none")
+        return(c(T1, T2))
+    uDTF <- colMeans(abs(Ds))
+    uDTF_percent <- uDTF[1L]/max_score * 100
+    sDTF <- colMeans(Ds)
+    sDTF_percent <- sDTF[1L]/max_score * 100
+    ret <- c(sDTF = sDTF[1L], `sDTF(%)` = sDTF_percent, sDTF = sDTF[-1L],
[…]

### 2020-02-24 R-work-complete/clustering_understanding.R

@@ -1,0 +1,5 @@
+emlg <- read_rds(here("data-documentation/emlg.rds"))
+
+add_clusters <- emlg %>% mutate(cluster = irt_cluster(difference_in_easy))
+
+add_clusters %>% left_join(get_cluster_stats(add_clusters))

### 2020-02-24 R-work-complete/cov_doesnt_change_when_shifting.R

@@ -1,0 +1,37 @@
+data <- out_3_biased_items$sim[[1]]$data
+groups <- out_3_biased_items$sim[[1]]$groups
+
+mod <-
+    multipleGroup(
+        data,
+        1,
+        itemtype = "Rasch",
+        group = groups,
+        invariance = c("free_var")
+    )
+
+coef_1f(mod)
+
+mod <-
+    multipleGroup(
+        data,
+        1,
+        itemtype = "Rasch",
+        group = groups,
+        invariance = c("free_var"),
+        pars = "values"
+    )
+
+mod$value[67] <- 3
+
+mod2 <-
+    multipleGroup(
+        data,
+        1,
+        itemtype = "Rasch",
+        group = groups,
+        invariance = c("free_var"),
+        pars = mod
+    )
+
+coef_1f(mod2)

### 2020-02-24 R-work-complete/dif_integral_understanding.R

@@ -1,0 +1,145 @@
+# i was trying to figure out why increasing theta didn't show linear increases in my between curves dif stat.
+# the reason is because of the shape of the inverse logit function which i show below
+# makes me slightly wonder if you'd want to switch to like a log likelihood or something that does increase linearly
+# and is a bit more in line with maximum likelihood estimation but probably not for now
+# basically i think what im suggesting is that to get around the identifiability problem
+# where we can always change both the item parameters and group ability and wind up with the same likelihood
+# is that we should prefer models where the differences in item parameters are the smallest (rmse is a simple solution and in the rasch case
+# total dif is another solution that performs similarly but i think will work better in other models)
+
+# show not linear ---------------------------------------------------------
+library(tidyverse)
+library(DescTools)
+
+mod <- out_3_biased_items$intuitive_mod[[1]]
+
+coef_1f(mod)
+
+
+f <- function(mod, foc_worse_by){
+    Theta <- seq(-6, 6, length.out = 61)
+
+    data <-
+        crossing(item = 1:length(mod@Data$K)) %>%
+        mutate(
+            a_ref = item %>% map(~ probtrace(extract.item(extract.group(mod, 1), .), Theta)[ , 2]),
+            b_foc = item %>% map(~ probtrace(extract.item(extract.group(mod, 2), .), Theta + foc_worse_by)[ , 2])
+        ) %>%
+        unnest(cols = c(a_ref, b_foc)) %>%
+        mutate(
+            theta = rep(Theta, length(mod@Data$K)),
+            weight = rep(dnorm(Theta) / sum(dnorm(Theta)), length(mod@Data$K))
+        ) %>%
+        group_by(item) %>%
+        summarize(dif = sum(abs(a_ref - b_foc) * weight))
+
+    sum(data$dif)
+}
+
+by_item <- coef_1f(mod)$items %>% transmute(item, foc_worse_by = a_ref_easy - b_foc_easy)
+
+a <- coef_1f(mod)$items$a_ref_easy
+b <- coef_1f(mod)$items$b_foc_easy
+
+out <-
+    tibble(
[…]

### 2020-02-24 R-work-in-progress/work.R

@@ -1,0 +1,58 @@
+# can i add some sort of sampling error somewhere
+
+out$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    View()
+
+# was playing with clustering for each draw..
+
+mod_intuitive <- out$intuitive_mod[[1]]
+
+
+
+x <- mod %>% mod_to_draws_df(10)
+
+hi <-
+    x %>%
+    gather(item, easy, -run) %>%
+    arrange(run) %>%
+    split(.$run) %>%
+    map(~ irt_cluster(.$easy))
+
+vec <- y$val[1:8]
+
+#
+
+library(tidyverse)
+library(mirt)
+
+pars <-
+    read_csv("data-parameters/kopf2015-8.csv") %>%
+    arrange(desc(easy)) %>%
+    mutate(item = row_number())
+
+n_ref <- 100000
+n_foc <- 100000
+
+sim <- sim_bias(pars = pars, n_ref = 100000, n_foc = 100000, mean_foc = -1)
+
+# this very much works, but requires that we know the anchor items
+mod1f <-
+    multipleGroup(
+        sim$data,
+        1,
+        itemtype = "2PL",
+        sim$groups,
+        invariance = c("free_means", paste0("Item_", 4:8)),
+        technical = list(NCYCLES = 100)
+    )
+
+coef_1f(mod1f)
+
+# not sure why all of the action happens in item 1 here
+# perhaps evidence that modeling multidimensionally is hard
+# and maybe something to do with variance of that second dimension being 0 but us not inputting that
+mod2f <- multipleGroup(sim$data, 2, itemtype = "2PL", sim$groups, invariance = c("free_means","slopes", "intercepts"), technical = list(NCYCLES = 50))
+coef(mod2f)
+
+summary(mod2f, rotate = "oblimin")

### 2020-02-24 R/clustering.R

@@ -25,8 +25,8 @@ get_cluster_stats <- function(cluster){
         group_by(cluster) %>%
         summarize(
             n = n(),
-            sd = sd(dif)
+            sd = sd(difference_in_easy)
         ) %>%
-        arrange(desc(n), sd) %>%
+        arrange(desc(n), sd) %>% # chooses based on the greatest number of items
         mutate(status = ifelse(row_number() == 1, "unknown", "flex"))
 }

### 2020-02-24 R/dif_AOAA.R

@@ -1,0 +1,87 @@
+dif_AOAA <- function(data, groups){
+    tibble(
+        item = 1:ncol(data)
+    ) %>%
+        mutate(
+            p = item %>% map_dbl(~ constrained_baseline_test_flex_items(data, groups, flex_items = .)),
+            status = ifelse(p > 0.05, "unknown", "flex")
+        )
+}
+
+get_next_status <- function(status, data, groups){
+    the_flex <- which(status$status == "flex")
+
+    status %>%
+        filter(status == "unknown") %>%
+        mutate(
+            p = item %>%
+                map_dbl(~ test_with_known_flex(data, groups, flex_known = the_flex, flex_test = .))
+        ) %>%
+        select(-status)
+}
+
+dif_AOAA_OAT <- function(data, groups, all_other){
+
+    if (all(all_other$p > 0.05)){
+        return(
+            all_other %>%
+                mutate(
+                    status = ifelse(item == which.min(p), "flex", "unknown")
+                )
+        )
+    }
+
+    status <-
+        all_other %>%
+        mutate(
+            status = ifelse(item == which.min(p), "flex", "unknown")
+        )
+
+    while(TRUE){
+        new_status <- get_next_status(status, data, groups)
+
+        status <-
+            status %>%
+            select(-p) %>%
+            left_join(new_status, by = "item")
+
+        if(all(status$p > 0.05, na.rm = TRUE)){break}
+
+        status$status[which.min(status$p)] <- "flex"
+    }
+
+    status
+}
+
+dif_AOAA_AS <- function(data, groups, all_other){
+
+    if (all(all_other$p > 0.05)){
+        return(
+            all_other %>%
+                mutate(
+                    status = ifelse(item == which.min(p), "flex", "unknown")
+                )
+        )
+    }
+
+    status <-
+        all_other %>%
+        mutate(
+            status = ifelse(item == which.min(p), "flex", "unknown")
+        )
+
+    while(TRUE){
+        new_status <- get_next_status(status, data, groups)
+
+        status <-
+            status %>%
+            select(-p) %>%
+            left_join(new_status, by = "item")
+
+        if(all(status$p > 0.05, na.rm = TRUE)){break}
+
+        status$status[which(status$p < 0.05)] <- "flex"
+    }
+
+    status
+}

### 2020-02-24 R/dif_all_other.R

@@ -1,9 +1,0 @@
-dif_all_other <- function(data, groups){
-    tibble(
-        item = 1:ncol(data)
-    ) %>%
-        mutate(
-            p = item %>% map_dbl(~ constrained_baseline_test_flex_items(data, groups, flex_items = .)),
-            status = ifelse(p > 0.05, "unknown", "flex")
-        )
-}

### 2020-02-24 R/dif_intuitive.R

@@ -14,7 +14,7 @@ mod_intuitive_to_draws_df <- function(mod){
     n_items <- ncol(par_draws) / 2
 
     for (i in 1:n_items) {
-        draws_df[[paste0("item", i)]] <- par_draws[ , i + n_items] - par_draws[ , i]
+        draws_df[[paste0("item", i)]] <- par_draws[ , i] - par_draws[ , i + n_items]
     }
 
     draws_df

### 2020-02-24 R/dif_iterative_backward.R

@@ -1,44 +1,0 @@
-get_next_status <- function(status, data, groups){
-    the_flex <- which(status$status == "flex")
-
-    status %>%
-        filter(status == "unknown") %>%
-        mutate(
-            p = item %>%
-                map_dbl(~ test_with_known_flex(data, groups, flex_known = the_flex, flex_test = .))
-        ) %>%
-        select(-status)
-}
-
-dif_iterative_backward <- function(data, groups, all_other){
-
-    if (all(all_other$p > 0.05)){
-        return(
-            all_other %>%
-                mutate(
-                    status = ifelse(item == which.min(p), "flex", "unknown")
-                )
-        )
-    }
-
-    status <-
-        all_other %>%
-        mutate(
-            status = ifelse(item == which.min(p), "flex", "unknown")
-        )
-
-    while(TRUE){
-        new_status <- get_next_status(status, data, groups)
-
-        status <-
-            status %>%
-            select(-p) %>%
-            left_join(new_status, by = "item")
-
-        if(all(status$p > 0.05, na.rm = TRUE)){break}
-
-        status$status[which.min(status$p)] <- "flex"
-    }
-
-    status
-}

### 2020-02-24 R/emlg_to_gini_l1.R

@@ -1,0 +1,19 @@
+emlg_to_gini_l1 <- function(emlg, lo = -2, hi = 2){
+
+    stopifnot(all(c("ref_easy", "foc_easy", "difference_in_easy") %in% names(emlg)))
+
+    grid <-
+        tibble(
+            anchor_point = seq(lo, hi, 0.0001)
+        ) %>%
+        mutate(
+            gini = anchor_point %>% map_dbl(~ Gini(abs(emlg$ref_easy - emlg$foc_easy + .))),
+            l1 = anchor_point %>% map_dbl(~ mean(abs((emlg$ref_easy - emlg$foc_easy + .))^1))
+        )
+
+    list(
+        grid = grid,
+        gini_anchor_points = grid$anchor_point[grid$gini == max(grid$gini)],
+        l1_anchor_points = grid$anchor_point[grid$l1 == min(grid$l1)]
+    )
+}

### 2020-02-24 R/min_between_curves.R

@@ -1,0 +1,65 @@
+# for walkthrough see doc_min_between_curves.R
+# see note on assuming cov stays constant below ...  think about if going to 2PL
+# see note on Theta + anchor_point below which needs thinking if I do the 2PL
+
+min_between_curves <- function(intuitive_mod, lo, hi, by){
+
+    n_items <- length(intuitive_mod@Data$K)
+    Theta <- seq(-6, 6, length.out = 61)
+
+    a_ref_probtrace <- 1:n_items %>% map(~ probtrace(extract.item(extract.group(intuitive_mod, 1), .), Theta)[ , 2])
+
+    mu_ref <- 0
+    sd_ref <- sqrt(coef(intuitive_mod, simplify = TRUE)$a_ref$cov[1])
+    weights_ref <- dnorm(Theta, mu_ref, sd_ref) / sum(dnorm(Theta, mu_ref, sd_ref))
+
+    f <- function(intuitive_mod, anchor_point){
+
+        mu_foc <- anchor_point
+        # assumes cov doesn't change when I shift mean
+        # which I mostly cerified for Rasch in
+        # R-work-complete > cov_doesnt_change_when_shifting
+        sd_foc <- sqrt(coef(intuitive_mod, simplify = TRUE)$b_foc$cov[1])
+        weights_foc <- dnorm(Theta, mu_foc, sd_foc) / sum(dnorm(Theta, mu_foc, sd_foc))
+
+        weights <- (weights_foc + weights_ref) / 2 # assumes 50% in each group which I could of course change
+
+        data <-
+            tibble(
+                item = 1:n_items,
+                a_ref = a_ref_probtrace
+            ) %>%
+            # Theta + anchor_point was throwing me for a loop
+            # remember that intuitive_mod had fixed means and independent item parameters across groups
+            # so this is a shortcut for shifting all of the item parameters for the focal group
+            # will need to rethink this if I move to the 2PL
+            mutate(
+                b_foc = item %>% map(~ probtrace(extract.item(extract.group(intuitive_mod, 2), .), Theta + anchor_point)[ , 2])
+            ) %>%
+            unnest(cols = c(a_ref, b_foc)) %>%
+            mutate(
+                theta = rep(Theta, n_items),
+                weight = rep(weights, n_items)
+            ) %>%
+            group_by(item) %>%
+            summarize(between_curves = sum(abs(a_ref - b_foc) * weight))
+
+        sum(data$between_curves)
+    }
+
+    grid <-
[…]

### 2020-02-24 R/mod_zelpers.R

@@ -4,9 +4,8 @@ mod_to_final_dif <- function(all_other_mod){
     the_coef$items %>%
         mutate(
             item = paste0("item", 1:8),
-            dif = b_foc_easy - a_ref_easy + the_coef$ability$b_foc[1]
-        ) %>%
-        select(item, anchor, dif)
+            difference_in_easy = a_ref_easy - b_foc_easy - the_coef$ability$b_foc[1]
+        )
 }
 
 # this works for a general model where some are fixed (as opposed to mod intuitive where all flex)

### 2020-02-24 gini_l1.R

@@ -1,32 +1,0 @@
-a <- c(0, 0, 1, 1)
-b <- c(0, 0, 0, 0)
-
-a <- c(0, 0, 0, 1, 1)
-b <- c(0, 0, 0, 0, 0)
-
-a <- c(0.1, 0.2, 0.3, 1, 1)
-b <- c(0, 0, 0, 0, 0)
-
-library(tidyverse)
-library(DescTools)
-
-tibble(
-    p = seq(0, 1, 0.05)
-) %>%
-    mutate(
-        gini = p %>% map_dbl(~ Gini(abs(a - b - .))),
-        #llog = p %>% map_dbl(~ mean(log(abs(a - b - .)))),
-        #l12 = p %>% map_dbl(~ mean(abs((a - b - .))^1/2)),
-        l1 = p %>% map_dbl(~ mean(abs((a - b - .))^1)),
-        #l2 = p %>% map_dbl(~ mean(abs((a - b - .))^2)),
-        #l3 = p %>% map_dbl(~ mean(abs((a - b - .))^3))
-    ) %>%
-    gather(var, val, -p) %>%
-    ggplot(aes(x = p, y = val, color = var)) +
-    geom_path()
-
-mean((values - 0.5)^2)
-
-Gini(values - 0)
-
-Gini

### 2020-02-24 paper/paper_2_sim.R

@@ -7,13 +7,13 @@ out_3_biased_items <-
         sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
         intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
     ) %>%
-    # dif: all other
+    # dif: AOAA
     mutate(
-        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
-        all_other_mod =
+        AOAA_status = sim %>% map(~ dif_AOAA(.$data, .$groups)),
+        AOAA_mod =
             map2(
                 sim,
-                all_other_status,
+                AOAA_status,
                 ~ mod_flexible(
                     .x$data,
                     .x$groups,
@@ -21,15 +21,15 @@ out_3_biased_items <-
                     se = TRUE
                 )
             ),
-        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
+        AOAA_final_dif = AOAA_mod %>% map(mod_to_final_dif)
     ) %>%
-    # dif: iterative backward
+    # dif: AOAA AS
     mutate(
-        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
-        it_back_mod =
+        AOAA_AS_status = map2(sim, AOAA_status, ~ dif_AOAA_AS(.x$data, .x$groups, .y)),
+        AOAA_AS_mod =
             map2(
                 sim,
-                it_back_status,
+                AOAA_AS_status,
                 ~ mod_flexible(
                     .x$data,
                     .x$groups,
@@ -37,14 +37,30 @@ out_3_biased_items <-
                     se = TRUE
                 )
             ),
-        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
+        AOAA_AS_final_dif = AOAA_AS_mod %>% map(mod_to_final_dif)
+    ) %>%
+    # dif: AOAA OAT
+    mutate(
+        AOAA_OAT_status = map2(sim, AOAA_status, ~ dif_AOAA_OAT(.x$data, .x$groups, .y)),
+        AOAA_OAT_mod =
+            map2(
+                sim,
+                AOAA_OAT_status,
+                ~ mod_flexible(
+                    .x$data,
+                    .x$groups,
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
[…]

### 2020-02-24 work_graph.R

@@ -1,7 +1,0 @@
-out$intuitive_mod[[1]] %>%
-    mod_intuitive_to_draws_df() %>%
-    draws_df_to_logit_plot() +
-    geom_point(data = out$all_other_final_dif[[1]], aes(x = dif, y = item), color = "red", shape = 1) +
-    geom_vline(xintercept = filter(out$all_other_final_dif[[1]], anchor)$dif[1], linetype = "dashed", color = "red") +
-    geom_point(data = out$it_back_final_dif[[1]], aes(x = dif, y = item), color = "blue", shape = 3) +
-    geom_vline(xintercept = filter(out$it_back_final_dif[[1]], anchor)$dif[1], linetype = "dotted", color = "blue")

### 2020-02-24 paper/paper.Rmd
okay! i have a decent start to this draft. nice!

@@
     includes:
-    keep_tex: yes
+    keep_tex: no
     latex_engine: xelatex
@@
     toc: yes
-    toc_depth: 2
+    toc_depth: 5
 header-includes: 
+- \usepackage{amsmath}
+- \DeclareMathOperator*{\argmax}{arg\,max}
 - \usepackage{float}
@@
 library(tidyverse)
+library(ggtext)
+library(here)
 library(mirt)
 library(cluster)
+library(DescTools)
 library(knitr)
@@
 set.seed(123)
-R.utils::sourceDirectory(here::here("R"))
+R.utils::sourceDirectory(here("R"))
 
@@
 \clearpage
+
+# TO DO
+
+- be sure all my directions match. i think i switched everything to be item easiness at some point but not sure if i did it throughout
 
@@
 
+# Next steps
+
+```{r child = "paper_3_next_steps.Rmd"}
+```
+
+\clearpage
+
+# Notes
+
+- adjust to all be easiness including multidimensional talk
+
 # References

### 2020-02-24 paper/paper_2_dif_methods.Rmd
okay! i have a decent start to this draft. nice!

@@
 
-Figure \ref{fig:workplease} shows these distributions in the EMLG.
+Figure \ref{fig:workplease} shows the gini coefficient at a variety of possible values for $\Theta$. We find that the anchor point is `r round(gini_results$gini_anchor_points, 2)`.
 
@@
 
-Strobl used the Gini coefficient
+### Minimizing between curves
 
-Also the method where we say okay we can always add 1 to group ability and subtract 1 from group difference so we need a way to select from all of these solutions which are equivalent from a likelihood perspective.
-An obvious solution is to min total between curves
+Imagine multigroup item response data where $\Theta^{\text{foc}} = \Theta^{\text{ref}}$ and $d_j = 0 \forall j$ so that there is no DIF.
+The fundamental identification problem of DIF in multipgroup IRT models is that
+there are an infinite number of models with the same likelihood from which to choose.
+For example, we could correctly assume that the focal group has the same ability as the reference group and fix $\Theta^{\text{foc}} = 0$.
+The model would then estimate $d_j \approx 0 \forall j$, and we would correctly conclude the groups have the same ability and there is no DIF.
+Alternatively, we could assume that the focal group has very high ability and fix $\Theta^{\text{foc}} = 3$.
+The model would then compensate by finding $d_j \approx -3 \forall j$, and we would incorrectly conclude that the the focal group is high ability but each item contains DIF against the focal group.
+Both of these models have the same likelihood, so how should we choose which model to believe?
+This is the fundamental identification problem.
 
-- min between curves perform similarly to minimizing l1 norm in Rasch case but has clearer solution for 2PL and on
+We put forth a method we call minimizing between curves (MBC) which chooses the model with least amount of total DIF on the test.
+"Curves" refers to the item characteristic curve which maps a student's ability to their probability of correct response.
+The amount of DIF on each item is calculated as
+
+$$
+\text{DIF}_j = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})| g(\theta)d\theta
+$$
+where $g(\theta)$ is a weighting function such that $\int g(\theta)d\theta = 1$.
+The total DIF on the test, then, is $\text{Total DIF} = \sum_{j} \text{DIF}_j$.
+Different choices for $\Theta^{\text{foc}}$ lead to different $\mathbf{b}^\text{foc}$, and we select the $\Theta^{\text{foc}}$ that minimizes $\text{Total DIF}$.
+The idea for considering the difference between item characteristic curves dates back to @stocking1983developing.
+More recently, @chalmers2016might use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
+
+The selection of $g(\theta)$ is important and determines where on the ability spectrum to consider DIF.
+-@chalmers2016might does not discuss how to choose $g(\theta)$ and in practice uses $g(\theta)$ uniform for $-6 \le \theta \le 6$, which seems like an odd choice.
+It might seem intuitive to choose $g(\theta) \sim N(0, 1)$ because $\mu^\text{ref} = 0$, but this choice doesn't take into account the ability distribution of the focal group.
+If $\mu^\text{foc} = 3$, wouldn't we also want to prioritize high $\theta$ values?
+We set $g(\theta)$ to be the average of the reference and focal group ability distributions:
+
+$$
+g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
+$$
 
@@
 ```
+
+## Bounds for DIF
+
+## Summary

### 2020-02-27 paper/TEST.Rmd
nice progress from morning session, slowly but surely

@@
 - \DeclareMathOperator*{\argmax}{arg\,max}
+- \AtBeginEnvironment{tabular}{\singlespacing}
+- \usepackage{natbib}
 fontsize: 11pt
@@
 
-The Gini index is maximized to select the anchor point.
-The intuition is that we want to find the anchor point such that a small minority of items have bias while the vast majority do not.
-Denote the easiness difference between the two groups when both group means are fixed to zero as $d_j = {b_j}^\text{ref} - {b_j}^\text{foc}$, a function that calculates the Gini index from a vector of non-negative elements as $G(\mathbf{x})$, and the mean of the focal group ability when the reference group ability is fixed to zero (i.e.
-the anchor point) as $\Theta$.
+@lord1980
 
-$$
-\argmax_\Theta G(\mathbf{d} + \Theta)
-$$
+@drasgow1987study
+
+\begin{center}
+ \begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|} 
+ \toprule
+ Method & Description &  Literature \\
+ \midrule
+ AOAA & Test if each item has DIF by using all of the other items as anchors. Not iterative. & Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\ 
+ \hline
+ AOAA-AS & The first iteration is AOAA.
+All items that test positive for DIF are removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& Proposed by \cite{drasgow1987study} \\
+ \hline
+ AOAA-OAT & The first iteration is AOAA.
+However, only the item that shows the most extreme DIF is removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& To our knowledge, not proposed or used previously.
+\\
+ \bottomrule
+ \end{tabular}
+\end{center}
+
+\bibliographystyle{apa}
+\bibliography{paper.bib}

### 2020-02-27 paper/TEST2.Rmd
nice progress from morning session, slowly but surely

--- 2020-02-27 paper/TEST.Rmd
@@
 output: 
-  bookdown::pdf_document2:
-    includes:
+  pdf_document:
     keep_tex: yes
-    latex_engine: xelatex
-    number_sections: yes
-    toc: yes
-    toc_depth: 2
 header-includes: 
-- \usepackage{float}
-- \usepackage[autostyle, english = american]{csquotes}
-- \setlength{\parindent}{4em}
-- \setlength{\parskip}{2em}
-- \usepackage{setspace}\doublespacing
-- \usepackage{amsmath}
-- \DeclareMathOperator*{\argmax}{arg\,max}
-- \AtBeginEnvironment{tabular}{\singlespacing}
 - \usepackage{natbib}
-fontsize: 11pt
-link-citations: yes
-title: "Measurement invariance"
-subtitle: |
-    Ben Stenhaug & Ben Domingue  
-    Stanford University
-abstract: |
-  ABSTRACT TO DO
-  <!-- https://tinyurl.com/ybremelq -->
-bibliography: paper.bib
+- \usepackage{booktabs}
 ---
 
-```{r setup, include = FALSE}
-# 
-# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
-
-knitr::opts_chunk$set(
-    fig.align='center',
-    echo=FALSE,
-    warning=TRUE,
-    fig.pos='H',
-    fig.retina = 5
-)
-
-library(tidyverse)
-library(ggtext)
-library(here)
-library(mirt)
-library(cluster)
-library(DescTools)
-library(knitr)
-library(kableExtra)
-set.seed(123)
-R.utils::sourceDirectory(here("R"))
-
-theme_set(theme_bw())
-```
-
-@lord1980
-
-@drasgow1987study
-
-\begin{center}
+\begin{table}
+\label{table:oos}
+\caption{EDIT THIS}
+\centering
 \begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
 \toprule
+
 Method & Description & Literature \\
+
 \midrule
- AOAA & Test if each item has DIF by using all of the other items as anchors. Not iterative. & Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\ 
- \hline
- AOAA-AS & The first iteration is AOAA.
-All items that test positive for DIF are removed from the anchor set.
-Continue iterating until no new items test positive for DIF.
-& Proposed by \cite{drasgow1987study} \\
- \hline
+
+AOAA & Test if each item has DIF by using all of the other items as anchors. Not iterative. & Originally proposed by  \\\hline
+
+AOAA-AS & The first iteration is AOAA. All items that test positive for DIF are removed from the anchor set. Continue iterating until no new items test positive for DIF. & Proposed by \\\hline
+
 AOAA-OAT & The first iteration is AOAA.
@@
 \\
+
 \bottomrule
 \end{tabular}
-\end{center}
-
-\bibliographystyle{apa}
-\bibliography{paper.bib}
+\end{table}

### 2020-02-27 paper/TEST2.tex
nice progress from morning session, slowly but surely

--- 2020-01-16 paper/paper.tex
@@
 \documentclass[
-  11pt,
 ]{article}
@@
 \hypersetup{
-  pdftitle={Measurement invariance},
   hidelinks,
@@
 \usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
 \usepackage{graphicx,grffile}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-\usepackage{float}
-\usepackage[autostyle, english = american]{csquotes}
-\setlength{\parindent}{4em}
-\setlength{\parskip}{2em}
-\usepackage{setspace}\doublespacing
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{natbib}
 
-\title{Measurement invariance}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Ben Stenhaug \& Ben Domingue\\
-Stanford University}
 \author{}
@@
 \begin{document}
-\maketitle
-\begin{abstract}
-ABSTRACT TO DO
-\end{abstract}
 
-{
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-\clearpage
+\begin{table}
+\label{table:oos}
+\caption{EDIT THIS}
+\centering
+\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
+\toprule
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+Method & Description & Literature \\
 
-Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we view differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
-From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is a property of the student, not the item.
-For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies the DIF.
+\midrule
 
-Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item.
-For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item difficulties for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
-They set item difficulties for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group and generate item responses according to the Rasch model.
-The Rasch model specifies that the probability of student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
-\end{align}
-where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
+AOAA & Test if each item has DIF by using all of the other items as anchors. Not iterative. & Originally proposed by  \\\hline
 
-We find it more informative to describe simulation conditions in terms of single-dimensional items and multidimensional abilities.
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item difficulty is set to what was previously \({b_j}^{\text{ref}}\) for all students.
-Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF and \(a_{2j} = 0\) otherwise.
-Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional 2PL model, the probability student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
-\end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students in the same group.
+AOAA-AS & The first iteration is AOAA. All items that test positive for DIF are removed from the anchor set. Continue iterating until no new items test positive for DIF. & Proposed by \\\hline
 
-If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model?
-Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}, 144) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities.} He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary.
-It's hard to imagine how one would have such knowledge.
-Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
+AOAA-OAT & The first iteration is AOAA.
+However, only the item that shows the most extreme DIF is removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& To our knowledge, not proposed or used previously.
+\\
 
-\hypertarget{dif-methods}{%
-\section{DIF methods}\label{dif-methods}}
-
-Psychometricians have long been in search of the perfect DIF detection method.
-Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method, where each item is tested one at a time for DIF using all other items as anchors.
-More specifically, testing is done using a nested model comparison between the baseline model, where all item parameters are fixed across groups, and the flexible model, where the parameters of the tested item are freed across groups.
-Anchor items have fixed parameters across groups and consequently are used to estimate the ability difference between groups.
-For example, Edelen et al.
-(\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can we be sure its those 10 items and not the other 11 items with DIF?
-Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}).
-In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
-
-Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
-A simple thought experiment illustrates how AOAA fails: Imagine a test with three items where the first item has DIF, and the other two do not.
-Using AOAA, all items test positive for DIF.
-The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until (Andrich and Hagquist \protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}
-
-One way to attempt to counter artificial DIF is with purification with iterative anchor selection.
-For example, (Drasgow \protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then continued iteratively to test the remaining items for DIF until no more items tested positively.
-(Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
-Iterative-backward-AOAA might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
-With a sufficient sample size, this will necessarily be the case.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-Woods suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items display DIF themselves.
-
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge, has not previously been proposed.
-AOAA-OAT starts with AOAA, and the single item exhibiting the most DIF is removed from the anchor set.
-Each item in the anchor set is then tested for DIF again (with items outside of the anchor set allowed to have free parameters across groups in both the baseline and flexible model).
-This process continues until no new items display DIF.
-Applying AOAA-OAT to our thought experiment demonstrates its logic.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
-In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
-
-DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human.
-In some cases, it might make sense to present information to the analyst in a way that empowers them to make decisions.
-Our preferred technique, which we refer to as the \enquote{equal means logit graph} (EMLG), is to set both groups means to 0 so that differences in performance to manifest in the item parameters.
-Relatedly, Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) fit a model with both groups means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
-
-We use a simple simulation to demonstrate EMLG.
-10,000 reference group students and 10,000 focal group students take an eight-item test.
-Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
-We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data.
-We estimate the item parameter covariance matrix using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of the difference in difficulty, \({b_j}^{\text{foc}} - {b_j}^{\text{ref}}\), for each item.
-Figure \ref{fig:emlg} shows these distributions.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emlg-1}
-
-}
-
-\caption{Equal means logit graph (EMLG)}\label{fig:emlg}
-\end{figure}
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ackerman1992didactic}{}%
-Ackerman, Terry A. 1992. ``A Didactic Explanation of Item Bias, Item Impact, and Item Validity from a Multidimensional Perspective.'' \emph{Journal of Educational Measurement} 29 (1): 67--91.
-
-\leavevmode\hypertarget{ref-andrich2012real}{}%
-Andrich, David, and Curt Hagquist. 2012. ``Real and Artificial Differential Item Functioning.'' \emph{Journal of Educational and Behavioral Statistics} 37 (3): 387--416.
-
-\leavevmode\hypertarget{ref-camilli1992conceptual}{}%
-Camilli, Gregory. 1992. ``A Conceptual Analysis of Differential Item Functioning in Terms of a Multidimensional Item Response Model.'' \emph{Applied Psychological Measurement} 16 (2): 129--47.
-
-\leavevmode\hypertarget{ref-chalmers2018numerical}{}%
-Chalmers, R Philip. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
-
-\leavevmode\hypertarget{ref-drasgow1987study}{}%
-Drasgow, Fritz. 1987. ``Study of the Measurement Bias of Two Standardized Psychological Tests.'' \emph{Journal of Applied Psychology} 72 (1): 19.
-
-\leavevmode\hypertarget{ref-edelen2006identification}{}%
-Edelen, Maria Orlando, David Thissen, Jeanne A Teresi, Marjorie Kleinman, and Katja Ocepek-Welikson.
-2006.
-``Identification of Differential Item Functioning Using Item Response Theory and the Likelihood-Based Model Comparison Approach: Application to the Mini-Mental State Examination.'' \emph{Medical Care}, S134--S142.
-
-\leavevmode\hypertarget{ref-kopf2015framework}{}%
-Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.
-
-\leavevmode\hypertarget{ref-meade2012solving}{}%
-Meade, Adam W, and Natalie A Wright. 2012. ``Solving the Measurement Invariance Anchor Item Problem in Item Response Theory.'' \emph{Journal of Applied Psychology} 97 (5): 1016.
-
-\leavevmode\hypertarget{ref-pohl2017cluster}{}%
-Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``CluStEr-baSEd anCHor Item Identification and Selection.'' NEPS Working Paper.
-
-\leavevmode\hypertarget{ref-stark2006detecting}{}%
-Stark, Stephen, Oleksandr S Chernyshenko, and Fritz Drasgow.
-2006.
-``Detecting Differential Item Functioning with Confirmatory Factor Analysis and Item Response Theory: Toward a Unified Strategy.'' \emph{Journal of Applied Psychology} 91 (6): 1292.
-
-\leavevmode\hypertarget{ref-strobl2018anchor}{}%
-Strobl, Carolin, Julia Kopf, Raphael Hartmann, and Achim Zeileis. 2018. ``Anchor Point Selection: An Approach for Anchoring Without Anchor Items.'' Working Papers in Economics; Statistics.
-
-\leavevmode\hypertarget{ref-talbot2013taking}{}%
-Talbot III, Robert M.
-2013.
-``Taking an Item-Level Approach to Measuring Change with the Force and Motion Conceptual Evaluation: An Application of Item Response Theory.'' \emph{School Science and Mathematics} 113 (7): 356--65.
-
-\leavevmode\hypertarget{ref-yang2012characterizing}{}%
-Yang, Ji Seung, Mark Hansen, and Li Cai. 2012. ``Characterizing Sources of Uncertainty in Item Response Theory Scale Scores.'' \emph{Educational and Psychological Measurement} 72 (2): 264--90.
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2020-02-27 paper/paper.Rmd
nice progress from morning session, slowly but surely

@@
     toc_depth: 5
+
+fontsize: 11pt
+bibliography: paper.bib
+link-citations: yes
+
+title: |
+    New and old ways to chase a rabbit:  
+    Differential item functioning detection methods
+subtitle: |
+    Ben Stenhaug  
+    Stanford University
+abstract: |
+  summarize propose extensions insist as dif as student property and on communicating uncertainty
+  <!-- https://tinyurl.com/ybremelq -->
 header-includes: 
 - \usepackage{amsmath}
-- \DeclareMathOperator*{\argmax}{arg\,max}
+- \DeclareMathOperator*{\argmax}{arg\,max} # for gini coefficient
+
 - \usepackage{float}
 - \usepackage[autostyle, english = american]{csquotes}
-- \setlength{\parindent}{4em}
+
+- \setlength{\parindent}{4em} # all of this for draft, papaja should take care of things
 - \setlength{\parskip}{2em}
 - \usepackage{setspace}\doublespacing
-fontsize: 11pt
-link-citations: yes
-title: "Measurement invariance"
-subtitle: |
-    Ben Stenhaug & Ben Domingue  
-    Stanford University
-abstract: |
-  ABSTRACT TO DO
-  <!-- https://tinyurl.com/ybremelq -->
-bibliography: paper.bib
+- \AtBeginEnvironment{tabular}{\singlespacing} # this makes that table single spaced
+
+- \usepackage{natbib} # citations in that table were a hassle. this is still a bit broken. remember can always change tex file.
 ---
@@
 
-<!-- weird formatting, solution is to use papaja (look at paper_apa.Rmd) or this might also be helpful https://stackoverflow.com/questions/4073674/how-to-disable-indentation-in-particular-section-in-latex-->
+<!-- weird formatting that isn't apa I don't think and indentation is weird, solution is to use papaja (look at paper_apa.Rmd) or this might also be helpful https://stackoverflow.com/questions/4073674/how-to-disable-indentation-in-particular-section-in-latex-->
+
+<!-- hopefully all of this below goes away with papaja and other solutions. but remember if it can't editing the latek file is always an option on the final draft! 
+https://github.com/haozhu233/kableExtra/issues/214
+https://stackoverflow.com/questions/51613152/citing-within-an-rmarkdown-table
+https://tex.stackexchange.com/questions/272475/citations-in-tabular-environment-not-working
+https://tex.stackexchange.com/questions/359872/is-it-possible-to-use-cite-inside-xtable
+https://github.com/yihui/rmarkdown-cookbook/issues/71
+-->
+
+\bibliographystyle{apa}
+\bibliography{paper.bib}

### 2020-02-27 paper/paper.tex
nice progress from morning session, slowly but surely

--- 2020-01-16 paper/paper.tex
@@
 \hypersetup{
-  pdftitle={Measurement invariance},
+  pdftitle={New and old ways to chase a rabbit: Differential item functioning detection methods},
   hidelinks,
@@
 \setcounter{secnumdepth}{5}
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \usepackage{float}
@@
 \usepackage{setspace}\doublespacing
+\AtBeginEnvironment{tabular}{\singlespacing}
+\usepackage{natbib}
 \usepackage{booktabs}
@@
 
-\title{Measurement invariance}
+\title{New and old ways to chase a rabbit:\\
+Differential item functioning detection methods}
 \usepackage{etoolbox}
@@
 \makeatother
-\subtitle{Ben Stenhaug \& Ben Domingue\\
+\subtitle{Ben Stenhaug\\
 Stanford University}
@@
 \begin{abstract}
-ABSTRACT TO DO
+summarize propose extensions insist as dif as student property and on communicating uncertainty
 \end{abstract}
@@
 {
-\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{5}
 \tableofcontents
@@
 \clearpage
+
+\hypertarget{to-do}{%
+\section{TO DO}\label{to-do}}
+
+\begin{itemize}
+\tightlist
+\item
+  be sure all my directions match. i think i switched everything to be item easiness at some point but not sure if i did it throughout
+\end{itemize}
 
@@
 For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies the DIF.
+In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.
 
@@
 For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item difficulties for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
-They set item difficulties for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group and generate item responses according to the Rasch model.
-The Rasch model specifies that the probability of student \(i\) responding correctly to item \(j\) is
+They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
+They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
+They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
+They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
 \begin{align}
-    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
+    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
 \end{align}
@@
 
-We find it more informative to describe simulation conditions in terms of single-dimensional items and multidimensional abilities.
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item difficulty is set to what was previously \({b_j}^{\text{ref}}\) for all students.
+We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups.
+For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students.
 Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
@@
 The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF and \(a_{2j} = 0\) otherwise.
+We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
@@
 \begin{align}
-    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
+    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
 \end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students in the same group.
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability for students in the same group.
 
-If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model?
-Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}, 144) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities.} He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary.
+If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model?
+Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities} {[}p.~144{]}.
+He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary {[}p.~144{]}.
 It's hard to imagine how one would have such knowledge.
@@
 Psychometricians have long been in search of the perfect DIF detection method.
-Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method, where each item is tested one at a time for DIF using all other items as anchors.
-More specifically, testing is done using a nested model comparison between the baseline model, where all item parameters are fixed across groups, and the flexible model, where the parameters of the tested item are freed across groups.
-Anchor items have fixed parameters across groups and consequently are used to estimate the ability difference between groups.
-For example, Edelen et al.
+In this section, we summarize existing DIF methods and propose extensions of those methods.
+We limit our discussion to IRT-based methods.
+As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).
+For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
+And, in search of a coherent framework, we sometimes edit names of existing methods.
+We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) and that we risk contributing to the proliferation of names.
+
+We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
+Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{ref}} = -1\).
+The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
+From the DIF-as-item-property view, there is no nuisance ability.
+Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.
+
+In general, we denote the mean ability for the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\).
+We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
+The fundamental challenge for a DIF method is to first estimate \(\mu^\text{foc}\).
+How can \(\mu^\text{foc}\) be disentangled from the possibility that items are biased?
+The most common approach is to identify a group of anchor items which are assumed to be DIF-free.
+The anchor items identify the model because they can be used to estimate \(\mu^\text{foc}\).
+
+\hypertarget{anchor-items}{%
+\subsection{Anchor items}\label{anchor-items}}
+
+\hypertarget{all-others-as-anchors}{%
+\subsubsection{All-others-as-anchors}\label{all-others-as-anchors}}
+
+Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA).
+AOAA tests each item for DIF one at a time using all of the other items as anchors.
+For example, when testing the first item for DIF, all of the other items are used as anchors.
+This is done using a likelihood ratio test that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}).
+Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
+The items for which the flexible model outperforms the baseline model (typically based on a chi-squared test) are identified as having DIF and the rest of the items become anchor items.
+
+Edelen et al.
 (\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
 How can we be sure its those 10 items and not the other 11 items with DIF?
+We can't be.
 Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}).
+More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group and others are biased against the reference group.
+Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method (Strobl et al.
+\protect\hyperlink{ref-strobl2018anchor}{2018}).
 In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
@@
 Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
-A simple thought experiment illustrates how AOAA fails: Imagine a test with three items where the first item has DIF, and the other two do not.
+A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not.
 Using AOAA, all items test positive for DIF.
 The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until (Andrich and Hagquist \protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}
 
-One way to attempt to counter artificial DIF is with purification with iterative anchor selection.
-For example, (Drasgow \protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then continued iteratively to test the remaining items for DIF until no more items tested positively.
-(Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
-Iterative-backward-AOAA might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
-With a sufficient sample size, this will necessarily be the case.
+\hypertarget{all-others-as-anchors-all-significant}{%
+\paragraph{All-others-as-anchors-all-significant}\label{all-others-as-anchors-all-significant}}
+
+One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
+For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively --- with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model --- until no more items tested positively.
+Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free.
+We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
+Appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF being removed from the anchor set.
+
+AOAA-AS might seem like an improvement but it doesn't solve a fundamental problem of AOAA: What do we do when all items test positive for DIF?
+With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
+In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF and there are no anchor items.
 Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-Woods suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
+Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
 It's unclear if one should proceed if those four items display DIF themselves.
 
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge, has not previously been proposed.
-AOAA-OAT starts with AOAA, and the single item exhibiting the most DIF is removed from the anchor set.
-Each item in the anchor set is then tested for DIF again (with items outside of the anchor set allowed to have free parameters across groups in both the baseline and flexible model).
-This process continues until no new items display DIF.
-Applying AOAA-OAT to our thought experiment demonstrates its logic.
+\hypertarget{all-other-as-anchors-one-at-a-time}{%
+\paragraph{All-other-as-anchors-one-at-a-time}\label{all-other-as-anchors-one-at-a-time}}
+
+We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been proposed.
+Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the \(\chi^2\) test statistic, is removed from the anchor set.
+The process continues iteratively until no new items display DIF.
+AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
+
+Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
 The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
@@
 
+\hypertarget{summary-and-performance}{%
+\paragraph{Summary and performance}\label{summary-and-performance}}
+
+Table \ref{table:themethods} summarizes the three all-others-as-anchors methods.
+It's useful to remember that AOAA is not an iterative procedure, while the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
+
+\begin{table}[H]
+\caption{EDIT THIS}
+\centering
+\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
+\toprule
+
+Method & Description & Literature \\
+
+\midrule
+
+AOAA & Test if each item has DIF by using all of the other items as anchors (not iterative). & Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
+
+AOAA-AS & The first iteration is AOAA.
+All items that test positive for DIF are removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& Proposed by \cite{drasgow1987study} \\\hline
+
+AOAA-OAT & The first iteration is AOAA.
+Only the item that shows the most extreme DIF is removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& To our knowledge, not proposed or used previously \\
+
+\bottomrule
+\end{tabular}
+\label{table:themethods}
+\end{table}
+
+{[}ADD IN SIMULATION RESULTS{]}
+
+\hypertarget{clustering-items}{%
+\subsubsection{Clustering items}\label{clustering-items}}
+
+Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster.} They pointed out that one way around the unidentifiability issue is to consider relative item parameters.
+For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\).
+The ultimate matrix of interest is \(\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}\) which is the \enquote{differences between groups in the pairwise differences in (easiness) between items} {[}p.~323{]}.
+
+The general idea of identifying clusters of items is interesting.
+However, their approach is needlessly complicated and they did not give specific suggestions for moving from \(\Delta \mathbf{R}\) to a set of anchor items.
+Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing an approach for identifying the anchor cluster.
+\(\Delta \mathbf{R}\) is skew-symmetric and of rank 1 so that all information is contained in a single row or column.
+Accordingly, they use k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC.
+They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
+However, in their simulation study, they find that BIC identifies only a single cluster, so they use all items as anchors.
+
+We propose a new cluster-based approach which we call \enquote{equal means clustering} (EMC).
+Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\) of differences in item easiness, \(\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}\), where the model is identified by setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), thus the name \enquote{equal means clustering}).
+This results in all differences in performance --- either from group ability differences or DIF --- manifesting in the item easiness difference parameter \(\tilde{d_j}\).
+The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\).
+
+Instead of choosing k with BIC, we use the gap statistic method recommended by Hastie, Tibshirani, and Walther (\protect\hyperlink{ref-hastie2001estimating}{2001}).
+The largest cluster is chosen as the anchor cluster.
+If there is a tie for the largest cluster, the cluster with the lowest standard deviation of item easiness differences is selected.
+
+OUR SIMULATION RESULTS
+
+\begin{tabular}{l|r|r|r|r|l}
+\hline
+item & a\_ref\_easy & b\_foc\_easy & difference\_in\_easy & cluster & anchor\\
+\hline
+item1 & -0.5638213 & -1.5520228 & 0.9882015 & 1 & TRUE\\
+\hline
+item2 & -0.2618317 & -1.2619495 & 1.0001178 & 1 & TRUE\\
+\hline
+item3 & 1.5747215 & 0.5456186 & 1.0291029 & 1 & TRUE\\
+\hline
+item4 & 0.0521493 & -0.9183001 & 0.9704494 & 1 & TRUE\\
+\hline
+item5 & 0.1335217 & -0.8599758 & 0.9934975 & 1 & TRUE\\
+\hline
+item6 & 1.6971526 & 0.1926118 & 1.5045408 & 2 & FALSE\\
+\hline
+item7 & 0.4629179 & -1.0572527 & 1.5201706 & 2 & FALSE\\
+\hline
+item8 & -1.2953061 & -2.8759098 & 1.5806036 & 2 & FALSE\\
+\hline
+\end{tabular}
+
+\hypertarget{analyst-chooses}{%
+\subsubsection{Analyst chooses}\label{analyst-chooses}}
+
 DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human.
-In some cases, it might make sense to present information to the analyst in a way that empowers them to make decisions.
-Our preferred technique, which we refer to as the \enquote{equal means logit graph} (EMLG), is to set both groups means to 0 so that differences in performance to manifest in the item parameters.
+We refer to DIF methods that do not use an analyst's judgement as \enquote{algorithmic DIF methods.} On the other hand, it's often useful to present information to the analyst in a way that empowers them to see DIF and potentially choose the anchor items.
+
+We propose a technique, the \enquote{equal means logit graph} (EMLG), where both group means are to 0 so that differences in performance manifest exclusively in the item parameters.
 Relatedly, Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) fit a model with both groups means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
 
-We use a simple simulation to demonstrate EMLG.
-10,000 reference group students and 10,000 focal group students take an eight-item test.
-Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
 We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data.
@@
 Multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of the difference in difficulty, \({b_j}^{\text{foc}} - {b_j}^{\text{ref}}\), for each item.
-Figure \ref{fig:emlg} shows these distributions.
+Figure \ref{fig:emlg} shows these distributions in the EMLG.
 
@@
 
+It cannot be stated strongly enough that the EMLG contains all possible information about the difference in group performance.
+The challenge, then, is to identify the difference in group ability.
+The analyst might assume that --- because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and should be used as anchor items.
+At this point, the analyst re-fits the model with the anchor items having fixed parameters across groups so that they are used to estimate the difference in group ability.
+The parameters of the other items are free to vary across groups, and, indeed, the new model finds that the item easiness difference, \({b_j}^{\text{ref}} - {b_j}^{\text{foc}}\), is about 0.5 logits for each of these items.
+
+One way to show this:
+
+\begin{verbatim}
+## Picking joint bandwidth of 0.00548
+\end{verbatim}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/who-1}
+
+}
+
+\caption{fix this}\label{fig:who}
+\end{figure}
+
+Another way to show this:
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/why-1}
+
+}
+
+\caption{fix this}\label{fig:why}
+\end{figure}
+
+One of our key concerns with algorithmic DIF methods is that they can lull the analyst into a false sense of security.
+Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
+The EMLG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst is may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even in cases where algorithmic DIF methods are used, the analyst should use the EMLG to give them a sense of their item response data.
+
+\hypertarget{anchor-points}{%
+\subsection{Anchor points}\label{anchor-points}}
+
+All of the previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
+The difference in group abilities is then estimated using exclusively the anchor items.
+An alternative strategy is to select an anchor point, which essentially specifies the difference in group abilities directly.
+Anchor point methods have the advantage of not assuming that any particular item is DIF-free and therefore, all items can be tested for DIF.
+
+\hypertarget{gini-index}{%
+\subsubsection{Gini index}\label{gini-index}}
+
+Strobl et al.
+(\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (\protect\hyperlink{ref-gini1912variabilita}{1912}) for selecting anchor points.
+The Gini index is typically used to measure the inequality of wealth distribution in a country.
+South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution Chitiga, Sekyere, and Tsoanamatsie (\protect\hyperlink{ref-chitiga2015income}{2015}).
+In general, the Gini index \enquote{takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little} (Strobl et al.
+\protect\hyperlink{ref-strobl2018anchor}{2018}, 7).
+
+The Gini index is maximized to select the anchor point.
+The intuition is that we want to find the anchor point such that a small minority of items have bias while the vast majority do not.
+Denote the easiness difference between the two groups when both group means are fixed to zero as \(d_j = {b_j}^\text{ref} - {b_j}^\text{foc}\), a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), and the mean of the focal group ability when the reference group ability is fixed to zero (i.e.~the anchor point) as \(\Theta\).
+The anchor point is set to
+\[
+\mathop\mathrm{arg\,max}_\Theta G(\mathbf{d} + \Theta)
+\]
+where \(\Theta \in (-\infty, \infty)\) and \(\Theta\) is added to each element of \(\mathbf{d}\).
+
+Figure \ref{fig:workplease} shows the gini coefficient at a variety of possible values for \(\Theta\). We find that the anchor point is -0.99.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/workplease-1}
+
+}
+
+\caption{always need a capt}\label{fig:workplease}
+\end{figure}
+
+\hypertarget{minimizing-between-curves}{%
+\subsubsection{Minimizing between curves}\label{minimizing-between-curves}}
+
+Race created a statistical test which was one of the most popular methods used until recently.
+
+Imagine multigroup item response data where \(\Theta^{\text{foc}} = \Theta^{\text{ref}}\) and \(d_j = 0 \forall j\) so that there is no DIF.
+The fundamental identification problem of DIF in multipgroup IRT models is that
+there are an infinite number of models with the same likelihood from which to choose.
+For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\Theta^{\text{foc}} = 0\).
+The model would then estimate \(d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF.
+Alternatively, we could assume that the focal group has very high ability and fix \(\Theta^{\text{foc}} = 3\).
+The model would then compensate by finding \(d_j \approx -3 \forall j\), and we would incorrectly conclude that the the focal group is high ability but each item contains DIF against the focal group.
+Both of these models have the same likelihood, so how should we choose which model to believe?
+This is the fundamental identification problem.
+
+We put forth a method we call minimizing between curves (MBC) which chooses the model with least amount of total DIF on the test.
+\enquote{Curves} refers to the item characteristic curve which maps a student's ability to their probability of correct response.
+The amount of DIF on each item is calculated as
+
+\[
+\text{DIF}_j = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})| g(\theta)d\theta
+\]
+where \(g(\theta)\) is a weighting function such that \(\int g(\theta)d\theta = 1\).
+The total DIF on the test, then, is \(\text{Total DIF} = \sum_{j} \text{DIF}_j\).
+Different choices for \(\Theta^{\text{foc}}\) lead to different \(\mathbf{b}^\text{foc}\), and we select the \(\Theta^{\text{foc}}\) that minimizes \(\text{Total DIF}\).
+The idea for considering the difference between item characteristic curves dates back to Stocking and Lord (\protect\hyperlink{ref-stocking1983developing}{1983}).
+More recently, Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}) use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
+
+The selection of \(g(\theta)\) is important and determines where on the ability spectrum to consider DIF.
+(\protect\hyperlink{ref-chalmers2016might}{2016}) does not discuss how to choose \(g(\theta)\) and in practice uses \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which seems like an odd choice.
+It might seem intuitive to choose \(g(\theta) \sim N(0, 1)\) because \(\mu^\text{ref} = 0\), but this choice doesn't take into account the ability distribution of the focal group.
+If \(\mu^\text{foc} = 3\), wouldn't we also want to prioritize high \(\theta\) values?
+We set \(g(\theta)\) to be the average of the reference and focal group ability distributions:
+
+\[
+g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
+\]
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/workplease2-1}
+
+}
+
+\caption{always need a capt}\label{fig:workplease2}
+\end{figure}
+
+\hypertarget{bounds-for-dif}{%
+\subsection{Bounds for DIF}\label{bounds-for-dif}}
+
+\hypertarget{summary}{%
+\subsection{Summary}\label{summary}}
+
+\hypertarget{next-steps}{%
+\section{Next steps}\label{next-steps}}
+
+\hypertarget{simulate-and-then-walk-through-as-the-analyst-would-with-each-method}{%
+\section{Simulate and then walk through as the analyst would with each method}\label{simulate-and-then-walk-through-as-the-analyst-would-with-each-method}}
+
+\hypertarget{mike-frank-example}{%
+\section{Mike Frank example}\label{mike-frank-example}}
+
+We know that this test contains DIF.
+Which items depends on ability difference which we can be sure of.
+We can, however, quantify the amount of DIF on this test.
+We can take the SD or we can assume between and calculate chalmers statistic for each.
+We might not know for sure the ability difference but we can quantify the potential for DIF.
+For each we calculate the SD across item parameters.
+
+Something with DTF?
+
+Do a simulation where the second dimension varies - a more real testing ground
+
+Example with real data perhaps the Mike Frank data
+
+Summer learning loss as applied DIF?
+
+\clearpage
+
+\hypertarget{notes}{%
+\section{Notes}\label{notes}}
+
+\begin{itemize}
+\tightlist
+\item
+  adjust to all be easiness including multidimensional talk
+\end{itemize}
+
 \hypertarget{references}{%
 \section{References}\label{references}}
+
+\bibliographystyle{apa}
+\bibliography{paper.bib}
 
@@
 
+\leavevmode\hypertarget{ref-bechger2015statistical}{}%
+Bechger, Timo M, and Gunter Maris. 2015. ``A Statistical Test for Differential Item Pair Functioning.'' \emph{Psychometrika} 80 (2): 317--40.
+
 \leavevmode\hypertarget{ref-camilli1992conceptual}{}%
@@
 Chalmers, R Philip. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
+
+\leavevmode\hypertarget{ref-chalmers2016might}{}%
+Chalmers, R Philip, Alyssa Counsell, and David B Flora.
+2016.
+``It Might Not Make a Big Dif: Improved Differential Test Functioning Statistics That Account for Sampling Variability.'' \emph{Educational and Psychological Measurement} 76 (1): 114--40.
+
+\leavevmode\hypertarget{ref-chitiga2015income}{}%
+Chitiga, Margaret, E Sekyere, and N Tsoanamatsie.
+2015.
+``Income Inequality and Limitations of the Gini Index: The Case of South Africa.'' \emph{Human Sciences Research Council (HSRC), Available at: Http://Www.
+Hsrc.
+Ac.
+Za/En/Review/Hsrc-Review-November-2014/Limitations-of-Gini-Index, Site Accessed} 2.
 
@@
 ``Identification of Differential Item Functioning Using Item Response Theory and the Likelihood-Based Model Comparison Approach: Application to the Mini-Mental State Examination.'' \emph{Medical Care}, S134--S142.
+
+\leavevmode\hypertarget{ref-gini1912variabilita}{}%
+Gini, Corrado. 1912. ``Variabilità E Mutabilità (Variability and Mutability).'' \emph{Tipografia Di Paolo Cuppini, Bologna, Italy}, 156.
+
+\leavevmode\hypertarget{ref-hastie2001estimating}{}%
+Hastie, Trevor, Robert Tibshirani, and Guenther Walther. 2001. ``Estimating the Number of Data Clusters via the Gap Statistic.'' \emph{J Roy Stat Soc B} 63: 411--23.
+
+\leavevmode\hypertarget{ref-holland1986differential}{}%
+Holland, Paul W, and Dorothy T Thayer. 1986. ``Differential Item Functioning and the Mantel-Haenszel Procedure.'' \emph{ETS Research Report Series} 1986 (2): i--24.
 
@@
 
+\leavevmode\hypertarget{ref-stocking1983developing}{}%
+Stocking, Martha L, and Frederic M Lord. 1983. ``Developing a Common Metric in Item Response Theory.'' \emph{Applied Psychological Measurement} 7 (2): 201--10.
+
 \leavevmode\hypertarget{ref-strobl2018anchor}{}%
 Strobl, Carolin, Julia Kopf, Raphael Hartmann, and Achim Zeileis. 2018. ``Anchor Point Selection: An Approach for Anchoring Without Anchor Items.'' Working Papers in Economics; Statistics.
+
+\leavevmode\hypertarget{ref-swaminathan1990detecting}{}%
+Swaminathan, Hariharan, and H Jane Rogers. 1990. ``Detecting Differential Item Functioning Using Logistic Regression Procedures.'' \emph{Journal of Educational Measurement} 27 (4): 361--70.
 
@@
 
+\leavevmode\hypertarget{ref-thissen1993detection}{}%
+Thissen, David, Lynne Steinberg, and Howard Wainer. 1993. ``Detection of Differential Item Functioning Using the Parameters of Item Response Models.''
+
+\leavevmode\hypertarget{ref-woods2009empirical}{}%
+Woods, Carol M. 2009. ``Empirical Selection of Anchors for Tests of Differential Item Functioning.'' \emph{Applied Psychological Measurement} 33 (1): 42--57.
+
 \leavevmode\hypertarget{ref-yang2012characterizing}{}%

### 2020-02-27 paper/paper_1_intro.Rmd
nice progress from morning session, slowly but surely

@@
 From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item.
-For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but their performance also depends on their verbal ability.
+For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
 In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.
@@
 For example, @kopf2015framework simulate students as belonging to either a reference or focal group.
-They fix the item difficulties for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study.
-They set item difficulties for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group and generate item responses according to the Rasch model.
-The Rasch model specifies that the probability of student $i$ responding correctly to item $j$ is
+They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study.
+They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits.
+They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
+They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is
 \begin{align}
-    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
+    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
 \end{align}
@@
 
-We find it more informative to describe simulation conditions in terms of single-dimensional items and multidimensional abilities.
-For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item difficulty is set to what was previously ${b_j}^{\text{ref}}$ for all students.
+We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups.
+For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students.
 Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
@@
 The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group.
-We now need to use a 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF and $a_{2j} = 0$ otherwise.
+We now need to use a 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
@@
 \begin{align}
-    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
+    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
 \end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students of the same group.
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability for students in the same group.
 
-If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model?
-@camilli1992conceptual [p.
-144] tested this idea with the goal of a "more satisfying description of the secondary abilities." He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that "a priori knowledge of the true factor structure" is necessary.
+If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model?
+@camilli1992conceptual tested this idea with the goal of a "more satisfying description of the secondary abilities" [p.
+144].
+He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that "a priori knowledge of the true factor structure" is necessary [p.
+144].
 It's hard to imagine how one would have such knowledge.

### 2020-02-27 paper/paper_2_dif_methods.Rmd
nice progress from morning session, slowly but surely

@@
 Psychometricians have long been in search of the perfect DIF detection method.
-In the following three sections, we summarize DIF methods and propose new DIF methods.
-We use a simple simulation to demonstrate DIF methods: 10,000 reference group students and 10,000 focal group students take an eight-item test.
+In this section, we summarize existing DIF methods and propose extensions of those methods.
+We limit our discussion to IRT-based methods.
+As a result, we avoid methods like the Mantel-Haenszel procedure [@holland1986differential], which muddies the waters by moving away from the IRT framework, and has been shown to perform no better — and in some cases worse — than IRT-based methods [@swaminathan1990detecting].
+For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
+And, in search of a coherent framework, we sometimes edit names of existing methods.
+We recognize that others have done the same [e.g.
+@kopf2015framework] and that we risk contributing to the proliferation of names.
+
+We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group.
+Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{ref}} = -1$.
 The slope on target ability is set to $a_{1j} = 1$ for all items, and the slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise.
+From the DIF-as-item-property view, there is no nuisance ability.
+Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.
 
-## All others as anchors
+In general, we denote the mean ability for the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$.
+We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
+The fundamental challenge for a DIF method is to first estimate $\mu^\text{foc}$.
+How can $\mu^\text{foc}$ be disentangled from the possibility that items are biased?
+The most common approach is to identify a group of anchor items which are assumed to be DIF-free.
+The anchor items identify the model because they can be used to estimate $\mu^\text{foc}$.
 
-@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method in which each item is tested one at a time for DIF using all other items as anchors.
-More specifically, testing is done using a nested model comparison between the baseline model, where all item parameters are fixed across groups, and the flexible model, where the parameters of the tested item are freed across groups.
-Anchor items have fixed parameters across groups and consequently are used to estimate the ability difference between groups.
-For example, @edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
+## Anchor items
+
+### All-others-as-anchors
+
+@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA).
+AOAA tests each item for DIF one at a time using all of the other items as anchors.
+For example, when testing the first item for DIF, all of the other items are used as anchors.
+This is done using a likelihood ratio test that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection].
+Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
+The items for which the flexible model outperforms the baseline model (typically based on a chi-squared test) are identified as having DIF and the rest of the items become anchor items.
+
+@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
 How can we be sure its those 10 items and not the other 11 items with DIF?
+We can't be.
 Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method @strobl2018anchor.
+More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group and others are biased against the reference group.
+Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method [@strobl2018anchor].
 In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
@@
 
-One way to attempt to counter artificial DIF is with purification with iterative anchor selection.
-For example, [@drasgow1987study] started with AOAA, removed items displaying DIF from the anchor set, then continued iteratively to test the remaining items for DIF until no more items tested positively.
-[@kopf2015framework] named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF free.
-We find this name confusing and instead refer to this method as all-others-as-anchors-all-significiant (AOAA-AS) where appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF are removed from the anchor set.
-AOAA-AS might seem like an improvement, but what does one do when all items display DIF in the first AOAA stage?
-With a sufficient sample size, this will necessarily be the case.
+#### All-others-as-anchors-all-significant
+
+One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
+For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively.
+@kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free.
+We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
+Appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF being removed from the anchor set.
+
+AOAA-AS might seem like an improvement but it doesn't solve a fundamental problem of AOAA: What do we do when all items test positive for DIF?
+With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
+In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF and there are no anchor items.
 @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
@@
 
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAAT), which, to our knowledge (and surprise), has not previously been proposed.
-AOAA-OAT starts with AOAA, and the single item exhibiting the most DIF is removed from the anchor set.
-Each item in the anchor set is then tested for DIF again (with items outside of the anchor set allowed to have free parameters across groups in both the baseline and flexible model).
-This process continues until no new items display DIF.
-AOAA-OAAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAAT takes the more conservative approach of removing only the item exhibiting the most DIF.
-Applying AOAA-OAT to our thought experiment demonstrates its logic.
+#### All-other-as-anchors-one-at-a-time
+
+We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been proposed.
+Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the $\chi^2$ test statistic, is removed from the anchor set.
+The process continues iteratively until no new items display DIF.
+AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
+
+Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
 The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
@@
 AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
-Table X summarizes these methods.
-In wrapping one's mind around these methods it's useful to remember that AOAA is not an iterative procedure, while the methods with a dash, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the most conservative.
 
-ALSO MAKE A TABLE OF SIMULATION RESULTS
+#### Summary and performance
+
+Table \ref{table:themethods} summarizes the three all-others-as-anchors methods.
+It's useful to remember that AOAA is not an iterative procedure, while the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
 
@@
 
-\begin{center}
- \begin{tabular}{||c c c c c||} 
- \hline
- Name & Abbreviation & Description & Assumptions & Literature \\ [0.5ex] 
- \hline\hline
- All-others-as-anchors & AOAA & 87837 & 787 & 1\\ 
- \hline
- All-others-as-anchors-all-significant & AOAA-AS & 78 & 5415 & 1\\
- \hline
- All-others-as-anchors-one-at-a-time & AOAA-OAAT & 778 & 7507 & 1\\
- \hline
- 4 & 545 & 18744 & 7560 & 1\\
- \hline
- 5 & 88 & 788 & 6344 & 1\\ [1ex] 
- \hline
+\begin{table}[H]
+\caption{EDIT THIS}
+\centering
+\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
+\toprule
+
+Method & Description & Literature \\
+
+\midrule
+
+AOAA & Test if each item has DIF by using all of the other items as anchors (not iterative). & Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
+
+AOAA-AS & The first iteration is AOAA.
+All items that test positive for DIF are removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& Proposed by \cite{drasgow1987study} \\\hline
+
+AOAA-OAT & The first iteration is AOAA.
+Only the item that shows the most extreme DIF is removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& To our knowledge, not proposed or used previously \\
+
+\bottomrule
 \end{tabular}
-\end{center}
+\label{table:themethods}
+\end{table}
 
-## Presenting information to the analyst
+[ADD IN SIMULATION RESULTS]
+
+### Clustering items
+
+```{r}
+# source(here("paper/paper_2_sim.R"))
+# out_3_biased_items %>% write_rds(here("paper", "paper_2_out_3_biased_items.rds"))
+out_3_biased_items <- read_rds(here("paper", "paper_2_out_3_biased_items.rds"))
+```
+
+@bechger2015statistical proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider relative item parameters.
+For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$.
+The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p.
+323].
+
+The general idea of identifying clusters of items is interesting.
+However, their approach is needlessly complicated and they did not give specific suggestions for moving from $\Delta \mathbf{R}$ to a set of anchor items.
+@pohl2017cluster extended their work by proposing an approach for identifying the anchor cluster.
+$\Delta \mathbf{R}$ is skew-symmetric and of rank 1 so that all information is contained in a single row or column.
+Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
+They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
+However, in their simulation study, they find that BIC identifies only a single cluster, so they use all items as anchors.
+
+We propose a new cluster-based approach which we call "equal means clustering" (EMC).
+Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$ of differences in item easiness, $\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}$, where the model is identified by setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, thus the name "equal means clustering").
+This results in all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$.
+The tilde is used to denote parameters estimated with $\mu^\text{foc}$ set to $0$.
+
+Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating.
+The largest cluster is chosen as the anchor cluster.
+If there is a tie for the largest cluster, the cluster with the lowest standard deviation of item easiness differences is selected.
+
+OUR SIMULATION RESULTS
+
+```{r}
+out_3_biased_items$cluster_status[[1]] %>% 
+    select(item, a_ref_easy, b_foc_easy, difference_in_easy, cluster) %>% 
+    mutate(anchor = ifelse(cluster == 1, TRUE, FALSE)) %>% 
+    kable()
+```
+
+### Analyst chooses
 
 DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human.
-We refer to DIF methods that do not use an analyst's judgement as "algorithmic DIF methods." On the other hand, it might make sense to present information to the analyst in a way that empowers them to make decisions in some cases.
+We refer to DIF methods that do not use an analyst's judgement as "algorithmic DIF methods." On the other hand, it's often useful to present information to the analyst in a way that empowers them to see DIF and potentially choose the anchor items.
+
 We propose a technique, the "equal means logit graph" (EMLG), where both group means are to 0 so that differences in performance manifest exclusively in the item parameters.
@@
 ```{r emlg, fig.cap = 'Equal means logit graph (EMLG)', out.width="70%", warning = FALSE, message = FALSE}
-# source(here("paper/paper_2_sim.R"))
-# out_3_biased_items %>% write_rds(here("paper", "paper_2_out_3_biased_items.rds"))
-out_3_biased_items <- read_rds(here("paper", "paper_2_out_3_biased_items.rds"))
-
 out_3_biased_items$intuitive_mod[[1]] %>%
@@
 
-## Clustering
-
-@bechger2015statistical proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue of item response models is to arbitrarily set an item easiness parameter to 0.
-For each group, the relative difficulties for each pair of items are stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$.
-The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in difficulty between items" [p.
-323].
-
-The general idea of identifying clusters of items is interesting.
-However, their approach is needlessly complicated and they did not give specific suggestions on how to move from $\Delta \mathbf{R}$ to a set of anchor items.
-@pohl2017cluster extended their work by proposing an approach for identifying anchor items.
-$\Delta \mathbf{R}$ is skew-symmetric and of rank 1 so that all information is contained in a single row or column.
-Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
-They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-However, in their simulation study, they find that BIC identifies only a single cluster, so they use all items as anchors.
-
-We propose a new cluster-based approach which we refer to as equal means clustering (EMC).
-Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector of differences in item easiness, ${b_j}^\text{ref} - {b_j}^\text{foc}$, when both groups have their means fixed to 0 as is the case in the EMLG.
-Instead of choosing k with BIC, we use the gap statistic method reccommended by @hastie2001estimating.
-The largest cluster is chosen as the anchor cluster.
-If there is a tie for the largest cluster, the cluster with the lowest standard deviation of item easiness differences is selected.
-
-Using this method, we find that _________.
-
-```{r}
-out_3_biased_items$cluster_status[[1]] %>% 
-    select(item, a_ref_easy, b_foc_easy, difference_in_easy, cluster) %>% 
-    mutate(anchor = ifelse(cluster == 1, TRUE, FALSE))
-```
-
 ## Anchor points
@@
 ### Minimizing between curves
+
+Race created a statistical test which was one of the most popular methods used until recently.
 

### 2020-02-29 paper/paper.Rmd
we're getting there!

@@
 - \DeclareMathOperator*{\argmax}{arg\,max} # for gini coefficient
+- \DeclareMathOperator*{\argmin}{arg\,min} # for gini coefficient
 
@@
 
-```{r child = "paper_2_dif_methods.Rmd"}
+```{r child = "paper_2a_anchor_items.Rmd"}
+```
+
+## Anchor points
+
+```{r child = "paper_2b_anchor_points.Rmd"}
 ```

### 2020-02-29 paper/paper.tex
we're getting there!

@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 \usepackage{float}
@@
 
-Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we view differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
+Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we think differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
 From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is a property of the student, not the item.
@@
 
-In general, we denote the mean ability for the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\).
+In general, we denote the mean ability of the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\), respectively.
 We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
-The fundamental challenge for a DIF method is to first estimate \(\mu^\text{foc}\).
-How can \(\mu^\text{foc}\) be disentangled from the possibility that items are biased?
-The most common approach is to identify a group of anchor items which are assumed to be DIF-free.
-The anchor items identify the model because they can be used to estimate \(\mu^\text{foc}\).
+The fundamental challenge for a DIF method is to first get an estimate for \(\mu^\text{foc}\), \(\hat\mu^\text{foc}\).
+How can group ability be disentangled from the possibility that items are biased?
+The most common approach is to identify a group of anchor items that are assumed to be DIF-free.
+These items are used to get \(\hat\mu^\text{foc}\).
 
@@
 AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
+As a result, AOAA-OAT is less likely to arrive at the wrong conclusion because of artifical DIF.
 
@@
 
-The general idea of identifying clusters of items is interesting.
-However, their approach is needlessly complicated and they did not give specific suggestions for moving from \(\Delta \mathbf{R}\) to a set of anchor items.
-Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing an approach for identifying the anchor cluster.
-\(\Delta \mathbf{R}\) is skew-symmetric and of rank 1 so that all information is contained in a single row or column.
+The general idea of identifying clusters of items is intriguing.
+However, their approach is needlessly complicated and they did not describe a process for moving from \(\Delta \mathbf{R}\) to a set of anchor items.
+Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing one such process.
+\(\Delta \mathbf{R}\) is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
 Accordingly, they use k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC.
 They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-However, in their simulation study, they find that BIC identifies only a single cluster, so they use all items as anchors.
+Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.
 
-We propose a new cluster-based approach which we call \enquote{equal means clustering} (EMC).
+We propose a new cluster-based approach, which we call \enquote{equal means clustering} (EMC).
 Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\) of differences in item easiness, \(\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}\), where the model is identified by setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), thus the name \enquote{equal means clustering}).
 This results in all differences in performance --- either from group ability differences or DIF --- manifesting in the item easiness difference parameter \(\tilde{d_j}\).
-The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\).
+The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) arbitrarily set to \(0\).
 
@@
 The largest cluster is chosen as the anchor cluster.
-If there is a tie for the largest cluster, the cluster with the lowest standard deviation of item easiness differences is selected.
+If there is a tie for the largest cluster, the cluster with the lowest standard deviation of \(\tilde{\mathbf{d}}\) is selected.
+This process makes the assumption that the largest cluster (but necessarily the majority) of items do not contain DIF.
 
@@
 
-DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human.
-We refer to DIF methods that do not use an analyst's judgement as \enquote{algorithmic DIF methods.} On the other hand, it's often useful to present information to the analyst in a way that empowers them to see DIF and potentially choose the anchor items.
+DIF methods are generally designed to automatically detect DIF without any human judgement.
+On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
 
-We propose a technique, the \enquote{equal means logit graph} (EMLG), where both group means are to 0 so that differences in performance manifest exclusively in the item parameters.
-Relatedly, Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) fit a model with both groups means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
+We propose a technique, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
+Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc}\) to \(0\).
+Again, the result is that all differences in performance manifest in \(\tilde{\mathbf{d}}\).
+To get a sense of variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
+Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of \(\tilde{d_j}\) for every item.
+These are the distributions displayed in the EM-MILG.
+The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
 
-We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data.
-We estimate the item parameter covariance matrix using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of the difference in difficulty, \({b_j}^{\text{foc}} - {b_j}^{\text{ref}}\), for each item.
-Figure \ref{fig:emlg} shows these distributions in the EMLG.
+Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data.
+It cannot be stated strongly enough that the EM-MILG contains all possible information about the difference in group performance.
+The challenge, then, is to select the anchor items.
+The analyst might assume that --- because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emlg-1}
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emmilg-1}
 
@@
 
-\caption{Equal means logit graph (EMLG)}\label{fig:emlg}
+\caption{The equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
 \end{figure}
 
-It cannot be stated strongly enough that the EMLG contains all possible information about the difference in group performance.
-The challenge, then, is to identify the difference in group ability.
-The analyst might assume that --- because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and should be used as anchor items.
-At this point, the analyst re-fits the model with the anchor items having fixed parameters across groups so that they are used to estimate the difference in group ability.
-The parameters of the other items are free to vary across groups, and, indeed, the new model finds that the item easiness difference, \({b_j}^{\text{ref}} - {b_j}^{\text{foc}}\), is about 0.5 logits for each of these items.
-
-One way to show this:
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.00548
-\end{verbatim}
+After choosing the anchor items, the model is refit.
+The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc}\) to \(0\).
+The same process of using multiple imputations to estimate the distribution of \(d_j\) can be used with the new model.
+Because the equal means assumption is not made, we refer to the resulting graph as a multiple imputations logit graph (MILG).
+In our simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/who-1}
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/milg-1}
 
@@
 
-\caption{fix this}\label{fig:who}
+\caption{The multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed to 0.}\label{fig:milg}
 \end{figure}
 
-Another way to show this:
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/why-1}
-
-}
-
-\caption{fix this}\label{fig:why}
-\end{figure}
-
-One of our key concerns with algorithmic DIF methods is that they can lull the analyst into a false sense of security.
+One of our key concerns with typical DIF methods is that they can lull the analyst into a false sense of security.
 Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
-The EMLG combats this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst is may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even in cases where algorithmic DIF methods are used, the analyst should use the EMLG to give them a sense of their item response data.
+The EM-MILG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even when other DIF methods are used, the analyst can use the EM-MILG as a first step in order to give them a sense of their item response data.
+And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
 
@@
 
-All of the previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
-The difference in group abilities is then estimated using exclusively the anchor items.
-An alternative strategy is to select an anchor point, which essentially specifies the difference in group abilities directly.
-Anchor point methods have the advantage of not assuming that any particular item is DIF-free and therefore, all items can be tested for DIF.
+The previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
+The anchor items are used to estimate \(\hat \mu^\text{foc}\).
+An alternative strategy is to directly select the anchor point, \(\mu^{\star\text{foc}}\).
+Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, all items can be tested for DIF.
+The question then becomes the following: How do we select the anchor point?
 
@@
 Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (\protect\hyperlink{ref-gini1912variabilita}{1912}) for selecting anchor points.
+(\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (\protect\hyperlink{ref-gini1912variabilita}{1912}) to select the anchor point.
 The Gini index is typically used to measure the inequality of wealth distribution in a country.
-South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution Chitiga, Sekyere, and Tsoanamatsie (\protect\hyperlink{ref-chitiga2015income}{2015}).
+For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution Chitiga, Sekyere, and Tsoanamatsie (\protect\hyperlink{ref-chitiga2015income}{2015}).
 In general, the Gini index \enquote{takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little} (Strobl et al.
@@
 
-The Gini index is maximized to select the anchor point.
-The intuition is that we want to find the anchor point such that a small minority of items have bias while the vast majority do not.
-Denote the easiness difference between the two groups when both group means are fixed to zero as \(d_j = {b_j}^\text{ref} - {b_j}^\text{foc}\), a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), and the mean of the focal group ability when the reference group ability is fixed to zero (i.e.~the anchor point) as \(\Theta\).
-The anchor point is set to
+The anchor point, \(\mu^{\star\text{foc}}\), is selected by maximizing the Gini index.
+The intuition and assumption is that some items might have bias but the majority do not.
+Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\),\\
 \[
-\mathop\mathrm{arg\,max}_\Theta G(\mathbf{d} + \Theta)
+\mu^{\star\text{foc}} = \mathop\mathrm{arg\,max}_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
 \]
-where \(\Theta \in (-\infty, \infty)\) and \(\Theta\) is added to each element of \(\mathbf{d}\).
+where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\mathbf{d}\).
 
-Figure \ref{fig:workplease} shows the gini coefficient at a variety of possible values for \(\Theta\). We find that the anchor point is -0.99.
+For the simulated data, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible values for \(\mu^\text{foc}\). The anchor point is \(\mu^\text{foc} = -0.99\).
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/workplease-1}
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginipath-1}
 
@@
 
-\caption{always need a capt}\label{fig:workplease}
+\caption{Selecting the anchor point by maximizing the Gini index.}\label{fig:ginipath}
+\end{figure}
+
+We then refit the model with the identifying assumption of \(\mu^{\star\text{foc}} = -0.99\). We test each item for DIF and display the results in a MILG as is shown in Figure \ref{fig:ginimilg}.
+
+\begin{verbatim}
+## Picking joint bandwidth of 0.00544
+\end{verbatim}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginimilg-1}
+
+}
+
+\caption{Selecting the anchor point by maximizing the Gini index.}\label{fig:ginimilg}
 \end{figure}
@@
 
-Race created a statistical test which was one of the most popular methods used until recently.
+Raju's area method measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group (Raju \protect\hyperlink{ref-raju1988area}{1988}):
 
-Imagine multigroup item response data where \(\Theta^{\text{foc}} = \Theta^{\text{ref}}\) and \(d_j = 0 \forall j\) so that there is no DIF.
-The fundamental identification problem of DIF in multipgroup IRT models is that
-there are an infinite number of models with the same likelihood from which to choose.
-For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\Theta^{\text{foc}} = 0\).
-The model would then estimate \(d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF.
-Alternatively, we could assume that the focal group has very high ability and fix \(\Theta^{\text{foc}} = 3\).
-The model would then compensate by finding \(d_j \approx -3 \forall j\), and we would incorrectly conclude that the the focal group is high ability but each item contains DIF against the focal group.
-Both of these models have the same likelihood, so how should we choose which model to believe?
-This is the fundamental identification problem.
+\begin{align}
+    \text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
+\end{align}
 
-We put forth a method we call minimizing between curves (MBC) which chooses the model with least amount of total DIF on the test.
-\enquote{Curves} refers to the item characteristic curve which maps a student's ability to their probability of correct response.
+Raju's area method, along with AOAA, have been cited as the two most commonly used IRT-based DIF detection methods (Magis et al.
+\protect\hyperlink{ref-magis2011generalized}{2011}).
+We see two weakness of Raju's area method: First, the item characteric curves still need to be linked by anchor items, which the method does not help to identify.
+Second, the area is unweighted so all values of \(\theta\) matter equally, despite some being much more realistic than others.
+
+Inspired by Raju's area method, we propose a new method, which we call \enquote{minimizing the area between curves} (MABC).
+To understand MABC, imagine multigroup item response data where \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\) so that there is no DIF.
+The fundamental identification problem of DIF in multigroup IRT models is that there are an infinite number of models with the same likelihood from which to choose.
+For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\mu^{\star\text{foc}} = 0\).
+The model would then estimate \(\hat d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF.
+Alternatively, we could assume that the focal group has \(\mu^{\star\text{foc}} = 3\).
+The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability but every item contains DIF against them.
+Both of these models have the same likelihood, so how should one choose which model to believe?
+MABC chooses the model with the least amount of total DIF, as measured by the total area between the item characteric curves.
+This is another, perhaps clearer, way of assuming that the majority of items do not contain DIF.
+
+Denote a function that estimates \(\hat b_j^\text{foc}\) by fitting a unidimensional Rasch model where the value of \(\mu^\text{foc}\) is fixed as \(m_j(\mu^\text{foc})\).
 The amount of DIF on each item is calculated as
-
 \[
-\text{DIF}_j = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})| g(\theta)d\theta
+\text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
 \]
-where \(g(\theta)\) is a weighting function such that \(\int g(\theta)d\theta = 1\).
-The total DIF on the test, then, is \(\text{Total DIF} = \sum_{j} \text{DIF}_j\).
-Different choices for \(\Theta^{\text{foc}}\) lead to different \(\mathbf{b}^\text{foc}\), and we select the \(\Theta^{\text{foc}}\) that minimizes \(\text{Total DIF}\).
-The idea for considering the difference between item characteristic curves dates back to Stocking and Lord (\protect\hyperlink{ref-stocking1983developing}{1983}).
-More recently, Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}) use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
-
+where \(g(\theta)\) is a weighting function such that \(\int g(\theta)d\theta = 1\). The total DIF on the test, then, is
+\[
+\text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc})
+\]
+In this way, \(\text{Total DIF}(\mu^\text{foc})\) is a function where the input is \(\mu^\text{foc}\) and the output is the total amount of DIF on the test. MABC sets
+\[
+\mu^{\star\text{foc}} = \mathop\mathrm{arg\,min}_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
+\]
+MABC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}) who uses the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
 The selection of \(g(\theta)\) is important and determines where on the ability spectrum to consider DIF.
@@
 We set \(g(\theta)\) to be the average of the reference and focal group ability distributions:
-
 \[
@@
 
+For the simulated data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\). The anchor point is \(\mu^\text{foc} = -1\).
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/workplease2-1}
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/mabc-1}
 
@@
 
-\caption{always need a capt}\label{fig:workplease2}
+\caption{Selecting the anchor point by minimizing the total amount of DIF.}\label{fig:mabc}
 \end{figure}
-
-\hypertarget{bounds-for-dif}{%
-\subsection{Bounds for DIF}\label{bounds-for-dif}}
 
@@
 
+{[}ADD A TABLE OF METHODS? OR THAT CAN SORT OF BE DONE IN THE SIMULATION RESULTS?{]}
+
 \hypertarget{next-steps}{%
 \section{Next steps}\label{next-steps}}
+
+\hypertarget{remember-my-bounds-idea-i-like-that}{%
+\section{Remember my bounds idea I like that}\label{remember-my-bounds-idea-i-like-that}}
 
@@
 
+\leavevmode\hypertarget{ref-magis2011generalized}{}%
+Magis, David, Gilles Raı̂che, Sébastien Béland, and Paul Gérard.
+2011.
+``A Generalized Logistic Regression Procedure to Detect Differential Item Functioning Among Multiple Groups.'' \emph{International Journal of Testing} 11 (4): 365--86.
+
 \leavevmode\hypertarget{ref-meade2012solving}{}%
@@
 
+\leavevmode\hypertarget{ref-raju1988area}{}%
+Raju, Nambury S. 1988. ``The Area Between Two Item Characteristic Curves.'' \emph{Psychometrika} 53 (4): 495--502.
+
 \leavevmode\hypertarget{ref-stark2006detecting}{}%
@@
 ``Detecting Differential Item Functioning with Confirmatory Factor Analysis and Item Response Theory: Toward a Unified Strategy.'' \emph{Journal of Applied Psychology} 91 (6): 1292.
-
-\leavevmode\hypertarget{ref-stocking1983developing}{}%
-Stocking, Martha L, and Frederic M Lord. 1983. ``Developing a Common Metric in Item Response Theory.'' \emph{Applied Psychological Measurement} 7 (2): 201--10.
 

### 2020-02-29 paper/paper_1_intro.Rmd
we're getting there!

@@
-Inspired by @camilli1992conceptual, we view differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
+Inspired by @camilli1992conceptual, we think differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
 From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item.
@@
 
+<!-- i might be a bit share on this here.
+it's basically do you have a q matrix you believe in.
+there must be some work with cross-validating different q matrices and then believing that.
+klint said he had some ideas here too.
+maybe things to think about in the future -->
+
 If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model?

### 2020-02-29 paper/paper_2a_anchor_items.Rmd
we're getting there!

--- 2020-02-27 paper/paper_2_dif_methods.Rmd
@@
 
-In general, we denote the mean ability for the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$.
+In general, we denote the mean ability of the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$, respectively.
 We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
-The fundamental challenge for a DIF method is to first estimate $\mu^\text{foc}$.
-How can $\mu^\text{foc}$ be disentangled from the possibility that items are biased?
-The most common approach is to identify a group of anchor items which are assumed to be DIF-free.
-The anchor items identify the model because they can be used to estimate $\mu^\text{foc}$.
+The fundamental challenge for a DIF method is to first get an estimate for $\mu^\text{foc}$, $\hat\mu^\text{foc}$.
+How can group ability be disentangled from the possibility that items are biased?
+The most common approach is to identify a group of anchor items that are assumed to be DIF-free.
+These items are used to get $\hat\mu^\text{foc}$.
 
@@
 
-The general idea of identifying clusters of items is interesting.
-However, their approach is needlessly complicated and they did not give specific suggestions for moving from $\Delta \mathbf{R}$ to a set of anchor items.
-@pohl2017cluster extended their work by proposing an approach for identifying the anchor cluster.
-$\Delta \mathbf{R}$ is skew-symmetric and of rank 1 so that all information is contained in a single row or column.
+The general idea of identifying clusters of items is intriguing.
+However, their approach is needlessly complicated and they did not describe a process for moving from $\Delta \mathbf{R}$ to a set of anchor items.
+@pohl2017cluster extended their work by proposing one such process.
+$\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
 Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
 They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-However, in their simulation study, they find that BIC identifies only a single cluster, so they use all items as anchors.
+Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.
 
-We propose a new cluster-based approach which we call "equal means clustering" (EMC).
+We propose a new cluster-based approach, which we call "equal means clustering" (EMC).
 Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$ of differences in item easiness, $\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}$, where the model is identified by setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, thus the name "equal means clustering").
 This results in all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$.
-The tilde is used to denote parameters estimated with $\mu^\text{foc}$ set to $0$.
+The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.
 
@@
 The largest cluster is chosen as the anchor cluster.
-If there is a tie for the largest cluster, the cluster with the lowest standard deviation of item easiness differences is selected.
+If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected.
+This process makes the assumption that the largest cluster (but necessarily the majority) of items do not contain DIF.
 
@@
 
-DIF methods, in general, seem to be designed to automatically detect items with DIF without any intervention from a human.
-We refer to DIF methods that do not use an analyst's judgement as "algorithmic DIF methods." On the other hand, it's often useful to present information to the analyst in a way that empowers them to see DIF and potentially choose the anchor items.
+DIF methods are generally designed to automatically detect DIF without any human judgement.
+On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
 
-We propose a technique, the "equal means logit graph" (EMLG), where both group means are to 0 so that differences in performance manifest exclusively in the item parameters.
-Relatedly, @pohl2017cluster fit a model with both groups means set to 0 in a pedagogical example, and @talbot2013taking fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
+We propose a technique, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
+Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$.
+Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$.
+To get a sense of variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical].
+Then, multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of $\tilde{d_j}$ for every item.
+These are the distributions displayed in the EM-MILG.
+The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
 
-We fit a unidimensional Rasch, with both group means fixed to 0, to the simulated item response data.
-We estimate the item parameter covariance matrix using Oakes' identity [@chalmers2018numerical].
-Multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of the difference in difficulty, ${b_j}^{\text{foc}} - {b_j}^{\text{ref}}$, for each item.
-Figure \ref{fig:emlg} shows these distributions in the EMLG.
+Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data.
+It cannot be stated strongly enough that the EM-MILG contains all possible information about the difference in group performance.
+The challenge, then, is to select the anchor items.
+The analyst might assume that — because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.
 
-```{r emlg, fig.cap = 'Equal means logit graph (EMLG)', out.width="70%", warning = FALSE, message = FALSE}
+```{r emmilg, fig.cap = 'The equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.', out.width="70%", warning = FALSE, message = FALSE}
 out_3_biased_items$intuitive_mod[[1]] %>%
@@
     labs(
-        x = latex2exp::TeX("$b_j^{ref} - b_j^{foc}$ (group means fixed to 0)")
+        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
     )
@@
 
-It cannot be stated strongly enough that the EMLG contains all possible information about the difference in group performance.
-The challenge, then, is to identify the difference in group ability.
-The analyst might assume that — because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and should be used as anchor items.
-At this point, the analyst re-fits the model with the anchor items having fixed parameters across groups so that they are used to estimate the difference in group ability.
-The parameters of the other items are free to vary across groups, and, indeed, the new model finds that the item easiness difference, ${b_j}^{\text{ref}} - {b_j}^{\text{foc}}$, is about 0.5 logits for each of these items.
+After choosing the anchor items, the model is refit.
+The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc}$ to $0$.
+The same process of using multiple imputations to estimate the distribution of $d_j$ can be used with the new model.
+Because the equal means assumption is not made, we refer to the resulting graph as a multiple imputations logit graph (MILG).
+In our simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
 
-One way to show this:
+```{r milg, fig.cap = 'The multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group.
+Anchor items are fixed to 0.', out.width="70%", warning = FALSE, message = FALSE}
+# CAREFUL - this code assumes that the analyst would choose the same anchor items as AOAA-OAT found
 
-```{r who, fig.cap = 'fix this', out.width="70%"}
-out_3_biased_items$intuitive_mod[[1]] %>%
-    mod_intuitive_to_draws_df() %>%
-    draws_df_to_logit_plot() +
-    geom_vline(xintercept = filter(out_3_biased_items$AOAA_OAT_final_dif[[1]], anchor)$difference_in_easy[1], linetype = "dashed", color = "#009E73") +
-    geom_point(
-        data = out_3_biased_items$AOAA_OAT_final_dif[[1]],
-        aes(x = difference_in_easy, y = item),
-        color = "#009E73",
-        size = 2
+out_3_biased_items$AOAA_OAT_final_dif[[1]] %>% 
+    mutate(difference_in_easy = difference_in_easy - difference_in_easy[anchor][1]) %>% 
+    ggplot(aes(x = difference_in_easy, y = item)) +
+    ggridges::geom_density_ridges(
+        data = 
+            out_3_biased_items$AOAA_OAT_mod[[1]] %>% 
+            mod_to_draws_df(n = 1000) %>% 
+            select(-run) %>%
+            gather(var, val) %>% 
+            filter(val != 0),
+        aes(x = val, y = var)
     ) +
+    geom_point() +
     labs(
-        title =
-            "<span style='color:#009E73'> Analyst decision based on EMLG <br>
-            Anchor on items 1-5 and freely estimate items 6-8 </span>",
-        x = latex2exp::TeX("$b_j^{ref} - b_j^{foc} - \\mu_j^{foc}$")
-    ) +
-    theme(
-        plot.title = element_markdown(lineheight = 1.1)
+        x = latex2exp::TeX("$d_j = b_j^{ref} - b_j^{foc}$"),
+        y = ""
     )
@@
 
-Another way to show this:
-
-```{r why, fig.cap = 'fix this', out.width="70%"}
-out_3_biased_items$AOAA_OAT_final_dif[[1]] %>%
-    select(item, anchor, a_ref_easy, b_foc_easy) %>%
-    gather(var, val, -item, -anchor) %>%
-    mutate(var = case_when(anchor ~ "anchor", var == "a_ref_easy" ~ "ref_easy", var == "b_foc_easy" ~ "foc_easy")) %>%
-    ggplot(aes(x = item, y = val, color = var)) +
-    geom_point() +
-    coord_flip() +
-    scale_color_manual(values = c("black", "blue", "red"))
-```
-
-One of our key concerns with algorithmic DIF methods is that they can lull the analyst into a false sense of security.
+One of our key concerns with typical DIF methods is that they can lull the analyst into a false sense of security.
 Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
-The EMLG combats this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst is may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even in cases where algorithmic DIF methods are used, the analyst should use the EMLG to give them a sense of their item response data.
-
-## Anchor points
-
-All of the previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
-The difference in group abilities is then estimated using exclusively the anchor items.
-An alternative strategy is to select an anchor point, which essentially specifies the difference in group abilities directly.
-Anchor point methods have the advantage of not assuming that any particular item is DIF-free and therefore, all items can be tested for DIF.
-
-### Gini index
-
-@strobl2018anchor suggest using the Gini index -@gini1912variabilita for selecting anchor points.
-The Gini index is typically used to measure the inequality of wealth distribution in a country.
-South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution @chitiga2015income.
-In general, the Gini index "takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little" [@strobl2018anchor, p.
-7].
-
-The Gini index is maximized to select the anchor point.
-The intuition is that we want to find the anchor point such that a small minority of items have bias while the vast majority do not.
-Denote the easiness difference between the two groups when both group means are fixed to zero as $d_j = {b_j}^\text{ref} - {b_j}^\text{foc}$, a function that calculates the Gini index from a vector of non-negative elements as $G(\mathbf{x})$, and the mean of the focal group ability when the reference group ability is fixed to zero (i.e.
-the anchor point) as $\Theta$.
-The anchor point is set to
-$$
-\argmax_\Theta G(\mathbf{d} + \Theta)
-$$
-where $\Theta \in (-\infty, \infty)$ and $\Theta$ is added to each element of $\mathbf{d}$.
-
-```{r}
-emlg <- 
-    coef_1f(out_3_biased_items$intuitive_mod[[1]])$items %>% 
-    select(item, ref_easy = a_ref_easy, foc_easy = b_foc_easy) %>% 
-    mutate(difference_in_easy = ref_easy - foc_easy)
-
-gini_results <- emlg_to_gini_l1(emlg, lo = -2, hi = 0)
-```
-
-Figure \ref{fig:workplease} shows the gini coefficient at a variety of possible values for $\Theta$. We find that the anchor point is `r round(gini_results$gini_anchor_points, 2)`.
-
-```{r workplease, fig.cap = 'always need a capt', out.width="70%"}
-gini_results$grid %>% 
-    ggplot(aes(x = anchor_point, y = gini)) +
-    geom_point()
-```
-
-### Minimizing between curves
-
-Race created a statistical test which was one of the most popular methods used until recently.
-
-Imagine multigroup item response data where $\Theta^{\text{foc}} = \Theta^{\text{ref}}$ and $d_j = 0 \forall j$ so that there is no DIF.
-The fundamental identification problem of DIF in multipgroup IRT models is that
-there are an infinite number of models with the same likelihood from which to choose.
-For example, we could correctly assume that the focal group has the same ability as the reference group and fix $\Theta^{\text{foc}} = 0$.
-The model would then estimate $d_j \approx 0 \forall j$, and we would correctly conclude the groups have the same ability and there is no DIF.
-Alternatively, we could assume that the focal group has very high ability and fix $\Theta^{\text{foc}} = 3$.
-The model would then compensate by finding $d_j \approx -3 \forall j$, and we would incorrectly conclude that the the focal group is high ability but each item contains DIF against the focal group.
-Both of these models have the same likelihood, so how should we choose which model to believe?
-This is the fundamental identification problem.
-
-We put forth a method we call minimizing between curves (MBC) which chooses the model with least amount of total DIF on the test.
-"Curves" refers to the item characteristic curve which maps a student's ability to their probability of correct response.
-The amount of DIF on each item is calculated as
-
-$$
-\text{DIF}_j = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})| g(\theta)d\theta
-$$
-where $g(\theta)$ is a weighting function such that $\int g(\theta)d\theta = 1$.
-The total DIF on the test, then, is $\text{Total DIF} = \sum_{j} \text{DIF}_j$.
-Different choices for $\Theta^{\text{foc}}$ lead to different $\mathbf{b}^\text{foc}$, and we select the $\Theta^{\text{foc}}$ that minimizes $\text{Total DIF}$.
-The idea for considering the difference between item characteristic curves dates back to @stocking1983developing.
-More recently, @chalmers2016might use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
-
-The selection of $g(\theta)$ is important and determines where on the ability spectrum to consider DIF.
--@chalmers2016might does not discuss how to choose $g(\theta)$ and in practice uses $g(\theta)$ uniform for $-6 \le \theta \le 6$, which seems like an odd choice.
-It might seem intuitive to choose $g(\theta) \sim N(0, 1)$ because $\mu^\text{ref} = 0$, but this choice doesn't take into account the ability distribution of the focal group.
-If $\mu^\text{foc} = 3$, wouldn't we also want to prioritize high $\theta$ values?
-We set $g(\theta)$ to be the average of the reference and focal group ability distributions:
-
-$$
-g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
-$$
-
-```{r workplease2, fig.cap = 'always need a capt', out.width="70%"}
-# min_bc <-
-#     out_3_biased_items$intuitive_mod[[1]] %>%
-#     min_between_curves(lo = 0.95, hi = 1.1, by = 0.001)
-
-# min_bc %>% write_rds(here("paper", "paper_2_min_bc.rds"))
-
-min_bc <- read_rds(here("paper", "paper_2_min_bc.rds"))
-
-min_bc$graph
-```
-
-## Bounds for DIF
-
-## Summary
+The EM-MILG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even when other DIF methods are used, the analyst can use the EM-MILG as a first step in order to give them a sense of their item response data.
+And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.

### 2020-02-29 paper/paper_2b_anchor_points.Rmd
we're getting there!

The previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst. The anchor items are used to estimate $\hat \mu^\text{foc}$. An alternative strategy is to directly select the anchor point, $\mu^{\star\text{foc}}$. Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, all items can be tested for DIF. The question then becomes the following: How do we select the anchor point?

### Gini index

@strobl2018anchor suggest using the Gini index -@gini1912variabilita to select the anchor point. The Gini index is typically used to measure the inequality of wealth distribution in a country. For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution @chitiga2015income. In general, the Gini index "takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little" [@strobl2018anchor, p. 7].

The anchor point, $\mu^{\star\text{foc}}$, is selected by maximizing the Gini index. The intuition and assumption is that some items might have bias but the majority do not. Denoting a function that calculates the Gini index from a vector of non-negative elements as $G(\mathbf{x})$,  
$$
\mu^{\star\text{foc}} = \argmax_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
$$
where $\mu^\text{foc} \in (-\infty, \infty)$ and $\mu^\text{foc}$ is added to each element of $\mathbf{d}$.

```{r}
emlg <- 
    coef_1f(out_3_biased_items$intuitive_mod[[1]])$items %>% 
    select(item, ref_easy = a_ref_easy, foc_easy = b_foc_easy) %>% 
    mutate(difference_in_easy = ref_easy - foc_easy)

gini_results <- emlg_to_gini_l1(emlg, lo = -2, hi = 0)
```

For the simulated data, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible values for $\mu^\text{foc}$. The anchor point is $\mu^\text{foc} = `r round(gini_results$gini_anchor_points, 2)`$.

```{r ginipath, fig.cap = 'Selecting the anchor point by maximizing the Gini index.', out.width="70%", warning = FALSE, message = FALSE}
gini_results$grid %>% 
    ggplot(aes(x = anchor_point, y = gini)) +
    geom_point() +
    labs(
        x = latex2exp::TeX("$\\mu^{foc}$"),
        y = latex2exp::TeX("$G(|\\mu^{foc} + \\tilde{\\mathbf{d}}|)$")
    ) +
    # geom_point(
    #     data = gini_results$grid %>% filter(gini == max(gini)),
    #     color = "green",
    #     size = 3
    # ) +
    geom_vline(
        xintercept = gini_results$grid %>% filter(gini == max(gini)) %>% pull(anchor_point),
        linetype = "dashed"
    )
```

We then refit the model with the identifying assumption of $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$. We test each item for DIF and display the results in a MILG as is shown in Figure \ref{fig:ginimilg}.

```{r ginimilg, fig.cap = 'Selecting the anchor point by maximizing the Gini index.', out.width="70%"}

data <- out_3_biased_items$sim[[1]]$data
groups <- out_3_biased_items$sim[[1]]$groups

tmp <- 
    multipleGroup(
        data,
        1,
        itemtype = "Rasch",
        groups,
        invariance = c("free_var"),
        pars = "values"
    ) %>% 
    as_tibble()

tmp$value[length(tmp$value) - 1] <- gini_results$gini_anchor_points

mod <- 
    multipleGroup(
        data,
        1,
        itemtype = "Rasch",
        groups,
        pars = tmp %>% as.data.frame(),
        SE = TRUE,
        technical = list(NCYCLES = 500),
        verbose = FALSE,
    )

mod %>% 
    mod_to_draws_df(n = 10000) %>% 
    draws_df_to_logit_plot() +
    labs(
        x = latex2exp::TeX("$d_j = b_j^{ref} - b_j^{foc}$"),
        y = ""
    )
```

### Minimizing between curves

Raju's area method measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group [@raju1988area]:
```{r}
# {-\infty}^{\infty}
```
\begin{align}
    \text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
\end{align}

Raju's area method, along with AOAA, have been cited as the two most commonly used IRT-based DIF detection methods [@magis2011generalized]. We see two weakness of Raju's area method: First, the item characteric curves still need to be linked by anchor items, which the method does not help to identify. Second, the area is unweighted so all values of $\theta$ matter equally, despite some being much more realistic than others.

Inspired by Raju's area method, we propose a new method, which we call "minimizing the area between curves" (MABC). To understand MABC, imagine multigroup item response data where $\mu^{\text{foc}} = \mu^{\text{ref}}$ and $d_j = 0 \forall j$ so that there is no DIF. The fundamental identification problem of DIF in multigroup IRT models is that there are an infinite number of models with the same likelihood from which to choose. For example, we could correctly assume that the focal group has the same ability as the reference group and fix $\mu^{\star\text{foc}} = 0$. The model would then estimate $\hat d_j \approx 0 \forall j$, and we would correctly conclude the groups have the same ability and there is no DIF. Alternatively, we could assume that the focal group has $\mu^{\star\text{foc}} = 3$. The model would then compensate by finding $\hat d_j \approx -3 \forall j$, and we would incorrectly conclude that the focal group is high ability but every item contains DIF against them. Both of these models have the same likelihood, so how should one choose which model to believe? MABC chooses the model with the least amount of total DIF, as measured by the total area between the item characteric curves. This is another, perhaps clearer, way of assuming that the majority of items do not contain DIF.

Denote a function that estimates $\hat b_j^\text{foc}$ by fitting a unidimensional Rasch model where the value of $\mu^\text{foc}$ is fixed as $m_j(\mu^\text{foc})$. The amount of DIF on each item is calculated as 
$$
\text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
$$
where $g(\theta)$ is a weighting function such that $\int g(\theta)d\theta = 1$. The total DIF on the test, then, is 
$$
\text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc})
$$
In this way, $\text{Total DIF}(\mu^\text{foc})$ is a function where the input is $\mu^\text{foc}$ and the output is the total amount of DIF on the test. MABC sets
$$
\mu^{\star\text{foc}} = \argmin_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
$$
MABC is inspired in part by @chalmers2016might who uses the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF). The selection of $g(\theta)$ is important and determines where on the ability spectrum to consider DIF. -@chalmers2016might does not discuss how to choose $g(\theta)$ and in practice uses $g(\theta)$ uniform for $-6 \le \theta \le 6$, which seems like an odd choice. It might seem intuitive to choose $g(\theta) \sim N(0, 1)$ because $\mu^\text{ref} = 0$, but this choice doesn't take into account the ability distribution of the focal group. If $\mu^\text{foc} = 3$, wouldn't we also want to prioritize high $\theta$ values? We set $g(\theta)$ to be the average of the reference and focal group ability distributions:
$$
g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
$$
```{r}
# not ideal that i input -anchor_point here but can fix later
# min_bc <-
#     out_3_biased_items$intuitive_mod[[1]] %>%
#     min_between_curves(lo = 0, hi = 2, by = 0.1) # future: 0.0001
# min_bc %>% write_rds(here("paper", "paper_2_min_bc.rds"))

min_bc <- read_rds(here("paper", "paper_2_min_bc.rds"))
```

For the simulated data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$. The anchor point is $\mu^\text{foc} = `r round(min_bc$between_curves_anchor_points, 2)`$.

```{r mabc, fig.cap = 'Selecting the anchor point by minimizing the total amount of DIF.', out.width="70%"}
min_bc$grid %>% 
    ggplot(aes(x = anchor_point, y = total_between_curves)) +
    geom_point() +
    geom_vline(
        xintercept = min_bc$between_curves_anchor_points,
        linetype = "dashed"
    ) +
    labs(
        x = latex2exp::TeX("$\\mu^{foc}$"),
        y = latex2exp::TeX("$Total \\, DIF(\\mu^{foc})$")
    )
```

## Summary

[ADD A TABLE OF METHODS? OR THAT CAN SORT OF BE DONE IN THE SIMULATION RESULTS?]

### 2020-02-29 R/dif_AOAA.R

@@ -1,3 +1,16 @@
+# the way i am thinking about this is that we always start wtih all as anchors
+
+# start by doing AOAA
+
+# AS: make all p < 0.05 flex and then keep going
+
+# OAT: make min(p) flex and then keep going
+
+# For both, whenever all p > 0.05 everything in the anchor set is anchors and you are done
+
+# Potential problem: what if in the first AOAA all p < 0.05? this can also happen
+# in future iterations of AS but can never happen with OAT
+
 # data <- out_3_biased_items$sim[[1]]$data
 # groups <- out_3_biased_items$sim[[1]]$groups
 # dif_aoaa <- dif_AOAA(data, groups)

### 2020-02-29 R/min_between_curves.R

@@ -14,8 +14,10 @@ min_between_curves <- function(intuitive_mod, lo, hi, by){
     weights_ref <- dnorm(Theta, mu_ref, sd_ref) / sum(dnorm(Theta, mu_ref, sd_ref))
 
     f <- function(intuitive_mod, anchor_point){
+        # was disoriented but the anchor point is what you boost the focal group by
+        # so if anchor_point = 1 then the mean of the focal group is -1
 
-        mu_foc <- anchor_point
+        mu_foc <- -anchor_point
         # assumes cov doesn't change when I shift mean
         # which I mostly cerified for Rasch in
         # R-work-complete > cov_doesnt_change_when_shifting
@@ -51,15 +53,13 @@ min_between_curves <- function(intuitive_mod, lo, hi, by){
         tibble(
             anchor_point = seq(lo, hi, by)
         )  %>%
-        mutate(total_between_curves = anchor_point %>% map_dbl(~ f(intuitive_mod, .)))
+        mutate(
+            total_between_curves = anchor_point %>% map_dbl(~ f(intuitive_mod, .)),
+            anchor_point = -anchor_point
+        )
 
     list(
         grid = grid,
-        graph =
-            grid %>%
-            ggplot(aes(x = anchor_point, y = total_between_curves)) +
-            geom_point() +
-            geom_path(),
         between_curves_anchor_points = grid$anchor_point[grid$total_between_curves == min(grid$total_between_curves)]
     )
 }

### 2020-02-29 R/mod_zelpers.R

@@ -33,7 +33,7 @@ mod_to_draws_df <- function(mod, n){
     n_items <- ncol(par_draws) / 2
 
     for (i in 1:n_items) {
-        draws_df[[paste0("item", i)]] <- par_draws[[i + n_items]] - par_draws[[i]]
+        draws_df[[paste0("item", i)]] <- par_draws[[i]] - par_draws[[i + n_items]]
     }
 
     draws_df

### 2020-03-02 paper/paper.Rmd
implementing non compensatory models which seems to be going well

@@
 
+<!-- the literature basically talks separately about anchor selection and DIF methods.
+much more literature is on DIF methods and less is on anchor selection.
+I'm basically writing a paper about anchor selection and just calling it DIF methods because that seems like where the action is.
+I'm taking fixed at that the likelihood ratio test is the way to go (and perhaps that deserves more detail)?
+-->
+
 \clearpage
-
-# TO DO
-
-- be sure all my directions match. i think i switched everything to be item easiness at some point but not sure if i did it throughout
 

### 2020-03-02 paper/paper.tex
implementing non compensatory models which seems to be going well

@@
 They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
-\end{align}
-where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
+They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is \[P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)\] where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
 
@@
 The intuition and assumption is that some items might have bias but the majority do not.
-Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\),\\
+Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\),
 \[
@@
 
-Raju's area method measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group (Raju \protect\hyperlink{ref-raju1988area}{1988}):
-
-\begin{align}
+Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group:
+\[
 \text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
-\end{align}
-
+\]
 Raju's area method, along with AOAA, have been cited as the two most commonly used IRT-based DIF detection methods (Magis et al.

### 2020-03-02 paper/paper_1_intro.Rmd
implementing non compensatory models which seems to be going well

@@
+<!-- need to specify which multidimensional model im talking about -->
+
 Inspired by @camilli1992conceptual, we think differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
@@
 They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
-\end{align}
-where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.
+They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is $$P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)$$ where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.
 
@@
 According to the two-dimensional 2PL model, the probability student $i$ responding correctly to item $j$ is
-\begin{align}
+$$
 P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
-\end{align}
+$$
 This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability for students in the same group.

### 2020-03-02 paper/paper_2a_anchor_items.Rmd
implementing non compensatory models which seems to be going well

@@
 In this section, we summarize existing DIF methods and propose extensions of those methods.
-We limit our discussion to IRT-based methods.
+We take as given that the IRT-based likelihood ratio test is the best way to test for DIF, which has been well demonstrated [@meade2012solving].
 As a result, we avoid methods like the Mantel-Haenszel procedure [@holland1986differential], which muddies the waters by moving away from the IRT framework, and has been shown to perform no better — and in some cases worse — than IRT-based methods [@swaminathan1990detecting].
+
+The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items contain bias — can be made.
 For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
@@
 We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
-The fundamental challenge for a DIF method is to first get an estimate for $\mu^\text{foc}$, $\hat\mu^\text{foc}$.
-How can group ability be disentangled from the possibility that items are biased?
+The fundamental challenge for DIF methods is to disentangle estimating $\hat\mu^\text{foc}$ from estimating $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$.
 The most common approach is to identify a group of anchor items that are assumed to be DIF-free.
-These items are used to get $\hat\mu^\text{foc}$.
+These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$.
 
@@
 
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been proposed.
+We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
+AOAA-OAT is inspired by @hagquist2017recent who, in general, assert that "items showing DIF initially
+should not be resolved simultaneously but sequentially" [p.
+6].
 Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the $\chi^2$ test statistic, is removed from the anchor set.

### 2020-03-02 paper/paper_2b_anchor_points.Rmd
implementing non compensatory models which seems to be going well

@@
 
-```{r ginimilg, fig.cap = 'Selecting the anchor point by maximizing the Gini index.', out.width="70%"}
+```{r ginimilg, fig.cap = 'Selecting the anchor point by maximizing the Gini index.', out.width="70%", warning = FALSE, message = FALSE}
 
@@
 
-Raju's area method measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group [@raju1988area]:
-```{r}
-# {-\infty}^{\infty}
-```
-\begin{align}
+Raju's area method [@raju1988area] measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group:
+$$
 \text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
-\end{align}
-
+$$
 Raju's area method, along with AOAA, have been cited as the two most commonly used IRT-based DIF detection methods [@magis2011generalized].

### 2020-03-02 R/create_pars_compensatory.R

@@ -1,4 +1,4 @@
-create_pars_1_biased_item <- function(){
+create_pars_compensatory_1_biased_item <- function(){
     tibble(
         item = 1:8,
         easy = rnorm(8, 0, 1),
@@ -7,7 +7,7 @@ create_pars_1_biased_item <- function(){
     )
 }
 
-create_pars_3_biased_items <- function(){
+create_pars_compensatory_3_biased_items <- function(){
     tibble(
         item = 1:8,
         easy = rnorm(8, 0, 1),
@@ -16,7 +16,7 @@ create_pars_3_biased_items <- function(){
     )
 }
 
-create_pars_4_biased_items_balanced <- function(){
+create_pars_4_compensatory_biased_items_balanced <- function(){
     tibble(
         item = 1:8,
         easy = rnorm(8, 0, 1),

### 2020-03-02 R/create_pars_noncompensatory.R

@@ -1,0 +1,46 @@
+# create_pars_noncompensatory(n_items = 10, n_dif_items = 5) %>%
+#     graph_pars_noncompensatory(0, -1)
+
+angle_to_a_nuisance <- function(angle){
+    # uses that cos formula. assumes a_target is 1. see p 949 of walker and sahin
+    # angle_to_a_nuisance(30)
+    sqrt((1 - cos(angle * pi / 180)^2) / cos(angle * pi / 180)^2)
+}
+
+create_pars_noncompensatory <- function(n_items, n_dif_items){
+    # create item parameters. all target disc is 1 and nuisance disc is 0 for non-dif items and has angle that
+    # is lattice from 30 to 60 for dif items
+    tibble(
+        item = 1:n_items,
+        dif = c(rep("no", n_items - n_dif_items), rep("yes", n_dif_items))
+    ) %>%
+        mutate(
+            a_target = 1,
+            angle = ifelse(dif == "no", 0, seq(30, 60, length.out = n_dif_items)),
+            a_nuisance = angle_to_a_nuisance(angle),
+            b_target = rnorm(n_items, 0.5, 0.5),
+            b_nuisance = 1
+        )
+}
+
+graph_pars_noncompensatory <- function(pars, ref_mean_ability_nuisance, foc_mean_ability_nuisance){
+    # graphs
+    crossing(
+        item = 1:nrow(pars),
+        ability_target = seq(-3, 3, by = 0.1)
+    ) %>%
+        left_join(pars) %>%
+        mutate(
+            ref =
+                sigmoid(a_target * ability_target + b_target) *
+                sigmoid(a_nuisance * ref_mean_ability_nuisance + b_nuisance),
+            foc =
+                sigmoid(a_target * ability_target + b_target) *
+                sigmoid(a_nuisance * foc_mean_ability_nuisance + b_nuisance)
+        ) %>%
+        dplyr::select(item, ability_target, ref, foc) %>%
+        gather(group, prob, -item, -ability_target) %>%
+        ggplot(aes(x = ability_target, y = prob, color = group)) +
+        geom_point() +
+        facet_wrap(~ item)
+}

### 2020-03-02 R/sim_students_noncomp.R

@@ -1,0 +1,67 @@
+# pars <- create_pars_noncompensatory(n_items = 10, n_dif_items = 5)
+#
+# students <-
+#     sim_students_2d(
+#         n_ref = 10,
+#         n_foc = 10,
+#         ref_target_ability_mu = 0,
+#         ref_nuisance_ability_mu = 0,
+#         foc_target_ability_mu = -0.5,
+#         foc_nuisance_ability_mu = -1
+#     )
+#
+# p <- pars_students_to_p_noncomp(pars, students)
+#
+# data <- p_to_data_byway_of_flipping(p)
+#
+# sim <- students_p_to_sim(students, p)
+
+sim_students_2d <- function(n_ref, n_foc, ref_target_ability_mu, ref_nuisance_ability_mu, foc_target_ability_mu, foc_nuisance_ability_mu){
+    bind_rows(
+        MASS::mvrnorm(n_ref, mu = c(ref_target_ability_mu, ref_nuisance_ability_mu), Sigma = sigma) %>% as.data.frame() %>% as_tibble(),
+        MASS::mvrnorm(n_foc, mu = c(foc_target_ability_mu, foc_nuisance_ability_mu), Sigma = sigma) %>% as.data.frame() %>% as_tibble()
+    ) %>%
+        mutate(groups = c(rep("a_ref", n_ref), rep("b_foc", n_foc))) %>%
+        select(groups, target_ability = V1, nuisance_ability = V2)
+}
+
+get_p_noncomp <- function(a_target, a_nuisance, b_target, b_nuisance, target_ability, nuisance_ability){
+    sigmoid(a_target * target_ability + b_target) *
+    sigmoid(a_nuisance * nuisance_ability + b_nuisance)
+}
+
+pars_students_to_p_noncomp <- function(pars, students){
+    p <- matrix(nrow = nrow(students), ncol = nrow(pars))
+
+    for (i in 1:nrow(students)){
+        for (j in 1:nrow(pars)){
+            p[i, j] <-
+                get_p_noncomp(
+                    a_target = pars$a_target[j],
+                    a_nuisance = pars$a_nuisance[j],
+                    b_target = pars$b_target[j],
+                    b_nuisance = pars$b_nuisance[j],
+                    target_ability = students$target_ability[i],
+                    nuisance_ability = students$nuisance_ability[i]
+                )
+        }
+    }
+
[…]

### 2020-03-02 R/zelpers.R

@@ -3,3 +3,8 @@ quiet <- function(x){
     on.exit(sink())
     invisible(force(x))
 }
+
+sigmoid <- function(x){
+    exp(x) / (1 + exp(x))
+}
+

### 2020-03-02 paper/paper_2_sim.R

@@ -1,10 +1,10 @@
 out_3_biased_items <-
     tibble(
-        pars = rerun(1, create_pars_3_biased_items())
+        pars = rerun(1, create_pars_compensatory_3_biased_items())
     ) %>%
     # sim and fit intuitive
     mutate(
-        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
+        sim = pars %>% map(~ sim_students_compensatory(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
         intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
     ) %>%
     # dif: AOAA

### 2020-03-02 paper/paper_4_sim.R

@@ -1,0 +1,19 @@
+tibble(
+    pars = rerun(1, create_pars_noncompensatory(n_items = 20, n_dif_items = 8)),
+    students =
+        rerun(
+            1,
+            sim_students_2d(
+                n_ref = 10,
+                n_foc = 10,
+                ref_target_ability_mu = 0,
+                ref_nuisance_ability_mu = 0,
+                foc_target_ability_mu = -0.5,
+                foc_nuisance_ability_mu = -1
+            )
+        )
+) %>%
+    mutate(
+        p = map2(pars, students, pars_students_to_p_noncomp),
+        sim = map2(students, p, students_p_to_sim)
+    )

### 2020-03-02 sim.R

@@ -1,0 +1,49 @@
+library(tidyverse)
+library(MASS)
+
+# abilities
+sigma <- cbind(c(1, 0.5), c(0.5, 1))
+
+ref <- mvrnorm(10000, mu=c(0, 0), Sigma = sigma)
+var(ref)
+cor(ref)
+colMeans(ref)
+
+foc <- mvrnorm(10000, mu=c(-0.5, -1), Sigma = sigma)
+var(foc)
+cor(foc)
+colMeans(foc)
+
+# items
+n_items <- 20
+n_dif_items <- 5
+
+# go from disc2 to angle
+(angle <- acos(1 / sqrt(1 + disc2^2)) * 180 / pi)
+
+# and from angle to disc2
+sqrt((1 - cos(angle * pi / 180)^2) / cos(angle * pi / 180)^2)
+
+angle_to_a_nuisance <- function(angle){
+    # uses that cos formula. assumes a_target is 1
+    sqrt((1 - cos(angle * pi / 180)^2) / cos(angle * pi / 180)^2)
+}
+
+angle_to_a_nuisance(30)
+
+items <-
+    tibble(
+        item = 1:n_items,
+        dif = c(rep("no", n_items - n_dif_items), rep("yes", n_dif_items))
+    ) %>%
+    mutate(
+        a_target = 1,
+        angle = ifelse(dif == "no", 0, seq(30, 60, length.out = n_dif_items)),
+        a_nuisance = angle_to_a_nuisance(angle),
+        b_target = rnorm(n_items, 0.5, 0.5),
+        b_nuisance = 1
+    )
+
+
+
+

### 2020-03-04 paper/paper_2b_anchor_points.Rmd
use git search

@@
 
-tmp <- 
-    multipleGroup(
-        data,
-        1,
-        itemtype = "Rasch",
-        groups,
-        invariance = c("free_var"),
-        pars = "values"
-    ) %>% 
-    as_tibble()
-
-tmp$value[length(tmp$value) - 1] <- gini_results$gini_anchor_points
-
-mod <- 
-    multipleGroup(
-        data,
-        1,
-        itemtype = "Rasch",
-        groups,
-        pars = tmp %>% as.data.frame(),
-        SE = TRUE,
-        technical = list(NCYCLES = 500),
-        verbose = FALSE,
-    )
+mod <- gini_results_to_final_mod(gini_results$gini_anchor_points, data, groups)
 

### 2020-03-04 R/___bigsim___.R

@@ -1,0 +1,106 @@
+bigsim <- function(runs, n_dif_items){
+    tibble(
+        pars = rerun(runs, create_pars_noncompensatory(n_items = 20, n_dif_items = n_dif_items, angle_start = 10, b_nuisance_with_dif_sd = 0)),
+        students =
+            rerun(
+                1,
+                sim_students_2d(
+                    n_ref = 1000,
+                    n_foc = 1000,
+                    ref_target_ability_mu = 0,
+                    ref_nuisance_ability_mu = 0,
+                    foc_target_ability_mu = -0.5,
+                    foc_nuisance_ability_mu = -1
+                )
+            )
+    ) %>%
+        # sim
+        mutate(
+            p = map2(pars, students, pars_students_to_p_noncomp),
+            sim = map2(students, p, students_p_to_sim)
+        ) %>%
+        # fit intuitive
+        mutate(
+            intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
+        ) %>%
+        # dif: AOAA
+        mutate(
+            AOAA_status = sim %>% map(~ dif_AOAA(.$data, .$groups)),
+            AOAA_mod =
+                map2(
+                    sim,
+                    AOAA_status,
+                    ~ mod_flexible(
+                        .x$data,
+                        .x$groups,
+                        which(.y$status == "flex"),
+                        se = TRUE
+                    )
+                ),
+            AOAA_final_dif = AOAA_mod %>% map(mod_to_final_dif)
+        ) %>%
+        # dif: AOAA AS
+        mutate(
+            AOAA_AS_status = map2(sim, AOAA_status, ~ dif_AOAA_AS(.x$data, .x$groups, .y)),
+            AOAA_AS_mod =
+                map2(
+                    sim,
+                    AOAA_AS_status,
+                    ~ mod_flexible(
+                        .x$data,
+                        .x$groups,
+                        which(.y$status == "flex"),
+                        se = TRUE
+                    )
+                ),
+            AOAA_AS_final_dif = AOAA_AS_mod %>% map(mod_to_final_dif)
+        ) %>%
+        # dif: AOAA OAT
+        mutate(
+            AOAA_OAT_status = map2(sim, AOAA_status, ~ dif_AOAA_OAT(.x$data, .x$groups, .y)),
+            AOAA_OAT_mod =
+                map2(
+                    sim,
+                    AOAA_OAT_status,
+                    ~ mod_flexible(
[…]

### 2020-03-04 R/clustering.R

@@ -2,11 +2,16 @@
 # https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92
 # https://github.com/pimentel/cluster/blob/master/R/clusGap.R
 
+# the fancy stuff below was just giving one cluster so change to this
+irt_cluster_fix3 <- function(vec){
+    pam1(vec, 4)$cluster
+}
+
 pam1 <- function(x,k) list(cluster = pam(x, k, cluster.only=TRUE))
 
 irt_cluster <- function(vec){
 
-    theGap <- clusGap(matrix(vec), FUNcluster = pam1, K.max = 4)
+    theGap <- clusGap(matrix(vec), FUNcluster = pam1, K.max = 4, spaceH0 = "original")
 
     n_clust <-
         maxSE(

### 2020-03-04 R/create_pars_noncompensatory.R

@@ -7,7 +7,7 @@ angle_to_a_nuisance <- function(angle){
     sqrt((1 - cos(angle * pi / 180)^2) / cos(angle * pi / 180)^2)
 }
 
-create_pars_noncompensatory <- function(n_items, n_dif_items){
+create_pars_noncompensatory <- function(n_items, n_dif_items, angle_start, b_nuisance_with_dif_sd){
     # create item parameters. all target disc is 1 and nuisance disc is 0 for non-dif items and has angle that
     # is lattice from 30 to 60 for dif items
     tibble(
@@ -16,10 +16,10 @@ create_pars_noncompensatory <- function(n_items, n_dif_items){
     ) %>%
         mutate(
             a_target = 1,
-            angle = ifelse(dif == "no", 0, seq(30, 60, length.out = n_dif_items)),
+            angle = c(rep(0, n_items - n_dif_items), seq(angle_start, 60, length.out = n_dif_items)),
             a_nuisance = angle_to_a_nuisance(angle),
-            b_target = rnorm(n_items, 0.5, 0.5),
-            b_nuisance = 1
+            b_target = rnorm(n_items, 0, 0.0001), # CHANGE THIS BACK
+            b_nuisance = c(rep(999, n_items - n_dif_items), rnorm(n_dif_items, 0, b_nuisance_with_dif_sd))
         )
 }
 

### 2020-03-04 R/dif_intuitive.R

@@ -23,8 +23,14 @@ mod_intuitive_to_draws_df <- function(mod){
 draws_df_to_logit_plot <- function(draws_df){
     draws_df %>%
         select(-run) %>%
-        gather(var, val) %>%
+        gather(var, val, factor_key = TRUE) %>%
         ggplot(aes(x = val, y = var)) +
         ggridges::geom_density_ridges() +
         labs(x = "", y = "")
+}
+
+intuitive_mod_to_emlg_summary <- function(intuitive_mod){
+    coef_1f(out_noncomp$intuitive_mod[[1]])$items %>%
+        select(item, ref_easy = a_ref_easy, foc_easy = b_foc_easy) %>%
+        mutate(difference_in_easy = ref_easy - foc_easy)
 }

### 2020-03-04 R/sim_students_noncomp.R

@@ -17,6 +17,9 @@
 # sim <- students_p_to_sim(students, p)
 
 sim_students_2d <- function(n_ref, n_foc, ref_target_ability_mu, ref_nuisance_ability_mu, foc_target_ability_mu, foc_nuisance_ability_mu){
+
+    sigma <- matrix(c(1, 0.5, 0.5, 1), ncol = 2)
+
     bind_rows(
         MASS::mvrnorm(n_ref, mu = c(ref_target_ability_mu, ref_nuisance_ability_mu), Sigma = sigma) %>% as.data.frame() %>% as_tibble(),
         MASS::mvrnorm(n_foc, mu = c(foc_target_ability_mu, foc_nuisance_ability_mu), Sigma = sigma) %>% as.data.frame() %>% as_tibble()
@@ -26,8 +29,10 @@ sim_students_2d <- function(n_ref, n_foc, ref_target_ability_mu, ref_nuisance_ab
 }
 
 get_p_noncomp <- function(a_target, a_nuisance, b_target, b_nuisance, target_ability, nuisance_ability){
-    sigmoid(a_target * target_ability + b_target) *
-    sigmoid(a_nuisance * nuisance_ability + b_nuisance)
+    target <- sigmoid(a_target * target_ability + b_target)
+    nuisance <- sigmoid(a_nuisance * nuisance_ability + b_nuisance)
+    nuisance <- ifelse(is.nan(nuisance), 1, nuisance)
+    target * nuisance
 }
 
 pars_students_to_p_noncomp <- function(pars, students){

### 2020-03-04 scratch.R

@@ -1,0 +1,8 @@
+
+eight %>%
+    select(ends_with("mod"))
+
+a <- eight$cluster_final_dif[[1]]
+a
+
+irt_cluster(a$difference_in_easy)

### 2020-03-04 sim.R

@@ -1,49 +1,0 @@
-library(tidyverse)
-library(MASS)
-
-# abilities
-sigma <- cbind(c(1, 0.5), c(0.5, 1))
-
-ref <- mvrnorm(10000, mu=c(0, 0), Sigma = sigma)
-var(ref)
-cor(ref)
-colMeans(ref)
-
-foc <- mvrnorm(10000, mu=c(-0.5, -1), Sigma = sigma)
-var(foc)
-cor(foc)
-colMeans(foc)
-
-# items
-n_items <- 20
-n_dif_items <- 5
-
-# go from disc2 to angle
-(angle <- acos(1 / sqrt(1 + disc2^2)) * 180 / pi)
-
-# and from angle to disc2
-sqrt((1 - cos(angle * pi / 180)^2) / cos(angle * pi / 180)^2)
-
-angle_to_a_nuisance <- function(angle){
-    # uses that cos formula. assumes a_target is 1
-    sqrt((1 - cos(angle * pi / 180)^2) / cos(angle * pi / 180)^2)
-}
-
-angle_to_a_nuisance(30)
-
-items <-
-    tibble(
-        item = 1:n_items,
-        dif = c(rep("no", n_items - n_dif_items), rep("yes", n_dif_items))
-    ) %>%
-    mutate(
-        a_target = 1,
-        angle = ifelse(dif == "no", 0, seq(30, 60, length.out = n_dif_items)),
-        a_nuisance = angle_to_a_nuisance(angle),
-        b_target = rnorm(n_items, 0.5, 0.5),
-        b_nuisance = 1
-    )
-
-
-
-

### 2020-03-09 paper/paper_2b_anchor_points.Rmd
after making progress before deleting stuff

@@
 $$
-MABC is inspired in part by @chalmers2016might who uses the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
+MABC is inspired in part by @chalmers2016might who use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
 The selection of $g(\theta)$ is important and determines where on the ability spectrum to consider DIF.

### 2020-03-09 paper/paper_3_sim.Rmd
after making progress before deleting stuff

To compare each of the methods in table TODO, we conducted a simulated study. Our goal was to make the data generating process as realistic to the scenario described by @ackerman1992didactic in which some items on a math test also depend on a student’s verbal ability (the target ability is math ability and the nuisance ability is verbal ability). As described in [the introduction](#intro), nearly all DIF simulation studies simply generate data where the item easiness parameters are different for the reference and focal groups. This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group. One exception is @walker2017using who draws each student's target and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.

In our simulation study, it was critical that student ability was drawn in a realistic way similar to @walker2017using. However, we don’t believe that a compensatory model is realistic in describing the a math test where some items depend on verbal ability. For example, it’s hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability. Accordingly, we generate item responses using an adapted version of Sympson’s -@sympson1978model noncompensatory item response model in which 
$$
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{2j}) = \sigma(\theta_i) \cdot \sigma(a_{2j}\eta_i)
$$
where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $\sigma(a_{2j})$ is the item's loading on nuisance ability [@demars2016partially].

## Drawing parameters

In each run, we simulate 10,000 students with half coming from each of the reference and focal groups. For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean [$\mu_\theta^\text{ref} = 0$, $\mu_\eta^\text{ref} = 0$] and covariance matrix $\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}$. Abilities for students from the focal group are drawn using the same covariance matrix but with means [$\mu_\theta^\text{foc} = -0.5$, $\mu_\eta^\text{foc} = -1$].

The test always has 20 items, but we vary the number of items with DIF from two to ten. For items without DIF, $a_{2j} = \infty$ so that the model reduces to $\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)$. For items with DIF, $a_{2j}$ is calculated based on Ackerman's -@ackerman1994using angle equation as described in @walker2017using:
$$
\angle_j = \arccos \dfrac{a_{1j}^2}{a_{1j}^2 + a_{2j}^2}.
$$
The angle for an item measure the relative loading of the item on the two dimensions. For example, an angle of 45$^\circ$ indicates that the item loads equally on the target and nuisance ability dimensions. We are using a simple compensatory model in which $a_{1j} = 1$ for all items so the angle equation reduces to 
$$
\angle_j = \arccos \dfrac{1}{1 + a_{2j}^2}.
$$
We are interested in specifying the angle of an item, so the relevant equation becomes 
$$
a_{2j} = \sqrt{\dfrac{1 - \cos(\angle_j)^2}{\cos(\angle_j)^2}}.
$$
For DIF items, we set $a_{2j}$ based on angles with equal intervals between 30$^\circ$ and 60$^\circ$. For example, for a test with four DIF items the angles are 30$^\circ$, 40$^\circ$, 50$^\circ$, and 60$^\circ$.

## Visualizing a run
```{r}
# ADD IN CODE THAT GENERATES THIS
out <- read_rds(here("paper/paper_3_out2.rds"))
```
Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean. The items are ordered by the amount of DIF such that $\angle_{j = 11} = 30^\circ$ up to $\angle_{j = 20} = 60^\circ$.
```{r difmap, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[2]]$pars[[4]] %>% 
  graph_pars_noncompensatory(0, -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = latex2exp::TeX("$\\theta_i$"),
    y = latex2exp::TeX("$Pr(y_{ij} = 1)$")
  )
```

Figure \ref{fig:simemmlg} shows the EM-MILG for one run using the same item parameters that generated Figure \ref{fig:difmap}. As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first ten items which are DIF free. For the last ten items, $\tilde{d_j}$ increases as $\angle_j$ increases. Of course, we cannot include the "analyst chooses" method in our simulation, but regardless, the EM-MILG is a useful way to visualize the data.
```{r simemmlg, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[2]]$intuitive_mod[[4]] %>%
    mod_intuitive_to_draws_df() %>%
    draws_df_to_logit_plot() +
    labs(
          x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
      )
```

## Outcomes

### Target ability achievement gap
An effective method, should find $\hat\mu_\theta^\text{foc}$ close to 0.5. This outcome measures a method’s ability to disentangle differences in target ability from nuisance ability at the group level.

### Individual abilities
Assessments are frequently used to make decisions about individual student abilities. We measure a method’s ability to do so by calculating the rank correlation between the vector estimated abilities and true abilities. This is done both for all students combined and just for students in the focal group. We use rank correlation as opposed to the more typically used RMSE as a result of our agreement with Lord's -@lord1986maximum argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.

### Anchor items. 
For the methods that choose a set of anchor items, we look directly at which anchor items are selected. An effective method should use most of the non-DIF items as anchors (the anchor hit rate) while avoiding including items with DIF in the anchor set (the false anchor rate).

## Results
In total, we executed 100 runs for each number of two to ten DIF items. This multiplies to a total of 900 runs. Figure \ref{fig:achievegap} shows each method's performance on estimating the target ability achievement gap.

As expected, artificial DIF causes AOAA and AOAA-AS to go off the rails as the number of DIF items increases. CLUSTERING typically performs well but has a higher variance because the anchor cluster sometimes contains as few as half of the DIF-free items.

AOAA-OAT, GINI, and MINBC all perform similarly well. MINBC has a tendency to slightly overestimate the achievement gap. There is an interesting pattern, not previously described in the literature, where AOAA-OAT and GINI of overestimate the achievement gap when there are few items with DIF and underestimate the achivement gap when there are many items with DIF. The cause is more easily understood with AOAA-OAT where we can inspect which items are included in the anchor set.

Figure \ref{fig:anchorfalse} shows the mean false anchor counts for each method and number of items with DIF. For example, when there are 10 items with DIF on the test, AOAA-AS incorrectly includes an average of 1.8 of those items in the anchor set. AOAA-OAT remarkably never includes an item with DIF in the anchor set (i.e., a false anchor rate of 0\%).

Performance, then, is driven by which items are included in the anchor set. Accordingly, Figure \ref{fig:anchorhit} shows the number of items without DIF that are included in the anchor set. In general, AOAA-OAT usually fails to include a couple of items in the anchor set that are indeed DIF free. When there are a few items with DIF, the omitted items are more often items in which the focal group happened to, as a result of sampling variability, perform well relative to the reference group. On the other hand, when there are many items with DIF, the omitted items are more often items in which the focal group happened to perform poorly relative to the reference group. This selective omission of DIF free anchors is what drives the positive association between number of DIF items and $\hat\mu_\theta^\text{foc}$.

```{r achievegap, fig.cap = 'Each method's estimation of the target ability achievement gap.', out.width="70%", warning = FALSE, message = FALSE}
final_model_to_anchor_summary <- function(final_model, n_dif_items){
    coef_1f(final_model)$items %>%
        mutate(
            true_anchor =
                c(
                    rep("no dif", n_items - n_dif_items),
                    rep("some dif", ceiling(n_dif_items / 2)),
                    rep("high dif", floor(n_dif_items / 2))
                )
        ) %>%
        group_by(true_anchor) %>%
        summarize(selected = sum(anchor), total = n())
}

getfull <- function(one){
    one %>%
        select(ends_with("mod"), -intuitive_mod) %>%
        mutate(run = row_number()) %>%
        gather(method, model, -run) %>%
        mutate(
            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
            foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
        ) %>%
        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod"))
}

out %>%
    mutate(achievegap = bigsim %>% map(getfull)) %>%
    select(-bigsim) %>%
    unnest(achievegap) %>%
    ggplot(aes(x = n_dif_items, y = foc_mean)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~ method) +
    geom_hline(yintercept = -0.5, linetype = "dashed") +
    labs(
      x = "Number of DIF items",
      y = latex2exp::TeX("$\\hat{\\mu}_\\theta^foc$")
    )
```

```{r anchorfalse, fig.cap = 'Mean false anchor counts', out.width="70%", warning = FALSE, message = FALSE}
one_to_anchors <- function(one, n_dif_items){
    one %>%
        select(ends_with("mod"), -intuitive_mod, -gini_mod, -minbc_mod) %>%
        mutate(run = row_number()) %>%
        gather(method, model, -run) %>%
        mutate(anchor_summary = model %>% map(final_model_to_anchor_summary, n_dif_items)) %>%
        select(-run, -model) %>%
        unnest(anchor_summary) %>%
        group_by(method, true_anchor, total) %>%
        summarize(mean_selected = mean(selected)) %>%
        ungroup()
}

temp <-
    out %>%
    mutate(anchors = map2(bigsim, n_dif_items, one_to_anchors)) %>%
    select(-bigsim) %>%
    unnest(anchors)

temp %>%
    filter(true_anchor != "no dif") %>%
    group_by(n_dif_items, method) %>%
    summarize(total = sum(total), mean_selected = sum(mean_selected)) %>%
    gather(var, val, -n_dif_items, -method) %>%
    mutate(var = factor(var, 
                        levels = c("total", "mean_selected"),
                        labels = c("Total possible", "Mean selected to anchor set"))) %>% 
  filter(var == "Mean selected to anchor set") %>% 
    ggplot(aes(x = n_dif_items, y = val)) +
    geom_point() +
    ylim(0, 20) +
    facet_wrap(~ method) +
    labs(
      x = "Number of items with DIF on the test",
      y = "Mean number of items with DIF in the anchor set",
      color = ""
    )
```

```{r anchorhit, fig.cap = 'Mean anchor hit counts', out.width="70%", warning = FALSE, message = FALSE}
temp %>%
    filter(true_anchor == "no dif") %>%
    gather(var, val, -n_dif_items, -method, -true_anchor) %>%
    mutate(var = factor(var, 
                        levels = c("total", "mean_selected"),
                        labels = c("Total possible", "Mean selected to anchor set"))) %>% 
    ggplot(aes(x = n_dif_items, y = val, color = var)) +
    geom_point() +
    ylim(0, 20) +
    facet_wrap(~ method) +
    labs(
      x = "Number of items with DIF on the test",
      y = "Number of items without DIF",
      color = ""
    )
```

### 2020-03-09 OUT.R

@@ -1,0 +1,111 @@
+out <- read_rds("paper/paper_4_out1.rds")
+
+n_items <- 20
+n_dif_items <- 6
+
+one <- out$bigsim[[n_dif_items - 1]]
+pars <- one$pars[[4]]
+
+# graph 1
+one$pars[[1]] %>% graph_pars_noncompensatory(0, -1)
+
+# graph 2
+one$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    draws_df_to_logit_plot()
+
+# function - one to achievement gap
+one_to_achievement_gap <- function(one){
+    one %>%
+        select(ends_with("mod"), -intuitive_mod) %>%
+        mutate(run = row_number()) %>%
+        gather(method, model, -run) %>%
+        mutate(
+            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
+            foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
+        ) %>%
+        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod")) %>%
+        group_by(method) %>%
+        summarize(out = mean(foc_mean), runs_with_nonzero_anchors = n())
+}
+
+getfull <- function(one){
+    one %>%
+        select(ends_with("mod"), -intuitive_mod) %>%
+        mutate(run = row_number()) %>%
+        gather(method, model, -run) %>%
+        mutate(
+            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
+            foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
+        ) %>%
+        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod"))
+}
+
+# graph 3
+out %>%
+    mutate(achievegap = bigsim %>% map(one_to_achievement_gap)) %>%
+    select(-bigsim) %>%
+    unnest(achievegap) %>%
+    ggplot(aes(x = n_dif_items, y = out)) +
+    geom_point() +
+    facet_wrap(~ method) +
+    geom_hline(yintercept = -0.5, linetype = "dashed")
+
+out %>%
+    mutate(achievegap = bigsim %>% map(getfull)) %>%
+    select(-bigsim) %>%
+    unnest(achievegap) %>%
+    ggplot(aes(x = n_dif_items, y = foc_mean)) +
+    geom_point(alpha = 0.5) +
[…]

### 2020-03-09 R/___bigsim___.R

@@ -16,7 +16,7 @@ bigsim <- function(runs, n_dif_items){
                     ref_target_ability_mu = 0,
                     ref_nuisance_ability_mu = 0,
                     foc_target_ability_mu = -0.5,
-                    foc_nuisance_ability_mu = -2
+                    foc_nuisance_ability_mu = -1
                 )
             )
     ) %>%
@@ -63,7 +63,7 @@ bigsim <- function(runs, n_dif_items){
         ) %>%
         # dif: AOAA OAT
         mutate(
-            AOAA_OAT_status = map2(sim, AOAA_status, ~ dif_AOAA_OAT(.x$data, .x$groups, .y)),
+            AOAA_OAT_status = map(sim, ~ dif_AOAA_OAT(.$data, .$groups)),
             AOAA_OAT_mod =
                 map2(
                     sim,

### 2020-03-09 R/create_pars_noncompensatory.R

@@ -1,4 +1,6 @@
-# create_pars_noncompensatory(n_items = 10, n_dif_items = 5) %>%
+# pars <- create_pars_noncompensatory(n_items = 10, n_dif_items = 5, angle_start = 30, angle_end = 60, b_nuisance_with_dif_sd = 0)
+#
+# create_pars_noncompensatory(n_items = 10, n_dif_items = 5, angle_start = 30, angle_end = 60, b_nuisance_with_dif_sd = 0) %>%
 #     graph_pars_noncompensatory(0, -1)
 
 angle_to_a_nuisance <- function(angle){
@@ -39,8 +41,12 @@ graph_pars_noncompensatory <- function(pars, ref_mean_ability_nuisance, foc_mean
                 sigmoid(a_nuisance * foc_mean_ability_nuisance + b_nuisance)
         ) %>%
         dplyr::select(item, ability_target, ref, foc) %>%
-        gather(group, prob, -item, -ability_target) %>%
+        mutate(same = ref == foc) %>%
+        gather(group, prob, -item, -ability_target, -same) %>%
+        filter((same & group == "ref") | !same) %>%
+        mutate(group = ifelse(same, "both", group)) %>%
         ggplot(aes(x = ability_target, y = prob, color = group)) +
         geom_point() +
-        facet_wrap(~ item)
+        facet_wrap(~ item) +
+        scale_color_manual(values = c("black", "darkgreen", "red"))
 }

### 2020-03-09 R/dif_AOAA.R

@@ -39,6 +39,19 @@ get_next_status <- function(status, data, groups){
         select(-status)
 }
 
+get_next_status_OAT <- function(status, data, groups){
+    the_flex <- which(status$status == "flex")
+
+    status %>%
+        filter(status == "unknown") %>%
+        mutate(
+            anova = item %>% map(~ quiet(anova(mod_flexible(data, groups, the_flex), mod_flexible(data, groups, flex_items = c(the_flex, .))))),
+            X2 = anova %>% map_dbl(~ .$X2[2]),
+            p = anova %>% map_dbl(~ .$p[2])
+        ) %>%
+        select(-status)
+}
+
 # AOAA - AS ---------------------------------------------------------------
 dif_AOAA_AS <- function(data, groups, dif_aoaa){
 
@@ -84,7 +97,19 @@ dif_AOAA_AS <- function(data, groups, dif_aoaa){
 }
 
 # AOAA - OAT --------------------------------------------------------------
-dif_AOAA_OAT <- function(data, groups, dif_aoaa){
+dif_AOAA_OAT <- function(data, groups){
+    # I was using p-values but those get to 0 so switched to X2 values
+    # It works but did so in not the absolute best code style
+
+    dif_aoaa <-
+        tibble(
+            item = 1:ncol(data)
+        ) %>%
+        mutate(
+            anova = item %>% map(~ quiet(anova(mod_constrained_baseline(data, groups), mod_flexible(data, groups, flex_items = .)))),
+            X2 = anova %>% map_dbl(~ .$X2[2]),
+            p = anova %>% map_dbl(~ .$p[2])
+        )
 
     # if no items are significant, then they are all anchors and we are done
     if (all(dif_aoaa$p > 0.05)){
@@ -95,24 +120,24 @@ dif_AOAA_OAT <- function(data, groups, dif_aoaa){
     status <-
         dif_aoaa %>%
         mutate(
-            status = ifelse(item == which.min(p), "flex", "unknown")
+            status = ifelse(item == which.max(X2), "flex", "unknown")
         )
 
     while(TRUE){
         # for each unknown item, we use all other unknown items to test it
-        new_status <- get_next_status(status, data, groups)
+        new_status <- get_next_status_OAT(status, data, groups)
 
         # need to update the status df
         status <-
             status %>%
-            select(-p) %>%
+            select(-anova, -X2, -p) %>%
             left_join(new_status, by = "item")
 
         # if no new significant tests then we're done
[…]

### 2020-03-09 R/zelpers.R

@@ -5,6 +5,9 @@ quiet <- function(x){
 }
 
 sigmoid <- function(x){
-    exp(x) / (1 + exp(x))
+    tmp <- exp(x) / (1 + exp(x))
+    tmp[is.nan(tmp)] <- 1
+    tmp
 }
+
 

### 2020-03-09 paper/paper_2_sim.R

@@ -41,7 +41,7 @@ out_3_biased_items <-
     ) %>%
     # dif: AOAA OAT
     mutate(
-        AOAA_OAT_status = map2(sim, AOAA_status, ~ dif_AOAA_OAT(.x$data, .x$groups, .y)),
+        AOAA_OAT_status = map(sim, ~ dif_AOAA_OAT(.$data, .$groups)),
         AOAA_OAT_mod =
             map2(
                 sim,

### 2020-03-09 paper/paper_3_sim.R

@@ -1,0 +1,22 @@
+# NEED TO ADD MORE DIF
+# DO AN EXTREME CASE THEN WORK BACK FROM THERE
+
+# (higher values of b_target are associated with less dif)
+
+# out_noncomp$intuitive_mod[[1]] %>%
+#     mod_intuitive_to_draws_df() %>%
+#     draws_df_to_logit_plot()
+
+library(furrr)
+options(future.fork.enable = TRUE)
+plan(multiprocess)
+
+out <-
+    tibble(n_dif_items = c(4, 7, 10)) %>%
+    mutate(
+        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 10, n_dif_items = .))
+    )
+
+Sys.time()
+
+out %>% write_rds("paper/paper_4_out2.rds")

### 2020-03-09 paper/paper_4_sim.R

@@ -1,51 +1,0 @@
-# NEED TO ADD MORE DIF
-# DO AN EXTREME CASE THEN WORK BACK FROM THERE
-
-# (higher values of b_target are associated with less dif)
-
-# out_noncomp$intuitive_mod[[1]] %>%
-#     mod_intuitive_to_draws_df() %>%
-#     draws_df_to_logit_plot()
-
-out <-
-    tibble(n_dif_items = c(4, 6, 8, 10)) %>%
-    mutate(
-        bigsim = n_dif_items %>% map(~ bigsim(runs = 2, n_dif_items = .))
-    )
-
-out %>% write_rds("paper/paper_4_out.rds")
-
-
-
-one$intuitive_mod[[2]] %>%
-    mod_intuitive_to_draws_df() %>%
-    draws_df_to_logit_plot()
-
-n_items <- 20
-n_dif_items <- 10
-one <- out$bigsim[[n_dif_items - 2]]
-
-one$pars[[1]]
-
-one %>%
-    select(ends_with("mod"), -intuitive_mod) %>%
-    mutate(run = row_number()) %>%
-    gather(method, model, -run) %>%
-    mutate(
-        foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
-    ) %>%
-    ggplot(aes(x = method, y = foc_mean)) +
-    geom_point()
-
-one %>%
-    select(ends_with("mod"), -intuitive_mod, -gini_mod, -minbc_mod) %>%
-    mutate(run = row_number()) %>%
-    gather(method, model, -run) %>%
-    mutate(anchor_summary = model %>% map(final_model_to_anchor_summary)) %>%
-    select(-run, -model) %>%
-    unnest(anchor_summary) %>%
-    count(method, true_anchor, selected) %>%
-    ggplot(aes(x = selected, y = n)) +
-    geom_col() +
-    facet_grid(true_anchor ~ method)
-

### 2020-03-09 paper/paper_3_sim.Rmd
marching along

@@
 # ADD IN CODE THAT GENERATES THIS
-out <- read_rds(here("paper/paper_3_out2.rds"))
+out <- read_rds(here("paper/paper_3_out1.rds"))
+
+out <-
+    out %>%
+    mutate(bigsim = map2(bigsim, n_dif_items, add_foc_mean))
+
+# i have no idea why this is like this...
+out %>% 
+  mutate(bigsim = bigsim %>% map(~ select(., foc_mean))) %>% 
+  unnest(bigsim) %>% 
+  ggplot(aes(x = n_dif_items, y = foc_mean)) +
+  geom_point()
+
+out$bigsim[[4]]$p[[1]] %>% colMeans()
+
 ```

### 2020-03-09 OUT.R

@@ -1,111 +1,0 @@
-out <- read_rds("paper/paper_4_out1.rds")
-
-n_items <- 20
-n_dif_items <- 6
-
-one <- out$bigsim[[n_dif_items - 1]]
-pars <- one$pars[[4]]
-
-# graph 1
-one$pars[[1]] %>% graph_pars_noncompensatory(0, -1)
-
-# graph 2
-one$intuitive_mod[[1]] %>%
-    mod_intuitive_to_draws_df() %>%
-    draws_df_to_logit_plot()
-
-# function - one to achievement gap
-one_to_achievement_gap <- function(one){
-    one %>%
-        select(ends_with("mod"), -intuitive_mod) %>%
-        mutate(run = row_number()) %>%
-        gather(method, model, -run) %>%
-        mutate(
-            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
-            foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
-        ) %>%
-        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod")) %>%
-        group_by(method) %>%
-        summarize(out = mean(foc_mean), runs_with_nonzero_anchors = n())
-}
-
-getfull <- function(one){
-    one %>%
-        select(ends_with("mod"), -intuitive_mod) %>%
-        mutate(run = row_number()) %>%
-        gather(method, model, -run) %>%
-        mutate(
-            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
-            foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
-        ) %>%
-        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod"))
-}
-
-# graph 3
-out %>%
-    mutate(achievegap = bigsim %>% map(one_to_achievement_gap)) %>%
-    select(-bigsim) %>%
-    unnest(achievegap) %>%
-    ggplot(aes(x = n_dif_items, y = out)) +
-    geom_point() +
-    facet_wrap(~ method) +
-    geom_hline(yintercept = -0.5, linetype = "dashed")
-
-out %>%
-    mutate(achievegap = bigsim %>% map(getfull)) %>%
-    select(-bigsim) %>%
-    unnest(achievegap) %>%
-    ggplot(aes(x = n_dif_items, y = foc_mean)) +
-    geom_point(alpha = 0.5) +
[…]

### 2020-03-09 R/___bigsim___.R

@@ -1,7 +1,8 @@
 bigsim <- function(runs, n_dif_items){
     tibble(
-        pars = rerun(runs,
-                     create_pars_noncompensatory(
+        pars = rerun(
+                    runs,
+                    create_pars_noncompensatory(
                          n_items = 20,
                          n_dif_items = n_dif_items,
                          angle_start = 30,
@@ -9,10 +10,10 @@ bigsim <- function(runs, n_dif_items){
                          b_nuisance_with_dif_sd = 0)),
         students =
             rerun(
-                1,
+                runs,
                 sim_students_2d(
-                    n_ref = 5000,
-                    n_foc = 5000,
+                    n_ref = 20000, # NEED TO EDIT THIS
+                    n_foc = 20000,
                     ref_target_ability_mu = 0,
                     ref_nuisance_ability_mu = 0,
                     foc_target_ability_mu = -0.5,
@@ -29,38 +30,38 @@ bigsim <- function(runs, n_dif_items){
         mutate(
             intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
         ) %>%
-        # dif: AOAA
-        mutate(
-            AOAA_status = sim %>% map(~ dif_AOAA(.$data, .$groups)),
-            AOAA_mod =
-                map2(
-                    sim,
-                    AOAA_status,
-                    ~ mod_flexible(
-                        .x$data,
-                        .x$groups,
-                        which(.y$status == "flex"),
-                        se = TRUE
-                    )
-                ),
-            AOAA_final_dif = AOAA_mod %>% map(mod_to_final_dif)
-        ) %>%
-        # dif: AOAA AS
-        mutate(
-            AOAA_AS_status = map2(sim, AOAA_status, ~ dif_AOAA_AS(.x$data, .x$groups, .y)),
-            AOAA_AS_mod =
-                map2(
-                    sim,
-                    AOAA_AS_status,
-                    ~ mod_flexible(
-                        .x$data,
-                        .x$groups,
-                        which(.y$status == "flex"),
-                        se = TRUE
-                    )
-                ),
-            AOAA_AS_final_dif = AOAA_AS_mod %>% map(mod_to_final_dif)
-        ) %>%
+        # # dif: AOAA
+        # mutate(
[…]

### 2020-03-09 R/sim_students_noncomp.R

@@ -31,7 +31,6 @@ sim_students_2d <- function(n_ref, n_foc, ref_target_ability_mu, ref_nuisance_ab
 get_p_noncomp <- function(a_target, a_nuisance, b_target, b_nuisance, target_ability, nuisance_ability){
     target <- sigmoid(a_target * target_ability + b_target)
     nuisance <- sigmoid(a_nuisance * nuisance_ability + b_nuisance)
-    nuisance <- ifelse(is.nan(nuisance), 1, nuisance)
     target * nuisance
 }
 

### 2020-03-09 _add_foc_mean.R

@@ -1,0 +1,25 @@
+sim_n_dif_items_to_foc_mean <- function(sim, n_dif_items){
+    mod <-
+        mod_flexible(
+            sim$data,
+            sim$groups,
+            flex_items = (ncol(sim$data) - n_dif_items + 1):ncol(sim$data)
+        )
+
+    coef_1f(mod)$ability$b_foc[1]
+}
+
+add_foc_mean <- function(one, n_dif_items){
+    one %>%
+    mutate(foc_mean = sim %>% map_dbl(sim_n_dif_items_to_foc_mean, n_dif_items))
+}
+
+out <-
+    out %>%
+    mutate(bigsim = map2(bigsim, n_dif_items, add_foc_mean))
+
+out$bigsim[[1]]$foc_mean
+
+out$bigsim[[2]]$foc_mean
+
+out$bigsim[[3]]$foc_mean

### 2020-03-09 _new_out.R

@@ -1,0 +1,13 @@
+# EDIT BACK FROM 20000 to 5000
+
+library(furrr)
+options(future.fork.enable = TRUE)
+plan(multiprocess)
+
+out <-
+    tibble(n_dif_items = c(4, 6, 8)) %>%
+    mutate(
+        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 4, n_dif_items = .))
+    )
+
+Sys.time()

### 2020-03-09 scratch.R

@@ -1,8 +1,0 @@
-
-eight %>%
-    select(ends_with("mod"))
-
-a <- eight$cluster_final_dif[[1]]
-a
-
-irt_cluster(a$difference_in_easy)

### 2020-03-10 R/___bigsim___.R
before deleting stuff

@@ -1,19 +1,21 @@
 bigsim <- function(runs, n_dif_items){
     tibble(
-        pars = rerun(
-                    runs,
-                    create_pars_noncompensatory(
-                         n_items = 20,
-                         n_dif_items = n_dif_items,
-                         angle_start = 30,
-                         angle_end = 60,
-                         b_nuisance_with_dif_sd = 0)),
+        pars =
+            rerun(
+                runs,
+                create_pars_noncompensatory(
+                    n_items = 20,
+                    n_dif_items = n_dif_items,
+                    angle_start = 30,
+                    angle_end = 60,
+                    b_nuisance_with_dif_sd = 0)
+                ),
         students =
             rerun(
                 runs,
                 sim_students_2d(
-                    n_ref = 20000, # NEED TO EDIT THIS
-                    n_foc = 20000,
+                    n_ref = 500, # CHANGE THIS
+                    n_foc = 500,
                     ref_target_ability_mu = 0,
                     ref_nuisance_ability_mu = 0,
                     foc_target_ability_mu = -0.5,
@@ -30,38 +32,38 @@ bigsim <- function(runs, n_dif_items){
         mutate(
             intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
         ) %>%
-        # # dif: AOAA
-        # mutate(
-        #     AOAA_status = sim %>% map(~ dif_AOAA(.$data, .$groups)),
-        #     AOAA_mod =
-        #         map2(
-        #             sim,
-        #             AOAA_status,
-        #             ~ mod_flexible(
-        #                 .x$data,
-        #                 .x$groups,
-        #                 which(.y$status == "flex"),
-        #                 se = TRUE
-        #             )
-        #         ),
-        #     AOAA_final_dif = AOAA_mod %>% map(mod_to_final_dif)
-        # ) %>%
-        # # dif: AOAA AS
-        # mutate(
-        #     AOAA_AS_status = map2(sim, AOAA_status, ~ dif_AOAA_AS(.x$data, .x$groups, .y)),
-        #     AOAA_AS_mod =
-        #         map2(
-        #             sim,
-        #             AOAA_AS_status,
-        #             ~ mod_flexible(
-        #                 .x$data,
-        #                 .x$groups,
-        #                 which(.y$status == "flex"),
[…]

### 2020-03-10 R/dif_zelpers_level_2.R

@@ -5,7 +5,7 @@ mod_constrained_baseline <- function(data, groups){
         itemtype = "Rasch",
         groups,
         invariance = c("free_means", "free_var", "intercepts", "slopes"),
-        technical = list(NCYCLES = 100),
+        technical = list(NCYCLES = 500),
         verbose = FALSE
     )
 }
@@ -17,7 +17,7 @@ mod_flexible <- function(data, groups, flex_items, se = FALSE){
         itemtype = "Rasch",
         groups,
         invariance = c("free_means", "free_var", paste0("Item_", (1:ncol(data))[-flex_items])),
-        technical = list(NCYCLES = 100),
+        technical = list(NCYCLES = 500),
         SE = se,
         verbose = FALSE
     )

### 2020-03-10 _add_foc_mean.R

@@ -14,12 +14,19 @@ add_foc_mean <- function(one, n_dif_items){
     mutate(foc_mean = sim %>% map_dbl(sim_n_dif_items_to_foc_mean, n_dif_items))
 }
 
-out <-
+out2 <-
     out %>%
     mutate(bigsim = map2(bigsim, n_dif_items, add_foc_mean))
 
-out$bigsim[[1]]$foc_mean
+out$bigsim[[1]]$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    draws_df_to_logit_plot() +
+    labs(
+        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
+    )
 
-out$bigsim[[2]]$foc_mean
+a <- dif_AOAA_OAT(out$bigsim[[1]]$sim[[1]]$data, out$bigsim[[1]]$sim[[1]]$groups)
 
-out$bigsim[[3]]$foc_mean
+
+
+

### 2020-03-10 _new_out.R

@@ -1,5 +1,3 @@
-# EDIT BACK FROM 20000 to 5000
-
 library(furrr)
 options(future.fork.enable = TRUE)
 plan(multiprocess)
@@ -7,7 +5,23 @@ plan(multiprocess)
 out <-
     tibble(n_dif_items = c(4, 6, 8)) %>%
     mutate(
-        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 4, n_dif_items = .))
+        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 2, n_dif_items = .))
     )
 
 Sys.time()
+
+
+out$bigsim[[1]]$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    draws_df_to_logit_plot() +
+    labs(
+        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
+    )
+
+out$bigsim[[1]]$pars[[1]]
+
+debug(dif)
+
+a <- dif_AOAA_OAT(out$bigsim[[1]]$sim[[1]]$data, out$bigsim[[1]]$sim[[1]]$groups)
+
+

### 2020-03-10 paper/paper_3_sim.R

@@ -12,11 +12,11 @@ options(future.fork.enable = TRUE)
 plan(multiprocess)
 
 out <-
-    tibble(n_dif_items = c(4, 7, 10)) %>%
+    tibble(n_dif_items = 2:10) %>%
     mutate(
-        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 10, n_dif_items = .))
+        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 5, n_dif_items = .))
     )
 
 Sys.time()
 
-out %>% write_rds("paper/paper_4_out2.rds")
+out %>% write_rds("paper/paper_3_out1.rds")

### 2020-03-10 paper/paper.Rmd
after making progress before deleting stuff
simulation went well and i think i have my final graph

@@
 ```{r setup, include = FALSE}
-# 
 # https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
@@
 library(kableExtra)
-set.seed(123)
+set.seed(12345)
 R.utils::sourceDirectory(here("R"))
@@
 
-# Introduction
+# Introduction {#intro}
 

### 2020-03-10 paper/paper_3_sim.Rmd
before deleting stuff
simulation went well and i think i have my final graph

@@
 
+sim_n_dif_items_to_foc_mean <- function(sim, n_dif_items){
+    mod <-
+        mod_flexible(
+            sim$data,
+            sim$groups,
+            flex_items = (ncol(sim$data) - n_dif_items + 1):ncol(sim$data)
+        )
+
+    coef_1f(mod)$ability$b_foc[1]
+}
+
+add_foc_mean <- function(one, n_dif_items){
+    one %>%
+    mutate(foc_mean = sim %>% map_dbl(sim_n_dif_items_to_foc_mean, n_dif_items))
+}
+
 out <-
@@
   geom_point()
-
-out$bigsim[[4]]$p[[1]] %>% colMeans()
-
 ```
@@
 ```{r difmap, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
-out$bigsim[[2]]$pars[[4]] %>% 
+out$bigsim[[2]]$pars[[1]] %>% 
   graph_pars_noncompensatory(0, -1) +
@@
 ```{r simemmlg, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
-out$bigsim[[2]]$intuitive_mod[[4]] %>%
+out$bigsim[[3]]$intuitive_mod[[1]] %>%
     mod_intuitive_to_draws_df() %>%
@@
 ```{r achievegap, fig.cap = 'Each method's estimation of the target ability achievement gap.', out.width="70%", warning = FALSE, message = FALSE}
+
+n_items <- 12
+
 final_model_to_anchor_summary <- function(final_model, n_dif_items){
@@
     one %>%
-        select(ends_with("mod"), -intuitive_mod) %>%
+        select(foc_mean, ends_with("mod"), -intuitive_mod) %>%
         mutate(run = row_number()) %>%
-        gather(method, model, -run) %>%
+        gather(method, model, -run, -foc_mean) %>%
         mutate(
             n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
-            foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
+            est_foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
         ) %>%
-        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod"))
+        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod")) %>% 
+    mutate(resid = est_foc_mean - foc_mean)
 }
 
+what <- 
   out %>%
@@
     select(-bigsim) %>%
-    unnest(achievegap) %>%
-    ggplot(aes(x = n_dif_items, y = foc_mean)) +
-    geom_point(alpha = 0.5) +
+    unnest(achievegap)
+
+what %>% 
+  ggplot(aes(x = resid, y = as.factor(n_dif_items))) +
+  ggridges::geom_density_ridges() +
   facet_wrap(~ method) +
+  coord_flip() +
   geom_hline(yintercept = -0.5, linetype = "dashed") +

### 2020-03-10 R/___bigsim___.R

@@ -1,12 +1,13 @@
 bigsim <- function(runs, n_dif_items){
+
     tibble(
         pars =
             rerun(
                 runs,
                 create_pars_noncompensatory(
-                    n_items = 20,
+                    n_items = 12,
                     n_dif_items = n_dif_items,
-                    angle_start = 30,
+                    angle_start = 20,
                     angle_end = 60,
                     b_nuisance_with_dif_sd = 0)
                 ),
@@ -14,8 +15,8 @@ bigsim <- function(runs, n_dif_items){
             rerun(
                 runs,
                 sim_students_2d(
-                    n_ref = 500, # CHANGE THIS
-                    n_foc = 500,
+                    n_ref = 5000,
+                    n_foc = 5000,
                     ref_target_ability_mu = 0,
                     ref_nuisance_ability_mu = 0,
                     foc_target_ability_mu = -0.5,
@@ -80,7 +81,7 @@ bigsim <- function(runs, n_dif_items){
                 ),
             AOAA_OAT_final_dif = AOAA_OAT_mod %>% map(mod_to_final_dif)
         ) %>%
-        # dif: tibs cluster
+        # dif: 3 clusters
         mutate(
             cluster =
                 intuitive_mod %>%
@@ -113,3 +114,4 @@ bigsim <- function(runs, n_dif_items){
             minbc_mod = map2(minbc, sim, ~ anchptfocmean_to_final_mod(.x$between_curves_anchor_points, .y$data, .y$groups))
         )
 }
+

### 2020-03-10 R/create_pars_noncompensatory.R

@@ -12,6 +12,7 @@ angle_to_a_nuisance <- function(angle){
 create_pars_noncompensatory <- function(n_items, n_dif_items, angle_start, angle_end, b_nuisance_with_dif_sd){
     # create item parameters. all target disc is 1 and nuisance disc is 0 for non-dif items and has angle that
     # is lattice from 30 to 60 for dif items
+
     tibble(
         item = 1:n_items,
         dif = c(rep("no", n_items - n_dif_items), rep("yes", n_dif_items))
@@ -20,7 +21,7 @@ create_pars_noncompensatory <- function(n_items, n_dif_items, angle_start, angle
             a_target = 1,
             angle = c(rep(0, n_items - n_dif_items), seq(angle_start, angle_end, length.out = n_dif_items)),
             a_nuisance = angle_to_a_nuisance(angle),
-            b_target = rnorm(n_items, 0, 0.0001), # CHANGE THIS BACK
+            b_target = rnorm(n_items, 0, 0.0001), # CHANGE THIS BACK?
             b_nuisance = c(rep(999, n_items - n_dif_items), rnorm(n_dif_items, 0, b_nuisance_with_dif_sd))
         )
 }

### 2020-03-10 _add_foc_mean.R

@@ -1,32 +1,0 @@
-sim_n_dif_items_to_foc_mean <- function(sim, n_dif_items){
-    mod <-
-        mod_flexible(
-            sim$data,
-            sim$groups,
-            flex_items = (ncol(sim$data) - n_dif_items + 1):ncol(sim$data)
-        )
-
-    coef_1f(mod)$ability$b_foc[1]
-}
-
-add_foc_mean <- function(one, n_dif_items){
-    one %>%
-    mutate(foc_mean = sim %>% map_dbl(sim_n_dif_items_to_foc_mean, n_dif_items))
-}
-
-out2 <-
-    out %>%
-    mutate(bigsim = map2(bigsim, n_dif_items, add_foc_mean))
-
-out$bigsim[[1]]$intuitive_mod[[1]] %>%
-    mod_intuitive_to_draws_df() %>%
-    draws_df_to_logit_plot() +
-    labs(
-        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
-    )
-
-a <- dif_AOAA_OAT(out$bigsim[[1]]$sim[[1]]$data, out$bigsim[[1]]$sim[[1]]$groups)
-
-
-
-

### 2020-03-10 paper/paper_3_sim.R

@@ -12,9 +12,9 @@ options(future.fork.enable = TRUE)
 plan(multiprocess)
 
 out <-
-    tibble(n_dif_items = 2:10) %>%
+    tibble(n_dif_items = 2:6) %>%
     mutate(
-        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 5, n_dif_items = .))
+        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 20, n_dif_items = .))
     )
 
 Sys.time()

### 2020-03-14 paper/paper.Rmd
printed second verison!

@@
 title: |
-    New and old ways to chase a rabbit:  
-    Differential item functioning detection methods
+    Promise and peril: Agnostic identification methods  
+    for detecting differential item functioning
 subtitle: |
-    Ben Stenhaug  
+    Ben Stenhaug, Ben Domingue, and Mike Frank  
     Stanford University
 abstract: |
-  summarize propose extensions insist as dif as student property and on communicating uncertainty
-  <!-- https://tinyurl.com/ybremelq -->
+  It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
+However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF.
+We summarize existing assumptionless identification (AI) methods and propose a variety of new methods.
+We conduct a simulation study — which we believe to be more realistic than most DIF simulation studies in the literature — and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that presents clearly all information about possible DIF, including sampling variability in item parameters, to the analyst.
+\clearpage <!-- https://tinyurl.com/ybremelq -->
 header-includes: 
@@
 
-# DIF methods
+# Assumptionless identification methods
 
@@
 
-# Next steps
+# Simulation study
 
-```{r child = "paper_3_next_steps.Rmd"}
+```{r child = "paper_3_sim.Rmd"}
 ```
@@
 \clearpage
-
-# Notes
-
-- adjust to all be easiness including multidimensional talk
-
-# References
 

### 2020-03-14 paper/paper.tex
printed second verison!

@@
 \hypersetup{
-  pdftitle={New and old ways to chase a rabbit: Differential item functioning detection methods},
+  pdftitle={Promise and peril: Agnostic identification methods for detecting differential item functioning},
   hidelinks,
@@
 
-\title{New and old ways to chase a rabbit:\\
-Differential item functioning detection methods}
+\title{Promise and peril: Agnostic identification methods\\
+for detecting differential item functioning}
 \usepackage{etoolbox}
@@
 \makeatother
-\subtitle{Ben Stenhaug\\
+\subtitle{Ben Stenhaug, Ben Domingue, and Mike Frank\\
 Stanford University}
@@
 \begin{abstract}
-summarize propose extensions insist as dif as student property and on communicating uncertainty
+It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
+However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF.
+We summarize existing assumptionless identification (AI) methods and propose a variety of new methods.
+We conduct a simulation study --- which we believe to be more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that presents clearly all information about possible DIF, including sampling variability in item parameters, to the analyst.
+\clearpage
 \end{abstract}
@@
 
-\hypertarget{to-do}{%
-\section{TO DO}\label{to-do}}
+\hypertarget{intro}{%
+\section{Introduction}\label{intro}}
 
-\begin{itemize}
-\tightlist
-\item
-  be sure all my directions match. i think i switched everything to be item easiness at some point but not sure if i did it throughout
-\end{itemize}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we think differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
-From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is a property of the student, not the item.
+Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we think of differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
+From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item.
 For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
@@
 They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
-They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
+They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
 They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
@@
 The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
+The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for all students in the reference group and \({\eta_i}^{\text{ref}} = -1\) for all students in the focal group.
+We now need to use a compensatory 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional 2PL model, the probability student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
-\end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability for students in the same group.
+According to the two-dimensional compensatory 2PL model, the probability student \(i\) responding correctly to item \(j\) is
+\[
+\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
+\]
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability across students in the same group.
 
@@
 
-\hypertarget{dif-methods}{%
-\section{DIF methods}\label{dif-methods}}
+\hypertarget{assumptionless-identification-methods}{%
+\section{Assumptionless identification methods}\label{assumptionless-identification-methods}}
 
 Psychometricians have long been in search of the perfect DIF detection method.
-In this section, we summarize existing DIF methods and propose extensions of those methods.
-We limit our discussion to IRT-based methods.
+It has become well known that the IRT-based likelihood ratio test (LRT) is the best way to test for DIF, which has been well demonstrated (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
 As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).
+
+The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions --- about either relative group ability or which items contain bias --- can be made.
+That is, we're interested in assumptionless identification methods (hereafter referred to as \enquote{AI methods}) so that the LRTs can be used to test items for DIF.
 For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
 And, in search of a coherent framework, we sometimes edit names of existing methods.
-We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) and that we risk contributing to the proliferation of names.
+We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to the proliferation of names.
 
+In this section, we summarize existing AI methods and propose extensions of those methods.
 We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
@@
 Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{ref}} = -1\).
-The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
+The slope on target ability is set to \(a_{1j} = 1\) for all items.
+The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
 From the DIF-as-item-property view, there is no nuisance ability.
@@
 We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
-The fundamental challenge for a DIF method is to first get an estimate for \(\mu^\text{foc}\), \(\hat\mu^\text{foc}\).
-How can group ability be disentangled from the possibility that items are biased?
-The most common approach is to identify a group of anchor items that are assumed to be DIF-free.
-These items are used to get \(\hat\mu^\text{foc}\).
+The fundamental challenge is to use an AI method to identify the model, which allows for the disentangling \(\hat\mu^\text{foc}\) from estimating \(\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\).
+The most common AI method is to select a group of anchor items that are assumed to be DIF-free.
+These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.
 
-\hypertarget{anchor-items}{%
-\subsection{Anchor items}\label{anchor-items}}
+Computing is done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the R package mirt (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the set of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}).
+Code is available at TO DO.
+
+\hypertarget{anchoritems}{%
+\subsection{Anchor items}\label{anchoritems}}
 
@@
 
-Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA).
+Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method.
 AOAA tests each item for DIF one at a time using all of the other items as anchors.
 For example, when testing the first item for DIF, all of the other items are used as anchors.
-This is done using a likelihood ratio test that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}).
+This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}).
 Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
-The items for which the flexible model outperforms the baseline model (typically based on a chi-squared test) are identified as having DIF and the rest of the items become anchor items.
+The items for which the flexible model outperforms the baseline model (typically based on a \(\chi^2\) test) are identified as having DIF, and the rest of the items become anchor items.
+Note that AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
 
@@
 (\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can we be sure its those 10 items and not the other 11 items with DIF?
-We can't be.
+How can they be sure that its those 10 items and not the other 11 items with DIF?
+They cannot be.
 Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group and others are biased against the reference group.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method (Strobl et al.
+More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
+Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method (Strobl et al.
 \protect\hyperlink{ref-strobl2018anchor}{2018}).
@@
 We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF being removed from the anchor set.
+Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
+Note that AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.
 
-AOAA-AS might seem like an improvement but it doesn't solve a fundamental problem of AOAA: What do we do when all items test positive for DIF?
+AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF?
 With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
-In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF and there are no anchor items.
+In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items.
 Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
 Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items display DIF themselves.
+It's unclear if one should proceed if those four items display DIF.
 
@@
 
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been proposed.
+We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
+AOAA-OAT is inspired by Hagquist and Andrich (\protect\hyperlink{ref-hagquist2017recent}{2017}), who, in general, assert that \enquote{items showing DIF initially should not be resolved simultaneously but sequentially} {[}p.~6{]}.
 Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the \(\chi^2\) test statistic, is removed from the anchor set.
@@
 AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
-As a result, AOAA-OAT is less likely to arrive at the wrong conclusion because of artifical DIF.
+As a result, we believe that AOAA-OAT is less likely to be \enquote{fooled} by artifical DIF.
+Note that AOAA-OAT is not currently implemented in the R package mirt.
 
@@
 Table \ref{table:themethods} summarizes the three all-others-as-anchors methods.
-It's useful to remember that AOAA is not an iterative procedure, while the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
+It's useful to remember that AOAA is not an iterative procedure.
+The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
+In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
+On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.
 
 \begin{table}[H]
-\caption{EDIT THIS}
+\caption{Summary of the three all-others-as-anchors assumptionless identification methods}
 \centering
@@
 
-{[}ADD IN SIMULATION RESULTS{]}
-
 \hypertarget{clustering-items}{%
@@
 
-Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster.} They pointed out that one way around the unidentifiability issue is to consider relative item parameters.
+Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster.} They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
 For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\).
@@
 The general idea of identifying clusters of items is intriguing.
-However, their approach is needlessly complicated and they did not describe a process for moving from \(\Delta \mathbf{R}\) to a set of anchor items.
+However, their approach is needlessly complicated, and they did not describe a process for moving from \(\Delta \mathbf{R}\) to an anchor cluster.
 Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing one such process.
@@
 Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\) of differences in item easiness, \(\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}\), where the model is identified by setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), thus the name \enquote{equal means clustering}).
-This results in all differences in performance --- either from group ability differences or DIF --- manifesting in the item easiness difference parameter \(\tilde{d_j}\).
+As a result, all differences in performance --- either from group ability differences or DIF --- manifesting in the item easiness difference parameter \(\tilde{d_j}\).
 The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) arbitrarily set to \(0\).
@@
 If there is a tie for the largest cluster, the cluster with the lowest standard deviation of \(\tilde{\mathbf{d}}\) is selected.
-This process makes the assumption that the largest cluster (but necessarily the majority) of items do not contain DIF.
+This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF.
+In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
 
-OUR SIMULATION RESULTS
+\hypertarget{the-equal-means-multiple-imputation-logit-graph}{%
+\subsubsection{The equal means, multiple imputation logit graph}\label{the-equal-means-multiple-imputation-logit-graph}}
 
-\begin{tabular}{l|r|r|r|r|l}
-\hline
-item & a\_ref\_easy & b\_foc\_easy & difference\_in\_easy & cluster & anchor\\
-\hline
-item1 & -0.5638213 & -1.5520228 & 0.9882015 & 1 & TRUE\\
-\hline
-item2 & -0.2618317 & -1.2619495 & 1.0001178 & 1 & TRUE\\
-\hline
-item3 & 1.5747215 & 0.5456186 & 1.0291029 & 1 & TRUE\\
-\hline
-item4 & 0.0521493 & -0.9183001 & 0.9704494 & 1 & TRUE\\
-\hline
-item5 & 0.1335217 & -0.8599758 & 0.9934975 & 1 & TRUE\\
-\hline
-item6 & 1.6971526 & 0.1926118 & 1.5045408 & 2 & FALSE\\
-\hline
-item7 & 0.4629179 & -1.0572527 & 1.5201706 & 2 & FALSE\\
-\hline
-item8 & -1.2953061 & -2.8759098 & 1.5806036 & 2 & FALSE\\
-\hline
-\end{tabular}
-
-\hypertarget{analyst-chooses}{%
-\subsubsection{Analyst chooses}\label{analyst-chooses}}
-
-DIF methods are generally designed to automatically detect DIF without any human judgement.
+AI methods are generally designed to automatically detect DIF without any human judgement.
 On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
 
-We propose a technique, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
+We propose a new method, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
 Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc}\) to \(0\).
 Again, the result is that all differences in performance manifest in \(\tilde{\mathbf{d}}\).
-To get a sense of variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of \(\tilde{d_j}\) for every item.
-These are the distributions displayed in the EM-MILG.
+To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
+Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of \(\tilde{d_j}\) for each item.
+These are the distributions displayed in a EM-MILG.
 The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
@@
 Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data.
-It cannot be stated strongly enough that the EM-MILG contains all possible information about the difference in group performance.
+We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
 The challenge, then, is to select the anchor items.
-The analyst might assume that --- because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.
+The analyst might assume that --- because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.
 
@@
 
-\caption{The equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
+\caption{A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
 \end{figure}
@@
 The same process of using multiple imputations to estimate the distribution of \(d_j\) can be used with the new model.
-Because the equal means assumption is not made, we refer to the resulting graph as a multiple imputations logit graph (MILG).
-In our simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
+Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
+As expected, in our one-run simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
 
@@
 
-\caption{The multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed to 0.}\label{fig:milg}
+\caption{A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed by setting $d_j = 0$.}\label{fig:milg}
 \end{figure}
 
-One of our key concerns with typical DIF methods is that they can lull the analyst into a false sense of security.
+One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security.
 Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
-The EM-MILG combats this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even when other DIF methods are used, the analyst can use the EM-MILG as a first step in order to give them a sense of their item response data.
+A EM-MILG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
 And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
+In particular, we could visualize DIF in the non-anchor items after using AOAA or CLUSTER to identify the model.
 
@@
 
-The previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
+The previously discussed AI methods select a set of anchor items, whether it is an algorithm or the analyst that makes that selection.
 The anchor items are used to estimate \(\hat \mu^\text{foc}\).
@@
 Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, all items can be tested for DIF.
-The question then becomes the following: How do we select the anchor point?
+The question then becomes the following: How is the anchor point selected?
 
@@
 Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (\protect\hyperlink{ref-gini1912variabilita}{1912}) to select the anchor point.
+(\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (Gini \protect\hyperlink{ref-gini1912variabilita}{1912}) to select the anchor point.
 The Gini index is typically used to measure the inequality of wealth distribution in a country.
-For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution Chitiga, Sekyere, and Tsoanamatsie (\protect\hyperlink{ref-chitiga2015income}{2015}).
+For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution (Chitiga, Sekyere, and Tsoanamatsie \protect\hyperlink{ref-chitiga2015income}{2015}).
 In general, the Gini index \enquote{takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little} (Strobl et al.
@@
 
-The anchor point, \(\mu^{\star\text{foc}}\), is selected by maximizing the Gini index.
+\(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (which is why we use the abbreviation MGI).
 The intuition and assumption is that some items might have bias but the majority do not.
-Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\),
+Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
 \[
@@
 
-For the simulated data, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible values for \(\mu^\text{foc}\). The anchor point is \(\mu^\text{foc} = -0.99\).
+For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values. The result of MGI is \(\mu^{\star\text{foc}} = -0.99\).
 
@@
 
-\caption{Selecting the anchor point by maximizing the Gini index.}\label{fig:ginipath}
+\caption{Maximizing the Gini index (MGI) to select the anchor point.}\label{fig:ginipath}
 \end{figure}
 
-We then refit the model with the identifying assumption of \(\mu^{\star\text{foc}} = -0.99\). We test each item for DIF and display the results in a MILG as is shown in Figure \ref{fig:ginimilg}.
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.00544
-\end{verbatim}
+The model is then refit with the model identifying assumption of \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\), and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
 
@@
 
-\caption{Selecting the anchor point by maximizing the Gini index.}\label{fig:ginimilg}
+\caption{The MILG with $\mu^{\text{foc}}$ set to $\mu^{\star\text{foc}} = -0.99$}\label{fig:ginimilg}
 \end{figure}
@@
 
-Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group:
+Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
 \[
@@
 \]
-Raju's area method, along with AOAA, have been cited as the two most commonly used IRT-based DIF detection methods (Magis et al.
+Raju's area method, along with AOAA, has been cited as the two most commonly used IRT-based DIF detection methods (Magis et al.
 \protect\hyperlink{ref-magis2011generalized}{2011}).
-We see two weakness of Raju's area method: First, the item characteric curves still need to be linked by anchor items, which the method does not help to identify.
-Second, the area is unweighted so all values of \(\theta\) matter equally, despite some being much more realistic than others.
+We see two weakness of Raju's area method: First, the item characteristic curves still need to be linked by anchor items or an anchor point, which the method does not help to identify.
+Second, the area is unweighted, so all values of \(\theta\) matter equally, despite some being much more realistic than others.
 
-Inspired by Raju's area method, we propose a new method, which we call \enquote{minimizing the area between curves} (MABC).
+To build off of Raju's area method, we propose a new method, which we call \enquote{minimizing the area between curves} (MABC).
 To understand MABC, imagine multigroup item response data where \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\) so that there is no DIF.
@@
 Alternatively, we could assume that the focal group has \(\mu^{\star\text{foc}} = 3\).
-The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability but every item contains DIF against them.
+The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability, but every item contains DIF against them.
 Both of these models have the same likelihood, so how should one choose which model to believe?
-MABC chooses the model with the least amount of total DIF, as measured by the total area between the item characteric curves.
-This is another, perhaps clearer, way of assuming that the majority of items do not contain DIF.
+MABC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves.
+As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences as opposed to DIF.
 
-Denote a function that estimates \(\hat b_j^\text{foc}\) by fitting a unidimensional Rasch model where the value of \(\mu^\text{foc}\) is fixed as \(m_j(\mu^\text{foc})\).
+Denote a function that takes \(\mu^\text{foc}\) as input and estimates \(\hat b_j^\text{foc}\) by fitting a unidimensional Rasch model as \(m_j(\mu^\text{foc})\).
 The amount of DIF on each item is calculated as
@@
 \]
-MABC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}) who uses the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
-The selection of \(g(\theta)\) is important and determines where on the ability spectrum to consider DIF.
-(\protect\hyperlink{ref-chalmers2016might}{2016}) does not discuss how to choose \(g(\theta)\) and in practice uses \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which seems like an odd choice.
+MABC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}), who use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
+The selection of \(g(\theta)\) results in the relative weighting of \(\theta\) values.
+Chalmers, Counsell, and Flora do not discuss how to choose \(g(\theta)\) and in their empirical examples use \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which seems like an odd choice.
 It might seem intuitive to choose \(g(\theta) \sim N(0, 1)\) because \(\mu^\text{ref} = 0\), but this choice doesn't take into account the ability distribution of the focal group.
 If \(\mu^\text{foc} = 3\), wouldn't we also want to prioritize high \(\theta\) values?
-We set \(g(\theta)\) to be the average of the reference and focal group ability distributions:
+Accordingly, we set \(g(\theta)\) to be the average of the reference and focal group ability distributions:
 \[
@@
 
-For the simulated data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\). The anchor point is \(\mu^\text{foc} = -1\).
+For our one-run simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\).
+The anchor point is found to be \(\mu^{\star\text{foc}} = -1\).
+As with MGI, the model should then be refit using the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\).
 
@@
 
-\caption{Selecting the anchor point by minimizing the total amount of DIF.}\label{fig:mabc}
+\caption{Minimizing the area between curves (MABC) to select the anchor point.}\label{fig:mabc}
 \end{figure}
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+\hypertarget{summary-of-ai-methods}{%
+\subsection{Summary of AI methods}\label{summary-of-ai-methods}}
 
-{[}ADD A TABLE OF METHODS? OR THAT CAN SORT OF BE DONE IN THE SIMULATION RESULTS?{]}
+We have described a variety of AI methods and their corresponding acronymns.
+Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MGI and MABC, select an anchor point.
+Table \ref{table:allmethods} summarizes all of these methods.
 
-\hypertarget{next-steps}{%
-\section{Next steps}\label{next-steps}}
+\begin{table}[H]
+\caption{Summary of assumptionless identification methods}
+\centering
+\begin{tabular}{|p{4cm}|p{6cm}|p{4cm}|}
+\toprule
 
-\hypertarget{remember-my-bounds-idea-i-like-that}{%
-\section{Remember my bounds idea I like that}\label{remember-my-bounds-idea-i-like-that}}
+Method & Description & Literature \\
 
-\hypertarget{simulate-and-then-walk-through-as-the-analyst-would-with-each-method}{%
-\section{Simulate and then walk through as the analyst would with each method}\label{simulate-and-then-walk-through-as-the-analyst-would-with-each-method}}
+\midrule
 
-\hypertarget{mike-frank-example}{%
-\section{Mike Frank example}\label{mike-frank-example}}
+all-others-as-anchors (AOAA) & See Table \ref{table:themethods} for a summary of all three AOAA methods. &  \\\hline
 
-We know that this test contains DIF.
-Which items depends on ability difference which we can be sure of.
-We can, however, quantify the amount of DIF on this test.
-We can take the SD or we can assume between and calculate chalmers statistic for each.
-We might not know for sure the ability difference but we can quantify the potential for DIF.
-For each we calculate the SD across item parameters.
+equal means clustering (EMC) & Cluster items based on differences in performance across groups and choose one of the clusters to be the anchor cluster.
+& Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} \\\hline
 
-Something with DTF?
+equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
+Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
 
-Do a simulation where the second dimension varies - a more real testing ground
+multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
 
-Example with real data perhaps the Mike Frank data
+maximizing Gini index (MGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
 
-Summer learning loss as applied DIF?
+minimizing the area between curves (MABC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
+
+\bottomrule
+\end{tabular}
+\label{table:allmethods}
+\end{table}
+
+\hypertarget{simulation-study}{%
+\section{Simulation study}\label{simulation-study}}
+
+To compare each of the methods in table TODO, we conducted a simulated study.
+Our goal was to make the data generating process as realistic to the scenario described by Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) in which some items on a math test also depend on a student's verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
+As described in \protect\hyperlink{intro}{the introduction}, nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
+This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
+One exception is Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}) who draws each student's target and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
+
+In our simulation study, it was critical that student ability was drawn in a realistic way similar to Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}).
+However, we don't believe that a compensatory model is realistic in describing the a math test where some items depend on verbal ability.
+For example, it's hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability.
+Accordingly, we generate item responses using a simplified version of Sympson's (\protect\hyperlink{ref-sympson1978model}{1978}) noncompensatory item response model in which
+\[
+\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{2j}) = \sigma(\theta_i) \cdot \sigma(a_{2j}\eta_i)
+\]
+where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(\sigma(a_{2j})\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
+
+\hypertarget{drawing-parameters}{%
+\subsection{Drawing parameters}\label{drawing-parameters}}
+
+In each run, we simulate 10,000 students with half coming from each of the reference and focal groups.
+For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean {[}\(\mu_\theta^\text{ref} = 0\), \(\mu_\eta^\text{ref} = 0\){]} and covariance matrix \(\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}\).
+Abilities for students from the focal group are drawn using the same covariance matrix, but with means {[}\(\mu_\theta^\text{foc} = -0.5\), \(\mu_\eta^\text{foc} = -1\){]}.
+
+The test always has 12 items, but we vary the number of items with DIF from two to six.
+For items without DIF, \(a_{2j} = \infty\) so that the model reduces to \(\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)\).
+For items with DIF, \(a_{2j}\) is calculated based on Ackerman's (\protect\hyperlink{ref-ackerman1994using}{1994}) angle equation as described in Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}):
+\[
+\angle_j = \arccos \dfrac{a_{1j}^2}{a_{1j}^2 + a_{2j}^2}.
+\]
+The angle for an item measure the relative loading of the item on the two dimensions.
+For example, an angle of 45\(^\circ\) indicates that the item loads equally on the target and nuisance abilities.
+Our simple compensatory model has \(a_{1j} = 1\) for all items so the angle equation reduces to
+\[
+\angle_j = \arccos \dfrac{1}{1 + a_{2j}^2}.
+\]
+We are interested in specifying the angle of an item, so the relevant equation becomes
+\[
+a_{2j} = \sqrt{\dfrac{1 - \cos(\angle_j)^2}{\cos(\angle_j)^2}}.
+\]
+For DIF items, we set \(a_{2j}\) based on angles with equal intervals between 20\(^\circ\) and 60\(^\circ\).
+For example, for a test with three DIF items the angles are 20\(^\circ\), 40\(^\circ\), and 60\(^\circ\).
+
+\hypertarget{visualizing-a-run}{%
+\subsection{Visualizing a run}\label{visualizing-a-run}}
+
+Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between \(\theta_i\) and \(Pr(y_{ij} = 1)\) with \(\eta_i\) set to the group mean for a test with six items with DIF.
+The items are ordered by the amount of DIF such that \(\angle_{j = 7} = 20^\circ\) up to \(\angle_{j = 12} = 60^\circ\).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/difmap-1}
+
+}
+
+\caption{For a 12 item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.}\label{fig:difmap}
+\end{figure}
+
+Figure \ref{fig:simemmlg} shows the EM-MILG --- generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the \protect\hyperlink{anchoritems}{the anchor items section} --- for one run using the same item parameters that generated Figure \ref{fig:difmap}.
+As expected, \(\tilde{d_j}\) is about \(\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5\) for the first six items which are DIF free.
+For the last six items, \(\tilde{d_j}\) increases as \(\angle_j\) increases.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/simemmlg-1}
+
+}
+
+\caption{For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.}\label{fig:simemmlg}
+\end{figure}
+
+\hypertarget{outcomes}{%
+\subsection{Outcomes}\label{outcomes}}
+
+\hypertarget{achievement-gap-residual}{%
+\subsubsection{Achievement gap residual}\label{achievement-gap-residual}}
+
+An effective method should accurately estimate the group-level target ability gap, which we refer to as the achievement gap.
+The data-generating value of \(\mu_\theta^\text{foc}\) is 0.5, but each run will include sampling variability.
+To get at the heart of how well a method is doing, we calculate the achievement gap residual as the method's estimated achievement gap, \(\hat\mu_\theta^\text{foc}\), minus the achievement gap estimated when using only the DIF-free items as anchors.
+In summary, this outcome measures a method's ability to disentangle differences in target ability from nuisance ability at the group level.
+
+\hypertarget{individual-abilities}{%
+\subsubsection{Individual abilities}\label{individual-abilities}}
+
+Assessments are frequently used to make decisions about individual student abilities.
+We measure a method's ability to do so by calculating the rank correlation between the vector estimated abilities and true abilities.
+This is done both for all students combined and just for students in the focal group.
+We use rank correlation as opposed to the more typically used RMSE as a result of our agreement with Lord's (\protect\hyperlink{ref-lord1986maximum}{1986}) argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
+\emph{(Note that this isn't included in this current draft.)}
+
+\hypertarget{anchor-items.}{%
+\subsubsection{Anchor items.}\label{anchor-items.}}
+
+For the methods that choose a set of anchor items, we look directly at which anchor items are selected.
+An effective method should use most of the non-DIF items as anchors (the anchor hit rate) while avoiding including items with DIF in the anchor set (the false anchor rate).
+
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
+
+In total, we executed 100 runs for each of two, three, four, five, and six DIF items.
+Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual.
+As expected, artificial DIF causes AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
+CLUSTER performs better than AOAA and AOAA-AS but worse than the other methods.
+GINI and MINBC perform similarly well with GINI having a higher variance, while MINBC has more bias, especially for tests with a greater number of DIF items.
+This bias results from MINBC considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.
+AOAA-OAT is the clear winner.
+It performs nearly perfectly for two, three, or four DIF items, and even on its worst run, underestimates \(\mu_\theta^\text{foc}\) by 0.05 standard deviations.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/achievegap-1}
+
+}
+
+\caption{Target ability achievement gap estimates for each method.}\label{fig:achievegap}
+\end{figure}
+
+Figure \ref{fig:anchorfalse} shows the mean false anchor rates for each method and number of items with DIF.
+For example, when there are two items with DIF on the test, those two items have \(\angle_{11} = 20^\circ\) and \(\angle_{12} = 60^\circ\).
+AOAA-OAT never includes item 12 in the anchor set, but incorrectly includes item 11 in 60 out of the 100 runs.
+Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method is
+\[
+\dfrac{\text{Total number of DIF items in the anchor set}}{\text{Number of DIF items on each test} \cdot \text{Number of runs}} = \dfrac{60}{2 \cdot 100} = 30\%.
+\]
+The fact that the item with 20\(^\circ\) of DIF is most commonly incorrectly included in the anchor set is what drives the counterintuitive result that the mean false anchor rate decreases with more DIF items.
+
+Similarly, Figure \ref{fig:anchorhit} shows the mean anchor hit rates.
+Remarkably, AOAA-OAT includes an average of over \(90\%\) of DIF-free items in the anchor set regardless of the number of DIF items on the test.
+Interestingly, CLUSTER has a better anchor hit rate on tests with more DIF items.
+This result appears to be driven by the clustering algorithm sometimes splitting all of the DIF-free items into two separate clusters, especially when most of the items are DIF-free.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/anchorfalse-1}
+
+}
+
+\caption{Mean false anchor rates across 100 runs for each number of DIF items}\label{fig:anchorfalse}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/anchorhit-1}
+
+}
+
+\caption{Mean anchor hit rates across 100 runs for each number of DIF items}\label{fig:anchorhit}
+\end{figure}
+
+\hypertarget{discussion-and-summary}{%
+\section{Discussion and summary}\label{discussion-and-summary}}
+
+Measurement instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
+Inspired by Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}), we have resurrected the DIF-as-student-property view and shown its connection to the typical way in which AI methods and DIF simulation studies have been conceptualized.
+
+More importantly, we have reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe is more realistic than the typical DIF simulation study.
+In particular, we simulated student ability as drawn from a two-dimensional distribution representing a student's target and nuisance ability, and then generated data using a noncompensatory item response model.
+Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
+On the other hand, selecting anchor items using either AOAA-OAT or clustering and selecting an anchor point using either GINI or MINBC all performed reasonably.
+
+However, AOAA-OAT was the clear winner in terms of performance and we recommend its use whenever possible.
+One reason AOAA-OAT might not always be possible is that it can be computationally expensive.
+For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows.
+To increase AOAA-OAT's use, we recommend its implementation (perhaps as the default) in popular IRT software such as the mirt R package.
+
+In addition to exploring and testing algorithmic AI methods, we introduced a method, the EM-MILG, that an analyst can use to visualize the amount of potential DIF in their data.
+This method can be used either to build their intuition or as a way in which they can select anchor items by hand.
+The EM-MILG's sibling method, the MILG, is, we believe, the best way to visualize the results of a DIF analysis after anchor items have been selected.
+
+Future work should test these methods' performance under a greater variety of data-generating conditions.
+For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
+Furthermore, our work focused on the Rasch model, and it will be of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.
 
 \clearpage
-
-\hypertarget{notes}{%
-\section{Notes}\label{notes}}
-
-\begin{itemize}
-\tightlist
-\item
-  adjust to all be easiness including multidimensional talk
-\end{itemize}
 
@@
 
+\leavevmode\hypertarget{ref-ackerman1994using}{}%
+---------. 1994. ``Using Multidimensional Item Response Theory to Understand What Items and Tests Are Measuring.'' \emph{Applied Measurement in Education} 7 (4): 255--78.
+
 \leavevmode\hypertarget{ref-andrich2012real}{}%
@@
 
+\leavevmode\hypertarget{ref-chalmers2012mirt}{}%
+Chalmers, R Philip. 2012. ``Mirt: A Multidimensional Item Response Theory Package for the R Environment.'' \emph{Journal of Statistical Software} 48 (6): 1--29.
+
 \leavevmode\hypertarget{ref-chalmers2018numerical}{}%
-Chalmers, R Philip. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
+---------. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
 
@@
 
+\leavevmode\hypertarget{ref-demars2016partially}{}%
+DeMars, Christine E. 2016. ``Partially Compensatory Multidimensional Item Response Theory Models: Two Alternate Model Forms.'' \emph{Educational and Psychological Measurement} 76 (2): 231--57.
+
 \leavevmode\hypertarget{ref-drasgow1987study}{}%
@@
 
+\leavevmode\hypertarget{ref-hagquist2017recent}{}%
+Hagquist, Curt, and David Andrich.
+2017.
+``Recent Advances in Analysis of Differential Item Functioning in Health Research Using the Rasch Model.'' \emph{Health and Quality of Life Outcomes} 15 (1): 181.
+
 \leavevmode\hypertarget{ref-hastie2001estimating}{}%
@@
 Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.
+
+\leavevmode\hypertarget{ref-lord1986maximum}{}%
+Lord, Frederic M. 1986. ``Maximum Likelihood and Bayesian Parameter Estimation in Item Response Theory.'' \emph{Journal of Educational Measurement} 23 (2): 157--62.
 
@@
 \leavevmode\hypertarget{ref-pohl2017cluster}{}%
-Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``CluStEr-baSEd anCHor Item Identification and Selection.'' NEPS Working Paper.
+Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``Cluster-Based Anchor Item Identification and Selection.''
 
@@
 Raju, Nambury S. 1988. ``The Area Between Two Item Characteristic Curves.'' \emph{Psychometrika} 53 (4): 495--502.
+
+\leavevmode\hypertarget{ref-rcore}{}%
+R Core Team. 2019. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
 
@@
 
+\leavevmode\hypertarget{ref-sympson1978model}{}%
+Sympson, James B. 1978. ``A Model for Testing with Multidimensional Items.'' In \emph{Proceedings of the 1977 Computerized Adaptive Testing Conference}. 00014.
+
 \leavevmode\hypertarget{ref-talbot2013taking}{}%
@@
 
+\leavevmode\hypertarget{ref-walker2017using}{}%
+Walker, Cindy M, and Sakine Gocer Sahin.
+2017.
+``Using a Multidimensional Irt Framework to Better Understand Differential Item Functioning (Dif): A Tale of Three Dif Detection Procedures.'' \emph{Educational and Psychological Measurement} 77 (6): 945--70.
+
+\leavevmode\hypertarget{ref-tidy}{}%
+Wickham, Hadley. 2017. \emph{Tidyverse: Easily Install and Load the 'Tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}.
+
 \leavevmode\hypertarget{ref-woods2009empirical}{}%

### 2020-03-14 paper/paper_1_intro.Rmd
printed second verison!

@@
-<!-- need to specify which multidimensional model im talking about -->
-
-Inspired by @camilli1992conceptual, we think differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
-From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item.
+Inspired by @camilli1992conceptual, we think of differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
+From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item.
 For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
@@
 They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study.
-They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits.
+They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits.
 They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
@@
 The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group.
-We now need to use a 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise.
+The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for all students in the reference group and ${\eta_i}^{\text{ref}} = -1$ for all students in the focal group.
+We now need to use a compensatory 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional 2PL model, the probability student $i$ responding correctly to item $j$ is
+According to the two-dimensional compensatory 2PL model, the probability student $i$ responding correctly to item $j$ is
 $$
-P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
+\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
 $$
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability for students in the same group.
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability across students in the same group.
 

### 2020-03-14 paper/paper_2a_anchor_items.Rmd
printed second verison!

@@
 Psychometricians have long been in search of the perfect DIF detection method.
-In this section, we summarize existing DIF methods and propose extensions of those methods.
-We take as given that the IRT-based likelihood ratio test is the best way to test for DIF, which has been well demonstrated [@meade2012solving].
+It has become well known that the IRT-based likelihood ratio test (LRT) is the best way to test for DIF, which has been well demonstrated [@meade2012solving].
 As a result, we avoid methods like the Mantel-Haenszel procedure [@holland1986differential], which muddies the waters by moving away from the IRT framework, and has been shown to perform no better — and in some cases worse — than IRT-based methods [@swaminathan1990detecting].
@@
 The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items contain bias — can be made.
+That is, we're interested in assumptionless identification methods (hereafter referred to as "AI methods") so that the LRTs can be used to test items for DIF.
 For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
@@
 We recognize that others have done the same [e.g.
-@kopf2015framework] and that we risk contributing to the proliferation of names.
+@kopf2015framework], and that we risk contributing to the proliferation of names.
 
+In this section, we summarize existing AI methods and propose extensions of those methods.
 We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
@@
 Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{ref}} = -1$.
-The slope on target ability is set to $a_{1j} = 1$ for all items, and the slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise.
+The slope on target ability is set to $a_{1j} = 1$ for all items.
+The slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise.
 From the DIF-as-item-property view, there is no nuisance ability.
@@
 We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
-The fundamental challenge for DIF methods is to disentangle estimating $\hat\mu^\text{foc}$ from estimating $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$.
-The most common approach is to identify a group of anchor items that are assumed to be DIF-free.
-These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$.
+The fundamental challenge is to use an AI method to identify the model, which allows for the disentangling $\hat\mu^\text{foc}$ from estimating $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$.
+The most common AI method is to select a group of anchor items that are assumed to be DIF-free.
+These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.
 
-## Anchor items
+Computing is done in R [@rcore], model fitting in the R package mirt [@chalmers2012mirt], and data wrangling/visualization in the set of R packages known as the tidyverse [@tidy].
+Code is available at TO DO.
+
+## Anchor items {#anchoritems}
 
@@
 
-@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA).
+@meade2012solving compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method.
 AOAA tests each item for DIF one at a time using all of the other items as anchors.
 For example, when testing the first item for DIF, all of the other items are used as anchors.
-This is done using a likelihood ratio test that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection].
+This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection].
 Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
-The items for which the flexible model outperforms the baseline model (typically based on a chi-squared test) are identified as having DIF and the rest of the items become anchor items.
+The items for which the flexible model outperforms the baseline model (typically based on a $\chi^2$ test) are identified as having DIF, and the rest of the items become anchor items.
+Note that AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
 
 @edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can we be sure its those 10 items and not the other 11 items with DIF?
-We can't be.
+How can they be sure that its those 10 items and not the other 11 items with DIF?
+They cannot be.
 Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group and others are biased against the reference group.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method [@strobl2018anchor].
+More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
+Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor].
 In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
@@
 We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF being removed from the anchor set.
+Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
+Note that AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.
 
-AOAA-AS might seem like an improvement but it doesn't solve a fundamental problem of AOAA: What do we do when all items test positive for DIF?
+AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF?
 With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
-In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF and there are no anchor items.
+In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items.
 @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
 @woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items display DIF themselves.
+It's unclear if one should proceed if those four items display DIF.
 
@@
 We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
-AOAA-OAT is inspired by @hagquist2017recent who, in general, assert that "items showing DIF initially
-should not be resolved simultaneously but sequentially" [p.
+AOAA-OAT is inspired by @hagquist2017recent, who, in general, assert that "items showing DIF initially should not be resolved simultaneously but sequentially" [p.
 6].
@@
 AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
-As a result, AOAA-OAT is less likely to arrive at the wrong conclusion because of artifical DIF.
+As a result, we believe that AOAA-OAT is less likely to be "fooled" by artifical DIF.
+Note that AOAA-OAT is not currently implemented in the R package mirt.
 
@@
 Table \ref{table:themethods} summarizes the three all-others-as-anchors methods.
-It's useful to remember that AOAA is not an iterative procedure, while the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
+It's useful to remember that AOAA is not an iterative procedure.
+The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
+In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
+On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.
 
@@
 \begin{table}[H]
-\caption{EDIT THIS}
+\caption{Summary of the three all-others-as-anchors assumptionless identification methods}
 \centering
@@
 
-[ADD IN SIMULATION RESULTS]
-
-### Clustering items
-
 ```{r}
@@
 
-@bechger2015statistical proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider relative item parameters.
+### Clustering items
+
+@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
 For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$.
@@
 The general idea of identifying clusters of items is intriguing.
-However, their approach is needlessly complicated and they did not describe a process for moving from $\Delta \mathbf{R}$ to a set of anchor items.
+However, their approach is needlessly complicated, and they did not describe a process for moving from $\Delta \mathbf{R}$ to an anchor cluster.
 @pohl2017cluster extended their work by proposing one such process.
@@
 Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$ of differences in item easiness, $\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}$, where the model is identified by setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, thus the name "equal means clustering").
-This results in all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$.
+As a result, all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$.
 The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.
@@
 If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected.
-This process makes the assumption that the largest cluster (but necessarily the majority) of items do not contain DIF.
+This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF.
+In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
 
-OUR SIMULATION RESULTS
+### The equal means, multiple imputation logit graph
 
-```{r}
-out_3_biased_items$cluster_status[[1]] %>% 
-    select(item, a_ref_easy, b_foc_easy, difference_in_easy, cluster) %>% 
-    mutate(anchor = ifelse(cluster == 1, TRUE, FALSE)) %>% 
-    kable()
-```
-
-### Analyst chooses
-
-DIF methods are generally designed to automatically detect DIF without any human judgement.
+AI methods are generally designed to automatically detect DIF without any human judgement.
 On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
 
-We propose a technique, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
+We propose a new method, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
 Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$.
 Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$.
-To get a sense of variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical].
-Then, multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of $\tilde{d_j}$ for every item.
-These are the distributions displayed in the EM-MILG.
+To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical].
+Then, multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of $\tilde{d_j}$ for each item.
+These are the distributions displayed in a EM-MILG.
 The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
@@
 Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data.
-It cannot be stated strongly enough that the EM-MILG contains all possible information about the difference in group performance.
+We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
 The challenge, then, is to select the anchor items.
-The analyst might assume that — because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.
+The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.
 
-```{r emmilg, fig.cap = 'The equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.', out.width="70%", warning = FALSE, message = FALSE}
+```{r emmilg, fig.cap = 'A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.', out.width="70%", warning = FALSE, message = FALSE}
 out_3_biased_items$intuitive_mod[[1]] %>%
@@
 The same process of using multiple imputations to estimate the distribution of $d_j$ can be used with the new model.
-Because the equal means assumption is not made, we refer to the resulting graph as a multiple imputations logit graph (MILG).
-In our simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
+Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
+As expected, in our one-run simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
 
-```{r milg, fig.cap = 'The multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group.
-Anchor items are fixed to 0.', out.width="70%", warning = FALSE, message = FALSE}
+```{r milg, fig.cap = 'A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group.
+Anchor items are fixed by setting $d_j = 0$.', out.width="70%", warning = FALSE, message = FALSE}
 # CAREFUL - this code assumes that the analyst would choose the same anchor items as AOAA-OAT found
@@
 
-One of our key concerns with typical DIF methods is that they can lull the analyst into a false sense of security.
+One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security.
 Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
-The EM-MILG combats this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even when other DIF methods are used, the analyst can use the EM-MILG as a first step in order to give them a sense of their item response data.
+A EM-MILG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
 And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
+In particular, we could visualize DIF in the non-anchor items after using AOAA or CLUSTER to identify the model.

### 2020-03-14 paper/paper_2b_anchor_points.Rmd
printed second verison!

@@
-The previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
+The previously discussed AI methods select a set of anchor items, whether it is an algorithm or the analyst that makes that selection.
 The anchor items are used to estimate $\hat \mu^\text{foc}$.
@@
 
-@strobl2018anchor suggest using the Gini index -@gini1912variabilita to select the anchor point.
+@strobl2018anchor suggest using the Gini index [@gini1912variabilita] to select the anchor point.
 The Gini index is typically used to measure the inequality of wealth distribution in a country.
-For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution @chitiga2015income.
+For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution [@chitiga2015income].
 In general, the Gini index "takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little" [@strobl2018anchor, p.
@@
 
-The anchor point, $\mu^{\star\text{foc}}$, is selected by maximizing the Gini index.
+$\mu^{\star\text{foc}}$ is selected by maximizing the Gini index (which is why we use the abbreviation MGI).
 The intuition and assumption is that some items might have bias but the majority do not.
-Denoting a function that calculates the Gini index from a vector of non-negative elements as $G(\mathbf{x})$,
+Denoting a function that calculates the Gini index from a vector of non-negative elements as  $G(\mathbf{x})$, MGI sets <!-- its weird to arg max over a parameter we already have a definition for - maybe switch to t or something -->
 $$
@@
 
-For the simulated data, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible values for $\mu^\text{foc}$.
-The anchor point is $\mu^\text{foc} = `r round(gini_results$gini_anchor_points, 2)`$.
+For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible $\mu^\text{foc}$ values.
+The result of MGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$.
 
-```{r ginipath, fig.cap = 'Selecting the anchor point by maximizing the Gini index.', out.width="70%", warning = FALSE, message = FALSE}
+```{r ginipath, fig.cap = 'Maximizing the Gini index (MGI) to select the anchor point.', out.width="70%", warning = FALSE, message = FALSE}
 gini_results$grid %>% 
@@
 
-We then refit the model with the identifying assumption of $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$.
-We test each item for DIF and display the results in a MILG as is shown in Figure \ref{fig:ginimilg}.
+The model is then refit with the model identifying assumption of $\mu^{\text{foc}} = \mu^{\star\text{foc}}$, and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
 
-```{r ginimilg, fig.cap = 'Selecting the anchor point by maximizing the Gini index.', out.width="70%", warning = FALSE, message = FALSE}
-
+```{r ginimilg, fig.cap = "The MILG with $\\mu^{\\text{foc}}$ set to $\\mu^{\\star\\text{foc}} = -0.99$", out.width="70%", warning = FALSE, message = FALSE}
+# notice hard coding in this caption
 data <- out_3_biased_items$sim[[1]]$data
@@
 
-mod <- gini_results_to_final_mod(gini_results$gini_anchor_points, data, groups)
+mod <- anchptfocmean_to_final_mod(gini_results$gini_anchor_points, data, groups)
 
@@
 
-Raju's area method [@raju1988area] measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group:
+Raju's area method [@raju1988area] measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
 $$
@@
 $$
-Raju's area method, along with AOAA, have been cited as the two most commonly used IRT-based DIF detection methods [@magis2011generalized].
-We see two weakness of Raju's area method: First, the item characteric curves still need to be linked by anchor items, which the method does not help to identify.
-Second, the area is unweighted so all values of $\theta$ matter equally, despite some being much more realistic than others.
+Raju's area method, along with AOAA, has been cited as the two most commonly used IRT-based DIF detection methods [@magis2011generalized].
+We see two weakness of Raju's area method: First, the item characteristic curves still need to be linked by anchor items or an anchor point, which the method does not help to identify.
+Second, the area is unweighted, so all values of $\theta$ matter equally, despite some being much more realistic than others.
 
-Inspired by Raju's area method, we propose a new method, which we call "minimizing the area between curves" (MABC).
+To build off of Raju's area method, we propose a new method, which we call "minimizing the area between curves" (MABC).
 To understand MABC, imagine multigroup item response data where $\mu^{\text{foc}} = \mu^{\text{ref}}$ and $d_j = 0 \forall j$ so that there is no DIF.
@@
 Alternatively, we could assume that the focal group has $\mu^{\star\text{foc}} = 3$.
-The model would then compensate by finding $\hat d_j \approx -3 \forall j$, and we would incorrectly conclude that the focal group is high ability but every item contains DIF against them.
+The model would then compensate by finding $\hat d_j \approx -3 \forall j$, and we would incorrectly conclude that the focal group is high ability, but every item contains DIF against them.
 Both of these models have the same likelihood, so how should one choose which model to believe?
-MABC chooses the model with the least amount of total DIF, as measured by the total area between the item characteric curves.
-This is another, perhaps clearer, way of assuming that the majority of items do not contain DIF.
+MABC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves.
+As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences as opposed to DIF.
 
-Denote a function that estimates $\hat b_j^\text{foc}$ by fitting a unidimensional Rasch model where the value of $\mu^\text{foc}$ is fixed as $m_j(\mu^\text{foc})$.
+<!-- again might want to choose a different variable for provisional values -->
+Denote a function that takes $\mu^\text{foc}$ as input and estimates $\hat b_j^\text{foc}$ by fitting a unidimensional Rasch model as $m_j(\mu^\text{foc})$.
 The amount of DIF on each item is calculated as
@@
 $$
-MABC is inspired in part by @chalmers2016might who use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
-The selection of $g(\theta)$ is important and determines where on the ability spectrum to consider DIF.
--@chalmers2016might does not discuss how to choose $g(\theta)$ and in practice uses $g(\theta)$ uniform for $-6 \le \theta \le 6$, which seems like an odd choice.
+MABC is inspired in part by @chalmers2016might, who use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
+The selection of $g(\theta)$ results in the relative weighting of $\theta$ values.
+Chalmers, Counsell, and Flora do not discuss how to choose $g(\theta)$ and in their empirical examples use $g(\theta)$ uniform for $-6 \le \theta \le 6$, which seems like an odd choice.
 It might seem intuitive to choose $g(\theta) \sim N(0, 1)$ because $\mu^\text{ref} = 0$, but this choice doesn't take into account the ability distribution of the focal group.
 If $\mu^\text{foc} = 3$, wouldn't we also want to prioritize high $\theta$ values?
-We set $g(\theta)$ to be the average of the reference and focal group ability distributions:
+Accordingly, we set $g(\theta)$ to be the average of the reference and focal group ability distributions:
 $$
@@
 $$
+
 ```{r}
@@
 #     out_3_biased_items$intuitive_mod[[1]] %>%
-#     min_between_curves(lo = 0, hi = 2, by = 0.1) # future: 0.0001
+#     min_between_curves(lo = 0, hi = 2, by = 0.001)
 # min_bc %>% write_rds(here("paper", "paper_2_min_bc.rds"))
@@
 
-For the simulated data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$.
-The anchor point is $\mu^\text{foc} = `r round(min_bc$between_curves_anchor_points, 2)`$.
+For our one-run simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$.
+The anchor point is found to be $\mu^{\star\text{foc}} = `r round(min_bc$between_curves_anchor_points, 2)`$.
+As with MGI, the model should then be refit using the identifying assumption that $\mu^{\text{foc}} = \mu^{\star\text{foc}}$.
 
-```{r mabc, fig.cap = 'Selecting the anchor point by minimizing the total amount of DIF.', out.width="70%"}
+```{r mabc, fig.cap = 'Minimizing the area between curves (MABC) to select the anchor point.', out.width="70%"}
 min_bc$grid %>% 
@@
 
-## Summary
+## Summary of AI methods
 
-[ADD A TABLE OF METHODS? OR THAT CAN SORT OF BE DONE IN THE SIMULATION RESULTS?]
+We have described a variety of AI methods and their corresponding acronymns.
+Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MGI and MABC, select an anchor point.
+Table \ref{table:allmethods} summarizes all of these methods.
+
+\begin{table}[H]
+\caption{Summary of assumptionless identification methods}
+\centering
+\begin{tabular}{|p{4cm}|p{6cm}|p{4cm}|}
+\toprule
+
+Method & Description & Literature \\
+
+\midrule
+
+all-others-as-anchors (AOAA) & See Table \ref{table:themethods} for a summary of all three AOAA methods. &  \\\hline
+
+equal means clustering (EMC) & Cluster items based on differences in performance across groups and choose one of the clusters to be the anchor cluster.
+& Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} \\\hline
+
+equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
+Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
+
+multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
+
+maximizing Gini index (MGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
+
+minimizing the area between curves (MABC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
+
+\bottomrule
+\end{tabular}
+\label{table:allmethods}
+\end{table}

### 2020-03-14 paper/paper_3_sim.Rmd
printed second verison!

@@
 To compare each of the methods in table TODO, we conducted a simulated study.
-Our goal was to make the data generating process as realistic to the scenario described by @ackerman1992didactic in which some items on a math test also depend on a student’s verbal ability (the target ability is math ability and the nuisance ability is verbal ability).
-As described in [the introduction](#intro), nearly all DIF simulation studies simply generate data where the item easiness parameters are different for the reference and focal groups.
+Our goal was to make the data generating process as realistic to the scenario described by @ackerman1992didactic in which some items on a math test also depend on a student’s verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
+As described in [the introduction](#intro), nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
 This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
@@
 For example, it’s hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability.
-Accordingly, we generate item responses using an adapted version of Sympson’s -@sympson1978model noncompensatory item response model in which
+Accordingly, we generate item responses using a simplified version of Sympson’s -@sympson1978model noncompensatory item response model in which
 $$
@@
 For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean [$\mu_\theta^\text{ref} = 0$, $\mu_\eta^\text{ref} = 0$] and covariance matrix $\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}$.
-Abilities for students from the focal group are drawn using the same covariance matrix but with means [$\mu_\theta^\text{foc} = -0.5$, $\mu_\eta^\text{foc} = -1$].
+Abilities for students from the focal group are drawn using the same covariance matrix, but with means [$\mu_\theta^\text{foc} = -0.5$, $\mu_\eta^\text{foc} = -1$].
 
-The test always has 20 items, but we vary the number of items with DIF from two to ten.
+The test always has 12 items, but we vary the number of items with DIF from two to six.
 For items without DIF, $a_{2j} = \infty$ so that the model reduces to $\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)$.
@@
 The angle for an item measure the relative loading of the item on the two dimensions.
-For example, an angle of 45$^\circ$ indicates that the item loads equally on the target and nuisance ability dimensions.
-We are using a simple compensatory model in which $a_{1j} = 1$ for all items so the angle equation reduces to
+For example, an angle of 45$^\circ$ indicates that the item loads equally on the target and nuisance abilities.
+Our simple compensatory model has $a_{1j} = 1$ for all items so the angle equation reduces to
 $$
@@
 $$
-For DIF items, we set $a_{2j}$ based on angles with equal intervals between 30$^\circ$ and 60$^\circ$.
-For example, for a test with four DIF items the angles are 30$^\circ$, 40$^\circ$, 50$^\circ$, and 60$^\circ$.
+For DIF items, we set $a_{2j}$ based on angles with equal intervals between 20$^\circ$ and 60$^\circ$.
+For example, for a test with three DIF items the angles are 20$^\circ$, 40$^\circ$, and 60$^\circ$.
 
 ## Visualizing a run
+
 ```{r}
@@
 
-# i have no idea why this is like this...
-out %>% 
-  mutate(bigsim = bigsim %>% map(~ select(., foc_mean))) %>% 
-  unnest(bigsim) %>% 
-  ggplot(aes(x = n_dif_items, y = foc_mean)) +
-  geom_point()
+# this use to seem off and is a good check but seems fine now
+# out %>% 
+#   mutate(bigsim = bigsim %>% map(~ select(., foc_mean))) %>% 
+#   unnest(bigsim) %>% 
+#   ggplot(aes(x = n_dif_items, y = foc_mean)) +
+#   geom_point()
 ```
-Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean.
-The items are ordered by the amount of DIF such that $\angle_{j = 11} = 30^\circ$ up to $\angle_{j = 20} = 60^\circ$.
-```{r difmap, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
-out$bigsim[[2]]$pars[[1]] %>% 
+
+Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean for a test with six items with DIF.
+The items are ordered by the amount of DIF such that $\angle_{j = 7} = 20^\circ$ up to $\angle_{j = 12} = 60^\circ$.
+
+```{r difmap, fig.cap = 'For a 12 item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.', out.width="70%", warning = FALSE, message = FALSE}
+out$bigsim[[5]]$pars[[1]] %>% 
   graph_pars_noncompensatory(0, -1) +
@@
 
-Figure \ref{fig:simemmlg} shows the EM-MILG for one run using the same item parameters that generated Figure \ref{fig:difmap}.
-As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first ten items which are DIF free.
-For the last ten items, $\tilde{d_j}$ increases as $\angle_j$ increases.
-Of course, we cannot include the "analyst chooses" method in our simulation, but regardless, the EM-MILG is a useful way to visualize the data.
+Figure \ref{fig:simemmlg} shows the EM-MILG — generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the [the anchor items section](#anchoritems) — for one run using the same item parameters that generated Figure \ref{fig:difmap}.
+As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first six items which are DIF free.
+For the last six items, $\tilde{d_j}$ increases as $\angle_j$ increases.
+
 ```{r simemmlg, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
-out$bigsim[[3]]$intuitive_mod[[1]] %>%
+out$bigsim[[5]]$intuitive_mod[[1]] %>%
     mod_intuitive_to_draws_df() %>%
@@
 
-### Target ability achievement gap
-An effective method, should find $\hat\mu_\theta^\text{foc}$ close to 0.5.
-This outcome measures a method’s ability to disentangle differences in target ability from nuisance ability at the group level.
+### Achievement gap residual
+
+An effective method should accurately estimate the group-level target ability gap, which we refer to as the achievement gap.
+The data-generating value of $\mu_\theta^\text{foc}$ is 0.5, but each run will include sampling variability.
+To get at the heart of how well a method is doing, we calculate the achievement gap residual as the method's estimated achievement gap, $\hat\mu_\theta^\text{foc}$, minus the achievement gap estimated when using only the DIF-free items as anchors.
+In summary, this outcome measures a method’s ability to disentangle differences in target ability from nuisance ability at the group level.
 
 ### Individual abilities
+
 Assessments are frequently used to make decisions about individual student abilities.
@@
 We use rank correlation as opposed to the more typically used RMSE as a result of our agreement with Lord's -@lord1986maximum argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
+*(Note that this isn't included in this current draft.)*
 
 ### Anchor items.
+
 For the methods that choose a set of anchor items, we look directly at which anchor items are selected.
@@
 ## Results
-In total, we executed 100 runs for each number of two to ten DIF items.
-This multiplies to a total of 900 runs.
-Figure \ref{fig:achievegap} shows each method's performance on estimating the target ability achievement gap.
 
-As expected, artificial DIF causes AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
-CLUSTERING typically performs well but has a higher variance because the anchor cluster sometimes contains as few as half of the DIF-free items.
-
-AOAA-OAT, GINI, and MINBC all perform similarly well.
-MINBC has a tendency to slightly overestimate the achievement gap.
-There is an interesting pattern, not previously described in the literature, where AOAA-OAT and GINI of overestimate the achievement gap when there are few items with DIF and underestimate the achivement gap when there are many items with DIF.
-The cause is more easily understood with AOAA-OAT where we can inspect which items are included in the anchor set.
-
-Figure \ref{fig:anchorfalse} shows the mean false anchor counts for each method and number of items with DIF.
-For example, when there are 10 items with DIF on the test, AOAA-AS incorrectly includes an average of 1.8 of those items in the anchor set.
-AOAA-OAT remarkably never includes an item with DIF in the anchor set (i.e., a false anchor rate of 0\%).
-
-Performance, then, is driven by which items are included in the anchor set.
-Accordingly, Figure \ref{fig:anchorhit} shows the number of items without DIF that are included in the anchor set.
-In general, AOAA-OAT usually fails to include a couple of items in the anchor set that are indeed DIF free.
-When there are a few items with DIF, the omitted items are more often items in which the focal group happened to, as a result of sampling variability, perform well relative to the reference group.
-On the other hand, when there are many items with DIF, the omitted items are more often items in which the focal group happened to perform poorly relative to the reference group.
-This selective omission of DIF free anchors is what drives the positive association between number of DIF items and $\hat\mu_\theta^\text{foc}$.
-
-```{r achievegap, fig.cap = 'Each method's estimation of the target ability achievement gap.', out.width="70%", warning = FALSE, message = FALSE}
-
+```{r}
 n_items <- 12
@@
     unnest(achievegap)
+```
 
+In total, we executed 100 runs for each of two, three, four, five, and six DIF items.
+Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual.
+As expected, artificial DIF causes AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
+CLUSTER performs better than AOAA and AOAA-AS but worse than the other methods.
+GINI and MINBC perform similarly well with GINI having a higher variance, while MINBC has more bias, especially for tests with a greater number of DIF items.
+This bias results from MINBC considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.
+AOAA-OAT is the clear winner.
+It performs nearly perfectly for two, three, or four DIF items, and even on its worst run, underestimates $\mu_\theta^\text{foc}$ by 0.05 standard deviations.
+
+```{r achievegap, fig.cap = 'Target ability achievement gap estimates for each method.', out.width="70%", warning = FALSE, message = FALSE}
 what %>% 
+  mutate(
+    method = 
+      factor(
+        method,
+        levels = c("AOAA_OAT_mod", "cluster_mod", "gini_mod", "AOAA_AS_mod", "AOAA_mod", "minbc_mod"),
+        labels = c("AOAA-OAT", "CLUSTER", "GINI", "AOAA-AS", "AOAA", "MINBC")
+      )
+  ) %>% 
   ggplot(aes(x = resid, y = as.factor(n_dif_items))) +
@@
   labs(
-      x = "Number of DIF items",
-      y = latex2exp::TeX("$\\hat{\\mu}_\\theta^foc$")
+      y = "Number of DIF items",
+      x = latex2exp::TeX("Achievement gap residual")
   )
@@
 
-```{r anchorfalse, fig.cap = 'Mean false anchor counts', out.width="70%", warning = FALSE, message = FALSE}
+Figure \ref{fig:anchorfalse} shows the mean false anchor rates for each method and number of items with DIF.
+For example, when there are two items with DIF on the test, those two items have $\angle_{11} = 20^\circ$ and $\angle_{12} = 60^\circ$.
+AOAA-OAT never includes item 12 in the anchor set, but incorrectly includes item 11 in 60 out of the 100 runs.
+Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method is
+$$
+\dfrac{\text{Total number of DIF items in the anchor set}}{\text{Number of DIF items on each test} \cdot \text{Number of runs}} = \dfrac{60}{2 \cdot 100} = 30\%.
+$$
+The fact that the item with 20$^\circ$ of DIF is most commonly incorrectly included in the anchor set is what drives the counterintuitive result that the mean false anchor rate decreases with more DIF items.
+
+Similarly, Figure \ref{fig:anchorhit} shows the mean anchor hit rates.
+Remarkably, AOAA-OAT includes an average of over $90\%$ of DIF-free items in the anchor set regardless of the number of DIF items on the test.
+Interestingly, CLUSTER has a better anchor hit rate on tests with more DIF items.
+This result appears to be driven by the clustering algorithm sometimes splitting all of the DIF-free items into two separate clusters, especially when most of the items are DIF-free.
+
+```{r anchorfalse, fig.cap = 'Mean false anchor rates across 100 runs for each number of DIF items', out.width="70%", warning = FALSE, message = FALSE}
 one_to_anchors <- function(one, n_dif_items){
@@
     summarize(total = sum(total), mean_selected = sum(mean_selected)) %>%
-    gather(var, val, -n_dif_items, -method) %>%
-    mutate(var = factor(var, 
-                        levels = c("total", "mean_selected"),
-                        labels = c("Total possible", "Mean selected to anchor set"))) %>% 
-  filter(var == "Mean selected to anchor set") %>% 
-    ggplot(aes(x = n_dif_items, y = val)) +
+    ungroup() %>% 
+    mutate(rate = mean_selected / total) %>% 
+    mutate(
+      method = 
+        factor(
+          method,
+          levels = c("AOAA_OAT_mod", "cluster_mod", "AOAA_AS_mod", "AOAA_mod"),
+          labels = c("AOAA-OAT", "CLUSTER", "AOAA-AS", "AOAA")
+        )
+    ) %>% 
+    ggplot(aes(x = n_dif_items, y = rate)) +
     geom_point() +
-    ylim(0, 20) +
     facet_wrap(~ method) +
     labs(
-      x = "Number of items with DIF on the test",
-      y = "Mean number of items with DIF in the anchor set",
-      color = ""
-    )
+      x = "Number of DIF items",
+      y = "Mean false anchor rate"
+    ) +
+    scale_y_continuous(labels = scales::percent, limits = c(0, 0.5))
 ```
 
-```{r anchorhit, fig.cap = 'Mean anchor hit counts', out.width="70%", warning = FALSE, message = FALSE}
+```{r anchorhit, fig.cap = 'Mean anchor hit rates across 100 runs for each number of DIF items', out.width="70%", warning = FALSE, message = FALSE}
 temp %>%
     filter(true_anchor == "no dif") %>%
-    gather(var, val, -n_dif_items, -method, -true_anchor) %>%
-    mutate(var = factor(var, 
-                        levels = c("total", "mean_selected"),
-                        labels = c("Total possible", "Mean selected to anchor set"))) %>% 
-    ggplot(aes(x = n_dif_items, y = val, color = var)) +
+    mutate(rate = mean_selected / total) %>% 
+    mutate(
+        method = 
+          factor(
+            method,
+            levels = c("AOAA_OAT_mod", "cluster_mod", "AOAA_AS_mod", "AOAA_mod"),
+            labels = c("AOAA-OAT", "CLUSTER", "AOAA-AS", "AOAA")
+          )
+    ) %>%  
+    ggplot(aes(x = n_dif_items, y = rate)) +
     geom_point() +
-    ylim(0, 20) +
     facet_wrap(~ method) +
     labs(
-      x = "Number of items with DIF on the test",
-      y = "Number of items without DIF",
+      x = "Number of DIF items",
+      y = "Mean anchor hit rate",
       color = ""
-    )
+    ) +
+    scale_y_continuous(labels = scales::percent, limits = c(0, 1))
 ```
+
+# Discussion and summary
+
+Measurement instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
+Inspired by @ackerman1992didactic, we have resurrected the DIF-as-student-property view and shown its connection to the typical way in which AI methods and DIF simulation studies have been conceptualized.
+
+More importantly, we have reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe is more realistic than the typical DIF simulation study.
+In particular, we simulated student ability as drawn from a two-dimensional distribution representing a student's target and nuisance ability, and then generated data using a noncompensatory item response model.
+Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
+On the other hand, selecting anchor items using either AOAA-OAT or clustering and selecting an anchor point using either GINI or MINBC all performed reasonably.
+
+However, AOAA-OAT was the clear winner in terms of performance and we recommend its use whenever possible.
+One reason AOAA-OAT might not always be possible is that it can be computationally expensive.
+For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows.
+To increase AOAA-OAT's use, we recommend its implementation (perhaps as the default) in popular IRT software such as the mirt R package.
+
+In addition to exploring and testing algorithmic AI methods, we introduced a method, the EM-MILG, that an analyst can use to visualize the amount of potential DIF in their data.
+This method can be used either to build their intuition or as a way in which they can select anchor items by hand.
+The EM-MILG's sibling method, the MILG, is, we believe, the best way to visualize the results of a DIF analysis after anchor items have been selected.
+
+Future work should test these methods' performance under a greater variety of data-generating conditions.
+For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
+Furthermore, our work focused on the Rasch model, and it will be of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.

### 2020-03-14 R-work-complete/mirt doesnt do AOAA-OAT.R

@@ -1,0 +1,13 @@
+data <- out_3_biased_items$sim[[1]]$data
+groups <- out_3_biased_items$sim[[1]]$groups
+
+
+model_constrained <-
+    multipleGroup(
+        data, 1, groups, SE = TRUE,
+        invariance = c(colnames(data), "free_means", "free_var")
+    )
+
+DIF(model_constrained, 'd', scheme = 'drop_sequential')
+
+coef_1f(model_constrained)

### 2020-03-14 paper/paper.tex
after knitting and sending that copy to ben d and just before doing a reorganization to try to keep this repo wieldy

@@
 However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF.
-We summarize existing assumptionless identification (AI) methods and propose a variety of new methods.
+We summarize existing agnostic identification (AI) methods and propose a variety of new methods.
 We conduct a simulation study --- which we believe to be more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
@@
 
-We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups.
+For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
 For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students.
@@
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional compensatory 2PL model, the probability student \(i\) responding correctly to item \(j\) is
+According to the two-dimensional compensatory 2PL model, the probability student \(i\) responds correctly to item \(j\) is
 \[
@@
 \]
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability across students in the same group.
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
 
@@
 
-\hypertarget{assumptionless-identification-methods}{%
-\section{Assumptionless identification methods}\label{assumptionless-identification-methods}}
+\hypertarget{agnostic-identification-methods}{%
+\section{Agnostic identification methods}\label{agnostic-identification-methods}}
 
 Psychometricians have long been in search of the perfect DIF detection method.
-It has become well known that the IRT-based likelihood ratio test (LRT) is the best way to test for DIF, which has been well demonstrated (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
+It has been well demonstrated that the IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
 As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).
 
-The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions --- about either relative group ability or which items contain bias --- can be made.
-That is, we're interested in assumptionless identification methods (hereafter referred to as \enquote{AI methods}) so that the LRTs can be used to test items for DIF.
+The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions --- about either relative group ability or which items might have DIF --- can be made.
+That is, we're interested in agnostic identification methods (hereafter referred to as \enquote{AI methods}) so that LRTs can be used to test items for DIF.
 For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
 And, in search of a coherent framework, we sometimes edit names of existing methods.
-We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to the proliferation of names.
+We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.
 
@@
 Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{ref}} = -1\).
+Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{foc}} = -1\).
 The slope on target ability is set to \(a_{1j} = 1\) for all items.
 The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
-From the DIF-as-item-property view, there is no nuisance ability.
+We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability.
 Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.
@@
 We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
-The fundamental challenge is to use an AI method to identify the model, which allows for the disentangling \(\hat\mu^\text{foc}\) from estimating \(\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\).
-The most common AI method is to select a group of anchor items that are assumed to be DIF-free.
+The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of \(\hat\mu^\text{foc}\) from the estimation of \(\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\).
+The most common AI approach is to select a group of anchor items that are assumed to be DIF-free.
 These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.
 
-Computing is done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the R package mirt (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the set of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}).
-Code is available at TO DO.
+Computing is done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the mirt R package (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the suite of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}).
+Code is available at \emph{(TODO)}.
 
@@
 
-\hypertarget{all-others-as-anchors}{%
-\subsubsection{All-others-as-anchors}\label{all-others-as-anchors}}
+\hypertarget{all-others-as-anchors-aoaa}{%
+\subsubsection{All-others-as-anchors (AOAA)}\label{all-others-as-anchors-aoaa}}
 
@@
 The items for which the flexible model outperforms the baseline model (typically based on a \(\chi^2\) test) are identified as having DIF, and the rest of the items become anchor items.
-Note that AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
+AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
 
@@
 
-\hypertarget{all-others-as-anchors-all-significant}{%
-\paragraph{All-others-as-anchors-all-significant}\label{all-others-as-anchors-all-significant}}
+\hypertarget{all-others-as-anchors-all-significant-aoaa-as}{%
+\paragraph{All-others-as-anchors-all-significant (AOAA-AS)}\label{all-others-as-anchors-all-significant-aoaa-as}}
 
@@
 Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
-Note that AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.
+AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.
 
@@
 Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items display DIF.
+It's unclear if one should proceed if those four items test positive for DIF, too.
 
-\hypertarget{all-other-as-anchors-one-at-a-time}{%
-\paragraph{All-other-as-anchors-one-at-a-time}\label{all-other-as-anchors-one-at-a-time}}
+\hypertarget{all-other-as-anchors-one-at-a-time-aoaa-oat}{%
+\paragraph{All-other-as-anchors-one-at-a-time (AOAA-OAT)}\label{all-other-as-anchors-one-at-a-time-aoaa-oat}}
 
@@
 In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
+To work, AOAA-OAT has two requirements: First, that at least two items do not have DIF, and second, that the set of items without DIF are more homogeneous than other sets of items.
 
@@
 It's useful to remember that AOAA is not an iterative procedure.
-The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
+The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method.
 In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
@@
 \begin{table}[H]
-\caption{Summary of the three all-others-as-anchors assumptionless identification methods}
+\caption{Summary of the three all-others-as-anchors agnostic identification methods}
 \centering
@@
 
-\hypertarget{clustering-items}{%
-\subsubsection{Clustering items}\label{clustering-items}}
+\hypertarget{equal-means-clustering-emc}{%
+\subsubsection{Equal means clustering (EMC)}\label{equal-means-clustering-emc}}
 
@@
 
-\hypertarget{the-equal-means-multiple-imputation-logit-graph}{%
-\subsubsection{The equal means, multiple imputation logit graph}\label{the-equal-means-multiple-imputation-logit-graph}}
+\hypertarget{the-equal-means-multiple-imputation-logit-graph-em-milg}{%
+\subsubsection{The equal means, multiple imputation logit graph (EM-MILG)}\label{the-equal-means-multiple-imputation-logit-graph-em-milg}}
 
@@
 To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of \(\tilde{d_j}\) for each item.
+Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are drawn to estimate the distribution of \(\tilde{d_j}\) for each item.
 These are the distributions displayed in a EM-MILG.
@@
 
-Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data.
+Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation.
 We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
@@
 After choosing the anchor items, the model is refit.
-The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc}\) to \(0\).
-The same process of using multiple imputations to estimate the distribution of \(d_j\) can be used with the new model.
+The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc} = 0\).
+The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with the new model.
 Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
@@
 And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
-In particular, we could visualize DIF in the non-anchor items after using AOAA or CLUSTER to identify the model.
+In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.
 
@@
 The anchor items are used to estimate \(\hat \mu^\text{foc}\).
-An alternative strategy is to directly select the anchor point, \(\mu^{\star\text{foc}}\).
-Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, all items can be tested for DIF.
+An alternative strategy is to directly set the anchor point, \(\mu^{\star\text{foc}}\).
+Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF.
 The question then becomes the following: How is the anchor point selected?
 
-\hypertarget{gini-index}{%
-\subsubsection{Gini index}\label{gini-index}}
+\hypertarget{maximizing-the-gini-index-maxgi}{%
+\subsubsection{Maximizing the Gini index (MAXGI)}\label{maximizing-the-gini-index-maxgi}}
 
@@
 
-\(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (which is why we use the abbreviation MGI).
-The intuition and assumption is that some items might have bias but the majority do not.
+\(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (thus the abbreviation MAXGI).
+The intuition and assumption is that the items without DIF are the most homogeneous.
 Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
@@
 
-For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values. The result of MGI is \(\mu^{\star\text{foc}} = -0.99\).
+For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values.
+The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is extraordinarily close to the data-generating value of -1.
 
@@
 
-\caption{Maximizing the Gini index (MGI) to select the anchor point.}\label{fig:ginipath}
+\caption{Maximizing the Gini index (MAXGI) to select the anchor point.}\label{fig:ginipath}
 \end{figure}
 
-The model is then refit with the model identifying assumption of \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\), and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
+The model is then refit with the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\), and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
 
@@
 
-\caption{The MILG with $\mu^{\text{foc}}$ set to $\mu^{\star\text{foc}} = -0.99$}\label{fig:ginimilg}
+\caption{The MILG with $\mu^{\text{foc}}$ set to $\mu^{\star\text{foc}} = -0.99$.}\label{fig:ginimilg}
 \end{figure}
 
-\hypertarget{minimizing-between-curves}{%
-\subsubsection{Minimizing between curves}\label{minimizing-between-curves}}
+\hypertarget{minimizing-between-curves-minbc}{%
+\subsubsection{Minimizing between curves (MINBC)}\label{minimizing-between-curves-minbc}}
 
@@
 \]
-Raju's area method, along with AOAA, has been cited as the two most commonly used IRT-based DIF detection methods (Magis et al.
+Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods (Magis et al.
 \protect\hyperlink{ref-magis2011generalized}{2011}).
-We see two weakness of Raju's area method: First, the item characteristic curves still need to be linked by anchor items or an anchor point, which the method does not help to identify.
-Second, the area is unweighted, so all values of \(\theta\) matter equally, despite some being much more realistic than others.
+However, Raju's area method is not an AI method because the item characteristic curves still need to be linked by anchor items or an anchor point.
+An additional weakness is that the area is unweighted, so all values of \(\theta\) matter equally, despite some being much more realistic than others.
 
-To build off of Raju's area method, we propose a new method, which we call \enquote{minimizing the area between curves} (MABC).
-To understand MABC, imagine multigroup item response data where \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\) so that there is no DIF.
-The fundamental identification problem of DIF in multigroup IRT models is that there are an infinite number of models with the same likelihood from which to choose.
+To adapt Raju's area method into an AI method, we propose a new method, which we call \enquote{minimizing the area between curves} (MINBC).
+To understand MINBC, imagine a scenario in which the data-generating process is \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\), so that there is no DIF.
+The fundamental identification problem is that there are an infinite number of models with the same likelihood from which to choose.
 For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\mu^{\star\text{foc}} = 0\).
@@
 Both of these models have the same likelihood, so how should one choose which model to believe?
-MABC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves.
-As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences as opposed to DIF.
+MINBC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves.
+As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences (as opposed to DIF).
 
@@
 \]
-In this way, \(\text{Total DIF}(\mu^\text{foc})\) is a function where the input is \(\mu^\text{foc}\) and the output is the total amount of DIF on the test. MABC sets
+In this way, \(\text{Total DIF}(\mu^\text{foc})\) is a function where the input is \(\mu^\text{foc}\) and the output is the total amount of DIF on the test. MINBC sets
 \[
@@
 \]
-MABC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}), who use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
+MINBC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}), who use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
 The selection of \(g(\theta)\) results in the relative weighting of \(\theta\) values.
-Chalmers, Counsell, and Flora do not discuss how to choose \(g(\theta)\) and in their empirical examples use \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which seems like an odd choice.
+Chalmers, Counsell, and Flora do not discuss how to choose \(g(\theta)\) and in their empirical examples use \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which may be suboptimal in some cases.
 It might seem intuitive to choose \(g(\theta) \sim N(0, 1)\) because \(\mu^\text{ref} = 0\), but this choice doesn't take into account the ability distribution of the focal group.
 If \(\mu^\text{foc} = 3\), wouldn't we also want to prioritize high \(\theta\) values?
-Accordingly, we set \(g(\theta)\) to be the average of the reference and focal group ability distributions:
+Accordingly, we set \(g(\theta)\) to be the average of the reference and focal group ability probability density functions:
 \[
@@
 For our one-run simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\).
-The anchor point is found to be \(\mu^{\star\text{foc}} = -1\).
-As with MGI, the model should then be refit using the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\).
+In this case, MINBC works perfectly and the anchor point is found to be \(\mu^{\star\text{foc}} = -1\).
+As with MAXGI, the model should then be refit using the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\).
 
@@
 
-\caption{Minimizing the area between curves (MABC) to select the anchor point.}\label{fig:mabc}
+\caption{Minimizing the area between curves (MINBC) to select the anchor point.}\label{fig:mabc}
 \end{figure}
@@
 We have described a variety of AI methods and their corresponding acronymns.
-Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MGI and MABC, select an anchor point.
+Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point.
 Table \ref{table:allmethods} summarizes all of these methods.
@@
 \begin{table}[H]
-\caption{Summary of assumptionless identification methods}
+\caption{Summary of agnostic identification methods}
 \centering
@@
 
-maximizing Gini index (MGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
+maximizing Gini index (MAXGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
 
-minimizing the area between curves (MABC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
+minimizing the area between curves (MINBC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
 
@@
 
-To compare each of the methods in table TODO, we conducted a simulated study.
+To compare each of the methods in Table \ref{table:allmethods}, we conducted a simulated study.
 Our goal was to make the data generating process as realistic to the scenario described by Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) in which some items on a math test also depend on a student's verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
@@
 This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
-One exception is Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}) who draws each student's target and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
+One exception is Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}) who draw each student's target ability and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
 
 In our simulation study, it was critical that student ability was drawn in a realistic way similar to Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}).
-However, we don't believe that a compensatory model is realistic in describing the a math test where some items depend on verbal ability.
+However, we don't believe that a compensatory model is realistic in describing a math test where some items depend on verbal ability.
 For example, it's hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability.
@@
 \]
-The angle for an item measure the relative loading of the item on the two dimensions.
-For example, an angle of 45\(^\circ\) indicates that the item loads equally on the target and nuisance abilities.
-Our simple compensatory model has \(a_{1j} = 1\) for all items so the angle equation reduces to
+An item's angle measures the relative loading of the item on the two dimensions.
+For example, an angle of 45\(^\circ\) indicates that the item loads equally on the target ability and nuisance ability.
+Our simple noncompensatory model has \(a_{1j} = 1\) for all items so the angle equation reduces to
 \[
@@
 
-Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between \(\theta_i\) and \(Pr(y_{ij} = 1)\) with \(\eta_i\) set to the group mean for a test with six items with DIF.
+Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between \(\theta_i\) and \(Pr(y_{ij} = 1)\) with \(\eta_i\) set to the group mean for a test with six DIF items.
 The items are ordered by the amount of DIF such that \(\angle_{j = 7} = 20^\circ\) up to \(\angle_{j = 12} = 60^\circ\).
@@
 
+For each run, we applied each AI method to find the method's identifying assumption.
+The method's identifying assumption was then used to fit a final model.
+We compared the performance of those final models according to the following outcomes.
+
 \hypertarget{achievement-gap-residual}{%
@@
 
-An effective method should accurately estimate the group-level target ability gap, which we refer to as the achievement gap.
+An effective AI method should lead to a final model that accurately estimates the difference between the reference group's mean target ability and the focal group's mean target ability.
+We refer to this quantity as the achievement gap.
+Recall that all models set \(\mu_\theta^\text{ref} = 0\), so the achievement gap reduces to \(\mu_\theta^\text{ref}\).
 The data-generating value of \(\mu_\theta^\text{foc}\) is 0.5, but each run will include sampling variability.
-To get at the heart of how well a method is doing, we calculate the achievement gap residual as the method's estimated achievement gap, \(\hat\mu_\theta^\text{foc}\), minus the achievement gap estimated when using only the DIF-free items as anchors.
+To get at the heart of how well a method is doing, we calculated the achievement gap residual as the method's estimated achievement gap, \(\hat\mu_\theta^\text{foc}\), minus the achievement gap estimated when using only the DIF-free items as anchors.
 In summary, this outcome measures a method's ability to disentangle differences in target ability from nuisance ability at the group level.
@@
 Assessments are frequently used to make decisions about individual student abilities.
-We measure a method's ability to do so by calculating the rank correlation between the vector estimated abilities and true abilities.
-This is done both for all students combined and just for students in the focal group.
-We use rank correlation as opposed to the more typically used RMSE as a result of our agreement with Lord's (\protect\hyperlink{ref-lord1986maximum}{1986}) argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
-\emph{(Note that this isn't included in this current draft.)}
+We measured a method's ability to do so by calculating the rank correlation between the vector of estimated abilities and the vector of true target abilities.
+We used rank correlation as opposed to the more typically used RMSE in agreement with Lord's (\protect\hyperlink{ref-lord1986maximum}{1986}) argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
+\emph{(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary.
+maybe this idea of rank order being better than RMSE is for another paper)}
 
@@
 
-For the methods that choose a set of anchor items, we look directly at which anchor items are selected.
-An effective method should use most of the non-DIF items as anchors (the anchor hit rate) while avoiding including items with DIF in the anchor set (the false anchor rate).
+For the methods that choose a set of anchor items, we looked directly at which anchor items were selected.
+An effective method should use most of the non-DIF items as anchors (the anchor hit rate) while avoiding using items with DIF as anchors (the false anchor rate).
 
@@
 Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual.
-As expected, artificial DIF causes AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
-CLUSTER performs better than AOAA and AOAA-AS but worse than the other methods.
-GINI and MINBC perform similarly well with GINI having a higher variance, while MINBC has more bias, especially for tests with a greater number of DIF items.
-This bias results from MINBC considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.
 AOAA-OAT is the clear winner.
-It performs nearly perfectly for two, three, or four DIF items, and even on its worst run, underestimates \(\mu_\theta^\text{foc}\) by 0.05 standard deviations.
+It performed nearly perfectly for two, three, or four DIF items.
+Even when six of the 12 items on the test contained DIF, AOAA-OAT underestimated \(\mu_\theta^\text{foc}\) by only 0.05 standard deviations on its worst run.
+As expected, artificial DIF caused AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
+
+EMC performed better than AOAA and AOAA-AS, but worse than the other methods.
+MINBC and MAXGI performed similarly well with MINBC estimating the achievement gap with more precision but more bias than MAXGI, especially for tests with more than four DIF items.
+We hypothesize that MINBC's susceptibility to bias results from considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.
 
@@
 
-\caption{Target ability achievement gap estimates for each method.}\label{fig:achievegap}
+\caption{Achievement gap residual distributions across 100 runs for each AI method and number of DIF items.}\label{fig:achievegap}
 \end{figure}
@@
 Figure \ref{fig:anchorfalse} shows the mean false anchor rates for each method and number of items with DIF.
-For example, when there are two items with DIF on the test, those two items have \(\angle_{11} = 20^\circ\) and \(\angle_{12} = 60^\circ\).
-AOAA-OAT never includes item 12 in the anchor set, but incorrectly includes item 11 in 60 out of the 100 runs.
-Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method is
+For example, when there were two items with DIF on the test, those two items had \(\angle_{11} = 20^\circ\) and \(\angle_{12} = 60^\circ\).
+AOAA-OAT never included item 12 in the anchor set, but incorrectly included item 11 in 60 out of the 100 runs.
+Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method was
 \[
@@
 \]
-The fact that the item with 20\(^\circ\) of DIF is most commonly incorrectly included in the anchor set is what drives the counterintuitive result that the mean false anchor rate decreases with more DIF items.
+The fact that the item with 20\(^\circ\) of DIF is most commonly incorrectly included in the anchor set is what drove the counterintuitive result that the mean false anchor rate decreases with more DIF items.
 
 Similarly, Figure \ref{fig:anchorhit} shows the mean anchor hit rates.
-Remarkably, AOAA-OAT includes an average of over \(90\%\) of DIF-free items in the anchor set regardless of the number of DIF items on the test.
-Interestingly, CLUSTER has a better anchor hit rate on tests with more DIF items.
+Remarkably, AOAA-OAT included an average of over \(90\%\) of DIF-free items in the anchor set regardless of the number of DIF items on the test.
+Interestingly, EMC had a better anchor hit rate on tests with more DIF items.
 This result appears to be driven by the clustering algorithm sometimes splitting all of the DIF-free items into two separate clusters, especially when most of the items are DIF-free.
@@
 
-\caption{Mean false anchor rates across 100 runs for each number of DIF items}\label{fig:anchorfalse}
+\caption{Mean false anchor rates across 100 runs for each AI method and number of DIF items.}\label{fig:anchorfalse}
 \end{figure}
@@
 
-\caption{Mean anchor hit rates across 100 runs for each number of DIF items}\label{fig:anchorhit}
+\caption{Mean anchor hit rates across 100 runs for each AI method and number of DIF items.}\label{fig:anchorhit}
 \end{figure}
@@
 
-More importantly, we have reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe is more realistic than the typical DIF simulation study.
+More importantly, we reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe to be more realistic than the typical DIF simulation study.
 In particular, we simulated student ability as drawn from a two-dimensional distribution representing a student's target and nuisance ability, and then generated data using a noncompensatory item response model.
 Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
-On the other hand, selecting anchor items using either AOAA-OAT or clustering and selecting an anchor point using either GINI or MINBC all performed reasonably.
+On the other hand, AOAA-OAT, EMC, MINGI, and MAXBC all performed reasonably.
 
-However, AOAA-OAT was the clear winner in terms of performance and we recommend its use whenever possible.
+In particular, AOAA-OAT was the clear winner, and we recommend its use whenever possible.
 One reason AOAA-OAT might not always be possible is that it can be computationally expensive.
@@
 \clearpage
-
-\hypertarget{references}{%
-\section{References}\label{references}}
 

### 2020-03-14 paper/paper_1_intro.Rmd
after knitting and sending that copy to ben d and just before doing a reorganization to try to keep this repo wieldy

@@
 
-We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups.
+For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
 For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students.
@@
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional compensatory 2PL model, the probability student $i$ responding correctly to item $j$ is
+According to the two-dimensional compensatory 2PL model, the probability student $i$ responds correctly to item $j$ is
 $$
@@
 $$
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability across students in the same group.
+This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
 

### 2020-03-14 paper/paper_2a_anchor_items.Rmd
after knitting and sending that copy to ben d and just before doing a reorganization to try to keep this repo wieldy

@@
 Psychometricians have long been in search of the perfect DIF detection method.
-It has become well known that the IRT-based likelihood ratio test (LRT) is the best way to test for DIF, which has been well demonstrated [@meade2012solving].
+It has been well demonstrated that the IRT-based likelihood ratio test (LRT) is effective at detecting DIF [@meade2012solving].
 As a result, we avoid methods like the Mantel-Haenszel procedure [@holland1986differential], which muddies the waters by moving away from the IRT framework, and has been shown to perform no better — and in some cases worse — than IRT-based methods [@swaminathan1990detecting].
 
-The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items contain bias — can be made.
-That is, we're interested in assumptionless identification methods (hereafter referred to as "AI methods") so that the LRTs can be used to test items for DIF.
+The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items might have DIF — can be made.
+That is, we're interested in agnostic identification methods (hereafter referred to as "AI methods") so that LRTs can be used to test items for DIF.
 For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
@@
 We recognize that others have done the same [e.g.
-@kopf2015framework], and that we risk contributing to the proliferation of names.
+@kopf2015framework], and that we risk contributing to a proliferation of names.
 
@@
 Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{ref}} = -1$.
+Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{foc}} = -1$.
 The slope on target ability is set to $a_{1j} = 1$ for all items.
 The slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise.
-From the DIF-as-item-property view, there is no nuisance ability.
+We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability.
 Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.
@@
 We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
-The fundamental challenge is to use an AI method to identify the model, which allows for the disentangling $\hat\mu^\text{foc}$ from estimating $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$.
-The most common AI method is to select a group of anchor items that are assumed to be DIF-free.
+The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of $\hat\mu^\text{foc}$ from the estimation of $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$.
+The most common AI approach is to select a group of anchor items that are assumed to be DIF-free.
 These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.
 
-Computing is done in R [@rcore], model fitting in the R package mirt [@chalmers2012mirt], and data wrangling/visualization in the set of R packages known as the tidyverse [@tidy].
-Code is available at TO DO.
+Computing is done in R [@rcore], model fitting in the mirt R package [@chalmers2012mirt], and data wrangling/visualization in the suite of R packages known as the tidyverse [@tidy].
+Code is available at *(TODO)*.
 
@@
 
-### All-others-as-anchors
+### All-others-as-anchors (AOAA)
 
@@
 The items for which the flexible model outperforms the baseline model (typically based on a $\chi^2$ test) are identified as having DIF, and the rest of the items become anchor items.
-Note that AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
+AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
 
@@
 
-#### All-others-as-anchors-all-significant
+#### All-others-as-anchors-all-significant (AOAA-AS)
 
@@
 Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
-Note that AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.
+AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.
 
@@
 @woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items display DIF.
+It's unclear if one should proceed if those four items test positive for DIF, too.
 
-#### All-other-as-anchors-one-at-a-time
+#### All-other-as-anchors-one-at-a-time (AOAA-OAT)
 
@@
 
-### Clustering items
+### Equal means clustering (EMC)
 
@@
 
-### The equal means, multiple imputation logit graph
+### The equal means, multiple imputation logit graph (EM-MILG)
 
@@
 To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical].
-Then, multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of $\tilde{d_j}$ for each item.
+Then, multiple imputations (MI) [@yang2012characterizing] are drawn to estimate the distribution of $\tilde{d_j}$ for each item.
 These are the distributions displayed in a EM-MILG.
@@
 After choosing the anchor items, the model is refit.
-The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc}$ to $0$.
-The same process of using multiple imputations to estimate the distribution of $d_j$ can be used with the new model.
+The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc} = 0$.
+The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the new model.
 Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
@@
     labs(
-        x = latex2exp::TeX("$d_j = b_j^{ref} - b_j^{foc}$"),
+        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$"),
         y = ""
@@
 And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
-In particular, we could visualize DIF in the non-anchor items after using AOAA or CLUSTER to identify the model.
+In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.

### 2020-03-14 paper/paper_2b_anchor_points.Rmd
after knitting and sending that copy to ben d and just before doing a reorganization to try to keep this repo wieldy

@@
 The anchor items are used to estimate $\hat \mu^\text{foc}$.
-An alternative strategy is to directly select the anchor point, $\mu^{\star\text{foc}}$.
-Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, all items can be tested for DIF.
+An alternative strategy is to directly set the anchor point, $\mu^{\star\text{foc}}$.
+Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF.
 The question then becomes the following: How is the anchor point selected?
 
-### Gini index
+### Maximizing the Gini index (MAXGI)
 
@@
 
-$\mu^{\star\text{foc}}$ is selected by maximizing the Gini index (which is why we use the abbreviation MGI).
-The intuition and assumption is that some items might have bias but the majority do not.
+$\mu^{\star\text{foc}}$ is selected by maximizing the Gini index (thus the abbreviation MAXGI).
+The intuition and assumption is that the items without DIF are the most homogeneous.
 Denoting a function that calculates the Gini index from a vector of non-negative elements as  $G(\mathbf{x})$, MGI sets <!-- its weird to arg max over a parameter we already have a definition for - maybe switch to t or something -->
@@
 For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible $\mu^\text{foc}$ values.
-The result of MGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$.
+The result of MAXGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$, which is extraordinarily close to the data-generating value of -1.
 
-```{r ginipath, fig.cap = 'Maximizing the Gini index (MGI) to select the anchor point.', out.width="70%", warning = FALSE, message = FALSE}
+```{r ginipath, fig.cap = 'Maximizing the Gini index (MAXGI) to select the anchor point.', out.width="70%", warning = FALSE, message = FALSE}
 gini_results$grid %>% 
@@
 
-The model is then refit with the model identifying assumption of $\mu^{\text{foc}} = \mu^{\star\text{foc}}$, and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
+The model is then refit with the identifying assumption that $\mu^{\text{foc}} = \mu^{\star\text{foc}}$, and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
 
-```{r ginimilg, fig.cap = "The MILG with $\\mu^{\\text{foc}}$ set to $\\mu^{\\star\\text{foc}} = -0.99$", out.width="70%", warning = FALSE, message = FALSE}
+```{r ginimilg, fig.cap = "The MILG with $\\mu^{\\text{foc}}$ set to $\\mu^{\\star\\text{foc}} = -0.99$.", out.width="70%", warning = FALSE, message = FALSE}
 # notice hard coding in this caption
@@
 
-### Minimizing between curves
+### Minimizing between curves (MINBC)
 
@@
 $$
-Raju's area method, along with AOAA, has been cited as the two most commonly used IRT-based DIF detection methods [@magis2011generalized].
-We see two weakness of Raju's area method: First, the item characteristic curves still need to be linked by anchor items or an anchor point, which the method does not help to identify.
-Second, the area is unweighted, so all values of $\theta$ matter equally, despite some being much more realistic than others.
+Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods [@magis2011generalized].
+However, Raju's area method is not an AI method because the item characteristic curves still need to be linked by anchor items or an anchor point.
+An additional weakness is that the area is unweighted, so all values of $\theta$ matter equally, despite some being much more realistic than others.
 
-To build off of Raju's area method, we propose a new method, which we call "minimizing the area between curves" (MABC).
-To understand MABC, imagine multigroup item response data where $\mu^{\text{foc}} = \mu^{\text{ref}}$ and $d_j = 0 \forall j$ so that there is no DIF.
-The fundamental identification problem of DIF in multigroup IRT models is that there are an infinite number of models with the same likelihood from which to choose.
+To adapt Raju's area method into an AI method, we propose a new method, which we call "minimizing the area between curves" (MINBC).
+To understand MINBC, imagine a scenario in which the data-generating process is $\mu^{\text{foc}} = \mu^{\text{ref}}$ and $d_j = 0 \forall j$, so that there is no DIF.
+The fundamental identification problem is that there are an infinite number of models with the same likelihood from which to choose.
 For example, we could correctly assume that the focal group has the same ability as the reference group and fix $\mu^{\star\text{foc}} = 0$.
@@
 $$
-In this way, $\text{Total DIF}(\mu^\text{foc})$ is a function where the input is $\mu^\text{foc}$ and the output is the total amount of DIF on the test. MABC sets
+In this way, $\text{Total DIF}(\mu^\text{foc})$ is a function where the input is $\mu^\text{foc}$ and the output is the total amount of DIF on the test. MINBC sets
 $$
@@
 $$
-MABC is inspired in part by @chalmers2016might, who use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
+MINBC is inspired in part by @chalmers2016might, who use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
 The selection of $g(\theta)$ results in the relative weighting of $\theta$ values.
-Chalmers, Counsell, and Flora do not discuss how to choose $g(\theta)$ and in their empirical examples use $g(\theta)$ uniform for $-6 \le \theta \le 6$, which seems like an odd choice.
+Chalmers, Counsell, and Flora do not discuss how to choose $g(\theta)$ and in their empirical examples use $g(\theta)$ uniform for $-6 \le \theta \le 6$, which may be suboptimal in some cases.
 It might seem intuitive to choose $g(\theta) \sim N(0, 1)$ because $\mu^\text{ref} = 0$, but this choice doesn't take into account the ability distribution of the focal group.
 If $\mu^\text{foc} = 3$, wouldn't we also want to prioritize high $\theta$ values?
-Accordingly, we set $g(\theta)$ to be the average of the reference and focal group ability distributions:
+Accordingly, we set $g(\theta)$ to be the average of the reference and focal group ability probability density functions:
 $$
@@
 For our one-run simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$.
-The anchor point is found to be $\mu^{\star\text{foc}} = `r round(min_bc$between_curves_anchor_points, 2)`$.
-As with MGI, the model should then be refit using the identifying assumption that $\mu^{\text{foc}} = \mu^{\star\text{foc}}$.
+In this case, MINBC works perfectly and the anchor point is found to be $\mu^{\star\text{foc}} = `r round(min_bc$between_curves_anchor_points, 2)`$.
+As with MAXGI, the model should then be refit using the identifying assumption that $\mu^{\text{foc}} = \mu^{\star\text{foc}}$.
 
-```{r mabc, fig.cap = 'Minimizing the area between curves (MABC) to select the anchor point.', out.width="70%"}
+```{r mabc, fig.cap = 'Minimizing the area between curves (MINBC) to select the anchor point.', out.width="70%"}
 min_bc$grid %>% 

### 2020-03-14 paper/paper_3_sim.Rmd
after knitting and sending that copy to ben d and just before doing a reorganization to try to keep this repo wieldy

@@
-To compare each of the methods in table TODO, we conducted a simulated study.
+To compare each of the methods in Table \ref{table:allmethods}, we conducted a simulated study.
 Our goal was to make the data generating process as realistic to the scenario described by @ackerman1992didactic in which some items on a math test also depend on a student’s verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
@@
 This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
-One exception is @walker2017using who draws each student's target and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
+One exception is @walker2017using who draw each student's target ability and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
 
 In our simulation study, it was critical that student ability was drawn in a realistic way similar to @walker2017using.
-However, we don’t believe that a compensatory model is realistic in describing the a math test where some items depend on verbal ability.
+However, we don’t believe that a compensatory model is realistic in describing a math test where some items depend on verbal ability.
 For example, it’s hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability.
@@
 $$
-The angle for an item measure the relative loading of the item on the two dimensions.
-For example, an angle of 45$^\circ$ indicates that the item loads equally on the target and nuisance abilities.
-Our simple compensatory model has $a_{1j} = 1$ for all items so the angle equation reduces to
+An item's angle measures the relative loading of the item on the two dimensions.
+For example, an angle of 45$^\circ$ indicates that the item loads equally on the target ability and nuisance ability.
+Our simple noncompensatory model has $a_{1j} = 1$ for all items so the angle equation reduces to
 $$
@@
 
-Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean for a test with six items with DIF.
+Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean for a test with six DIF items.
 The items are ordered by the amount of DIF such that $\angle_{j = 7} = 20^\circ$ up to $\angle_{j = 12} = 60^\circ$.
@@
 
+For each run, we applied each AI method to find the method's identifying assumption.
+The method's identifying assumption was then used to fit a final model.
+We compared the performance of those final models according to the following outcomes.
+
 ### Achievement gap residual
 
-An effective method should accurately estimate the group-level target ability gap, which we refer to as the achievement gap.
+An effective AI method should lead to a final model that accurately estimates the difference between the reference group's mean target ability and the focal group's mean target ability.
+We refer to this quantity as the achievement gap.
+Recall that all models set $\mu_\theta^\text{ref} = 0$, so the achievement gap reduces to $\mu_\theta^\text{ref}$.
 The data-generating value of $\mu_\theta^\text{foc}$ is 0.5, but each run will include sampling variability.
-To get at the heart of how well a method is doing, we calculate the achievement gap residual as the method's estimated achievement gap, $\hat\mu_\theta^\text{foc}$, minus the achievement gap estimated when using only the DIF-free items as anchors.
+To get at the heart of how well a method is doing, we calculated the achievement gap residual as the method's estimated achievement gap, $\hat\mu_\theta^\text{foc}$, minus the achievement gap estimated when using only the DIF-free items as anchors.
 In summary, this outcome measures a method’s ability to disentangle differences in target ability from nuisance ability at the group level.
@@
 Assessments are frequently used to make decisions about individual student abilities.
-We measure a method’s ability to do so by calculating the rank correlation between the vector estimated abilities and true abilities.
-This is done both for all students combined and just for students in the focal group.
-We use rank correlation as opposed to the more typically used RMSE as a result of our agreement with Lord's -@lord1986maximum argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
-*(Note that this isn't included in this current draft.)*
+We measured a method’s ability to do so by calculating the rank correlation between the vector of estimated abilities and the vector of true target abilities.
+We used rank correlation as opposed to the more typically used RMSE in agreement with Lord's -@lord1986maximum argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
+*(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary.
+maybe this idea of rank order being better than RMSE is for another paper)*
 
@@
 Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual.
-As expected, artificial DIF causes AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
-CLUSTER performs better than AOAA and AOAA-AS but worse than the other methods.
-GINI and MINBC perform similarly well with GINI having a higher variance, while MINBC has more bias, especially for tests with a greater number of DIF items.
-This bias results from MINBC considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.
 AOAA-OAT is the clear winner.
-It performs nearly perfectly for two, three, or four DIF items, and even on its worst run, underestimates $\mu_\theta^\text{foc}$ by 0.05 standard deviations.
+It performed nearly perfectly for two, three, or four DIF items.
+Even when six of the 12 items on the test contained DIF, AOAA-OAT underestimated $\mu_\theta^\text{foc}$ by only 0.05 standard deviations on its worst run.
+As expected, artificial DIF caused AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
 
-```{r achievegap, fig.cap = 'Target ability achievement gap estimates for each method.', out.width="70%", warning = FALSE, message = FALSE}
+EMC performed better than AOAA and AOAA-AS, but worse than the other methods.
+MINBC and MAXGI performed similarly well with MINBC estimating the achievement gap with more precision but more bias than MAXGI, especially for tests with more than four DIF items.
+We hypothesize that MINBC's susceptibility to bias results from considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.
+
+```{r achievegap, fig.cap = 'Achievement gap residual distributions across 100 runs for each AI method and number of DIF items.', out.width="70%", warning = FALSE, message = FALSE}
 what %>% 
@@
         levels = c("AOAA_OAT_mod", "cluster_mod", "gini_mod", "AOAA_AS_mod", "AOAA_mod", "minbc_mod"),
-        labels = c("AOAA-OAT", "CLUSTER", "GINI", "AOAA-AS", "AOAA", "MINBC")
+        labels = c("AOAA-OAT", "EMC", "MAXGI", "AOAA-AS", "AOAA", "MINBC")
       )
@@
 Figure \ref{fig:anchorfalse} shows the mean false anchor rates for each method and number of items with DIF.
-For example, when there are two items with DIF on the test, those two items have $\angle_{11} = 20^\circ$ and $\angle_{12} = 60^\circ$.
-AOAA-OAT never includes item 12 in the anchor set, but incorrectly includes item 11 in 60 out of the 100 runs.
-Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method is
+For example, when there were two items with DIF on the test, those two items had $\angle_{11} = 20^\circ$ and $\angle_{12} = 60^\circ$.
+AOAA-OAT never included item 12 in the anchor set, but incorrectly included item 11 in 60 out of the 100 runs.
+Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method was
 $$
@@
 $$
-The fact that the item with 20$^\circ$ of DIF is most commonly incorrectly included in the anchor set is what drives the counterintuitive result that the mean false anchor rate decreases with more DIF items.
+The fact that the item with 20$^\circ$ of DIF is most commonly incorrectly included in the anchor set is what drove the counterintuitive result that the mean false anchor rate decreases with more DIF items.
 
 Similarly, Figure \ref{fig:anchorhit} shows the mean anchor hit rates.
-Remarkably, AOAA-OAT includes an average of over $90\%$ of DIF-free items in the anchor set regardless of the number of DIF items on the test.
-Interestingly, CLUSTER has a better anchor hit rate on tests with more DIF items.
+Remarkably, AOAA-OAT included an average of over $90\%$ of DIF-free items in the anchor set regardless of the number of DIF items on the test.
+Interestingly, EMC had a better anchor hit rate on tests with more DIF items.
 This result appears to be driven by the clustering algorithm sometimes splitting all of the DIF-free items into two separate clusters, especially when most of the items are DIF-free.
 
-```{r anchorfalse, fig.cap = 'Mean false anchor rates across 100 runs for each number of DIF items', out.width="70%", warning = FALSE, message = FALSE}
+```{r anchorfalse, fig.cap = 'Mean false anchor rates across 100 runs for each AI method and number of DIF items.', out.width="70%", warning = FALSE, message = FALSE}
 one_to_anchors <- function(one, n_dif_items){
@@
           levels = c("AOAA_OAT_mod", "cluster_mod", "AOAA_AS_mod", "AOAA_mod"),
-          labels = c("AOAA-OAT", "CLUSTER", "AOAA-AS", "AOAA")
+          labels = c("AOAA-OAT", "EMC", "AOAA-AS", "AOAA")
         )
@@
 
-```{r anchorhit, fig.cap = 'Mean anchor hit rates across 100 runs for each number of DIF items', out.width="70%", warning = FALSE, message = FALSE}
+```{r anchorhit, fig.cap = 'Mean anchor hit rates across 100 runs for each AI method and number of DIF items.', out.width="70%", warning = FALSE, message = FALSE}
 temp %>%
@@
             levels = c("AOAA_OAT_mod", "cluster_mod", "AOAA_AS_mod", "AOAA_mod"),
-            labels = c("AOAA-OAT", "CLUSTER", "AOAA-AS", "AOAA")
+            labels = c("AOAA-OAT", "EMC", "AOAA-AS", "AOAA")
           )

### 2020-03-14 R-understanding-complete/gini_l1_understanding.Rmd
after deleting things and reorganizing - im happy with the result!

goal is to understand what types of selections gini and l1 are making

# simulated

```{r}
emlg <- read_rds(here("data-documentation/emlg.rds"))
```

the thing to notice is that

- gini takes the median of the dominant group (not totally clear how it chooses)

- l1 anywhere in the median (here its even so its a range)

```{r}
emlg %>% emlg_to_gini_l1()

sort(emlg$difference_in_easy)
```

# create

```{r}
emlg <- 
    tibble(
        ref_easy = c(0, 0, 0, 0, 0.5, 1, 1, 1, 1),
        foc_easy = rep(0, 9)
    ) %>% 
    mutate(difference_in_easy = ref_easy - foc_easy)
```

- gini again the median of the dominant group

- l1 chooses the median

```{r}
emlg %>% emlg_to_gini_l1()

sort(emlg$difference_in_easy)
```

# create

```{r}
emlg <- 
    tibble(
        ref_easy = seq(0, 1, length.out = 9),
        foc_easy = rep(0, 9)
    ) %>% 
    mutate(difference_in_easy = ref_easy - foc_easy)
```

- gini again the median of the dominant group but its not clear how it finds it here but it does

- l1 chooses the median as always

```{r}
emlg %>% emlg_to_gini_l1()

sort(emlg$difference_in_easy)
```

### 2020-03-14 _out_scratch.R

@@ -1,34 +1,0 @@
-# test run ----------------------------------------------------------------
-library(furrr)
-options(future.fork.enable = TRUE)
-plan(multiprocess)
-
-out <-
-    tibble(n_dif_items = c(4, 6, 8)) %>%
-    mutate(
-        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 2, n_dif_items = .))
-    )
-
-Sys.time()
-
-# dive into test ----------------------------------------------------------
-out$bigsim[[1]]$intuitive_mod[[1]] %>%
-    mod_intuitive_to_draws_df() %>%
-    draws_df_to_logit_plot() +
-    labs(
-        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
-    )
-
-out$bigsim[[1]]$pars[[1]]
-
-out$bigsim[[3]]$AOAA_OAT_final_dif[[1]] %>% View()
-
-debug(dif_AOAA_OAT)
-
-a <-
-    dif_AOAA_OAT(
-        data = out$bigsim[[3]]$sim[[1]]$data,
-        groups = out$bigsim[[3]]$sim[[1]]$groups
-    )
-
-

### 2020-03-14 data-documentation/data-documentation.R

@@ -1,30 +1,0 @@
-library(tidyverse)
-library(mirt)
-set.seed(123)
-R.utils::sourceDirectory("R")
-
-# simulate ----------------------------------------------------------------
-pars <-
-    tibble(
-        item = 1:8,
-        easy = rnorm(8, 0, 1),
-        target_disc = 1,
-        nuisance_disc = c(rep(0.5, 3), rep(0, 5)) # 1 + 7 = 8
-    ) %>%
-    arrange(desc(nuisance_disc), easy)
-
-sim <- sim_bias(pars = pars, n_ref = 10000, n_foc = 10000, mean_foc = -1)
-
-sim %>% write_rds("data-documentation/sim.rds")
-
-# dif ---------------------------------------------------------------------
-
-# intuitive
-mod <- multipleGroup(sim$data, 1, itemtype = "Rasch", sim$groups, invariance = "free_var", SE = TRUE)
-mod %>% mod_intuitive_to_draws_df() %>% draws_df_to_logit_plot()
-
-# all other
-all_other <- dif_all_other(sim$data, sim$groups)
-
-# iterative backward
-iterative_backward <- dif_iterative_backward(sim$data, sim$groups, all_other)

### 2020-03-19 paper/CopyOfpaper.Rmd
getting there

--- 2020-03-14 paper/paper.Rmd
@@
 However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF.
-We summarize existing assumptionless identification (AI) methods and propose a variety of new methods.
+We summarize existing agnostic identification (AI) methods and propose a variety of new methods.
 We conduct a simulation study — which we believe to be more realistic than most DIF simulation studies in the literature — and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
@@
 header-includes: 
+- \usepackage{lineno}
+- \linenumbers
+
 - \usepackage{amsmath}
@@
 
-\clearpage
+\begin{table}[H]
+\caption{Summary of agnostic identification (AI) methods}
+\centering
+\begin{tabular}{|p{4cm}|p{6cm}|p{4cm}|}
+\toprule
 
-# Introduction {#intro}
+Method & Description & Literature \\
 
-```{r child = "paper_1_intro.Rmd"}
-```
+\midrule
 
-# Assumptionless identification methods
+all-others-as-anchors (AOAA) & Test if each item has DIF by using all of the other items as anchors (not iterative).
+& Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
 
-```{r child = "paper_2a_anchor_items.Rmd"}
-```
+all-others-as-anchors-all-significant (AOAA-AS) & The first iteration is AOAA.
+All items that test positive for DIF are removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& Proposed by \cite{drasgow1987study} \\\hline
 
-## Anchor points
+all-others-as-anchors-one-at-a-time (AOAA-OAT) & The first iteration is AOAA.
+Only the item that shows the most extreme DIF is removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& To our knowledge, not proposed or used previously \\\hline
 
-```{r child = "paper_2b_anchor_points.Rmd"}
-```
+equal means clustering (EMC) & Cluster items based on differences in performance across groups and choose one of the clusters to be the anchor cluster.
+& Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} \\\hline
 
-# Simulation study
+equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
+Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
 
-```{r child = "paper_3_sim.Rmd"}
-```
+multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
 
-\clearpage
+maximizing Gini index (MAXGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
 
-<!-- weird formatting that isn't apa I don't think and indentation is weird, solution is to use papaja (look at paper_apa.Rmd) or this might also be helpful https://stackoverflow.com/questions/4073674/how-to-disable-indentation-in-particular-section-in-latex-->
+minimizing the area between curves (MINBC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
 
-<!-- hopefully all of this below goes away with papaja and other solutions. but remember if it can't editing the latek file is always an option on the final draft! 
-https://github.com/haozhu233/kableExtra/issues/214
-https://stackoverflow.com/questions/51613152/citing-within-an-rmarkdown-table
-https://tex.stackexchange.com/questions/272475/citations-in-tabular-environment-not-working
-https://tex.stackexchange.com/questions/359872/is-it-possible-to-use-cite-inside-xtable
-https://github.com/yihui/rmarkdown-cookbook/issues/71
--->
+\bottomrule
+\end{tabular}
+\label{table:allmethods}
+\end{table}
 

### 2020-03-19 paper/CopyOfpaper.tex
getting there

--- 2020-03-02 paper/paper.tex
@@
 \setcounter{secnumdepth}{5}
+\usepackage{lineno}
+\linenumbers
 \usepackage{amsmath}
@@
 \begin{abstract}
-summarize propose extensions insist as dif as student property and on communicating uncertainty
+It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
+However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF.
+We summarize existing agnostic identification (AI) methods and propose a variety of new methods.
+We conduct a simulation study --- which we believe to be more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that presents clearly all information about possible DIF, including sampling variability in item parameters, to the analyst.
+\clearpage
 \end{abstract}
@@
 }
-\clearpage
-
-\hypertarget{to-do}{%
-\section{TO DO}\label{to-do}}
-
-\begin{itemize}
-\tightlist
-\item
-  be sure all my directions match. i think i switched everything to be item easiness at some point but not sure if i did it throughout
-\end{itemize}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we think differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed.
-From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is a property of the student, not the item.
-For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.
-
-Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item.
-For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
-They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is \[P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)\] where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
-
-We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups.
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students.
-Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for the reference group and \({\eta_i}^{\text{ref}} = -1\) for the focal group.
-We now need to use a 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
-Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional 2PL model, the probability student \(i\) responding correctly to item \(j\) is
-\begin{align}
-    P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
-\end{align}
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability for students in the same group.
-
-If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model?
-Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities} {[}p.~144{]}.
-He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary {[}p.~144{]}.
-It's hard to imagine how one would have such knowledge.
-Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
-
-\hypertarget{dif-methods}{%
-\section{DIF methods}\label{dif-methods}}
-
-Psychometricians have long been in search of the perfect DIF detection method.
-In this section, we summarize existing DIF methods and propose extensions of those methods.
-We limit our discussion to IRT-based methods.
-As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).
-For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
-And, in search of a coherent framework, we sometimes edit names of existing methods.
-We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}) and that we risk contributing to the proliferation of names.
-
-We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
-Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{ref}} = -1\).
-The slope on target ability is set to \(a_{1j} = 1\) for all items, and the slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
-From the DIF-as-item-property view, there is no nuisance ability.
-Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.
-
-In general, we denote the mean ability of the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\), respectively.
-We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
-The fundamental challenge for a DIF method is to first get an estimate for \(\mu^\text{foc}\), \(\hat\mu^\text{foc}\).
-How can group ability be disentangled from the possibility that items are biased?
-The most common approach is to identify a group of anchor items that are assumed to be DIF-free.
-These items are used to get \(\hat\mu^\text{foc}\).
-
-\hypertarget{anchor-items}{%
-\subsection{Anchor items}\label{anchor-items}}
-
-\hypertarget{all-others-as-anchors}{%
-\subsubsection{All-others-as-anchors}\label{all-others-as-anchors}}
-
-Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA).
-AOAA tests each item for DIF one at a time using all of the other items as anchors.
-For example, when testing the first item for DIF, all of the other items are used as anchors.
-This is done using a likelihood ratio test that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}).
-Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
-The items for which the flexible model outperforms the baseline model (typically based on a chi-squared test) are identified as having DIF and the rest of the items become anchor items.
-
-Edelen et al.
-(\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can we be sure its those 10 items and not the other 11 items with DIF?
-We can't be.
-Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group and others are biased against the reference group.
-Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method (Strobl et al.
-\protect\hyperlink{ref-strobl2018anchor}{2018}).
-In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
-
-Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
-A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not.
-Using AOAA, all items test positive for DIF.
-The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}
-
-\hypertarget{all-others-as-anchors-all-significant}{%
-\paragraph{All-others-as-anchors-all-significant}\label{all-others-as-anchors-all-significant}}
-
-One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
-For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively --- with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model --- until no more items tested positively.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free.
-We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF being removed from the anchor set.
-
-AOAA-AS might seem like an improvement but it doesn't solve a fundamental problem of AOAA: What do we do when all items test positive for DIF?
-With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
-In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF and there are no anchor items.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items display DIF themselves.
-
-\hypertarget{all-other-as-anchors-one-at-a-time}{%
-\paragraph{All-other-as-anchors-one-at-a-time}\label{all-other-as-anchors-one-at-a-time}}
-
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been proposed.
-Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the \(\chi^2\) test statistic, is removed from the anchor set.
-The process continues iteratively until no new items display DIF.
-AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
-As a result, AOAA-OAT is less likely to arrive at the wrong conclusion because of artifical DIF.
-
-Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
-In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-AOAA-OAT has two assumptions: First, that at least one item does not have DIF, and second, that the majority of items do not have DIF.
-
-\hypertarget{summary-and-performance}{%
-\paragraph{Summary and performance}\label{summary-and-performance}}
-
-Table \ref{table:themethods} summarizes the three all-others-as-anchors methods.
-It's useful to remember that AOAA is not an iterative procedure, while the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.
-
 \begin{table}[H]
-\caption{EDIT THIS}
+\caption{Summary of agnostic identification methods}
 \centering
-\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
+\begin{tabular}{|p{4cm}|p{6cm}|p{4cm}|}
 \toprule
@@
 
+equal means clustering (EMC) & Cluster items based on differences in performance across groups and choose one of the clusters to be the anchor cluster.
+& Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} \\\hline
+
+equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
+Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
+
+multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
+
+maximizing Gini index (MAXGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
+
+minimizing the area between curves (MINBC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
+
 \bottomrule
 \end{tabular}
-\label{table:themethods}
+\label{table:allmethods}
 \end{table}
-
-{[}ADD IN SIMULATION RESULTS{]}
-
-\hypertarget{clustering-items}{%
-\subsubsection{Clustering items}\label{clustering-items}}
-
-Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster.} They pointed out that one way around the unidentifiability issue is to consider relative item parameters.
-For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\).
-The ultimate matrix of interest is \(\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}\) which is the \enquote{differences between groups in the pairwise differences in (easiness) between items} {[}p.~323{]}.
-
-The general idea of identifying clusters of items is intriguing.
-However, their approach is needlessly complicated and they did not describe a process for moving from \(\Delta \mathbf{R}\) to a set of anchor items.
-Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing one such process.
-\(\Delta \mathbf{R}\) is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
-Accordingly, they use k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC.
-They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.
-
-We propose a new cluster-based approach, which we call \enquote{equal means clustering} (EMC).
-Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\) of differences in item easiness, \(\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}\), where the model is identified by setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), thus the name \enquote{equal means clustering}).
-This results in all differences in performance --- either from group ability differences or DIF --- manifesting in the item easiness difference parameter \(\tilde{d_j}\).
-The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) arbitrarily set to \(0\).
-
-Instead of choosing k with BIC, we use the gap statistic method recommended by Hastie, Tibshirani, and Walther (\protect\hyperlink{ref-hastie2001estimating}{2001}).
-The largest cluster is chosen as the anchor cluster.
-If there is a tie for the largest cluster, the cluster with the lowest standard deviation of \(\tilde{\mathbf{d}}\) is selected.
-This process makes the assumption that the largest cluster (but necessarily the majority) of items do not contain DIF.
-
-OUR SIMULATION RESULTS
-
-\begin{tabular}{l|r|r|r|r|l}
-\hline
-item & a\_ref\_easy & b\_foc\_easy & difference\_in\_easy & cluster & anchor\\
-\hline
-item1 & -0.5638213 & -1.5520228 & 0.9882015 & 1 & TRUE\\
-\hline
-item2 & -0.2618317 & -1.2619495 & 1.0001178 & 1 & TRUE\\
-\hline
-item3 & 1.5747215 & 0.5456186 & 1.0291029 & 1 & TRUE\\
-\hline
-item4 & 0.0521493 & -0.9183001 & 0.9704494 & 1 & TRUE\\
-\hline
-item5 & 0.1335217 & -0.8599758 & 0.9934975 & 1 & TRUE\\
-\hline
-item6 & 1.6971526 & 0.1926118 & 1.5045408 & 2 & FALSE\\
-\hline
-item7 & 0.4629179 & -1.0572527 & 1.5201706 & 2 & FALSE\\
-\hline
-item8 & -1.2953061 & -2.8759098 & 1.5806036 & 2 & FALSE\\
-\hline
-\end{tabular}
-
-\hypertarget{analyst-chooses}{%
-\subsubsection{Analyst chooses}\label{analyst-chooses}}
-
-DIF methods are generally designed to automatically detect DIF without any human judgement.
-On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
-
-We propose a technique, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
-Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc}\) to \(0\).
-Again, the result is that all differences in performance manifest in \(\tilde{\mathbf{d}}\).
-To get a sense of variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are used to estimate the distribution of \(\tilde{d_j}\) for every item.
-These are the distributions displayed in the EM-MILG.
-The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
-
-Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data.
-It cannot be stated strongly enough that the EM-MILG contains all possible information about the difference in group performance.
-The challenge, then, is to select the anchor items.
-The analyst might assume that --- because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emmilg-1}
-
-}
-
-\caption{The equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
-\end{figure}
-
-After choosing the anchor items, the model is refit.
-The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc}\) to \(0\).
-The same process of using multiple imputations to estimate the distribution of \(d_j\) can be used with the new model.
-Because the equal means assumption is not made, we refer to the resulting graph as a multiple imputations logit graph (MILG).
-In our simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/milg-1}
-
-}
-
-\caption{The multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed to 0.}\label{fig:milg}
-\end{figure}
-
-One of our key concerns with typical DIF methods is that they can lull the analyst into a false sense of security.
-Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
-The EM-MILG combats this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even when other DIF methods are used, the analyst can use the EM-MILG as a first step in order to give them a sense of their item response data.
-And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
-
-\hypertarget{anchor-points}{%
-\subsection{Anchor points}\label{anchor-points}}
-
-The previously discussed DIF methods select a set of anchor items, whether that selection is done by an algorithm or the analyst.
-The anchor items are used to estimate \(\hat \mu^\text{foc}\).
-An alternative strategy is to directly select the anchor point, \(\mu^{\star\text{foc}}\).
-Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, all items can be tested for DIF.
-The question then becomes the following: How do we select the anchor point?
-
-\hypertarget{gini-index}{%
-\subsubsection{Gini index}\label{gini-index}}
-
-Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (\protect\hyperlink{ref-gini1912variabilita}{1912}) to select the anchor point.
-The Gini index is typically used to measure the inequality of wealth distribution in a country.
-For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution Chitiga, Sekyere, and Tsoanamatsie (\protect\hyperlink{ref-chitiga2015income}{2015}).
-In general, the Gini index \enquote{takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little} (Strobl et al.
-\protect\hyperlink{ref-strobl2018anchor}{2018}, 7).
-
-The anchor point, \(\mu^{\star\text{foc}}\), is selected by maximizing the Gini index.
-The intuition and assumption is that some items might have bias but the majority do not.
-Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\),
-\[
-\mu^{\star\text{foc}} = \mathop\mathrm{arg\,max}_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
-\]
-where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\mathbf{d}\).
-
-For the simulated data, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible values for \(\mu^\text{foc}\). The anchor point is \(\mu^\text{foc} = -0.99\).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginipath-1}
-
-}
-
-\caption{Selecting the anchor point by maximizing the Gini index.}\label{fig:ginipath}
-\end{figure}
-
-We then refit the model with the identifying assumption of \(\mu^{\star\text{foc}} = -0.99\). We test each item for DIF and display the results in a MILG as is shown in Figure \ref{fig:ginimilg}.
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.00544
-\end{verbatim}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginimilg-1}
-
-}
-
-\caption{Selecting the anchor point by maximizing the Gini index.}\label{fig:ginimilg}
-\end{figure}
-
-\hypertarget{minimizing-between-curves}{%
-\subsubsection{Minimizing between curves}\label{minimizing-between-curves}}
-
-Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curve, the function that maps ability to probability correct response, for each group:
-\[
-\text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
-\]
-Raju's area method, along with AOAA, have been cited as the two most commonly used IRT-based DIF detection methods (Magis et al.
-\protect\hyperlink{ref-magis2011generalized}{2011}).
-We see two weakness of Raju's area method: First, the item characteric curves still need to be linked by anchor items, which the method does not help to identify.
-Second, the area is unweighted so all values of \(\theta\) matter equally, despite some being much more realistic than others.
-
-Inspired by Raju's area method, we propose a new method, which we call \enquote{minimizing the area between curves} (MABC).
-To understand MABC, imagine multigroup item response data where \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\) so that there is no DIF.
-The fundamental identification problem of DIF in multigroup IRT models is that there are an infinite number of models with the same likelihood from which to choose.
-For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\mu^{\star\text{foc}} = 0\).
-The model would then estimate \(\hat d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF.
-Alternatively, we could assume that the focal group has \(\mu^{\star\text{foc}} = 3\).
-The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability but every item contains DIF against them.
-Both of these models have the same likelihood, so how should one choose which model to believe?
-MABC chooses the model with the least amount of total DIF, as measured by the total area between the item characteric curves.
-This is another, perhaps clearer, way of assuming that the majority of items do not contain DIF.
-
-Denote a function that estimates \(\hat b_j^\text{foc}\) by fitting a unidimensional Rasch model where the value of \(\mu^\text{foc}\) is fixed as \(m_j(\mu^\text{foc})\).
-The amount of DIF on each item is calculated as
-\[
-\text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
-\]
-where \(g(\theta)\) is a weighting function such that \(\int g(\theta)d\theta = 1\). The total DIF on the test, then, is
-\[
-\text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc})
-\]
-In this way, \(\text{Total DIF}(\mu^\text{foc})\) is a function where the input is \(\mu^\text{foc}\) and the output is the total amount of DIF on the test. MABC sets
-\[
-\mu^{\star\text{foc}} = \mathop\mathrm{arg\,min}_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
-\]
-MABC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}) who uses the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
-The selection of \(g(\theta)\) is important and determines where on the ability spectrum to consider DIF.
-(\protect\hyperlink{ref-chalmers2016might}{2016}) does not discuss how to choose \(g(\theta)\) and in practice uses \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which seems like an odd choice.
-It might seem intuitive to choose \(g(\theta) \sim N(0, 1)\) because \(\mu^\text{ref} = 0\), but this choice doesn't take into account the ability distribution of the focal group.
-If \(\mu^\text{foc} = 3\), wouldn't we also want to prioritize high \(\theta\) values?
-We set \(g(\theta)\) to be the average of the reference and focal group ability distributions:
-\[
-g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
-\]
-
-For the simulated data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\). The anchor point is \(\mu^\text{foc} = -1\).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/mabc-1}
-
-}
-
-\caption{Selecting the anchor point by minimizing the total amount of DIF.}\label{fig:mabc}
-\end{figure}
-
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
-
-{[}ADD A TABLE OF METHODS? OR THAT CAN SORT OF BE DONE IN THE SIMULATION RESULTS?{]}
-
-\hypertarget{next-steps}{%
-\section{Next steps}\label{next-steps}}
-
-\hypertarget{remember-my-bounds-idea-i-like-that}{%
-\section{Remember my bounds idea I like that}\label{remember-my-bounds-idea-i-like-that}}
-
-\hypertarget{simulate-and-then-walk-through-as-the-analyst-would-with-each-method}{%
-\section{Simulate and then walk through as the analyst would with each method}\label{simulate-and-then-walk-through-as-the-analyst-would-with-each-method}}
-
-\hypertarget{mike-frank-example}{%
-\section{Mike Frank example}\label{mike-frank-example}}
-
-We know that this test contains DIF.
-Which items depends on ability difference which we can be sure of.
-We can, however, quantify the amount of DIF on this test.
-We can take the SD or we can assume between and calculate chalmers statistic for each.
-We might not know for sure the ability difference but we can quantify the potential for DIF.
-For each we calculate the SD across item parameters.
-
-Something with DTF?
-
-Do a simulation where the second dimension varies - a more real testing ground
-
-Example with real data perhaps the Mike Frank data
-
-Summer learning loss as applied DIF?
-
-\clearpage
-
-\hypertarget{notes}{%
-\section{Notes}\label{notes}}
-
-\begin{itemize}
-\tightlist
-\item
-  adjust to all be easiness including multidimensional talk
-\end{itemize}
-
-\hypertarget{references}{%
-\section{References}\label{references}}
 
@@
 
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ackerman1992didactic}{}%
-Ackerman, Terry A. 1992. ``A Didactic Explanation of Item Bias, Item Impact, and Item Validity from a Multidimensional Perspective.'' \emph{Journal of Educational Measurement} 29 (1): 67--91.
-
-\leavevmode\hypertarget{ref-andrich2012real}{}%
-Andrich, David, and Curt Hagquist. 2012. ``Real and Artificial Differential Item Functioning.'' \emph{Journal of Educational and Behavioral Statistics} 37 (3): 387--416.
-
-\leavevmode\hypertarget{ref-bechger2015statistical}{}%
-Bechger, Timo M, and Gunter Maris. 2015. ``A Statistical Test for Differential Item Pair Functioning.'' \emph{Psychometrika} 80 (2): 317--40.
-
-\leavevmode\hypertarget{ref-camilli1992conceptual}{}%
-Camilli, Gregory. 1992. ``A Conceptual Analysis of Differential Item Functioning in Terms of a Multidimensional Item Response Model.'' \emph{Applied Psychological Measurement} 16 (2): 129--47.
-
-\leavevmode\hypertarget{ref-chalmers2018numerical}{}%
-Chalmers, R Philip. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
-
-\leavevmode\hypertarget{ref-chalmers2016might}{}%
-Chalmers, R Philip, Alyssa Counsell, and David B Flora.
-2016.
-``It Might Not Make a Big Dif: Improved Differential Test Functioning Statistics That Account for Sampling Variability.'' \emph{Educational and Psychological Measurement} 76 (1): 114--40.
-
-\leavevmode\hypertarget{ref-chitiga2015income}{}%
-Chitiga, Margaret, E Sekyere, and N Tsoanamatsie.
-2015.
-``Income Inequality and Limitations of the Gini Index: The Case of South Africa.'' \emph{Human Sciences Research Council (HSRC), Available at: Http://Www.
-Hsrc.
-Ac.
-Za/En/Review/Hsrc-Review-November-2014/Limitations-of-Gini-Index, Site Accessed} 2.
-
-\leavevmode\hypertarget{ref-drasgow1987study}{}%
-Drasgow, Fritz. 1987. ``Study of the Measurement Bias of Two Standardized Psychological Tests.'' \emph{Journal of Applied Psychology} 72 (1): 19.
-
-\leavevmode\hypertarget{ref-edelen2006identification}{}%
-Edelen, Maria Orlando, David Thissen, Jeanne A Teresi, Marjorie Kleinman, and Katja Ocepek-Welikson.
-2006.
-``Identification of Differential Item Functioning Using Item Response Theory and the Likelihood-Based Model Comparison Approach: Application to the Mini-Mental State Examination.'' \emph{Medical Care}, S134--S142.
-
-\leavevmode\hypertarget{ref-gini1912variabilita}{}%
-Gini, Corrado. 1912. ``Variabilità E Mutabilità (Variability and Mutability).'' \emph{Tipografia Di Paolo Cuppini, Bologna, Italy}, 156.
-
-\leavevmode\hypertarget{ref-hastie2001estimating}{}%
-Hastie, Trevor, Robert Tibshirani, and Guenther Walther. 2001. ``Estimating the Number of Data Clusters via the Gap Statistic.'' \emph{J Roy Stat Soc B} 63: 411--23.
-
-\leavevmode\hypertarget{ref-holland1986differential}{}%
-Holland, Paul W, and Dorothy T Thayer. 1986. ``Differential Item Functioning and the Mantel-Haenszel Procedure.'' \emph{ETS Research Report Series} 1986 (2): i--24.
-
-\leavevmode\hypertarget{ref-kopf2015framework}{}%
-Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.
-
-\leavevmode\hypertarget{ref-magis2011generalized}{}%
-Magis, David, Gilles Raı̂che, Sébastien Béland, and Paul Gérard.
-2011.
-``A Generalized Logistic Regression Procedure to Detect Differential Item Functioning Among Multiple Groups.'' \emph{International Journal of Testing} 11 (4): 365--86.
-
-\leavevmode\hypertarget{ref-meade2012solving}{}%
-Meade, Adam W, and Natalie A Wright. 2012. ``Solving the Measurement Invariance Anchor Item Problem in Item Response Theory.'' \emph{Journal of Applied Psychology} 97 (5): 1016.
-
-\leavevmode\hypertarget{ref-pohl2017cluster}{}%
-Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``CluStEr-baSEd anCHor Item Identification and Selection.'' NEPS Working Paper.
-
-\leavevmode\hypertarget{ref-raju1988area}{}%
-Raju, Nambury S. 1988. ``The Area Between Two Item Characteristic Curves.'' \emph{Psychometrika} 53 (4): 495--502.
-
-\leavevmode\hypertarget{ref-stark2006detecting}{}%
-Stark, Stephen, Oleksandr S Chernyshenko, and Fritz Drasgow.
-2006.
-``Detecting Differential Item Functioning with Confirmatory Factor Analysis and Item Response Theory: Toward a Unified Strategy.'' \emph{Journal of Applied Psychology} 91 (6): 1292.
-
-\leavevmode\hypertarget{ref-strobl2018anchor}{}%
-Strobl, Carolin, Julia Kopf, Raphael Hartmann, and Achim Zeileis. 2018. ``Anchor Point Selection: An Approach for Anchoring Without Anchor Items.'' Working Papers in Economics; Statistics.
-
-\leavevmode\hypertarget{ref-swaminathan1990detecting}{}%
-Swaminathan, Hariharan, and H Jane Rogers. 1990. ``Detecting Differential Item Functioning Using Logistic Regression Procedures.'' \emph{Journal of Educational Measurement} 27 (4): 361--70.
-
-\leavevmode\hypertarget{ref-talbot2013taking}{}%
-Talbot III, Robert M.
-2013.
-``Taking an Item-Level Approach to Measuring Change with the Force and Motion Conceptual Evaluation: An Application of Item Response Theory.'' \emph{School Science and Mathematics} 113 (7): 356--65.
-
-\leavevmode\hypertarget{ref-thissen1993detection}{}%
-Thissen, David, Lynne Steinberg, and Howard Wainer. 1993. ``Detection of Differential Item Functioning Using the Parameters of Item Response Models.''
-
-\leavevmode\hypertarget{ref-woods2009empirical}{}%
-Woods, Carol M. 2009. ``Empirical Selection of Anchors for Tests of Differential Item Functioning.'' \emph{Applied Psychological Measurement} 33 (1): 42--57.
-
-\leavevmode\hypertarget{ref-yang2012characterizing}{}%
-Yang, Ji Seung, Mark Hansen, and Li Cai. 2012. ``Characterizing Sources of Uncertainty in Item Response Theory Scale Scores.'' \emph{Educational and Psychological Measurement} 72 (2): 264--90.
-
 \end{document}

### 2020-03-19 paper/CopyOfpaper_2a_anchor_items.Rmd
getting there

--- 2020-03-14 paper/paper_2a_anchor_items.Rmd
@@
 Code is available at *(TODO)*.
+
+## The equal means, multiple imputation logit graph (EM-MILG)
+
+AI methods are generally designed to automatically detect DIF without any human judgement.
+On the other hand, it can be useful to present information to the analyst in a way that empowers them to visualize their data.
+
+We propose a new method, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
+Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$.
+Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$.
+To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical].
+Then, multiple imputations (MI) [@yang2012characterizing] are drawn to estimate the distribution of $\tilde{d_j}$ for each item.
+These are the distributions displayed in a EM-MILG.
+The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
+
+Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation.
+We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
+The challenge, then, is to select the anchor items.
+The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.
+
+```{r emmilg, fig.cap = 'A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.', out.width="70%", warning = FALSE, message = FALSE}
+out_3_biased_items$intuitive_mod[[1]] %>%
+    mod_intuitive_to_draws_df() %>%
+    draws_df_to_logit_plot() +
+    labs(
+        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
+    )
+
+# labeling good enough for now maybe check here to fiddle more:
+# https://stackoverflow.com/questions/8190087/italic-greek-letters-latex-style-math-in-plot-titles
+```
+
+After choosing the anchor items, the model is refit.
+The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc} = 0$.
+The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the new model.
+Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
+As expected, in our one-run simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
+
+```{r milg, fig.cap = 'A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group.
+Anchor items are fixed by setting $d_j = 0$.', out.width="70%", warning = FALSE, message = FALSE}
+# CAREFUL - this code assumes that the analyst would choose the same anchor items as AOAA-OAT found
+
+out_3_biased_items$AOAA_OAT_final_dif[[1]] %>% 
+    mutate(difference_in_easy = difference_in_easy - difference_in_easy[anchor][1]) %>% 
+    ggplot(aes(x = difference_in_easy, y = item)) +
+    ggridges::geom_density_ridges(
+        data = 
+            out_3_biased_items$AOAA_OAT_mod[[1]] %>% 
+            mod_to_draws_df(n = 1000) %>% 
+            select(-run) %>%
+            gather(var, val) %>% 
+            filter(val != 0),
+        aes(x = val, y = var)
+    ) +
+    geom_point() +
+    labs(
+        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$"),
+        y = ""
+    )
+```
+
+One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security.
+Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
+A EM-MILG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
+And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
+In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.
 
@@
 In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
-
-### The equal means, multiple imputation logit graph (EM-MILG)
-
-AI methods are generally designed to automatically detect DIF without any human judgement.
-On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
-
-We propose a new method, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
-Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$.
-Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$.
-To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical].
-Then, multiple imputations (MI) [@yang2012characterizing] are drawn to estimate the distribution of $\tilde{d_j}$ for each item.
-These are the distributions displayed in a EM-MILG.
-The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
-
-Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation.
-We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
-The challenge, then, is to select the anchor items.
-The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.
-
-```{r emmilg, fig.cap = 'A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.', out.width="70%", warning = FALSE, message = FALSE}
-out_3_biased_items$intuitive_mod[[1]] %>%
-    mod_intuitive_to_draws_df() %>%
-    draws_df_to_logit_plot() +
-    labs(
-        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
-    )
-
-# labeling good enough for now maybe check here to fiddle more:
-# https://stackoverflow.com/questions/8190087/italic-greek-letters-latex-style-math-in-plot-titles
-```
-
-After choosing the anchor items, the model is refit.
-The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc} = 0$.
-The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the new model.
-Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
-As expected, in our one-run simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
-
-```{r milg, fig.cap = 'A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group.
-Anchor items are fixed by setting $d_j = 0$.', out.width="70%", warning = FALSE, message = FALSE}
-# CAREFUL - this code assumes that the analyst would choose the same anchor items as AOAA-OAT found
-
-out_3_biased_items$AOAA_OAT_final_dif[[1]] %>% 
-    mutate(difference_in_easy = difference_in_easy - difference_in_easy[anchor][1]) %>% 
-    ggplot(aes(x = difference_in_easy, y = item)) +
-    ggridges::geom_density_ridges(
-        data = 
-            out_3_biased_items$AOAA_OAT_mod[[1]] %>% 
-            mod_to_draws_df(n = 1000) %>% 
-            select(-run) %>%
-            gather(var, val) %>% 
-            filter(val != 0),
-        aes(x = val, y = var)
-    ) +
-    geom_point() +
-    labs(
-        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$"),
-        y = ""
-    )
-```
-
-One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security.
-Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
-A EM-MILG combats this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
-And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
-In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.

### 2020-03-19 paper/paper.Rmd
after knitting and sending that copy to ben d and just before doing a reorganization to try to keep this repo wieldy
getting there

--- 2020-03-10 paper/paper.Rmd
@@
 title: |
-    New and old ways to chase a rabbit:  
-    Differential item functioning detection methods
+    Promise and peril: Agnostic identification methods  
+    for detecting differential item functioning
 subtitle: |
-    Ben Stenhaug  
+    Ben Stenhaug, Ben Domingue, and Mike Frank  
     Stanford University
 abstract: |
-  summarize propose extensions insist as dif as student property and on communicating uncertainty
-  <!-- https://tinyurl.com/ybremelq -->
+  It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
+However, to use an LRT, a model identifying assumption needs to be made so that differences in group ability can be disentangled from potential DIF.
+We use the term "agnostic identification" (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
+We summarize existing AI methods and propose a variety of new methods.
+We conduct a simulation study — which we argue is more realistic than most DIF simulation studies in the literature — and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
+\clearpage <!-- https://tinyurl.com/ybremelq -->
+
 header-includes: 
@@
 
-# DIF methods
+# Agnostic identification methods
 
@@
 
-# Next steps
+# Simulation study {#simstudy}
 
-```{r child = "paper_3_next_steps.Rmd"}
+```{r child = "paper_3_sim.Rmd"}
 ```

### 2020-03-19 paper/paper.tex
getting there

@@
 It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
-However, to use an LRT, the model needs to be identified so that differences in group ability can be disentangled from potential DIF.
-We summarize existing agnostic identification (AI) methods and propose a variety of new methods.
-We conduct a simulation study --- which we believe to be more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
-We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that presents clearly all information about possible DIF, including sampling variability in item parameters, to the analyst.
+However, to use an LRT, a model identifying assumption needs to be made so that differences in group ability can be disentangled from potential DIF.
+We use the term ``agnostic identification'' (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
+We summarize existing AI methods and propose a variety of new methods.
+We conduct a simulation study --- which we argue is more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
 \clearpage
@@
 
-Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we think of differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
+Following Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
 From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item.
 For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.
+In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies \enquote{DIF}.
 
-Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item.
+Despite this, the usual setup of DIF simulation studies frames DIF as a property of the item.
 For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
@@
 They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is \[P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)\] where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
+They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
+\begin{align}
+P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
+\end{align}
+where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
 
@@
 The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for all students in the reference group and \({\eta_i}^{\text{ref}} = -1\) for all students in the focal group.
+The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for all students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for all students in the focal group.
 We now need to use a compensatory 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional compensatory 2PL model, the probability student \(i\) responds correctly to item \(j\) is
-\[
+According to the two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}), the probability student \(i\) responds correctly to item \(j\) is
+\begin{align}
 \text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
-\]
+\end{align}
 This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
@@
 
-The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions --- about either relative group ability or which items might have DIF --- can be made.
+The unsolved problem that we address here is how to link groups in the typical setting in which no a prior assumptions --- about either relative group ability or which items might have DIF --- can be made.
 That is, we're interested in agnostic identification methods (hereafter referred to as \enquote{AI methods}) so that LRTs can be used to test items for DIF.
@@
 
-In this section, we summarize existing AI methods and propose extensions of those methods.
-We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+In this section, we summarize existing AI methods and propose a variety of extensions.
+We use a simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
@@
 Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.
+To be sure, this data is used only for demonstrative purposes.
+We compare the performance of the AI methods in a more realistic scenario in the \protect\hyperlink{simstudy}{the simulation study}.
 
@@
 We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
-The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of \(\hat\mu^\text{foc}\) from the estimation of \(\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\).
+The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of \(\hat\mu^\text{foc}\) from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness.
 The most common AI approach is to select a group of anchor items that are assumed to be DIF-free.
@@
 Code is available at \emph{(TODO)}.
+
+\hypertarget{emmilg}{%
+\subsection{The equal means, multiple imputation logit graph (EM-MILG)}\label{emmilg}}
+
+AI methods typically work like a black box.
+The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting the final model.
+These types of AI methods have their role --- and the main focus of this paper is on them --- but wouldn't it be nice if there was a way for the analyst to visualize the possibility of DIF in their data?
+
+We think so and, in response, propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
+EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc} = 0\).
+Recall that \(\mu^\text{ref} = 0\), so the assumption is that the groups have equal mean ability.
+The result is that all differences in performance --- either from group ability differences or DIF --- manifest in the item easiness difference parameter \(\tilde{d_j}\).
+The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) arbitrarily set to \(0\).
+
+To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
+Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are drawn to estimate the distribution of \(\tilde{d_j}\) for each item.
+These are the distributions displayed in a EM-MILG.
+The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
+
+Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data.
+We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
+The analyst might assume that --- because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased.
+These unbiased items are known as anchor items.
+After anchor items have been selected, the model is refit where the assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the assumption that the anchor items are DIF-free (i.e., \(d_j = 0\)).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emmilg-1}
+
+}
+
+\caption{A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
+\end{figure}
+
+The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with the final model.
+Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
+As expected, in our demonstration data, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/milg-1}
+
+}
+
+\caption{A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed by setting $d_j = 0$.}\label{fig:milg}
+\end{figure}
+
+In general, one of our key concerns with traditional AI methods is that they can lull the analyst into a false sense of security.
+Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality.
+A EM-MILG combats this concern by presenting all information clearly to the analyst.
+In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
+Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
+And, of course, the MILG can be used to visualize DIF regardless of how the model is identified.
+We now move on to more traditional AI methods.
 
@@
 How can they be sure that its those 10 items and not the other 11 items with DIF?
-They cannot be.
+We argue that this method cannot tell us with certainty that it is those 10 items that have DIF while the other 11 do not.
 Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
@@
 Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
-A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not.
+A simple thought experiment illustrates how AOAA fails: Imagine a three item test with a sufficiently large number of students where the first item has DIF, and the other two do not.
 Using AOAA, all items test positive for DIF.
@@
 For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively --- with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model --- until no more items tested positively.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free.
+Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse) referring to beginning with the assumption that all items are DIF-free.
 We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
@@
 Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items test positive for DIF, too.
+It's unclear if one should proceed if those four items also test positive for DIF.
 
@@
 
-Table \ref{table:themethods} summarizes the three all-others-as-anchors methods.
 It's useful to remember that AOAA is not an iterative procedure.
-The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method.
-In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
+In constrast, the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method.
+All of the methods described in this paper are summarized in Table \ref{table:allmethods}.
+For our demonstration data, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
 On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.
-
-\begin{table}[H]
-\caption{Summary of the three all-others-as-anchors agnostic identification methods}
-\centering
-\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
-\toprule
-
-Method & Description & Literature \\
-
-\midrule
-
-AOAA & Test if each item has DIF by using all of the other items as anchors (not iterative). & Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
-
-AOAA-AS & The first iteration is AOAA.
-All items that test positive for DIF are removed from the anchor set.
-Continue iterating until no new items test positive for DIF.
-& Proposed by \cite{drasgow1987study} \\\hline
-
-AOAA-OAT & The first iteration is AOAA.
-Only the item that shows the most extreme DIF is removed from the anchor set.
-Continue iterating until no new items test positive for DIF.
-& To our knowledge, not proposed or used previously \\
-
-\bottomrule
-\end{tabular}
-\label{table:themethods}
-\end{table}
 
@@
 Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster.} They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
-For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\).
+For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\), where \(x\) and \(y\) both index the items.
 The ultimate matrix of interest is \(\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}\) which is the \enquote{differences between groups in the pairwise differences in (easiness) between items} {[}p.~323{]}.
@@
 We propose a new cluster-based approach, which we call \enquote{equal means clustering} (EMC).
-Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\) of differences in item easiness, \(\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}\), where the model is identified by setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), thus the name \enquote{equal means clustering}).
-As a result, all differences in performance --- either from group ability differences or DIF --- manifesting in the item easiness difference parameter \(\tilde{d_j}\).
-The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) arbitrarily set to \(0\).
+Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\), which, as described in the \protect\hyperlink{emmilg}{EM-MILG section}, comes from setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), and thus the name \enquote{equal means clustering}).
 
@@
 This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF.
-In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
-
-\hypertarget{the-equal-means-multiple-imputation-logit-graph-em-milg}{%
-\subsubsection{The equal means, multiple imputation logit graph (EM-MILG)}\label{the-equal-means-multiple-imputation-logit-graph-em-milg}}
-
-AI methods are generally designed to automatically detect DIF without any human judgement.
-On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
-
-We propose a new method, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
-Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc}\) to \(0\).
-Again, the result is that all differences in performance manifest in \(\tilde{\mathbf{d}}\).
-To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are drawn to estimate the distribution of \(\tilde{d_j}\) for each item.
-These are the distributions displayed in a EM-MILG.
-The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
-
-Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation.
-We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
-The challenge, then, is to select the anchor items.
-The analyst might assume that --- because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emmilg-1}
-
-}
-
-\caption{A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
-\end{figure}
-
-After choosing the anchor items, the model is refit.
-The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc} = 0\).
-The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with the new model.
-Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
-As expected, in our one-run simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/milg-1}
-
-}
-
-\caption{A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed by setting $d_j = 0$.}\label{fig:milg}
-\end{figure}
-
-One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security.
-Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
-A EM-MILG combats this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
-And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
-In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.
+For our demonstration data, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
 
@@
 The anchor items are used to estimate \(\hat \mu^\text{foc}\).
-An alternative strategy is to directly set the anchor point, \(\mu^{\star\text{foc}}\).
+An alternative strategy is to directly set the anchor point (Strobl et al.
+\protect\hyperlink{ref-strobl2018anchor}{2018}).
 Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF.
@@
 Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
-\[
+\begin{align}
 \mu^{\star\text{foc}} = \mathop\mathrm{arg\,max}_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
-\]
+\end{align}
 where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\mathbf{d}\).
 
-For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values.
+For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values.
 The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is extraordinarily close to the data-generating value of -1.
@@
 Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
-\[
+\begin{align}
 \text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
-\]
+\end{align}
 Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods (Magis et al.
@@
 The amount of DIF on each item is calculated as
-\[
+\begin{align}
 \text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
-\]
+\end{align}
 where \(g(\theta)\) is a weighting function such that \(\int g(\theta)d\theta = 1\). The total DIF on the test, then, is
-\[
+\begin{align}
 \text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc})
-\]
+\end{align}
 In this way, \(\text{Total DIF}(\mu^\text{foc})\) is a function where the input is \(\mu^\text{foc}\) and the output is the total amount of DIF on the test. MINBC sets
-\[
+\begin{align}
 \mu^{\star\text{foc}} = \mathop\mathrm{arg\,min}_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
-\]
+\end{align}
+
 MINBC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}), who use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
@@
 Accordingly, we set \(g(\theta)\) to be the average of the reference and focal group ability probability density functions:
-\[
+\begin{align}
 g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
-\]
+\end{align}
 
-For our one-run simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\).
+For our demonstration data simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\).
 In this case, MINBC works perfectly and the anchor point is found to be \(\mu^{\star\text{foc}} = -1\).
@@
 
-We have described a variety of AI methods and their corresponding acronymns.
+We described a variety of AI methods and their corresponding acronymns.
 Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point.
+MILG and EM-MILG are somewhat of outliers in that they are methods for visualizing DIF or potential DIF.
 Table \ref{table:allmethods} summarizes all of these methods.
@@
 \begin{table}[H]
-\caption{Summary of agnostic identification methods}
+\caption{Summary of agnostic identification (AI) methods}
 \centering
@@
 
-all-others-as-anchors (AOAA) & See Table \ref{table:themethods} for a summary of all three AOAA methods. &  \\\hline
+equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
+Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples by \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
+
+multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
+
+all-others-as-anchors (AOAA) & Test if each item has DIF by using all of the other items as anchors (not iterative).
+& Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
+
+all-others-as-anchors-all-significant (AOAA-AS) & The first iteration is AOAA.
+All items that test positive for DIF are removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& Proposed by \cite{drasgow1987study} \\\hline
+
+all-others-as-anchors-one-at-a-time (AOAA-OAT) & The first iteration is AOAA.
+Only the item that shows the most extreme DIF is removed from the anchor set.
+Continue iterating until no new items test positive for DIF.
+& To our knowledge, not proposed or used previously \\\hline
 
@@
 & Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} \\\hline
-
-equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
-Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
-
-multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
 
@@
 
-\hypertarget{simulation-study}{%
-\section{Simulation study}\label{simulation-study}}
+\hypertarget{simstudy}{%
+\section{Simulation study}\label{simstudy}}
 
@@
 For example, it's hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability.
-Accordingly, we generate item responses using a simplified version of Sympson's (\protect\hyperlink{ref-sympson1978model}{1978}) noncompensatory item response model in which
-\[
+Accordingly, we generated item responses using a simplified version of Sympson's (\protect\hyperlink{ref-sympson1978model}{1978}) noncompensatory item response model in which
+\begin{align}
 \text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{2j}) = \sigma(\theta_i) \cdot \sigma(a_{2j}\eta_i)
-\]
+\end{align}
 where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(\sigma(a_{2j})\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
@@
 
-In each run, we simulate 10,000 students with half coming from each of the reference and focal groups.
+In each run, we simulated 10,000 students with half coming from each of the reference and focal groups.
 For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean {[}\(\mu_\theta^\text{ref} = 0\), \(\mu_\eta^\text{ref} = 0\){]} and covariance matrix \(\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}\).
@@
 
-The test always has 12 items, but we vary the number of items with DIF from two to six.
+The test always has 12 items, but we varied the number of items with DIF from two to six.
 For items without DIF, \(a_{2j} = \infty\) so that the model reduces to \(\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)\).
 For items with DIF, \(a_{2j}\) is calculated based on Ackerman's (\protect\hyperlink{ref-ackerman1994using}{1994}) angle equation as described in Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}):
-\[
+\begin{align}
 \angle_j = \arccos \dfrac{a_{1j}^2}{a_{1j}^2 + a_{2j}^2}.
-\]
+\end{align}
 An item's angle measures the relative loading of the item on the two dimensions.
@@
 Our simple noncompensatory model has \(a_{1j} = 1\) for all items so the angle equation reduces to
-\[
+\begin{align}
 \angle_j = \arccos \dfrac{1}{1 + a_{2j}^2}.
-\]
+\end{align}
 We are interested in specifying the angle of an item, so the relevant equation becomes
-\[
+\begin{align}
 a_{2j} = \sqrt{\dfrac{1 - \cos(\angle_j)^2}{\cos(\angle_j)^2}}.
-\]
+\end{align}
 For DIF items, we set \(a_{2j}\) based on angles with equal intervals between 20\(^\circ\) and 60\(^\circ\).
@@
 Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method was
-\[
+\begin{align}
 \dfrac{\text{Total number of DIF items in the anchor set}}{\text{Number of DIF items on each test} \cdot \text{Number of runs}} = \dfrac{60}{2 \cdot 100} = 30\%.
-\]
+\end{align}
 The fact that the item with 20\(^\circ\) of DIF is most commonly incorrectly included in the anchor set is what drove the counterintuitive result that the mean false anchor rate decreases with more DIF items.
@@
 
+\leavevmode\hypertarget{ref-thissen1986taxonomy}{}%
+Thissen, David, and Lynne Steinberg. 1986. ``A Taxonomy of Item Response Models.'' \emph{Psychometrika} 51 (4): 567--77.
+
 \leavevmode\hypertarget{ref-thissen1993detection}{}%

### 2020-03-19 paper/paper_1_intro.Rmd
getting there

@@
-Inspired by @camilli1992conceptual, we think of differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
+Following @camilli1992conceptual, we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
 From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item.
 For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.
+In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies "DIF".
 
-Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item.
+Despite this, the usual setup of DIF simulation studies frames DIF as a property of the item.
 For example, @kopf2015framework simulate students as belonging to either a reference or focal group.
@@
 They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is $$P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)$$ where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.
+They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is
+\begin{align}
+P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
+\end{align}
+where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.
+
+<!-- future improvement would be 1) show that for some student they equal just to make it ultra clear and 2) note that there are infinite possibilities the key is that the second term is -0.6 for students from the focal group -->
 
@@
 The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for all students in the reference group and ${\eta_i}^{\text{ref}} = -1$ for all students in the focal group.
+The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for all students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for all students in the focal group.
 We now need to use a compensatory 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional compensatory 2PL model, the probability student $i$ responds correctly to item $j$ is
-$$
+According to the two-dimensional compensatory 2PL model [@thissen1986taxonomy], the probability student $i$ responds correctly to item $j$ is
+\begin{align}
 \text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
-$$
+\end{align}
 This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.

### 2020-03-19 paper/paper_2a_anchor_items.Rmd
getting there

@@
 
-The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items might have DIF — can be made.
+The unsolved problem that we address here is how to link groups in the typical setting in which no a prior assumptions — about either relative group ability or which items might have DIF — can be made.
 That is, we're interested in agnostic identification methods (hereafter referred to as "AI methods") so that LRTs can be used to test items for DIF.
@@
 
-In this section, we summarize existing AI methods and propose extensions of those methods.
-We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+In this section, we summarize existing AI methods and propose a variety of extensions.
+We use a simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
@@
 Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.
+To be sure, this data is used only for demonstrative purposes.
+We compare the performance of the AI methods in a more realistic scenario in the [the simulation study](#simstudy).
 
@@
 We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
-The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of $\hat\mu^\text{foc}$ from the estimation of $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$.
+The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of $\hat\mu^\text{foc}$ from the estimation of $\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$, the difference in item easiness.
 The most common AI approach is to select a group of anchor items that are assumed to be DIF-free.
@@
 Code is available at *(TODO)*.
-
-## Anchor items {#anchoritems}
-
-### All-others-as-anchors (AOAA)
-
-@meade2012solving compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method.
-AOAA tests each item for DIF one at a time using all of the other items as anchors.
-For example, when testing the first item for DIF, all of the other items are used as anchors.
-This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection].
-Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
-The items for which the flexible model outperforms the baseline model (typically based on a $\chi^2$ test) are identified as having DIF, and the rest of the items become anchor items.
-AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
-
-@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can they be sure that its those 10 items and not the other 11 items with DIF?
-They cannot be.
-Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
-Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor].
-In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
-
-Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies [@stark2006detecting].
-A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not.
-Using AOAA, all items test positive for DIF.
-The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF."
-
-#### All-others-as-anchors-all-significant (AOAA-AS)
-
-One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
-For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively.
-@kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free.
-We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
-AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.
-
-AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF?
-With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
-In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items.
-@kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-@woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
-It's unclear if one should proceed if those four items test positive for DIF, too.
-
-#### All-other-as-anchors-one-at-a-time (AOAA-OAT)
-
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
-AOAA-OAT is inspired by @hagquist2017recent, who, in general, assert that "items showing DIF initially should not be resolved simultaneously but sequentially" [p.
-6].
-Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the $\chi^2$ test statistic, is removed from the anchor set.
-The process continues iteratively until no new items display DIF.
-AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
-As a result, we believe that AOAA-OAT is less likely to be "fooled" by artifical DIF.
-Note that AOAA-OAT is not currently implemented in the R package mirt.
-
-Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
-In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-To work, AOAA-OAT has two requirements: First, that at least two items do not have DIF, and second, that the set of items without DIF are more homogeneous than other sets of items.
-
-#### Summary and performance
-
-Table \ref{table:themethods} summarizes the three all-others-as-anchors methods.
-It's useful to remember that AOAA is not an iterative procedure.
-The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method.
-In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
-On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.
-
-<!-- the two kopf 2015 papers love the iterative forward SA. I bet AOAA-OAT is as good or better while fitting fewer models -->
-
-\begin{table}[H]
-\caption{Summary of the three all-others-as-anchors agnostic identification methods}
-\centering
-\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
-\toprule
-
-Method & Description & Literature \\
-
-\midrule
-
-AOAA & Test if each item has DIF by using all of the other items as anchors (not iterative). & Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
-
-AOAA-AS & The first iteration is AOAA.
-All items that test positive for DIF are removed from the anchor set.
-Continue iterating until no new items test positive for DIF.
-& Proposed by \cite{drasgow1987study} \\\hline
-
-AOAA-OAT & The first iteration is AOAA.
-Only the item that shows the most extreme DIF is removed from the anchor set.
-Continue iterating until no new items test positive for DIF.
-& To our knowledge, not proposed or used previously \\
-
-\bottomrule
-\end{tabular}
-\label{table:themethods}
-\end{table}
 
@@
 
-### Equal means clustering (EMC)
+## The equal means, multiple imputation logit graph (EM-MILG) {#emmilg}
 
-@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
-For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$.
-The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p.
-323].
+AI methods typically work like a black box.
+The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting the final model.
+These types of AI methods have their role — and the main focus of this paper is on them — but wouldn't it be nice if there was a way for the analyst to visualize the possibility of DIF in their data?
 
-The general idea of identifying clusters of items is intriguing.
-However, their approach is needlessly complicated, and they did not describe a process for moving from $\Delta \mathbf{R}$ to an anchor cluster.
-@pohl2017cluster extended their work by proposing one such process.
-$\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
-Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
-They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.
-
-We propose a new cluster-based approach, which we call "equal means clustering" (EMC).
-Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$ of differences in item easiness, $\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}$, where the model is identified by setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, thus the name "equal means clustering").
-As a result, all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$.
+We think so and, in response, propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
+EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc} = 0$.
+Recall that $\mu^\text{ref} = 0$, so the assumption is that the groups have equal mean ability.
+The result is that all differences in performance — either from group ability differences or DIF — manifest in the item easiness difference parameter $\tilde{d_j}$.
 The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.
 
-Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating.
-The largest cluster is chosen as the anchor cluster.
-If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected.
-This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF.
-In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
-
-### The equal means, multiple imputation logit graph (EM-MILG)
-
-AI methods are generally designed to automatically detect DIF without any human judgement.
-On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.
-
-We propose a new method, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
-Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$.
-Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$.
 To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical].
@@
 
-Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation.
+Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data.
 We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
-The challenge, then, is to select the anchor items.
-The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.
+The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased.
+These unbiased items are known as anchor items.
+After anchor items have been selected, the model is refit where the assumption that $\mu^\text{foc} = 0$ is dropped and replaced with the assumption that the anchor items are DIF-free (i.e., $d_j = 0$).
 
@@
 
-After choosing the anchor items, the model is refit.
-The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc} = 0$.
-The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the new model.
+The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the final model.
 Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
-As expected, in our one-run simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
+As expected, in our demonstration data, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
 
@@
 
-One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security.
-Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items.
+In general, one of our key concerns with traditional AI methods is that they can lull the analyst into a false sense of security.
+Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality.
 A EM-MILG combats this concern by presenting all information clearly to the analyst.
@@
 Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
-And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.
-In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.
+And, of course, the MILG can be used to visualize DIF regardless of how the model is identified.
+We now move on to more traditional AI methods.
+
+## Anchor items {#anchoritems}
+
+### All-others-as-anchors (AOAA)
+
+@meade2012solving compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method.
+AOAA tests each item for DIF one at a time using all of the other items as anchors.
+For example, when testing the first item for DIF, all of the other items are used as anchors.
+This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection].
+Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
+The items for which the flexible model outperforms the baseline model (typically based on a $\chi^2$ test) are identified as having DIF, and the rest of the items become anchor items.
+AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
+
+@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
+How can they be sure that its those 10 items and not the other 11 items with DIF?
+We argue that this method cannot tell us with certainty that it is those 10 items that have DIF while the other 11 do not.
+Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
+More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
+Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor].
+In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
+
+Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies [@stark2006detecting].
+A simple thought experiment illustrates how AOAA fails: Imagine a three item test with a sufficiently large number of students where the first item has DIF, and the other two do not.
+Using AOAA, all items test positive for DIF.
+The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF."
+
+#### All-others-as-anchors-all-significant (AOAA-AS)
+
+One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
+For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively.
+@kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse) referring to beginning with the assumption that all items are DIF-free.
+We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
+Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
+AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.
+
+AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF?
+With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
+In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items.
+@kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
+@woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
+It's unclear if one should proceed if those four items also test positive for DIF.
+
+#### All-other-as-anchors-one-at-a-time (AOAA-OAT)
+
+We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
+AOAA-OAT is inspired by @hagquist2017recent, who, in general, assert that "items showing DIF initially should not be resolved simultaneously but sequentially" [p.
+6].
+Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the $\chi^2$ test statistic, is removed from the anchor set.
+The process continues iteratively until no new items display DIF.
+AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
+As a result, we believe that AOAA-OAT is less likely to be "fooled" by artifical DIF.
+Note that AOAA-OAT is not currently implemented in the R package mirt.
+
+<!-- improvement would be to explain what more homogeneous means in more detail -->
+
+Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
+The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
+In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
+To work, AOAA-OAT has two requirements: First, that at least two items do not have DIF, and second, that the set of items without DIF are more homogeneous than other sets of items.
+
+#### Summary and performance
+
+<!-- the two kopf 2015 papers love the iterative forward SA. I bet AOAA-OAT is as good or better while fitting fewer models -->
+
+It's useful to remember that AOAA is not an iterative procedure.
+In constrast, the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method.
+All of the methods described in this paper are summarized in Table \ref{table:allmethods}.
+For our demonstration data, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
+On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.
+
+<!-- probably should add transition here -->
+
+### Equal means clustering (EMC)
+
+<!-- ben d rightly points out that it would be good to say more what we mean by "similarly" - could look at how they describe it -->
+
+@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
+For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$, where $x$ and $y$ both index the items.
+The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p.
+323].
+
+The general idea of identifying clusters of items is intriguing.
+However, their approach is needlessly complicated, and they did not describe a process for moving from $\Delta \mathbf{R}$ to an anchor cluster.
+@pohl2017cluster extended their work by proposing one such process.
+$\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
+Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
+They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
+Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.
+
+We propose a new cluster-based approach, which we call "equal means clustering" (EMC).
+Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$, which, as described in the [EM-MILG section](#emmilg), comes from setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, and thus the name "equal means clustering").
+
+Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating.
+The largest cluster is chosen as the anchor cluster.
+If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected.
+This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF.
+For our demonstration data, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.

### 2020-03-19 paper/paper_2b_anchor_points.Rmd
getting there

@@
 The anchor items are used to estimate $\hat \mu^\text{foc}$.
-An alternative strategy is to directly set the anchor point, $\mu^{\star\text{foc}}$.
+An alternative strategy is to directly set the anchor point [@strobl2018anchor].
 Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF.
@@
 Denoting a function that calculates the Gini index from a vector of non-negative elements as  $G(\mathbf{x})$, MGI sets <!-- its weird to arg max over a parameter we already have a definition for - maybe switch to t or something -->
-$$
+\begin{align}
 \mu^{\star\text{foc}} = \argmax_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
-$$
+\end{align}
 where $\mu^\text{foc} \in (-\infty, \infty)$ and $\mu^\text{foc}$ is added to each element of $\mathbf{d}$.
@@
 
-For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible $\mu^\text{foc}$ values.
+For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible $\mu^\text{foc}$ values.
 The result of MAXGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$, which is extraordinarily close to the data-generating value of -1.
@@
 
+```{r}
+# not ideal that i input -anchor_point here but can fix later
+# min_bc <-
+#     out_3_biased_items$intuitive_mod[[1]] %>%
+#     min_between_curves(lo = 0, hi = 2, by = 0.001)
+# min_bc %>% write_rds(here("paper", "paper_2_min_bc.rds"))
+
+min_bc <- read_rds(here("paper", "paper_2_min_bc.rds"))
+```
+
 Raju's area method [@raju1988area] measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
-$$
+\begin{align}
 \text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
-$$
+\end{align}
 Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods [@magis2011generalized].
@@
 The amount of DIF on each item is calculated as
-$$
+\begin{align}
 \text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
-$$
+\end{align}
 where $g(\theta)$ is a weighting function such that $\int g(\theta)d\theta = 1$. The total DIF on the test, then, is
-$$
+\begin{align}
 \text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc})
-$$
+\end{align}
 In this way, $\text{Total DIF}(\mu^\text{foc})$ is a function where the input is $\mu^\text{foc}$ and the output is the total amount of DIF on the test. MINBC sets
-$$
+\begin{align}
 \mu^{\star\text{foc}} = \argmin_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
-$$
+\end{align}
+
 MINBC is inspired in part by @chalmers2016might, who use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF).
@@
 Accordingly, we set $g(\theta)$ to be the average of the reference and focal group ability probability density functions:
-$$
+\begin{align}
 g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
-$$
+\end{align}
 
-```{r}
-# not ideal that i input -anchor_point here but can fix later
-# min_bc <-
-#     out_3_biased_items$intuitive_mod[[1]] %>%
-#     min_between_curves(lo = 0, hi = 2, by = 0.001)
-# min_bc %>% write_rds(here("paper", "paper_2_min_bc.rds"))
-
-min_bc <- read_rds(here("paper", "paper_2_min_bc.rds"))
-```
-
-For our one-run simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$.
+For our demonstration data simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$.
 In this case, MINBC works perfectly and the anchor point is found to be $\mu^{\star\text{foc}} = `r round(min_bc$between_curves_anchor_points, 2)`$.

### 2020-03-19 paper/paper_3_sim.Rmd
getting there

@@
 For example, it’s hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability.
-Accordingly, we generate item responses using a simplified version of Sympson’s -@sympson1978model noncompensatory item response model in which
-$$
+Accordingly, we generated item responses using a simplified version of Sympson’s -@sympson1978model noncompensatory item response model in which
+\begin{align}
 \text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{2j}) = \sigma(\theta_i) \cdot \sigma(a_{2j}\eta_i)
-$$
+\end{align}
 where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $\sigma(a_{2j})$ is the item's loading on nuisance ability [@demars2016partially].
@@
 
-In each run, we simulate 10,000 students with half coming from each of the reference and focal groups.
+In each run, we simulated 10,000 students with half coming from each of the reference and focal groups.
 For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean [$\mu_\theta^\text{ref} = 0$, $\mu_\eta^\text{ref} = 0$] and covariance matrix $\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}$.
@@
 
-The test always has 12 items, but we vary the number of items with DIF from two to six.
+The test always has 12 items, but we varied the number of items with DIF from two to six.
 For items without DIF, $a_{2j} = \infty$ so that the model reduces to $\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)$.
 For items with DIF, $a_{2j}$ is calculated based on Ackerman's -@ackerman1994using angle equation as described in @walker2017using:
-$$
+\begin{align}
 \angle_j = \arccos \dfrac{a_{1j}^2}{a_{1j}^2 + a_{2j}^2}.
-$$
+\end{align}
 An item's angle measures the relative loading of the item on the two dimensions.
@@
 Our simple noncompensatory model has $a_{1j} = 1$ for all items so the angle equation reduces to
-$$
+\begin{align}
 \angle_j = \arccos \dfrac{1}{1 + a_{2j}^2}.
-$$
+\end{align}
 We are interested in specifying the angle of an item, so the relevant equation becomes
-$$
+\begin{align}
 a_{2j} = \sqrt{\dfrac{1 - \cos(\angle_j)^2}{\cos(\angle_j)^2}}.
-$$
+\end{align}
 For DIF items, we set $a_{2j}$ based on angles with equal intervals between 20$^\circ$ and 60$^\circ$.
@@
 Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method was
-$$
+\begin{align}
 \dfrac{\text{Total number of DIF items in the anchor set}}{\text{Number of DIF items on each test} \cdot \text{Number of runs}} = \dfrac{60}{2 \cdot 100} = 30\%.
-$$
+\end{align}
 The fact that the item with 20$^\circ$ of DIF is most commonly incorrectly included in the anchor set is what drove the counterintuitive result that the mean false anchor rate decreases with more DIF items.

### 2020-03-20 paper/paper.tex
got another draft to ben

@@
 It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
-However, to use an LRT, a model identifying assumption needs to be made so that differences in group ability can be disentangled from potential DIF.
+However, to use an LRT, a model identifying assumption must be made so that differences in group ability can be disentangled from potential DIF.
 We use the term ``agnostic identification'' (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
@@
 And, in search of a coherent framework, we sometimes edit names of existing methods.
-We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.
+We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.
 
 In this section, we summarize existing AI methods and propose a variety of extensions.
-We use a simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+We use simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
@@
 We think so and, in response, propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
-EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc} = 0\).
-Recall that \(\mu^\text{ref} = 0\), so the assumption is that the groups have equal mean ability.
+EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting \(\mu^\text{foc} = 0\).
+Recall that \(\mu^\text{ref} = 0\), so this implies that the groups have equal mean ability.
 The result is that all differences in performance --- either from group ability differences or DIF --- manifest in the item easiness difference parameter \(\tilde{d_j}\).
-The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) arbitrarily set to \(0\).
+The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\) --- as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from the final model.
 
@@
 And, of course, the MILG can be used to visualize DIF regardless of how the model is identified.
-We now move on to more traditional AI methods.
+We now move on to more traditional AI methods, which don't require judgment from the analyst.
 
@@
 We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
+Appending \enquote{all-significant} indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
 AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.
@@
 The process continues iteratively until no new items display DIF.
-AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
+AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only \emph{one} item in each iteration as opposed to \emph{all} items that test positive for DIF.
 As a result, we believe that AOAA-OAT is less likely to be \enquote{fooled} by artifical DIF.
-Note that AOAA-OAT is not currently implemented in the R package mirt.
+AOAA-OAT is not currently implemented in the R package mirt.
 
@@
 For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\), where \(x\) and \(y\) both index the items.
+To be sure, \(\mathbf{R}^{\text{ref}}\) is a square with the number of rows and columns equal to the number of items.
 The ultimate matrix of interest is \(\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}\) which is the \enquote{differences between groups in the pairwise differences in (easiness) between items} {[}p.~323{]}.
@@
 \(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (thus the abbreviation MAXGI).
-The intuition and assumption is that the items without DIF are the most homogeneous.
+The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF.
 Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
@@
 \end{align}
-where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\mathbf{d}\).
+where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\tilde{\mathbf{d}}\).
 
@@
 The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is extraordinarily close to the data-generating value of -1.
+Moreover, there is a local maximum at \(\mu^{\text{foc}} \approx 1.5\), which corresponds to the items with DIF.
+This illustrates that --- as Strobl et al.
+(\protect\hyperlink{ref-strobl2018anchor}{2018}) point out --- the search path is informative.
 
@@
 To adapt Raju's area method into an AI method, we propose a new method, which we call \enquote{minimizing the area between curves} (MINBC).
-To understand MINBC, imagine a scenario in which the data-generating process is \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\), so that there is no DIF.
+To understand MINBC, imagine a scenario in which the data-generating process is \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\), so that the groups have equal ability and there is no DIF.
 The fundamental identification problem is that there are an infinite number of models with the same likelihood from which to choose.
@@
 \end{align}
-where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(\sigma(a_{2j})\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
+where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(a_{2j}\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
 
@@
 An item's angle measures the relative loading of the item on the two dimensions.
-For example, an angle of 45\(^\circ\) indicates that the item loads equally on the target ability and nuisance ability.
+The lower the angle, the more the item loads on target ability as compared to nuisance ability, which corresponds to less DIF.
+An angle of 45\(^\circ\) indicates that the item loads equally on the target ability and nuisance ability.
 Our simple noncompensatory model has \(a_{1j} = 1\) for all items so the angle equation reduces to
@@
 
-\caption{For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.}\label{fig:simemmlg}
+\caption{For a 12 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.}\label{fig:simemmlg}
 \end{figure}
@@
 We refer to this quantity as the achievement gap.
-Recall that all models set \(\mu_\theta^\text{ref} = 0\), so the achievement gap reduces to \(\mu_\theta^\text{ref}\).
-The data-generating value of \(\mu_\theta^\text{foc}\) is 0.5, but each run will include sampling variability.
+Recall that all models set \(\mu_\theta^\text{ref} = 0\), so the achievement gap reduces to \(\mu_\theta^\text{foc}\).
+The data-generating value of \(\mu_\theta^\text{foc}\) is 0.5, but each run, of course, includes sampling variability.
 To get at the heart of how well a method is doing, we calculated the achievement gap residual as the method's estimated achievement gap, \(\hat\mu_\theta^\text{foc}\), minus the achievement gap estimated when using only the DIF-free items as anchors.
@@
 We used rank correlation as opposed to the more typically used RMSE in agreement with Lord's (\protect\hyperlink{ref-lord1986maximum}{1986}) argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
-\emph{(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary.
-maybe this idea of rank order being better than RMSE is for another paper)}
+\emph{(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary)}
 
@@
 
-In particular, AOAA-OAT was the clear winner, and we recommend its use whenever possible.
+The fact that the AOAA-OAT exhibited superior performance is an important finding given that it is, to our knowledge, not used currently.
+We recommend its use whenever possible.
 One reason AOAA-OAT might not always be possible is that it can be computationally expensive.
@@
 For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
-Furthermore, our work focused on the Rasch model, and it will be of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.
+Furthermore, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.
 

### 2020-03-20 paper/paper_2a_anchor_items.Rmd
got another draft to ben

@@
 And, in search of a coherent framework, we sometimes edit names of existing methods.
-We recognize that others have done the same [e.g.
-@kopf2015framework], and that we risk contributing to a proliferation of names.
+We recognize that others have done the same [e.g., @kopf2015framework], and that we risk contributing to a proliferation of names.
 
 In this section, we summarize existing AI methods and propose a variety of extensions.
-We use a simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+We use simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
@@
 We think so and, in response, propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
-EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc} = 0$.
-Recall that $\mu^\text{ref} = 0$, so the assumption is that the groups have equal mean ability.
+EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting $\mu^\text{foc} = 0$.
+Recall that $\mu^\text{ref} = 0$, so this implies that the groups have equal mean ability.
 The result is that all differences in performance — either from group ability differences or DIF — manifest in the item easiness difference parameter $\tilde{d_j}$.
-The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.
+The tilde (e.g., $\tilde{d_j}$) is used to denote parameters estimated with $\mu^\text{foc}$ set to $0$ — as compared to the hat (e.g., $\hat{d_j}$), which is used for parameter estimates from the final model.
 
@@
     labs(
-        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$"),
+        x = latex2exp::TeX("$\\hat{d_j} = \\hat{b_j}^{ref} - \\hat{b_j}^{foc}$"),
         y = ""
@@
 We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
+Appending "all-significant" indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
 AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.
@@
 The process continues iteratively until no new items display DIF.
-AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF.
+AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only *one* item in each iteration as opposed to *all* items that test positive for DIF.
 As a result, we believe that AOAA-OAT is less likely to be "fooled" by artifical DIF.
-Note that AOAA-OAT is not currently implemented in the R package mirt.
+AOAA-OAT is not currently implemented in the R package mirt.
 
@@
 For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$, where $x$ and $y$ both index the items.
+To be sure, $\mathbf{R}^{\text{ref}}$ is a square with the number of rows and columns equal to the number of items.
 The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p.

### 2020-03-20 paper/paper_2b_anchor_points.Rmd
got another draft to ben

@@
 $\mu^{\star\text{foc}}$ is selected by maximizing the Gini index (thus the abbreviation MAXGI).
-The intuition and assumption is that the items without DIF are the most homogeneous.
+The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF.
 Denoting a function that calculates the Gini index from a vector of non-negative elements as  $G(\mathbf{x})$, MGI sets <!-- its weird to arg max over a parameter we already have a definition for - maybe switch to t or something -->
@@
 \end{align}
-where $\mu^\text{foc} \in (-\infty, \infty)$ and $\mu^\text{foc}$ is added to each element of $\mathbf{d}$.
+where $\mu^\text{foc} \in (-\infty, \infty)$ and $\mu^\text{foc}$ is added to each element of $\tilde{\mathbf{d}}$.
 
@@
 The result of MAXGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$, which is extraordinarily close to the data-generating value of -1.
+Moreover, there is a local maximum at $\mu^{\text{foc}} \approx 1.5$, which corresponds to the items with DIF.
+This illustrates that — as @strobl2018anchor point out — the search path is informative.
 
@@
 To adapt Raju's area method into an AI method, we propose a new method, which we call "minimizing the area between curves" (MINBC).
-To understand MINBC, imagine a scenario in which the data-generating process is $\mu^{\text{foc}} = \mu^{\text{ref}}$ and $d_j = 0 \forall j$, so that there is no DIF.
+To understand MINBC, imagine a scenario in which the data-generating process is $\mu^{\text{foc}} = \mu^{\text{ref}}$ and $d_j = 0 \forall j$, so that the groups have equal ability and there is no DIF.
 The fundamental identification problem is that there are an infinite number of models with the same likelihood from which to choose.

### 2020-03-20 paper/paper_3_sim.Rmd
got another draft to ben

@@
 \end{align}
-where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $\sigma(a_{2j})$ is the item's loading on nuisance ability [@demars2016partially].
+where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $a_{2j}$ is the item's loading on nuisance ability [@demars2016partially].
 
@@
 An item's angle measures the relative loading of the item on the two dimensions.
-For example, an angle of 45$^\circ$ indicates that the item loads equally on the target ability and nuisance ability.
+The lower the angle, the more the item loads on target ability as compared to nuisance ability, which corresponds to less DIF.
+An angle of 45$^\circ$ indicates that the item loads equally on the target ability and nuisance ability.
 Our simple noncompensatory model has $a_{1j} = 1$ for all items so the angle equation reduces to
@@
 
-```{r simemmlg, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
+```{r simemmlg, fig.cap = 'For a 12 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
 out$bigsim[[5]]$intuitive_mod[[1]] %>%
@@
 
+<!-- Ben didn't totally foll -->
+
 An effective AI method should lead to a final model that accurately estimates the difference between the reference group's mean target ability and the focal group's mean target ability.
 We refer to this quantity as the achievement gap.
-Recall that all models set $\mu_\theta^\text{ref} = 0$, so the achievement gap reduces to $\mu_\theta^\text{ref}$.
-The data-generating value of $\mu_\theta^\text{foc}$ is 0.5, but each run will include sampling variability.
+Recall that all models set $\mu_\theta^\text{ref} = 0$, so the achievement gap reduces to $\mu_\theta^\text{foc}$.
+The data-generating value of $\mu_\theta^\text{foc}$ is 0.5, but each run, of course, includes sampling variability.
 To get at the heart of how well a method is doing, we calculated the achievement gap residual as the method's estimated achievement gap, $\hat\mu_\theta^\text{foc}$, minus the achievement gap estimated when using only the DIF-free items as anchors.
@@
 We used rank correlation as opposed to the more typically used RMSE in agreement with Lord's -@lord1986maximum argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
-*(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary.
-maybe this idea of rank order being better than RMSE is for another paper)*
+*(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary)*
 
@@
 For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
-Furthermore, our work focused on the Rasch model, and it will be of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.
+Furthermore, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.

### 2020-03-21 paper/paper.Rmd
got another draft to ben
after knitting revisions from ben d

@@
   It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
-However, to use an LRT, a model identifying assumption needs to be made so that differences in group ability can be disentangled from potential DIF.
+However, to use a LRT, we require an identifying assumption to disentangle differences in group ability from potential DIF.
 We use the term "agnostic identification" (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
-We summarize existing AI methods and propose a variety of new methods.
-We conduct a simulation study — which we argue is more realistic than most DIF simulation studies in the literature — and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
-We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
+We first summarize existing AI methods and propose a variety of new methods.
+We then conduct a simulation study—which we argue is more realistic than most DIF simulation studies in the literature—and find that one of the proposed new AI methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
 \clearpage <!-- https://tinyurl.com/ybremelq -->
@@
 
+# Ben D Notes
+
+- Switch 1 and 2 to theta and eta
+
+- Ben worries that in anchor items in particular having the demonstration data and the the thought experiment is too many threads
+
+- Be clear upfront that there are two ways to identify: anchor items and anchor points (and be sure to define both)
+
+- Explain AOAA-OAT inuition better. Perhaps using an EM-MILG or something? (in reference to the "least homogeneous in the sense of" part)
+
+- Compare performance across all methods at once, not one at a time
+
+- Show a paneled MILG for every method all at once. How cool.
+
+- About the simulation difmap figure: "It seems to me like more could be set about this.
+you might want to note, for example, that it need not be the case in general that we have 6 items 'clustering' around the same location of the scale.
+you could have items 1-6 spread out in same way as items 7-12.
+much harder in that case.
+in fact, it seems like there is a geometry problem here we could attempt to solve.
+there are some item-easiness configureations (in sense of MILG) that we really are probably in bad shape with.
+if the item easiness params are uniformly distributed between -1 and 1 for example, we're stuck!
+maybe we could try to think about how to characterize such cases and write about what to do there?"
+
+- Fix that equation to take into account sample sizes
+
 # Introduction {#intro}

### 2020-03-21 paper/paper.tex
after knitting revisions from ben d

@@
 It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
-However, to use an LRT, a model identifying assumption must be made so that differences in group ability can be disentangled from potential DIF.
+However, to use a LRT, we require an identifying assumption to disentangle differences in group ability from potential DIF.
 We use the term ``agnostic identification'' (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
-We summarize existing AI methods and propose a variety of new methods.
-We conduct a simulation study --- which we argue is more realistic than most DIF simulation studies in the literature --- and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
-We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
+We first summarize existing AI methods and propose a variety of new methods.
+We then conduct a simulation study---which we argue is more realistic than most DIF simulation studies in the literature---and find that one of the proposed new AI methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
 \clearpage
@@
 
+\hypertarget{ben-d-notes}{%
+\section{Ben D Notes}\label{ben-d-notes}}
+
+\begin{itemize}
+\item
+  Switch 1 and 2 to theta and eta
+\item
+  Ben worries that in anchor items in particular having the demonstration data and the the thought experiment is too many threads
+\item
+  Be clear upfront that there are two ways to identify: anchor items and anchor points (and be sure to define both)
+\item
+  Explain AOAA-OAT inuition better. Perhaps using an EM-MILG or something? (in reference to the \enquote{least homogeneous in the sense of} part)
+\item
+  Compare performance across all methods at once, not one at a time
+\item
+  Show a paneled MILG for every method all at once. How cool.
+\item
+  About the simulation difmap figure: \enquote{It seems to me like more could be set about this.
+you might want to note, for example, that it need not be the case in general that we have 6 items \enquote{clustering} around the same location of the scale.
+you could have items 1-6 spread out in same way as items 7-12.
+much harder in that case.
+in fact, it seems like there is a geometry problem here we could attempt to solve.
+there are some item-easiness configureations (in sense of MILG) that we really are probably in bad shape with.
+if the item easiness params are uniformly distributed between -1 and 1 for example, we're stuck!
+maybe we could try to think about how to characterize such cases and write about what to do there?}
+\item
+  Fix that equation to take into account sample sizes
+\end{itemize}
+
 \hypertarget{intro}{%
@@
 
-Following Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
+Following Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when one imposes an item response model with too few ability dimensions.
 From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item.
@@
 For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study.
+They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values that they obtained from a previous study.
 They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
+They then simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
+Finally, they generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
 \begin{align}
-P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
+P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j) \label{zopf}
 \end{align}
@@
 For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students.
+For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students.
 Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for all students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for all students in the focal group.
-We now need to use a compensatory 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
+The target ability is the same as unidimensional ability above where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
+The nuisance ability takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group.
+We then reply on a 2PL compensatory 2PL model; slopes on the target ability are \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability is \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}), the probability student \(i\) responds correctly to item \(j\) is
+According to the two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}), the probability that student \(i\) responds correctly to item \(j\) is
 \begin{align}
@@
 \end{align}
+For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
 This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
@@
 Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
+This approach depends upon a crucial identifying assumption; this assumption is the target of our investigation.
 
@@
 
-Psychometricians have long been in search of the perfect DIF detection method.
-It has been well demonstrated that the IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
-As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).
+Psychometricians have long been in search of the perfect method for detecting differential functioning of \emph{individual} items.
+The IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
+Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}) (which, for example, has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})).
 
-The unsolved problem that we address here is how to link groups in the typical setting in which no a prior assumptions --- about either relative group ability or which items might have DIF --- can be made.
-That is, we're interested in agnostic identification methods (hereafter referred to as \enquote{AI methods}) so that LRTs can be used to test items for DIF.
-For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
-And, in search of a coherent framework, we sometimes edit names of existing methods.
-We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.
+However, use of LRT requires an identifying assumption.
+The assumption is needed so as to link groups.
+In the typical setting, no a prior assumptions---about either relative group ability or which items might have DIF---can be made.
+Thus, we're interested in \emph{agnostic identification} methods (hereafter referred to as \enquote{AI methods}).
+For simplicity, we focus on the Rasch model to isolate the fundamental issues in \enquote{the AI problem}.
+Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.}
 
 In this section, we summarize existing AI methods and propose a variety of extensions.
-We use simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
-Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
+We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+Our simulation is similar to our multidimensional alternative to the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) approach discussed above.
+Target ability is simulated as \({\theta_i}^{\text{ref}} \sim N(\mu^\text{ref} = 0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(\mu^\text{foc} = -1,1)\) for students in the focal group.
 Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{foc}} = -1\).
@@
 
-In general, we denote the mean ability of the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\), respectively.
-We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
-The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of \(\hat\mu^\text{foc}\) from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness.
-The most common AI approach is to select a group of anchor items that are assumed to be DIF-free.
+The fundamental challenge is to use an AI method to disentangle estimation of \(\hat\mu^\text{foc} - \mu^\text{ref}\), the difference in mean abilities, from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness.
+The most common AI approaches identify a group of anchor items that are assumed to be DIF-free.
 These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.
 
-Computing is done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the mirt R package (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the suite of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}).
-Code is available at \emph{(TODO)}.
+Throughout this paper, we estimate models using marginal maximum likelihood estimation (Bock and Aitkin \protect\hyperlink{ref-bock1981marginal}{1981}).
+In addition, we follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
 
@@
 
-AI methods typically work like a black box.
-The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting the final model.
-These types of AI methods have their role --- and the main focus of this paper is on them --- but wouldn't it be nice if there was a way for the analyst to visualize the possibility of DIF in their data?
+AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.
 
-We think so and, in response, propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
+While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position for understanding the scale of the problem in their data.
+We propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
 EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting \(\mu^\text{foc} = 0\).
-Recall that \(\mu^\text{ref} = 0\), so this implies that the groups have equal mean ability.
-The result is that all differences in performance --- either from group ability differences or DIF --- manifest in the item easiness difference parameter \(\tilde{d_j}\).
-The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\) --- as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from the final model.
+Given that we also assume \(\mu^\text{ref} = 0\), this is equivalent to the assumption that the groups have equal mean ability.
+As a result, all differences in performance---either from group ability differences or DIF---manifest in the item easiness difference parameter \(\tilde{d_j}\).
+The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\)---as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from the properly identified model.
 
@@
 Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data.
-We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
-The analyst might assume that --- because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased.
-These unbiased items are known as anchor items.
+We emphasize that the EM-MILG contains all possible information about the difference in group performance.
+The analyst might assume that---because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits---items 1-5 are unbiased.
+Indeed, all AI methods rest on the logic that its the minority of items that contain DIF.
+These unbiased items are known as anchor items; setting anchor items is the most common identification assumption.
 After anchor items have been selected, the model is refit where the assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the assumption that the anchor items are DIF-free (i.e., \(d_j = 0\)).
+Of course, the demonstration data is designed such that there are obvious groups; this will not typically be the case.
 
@@
 In general, one of our key concerns with traditional AI methods is that they can lull the analyst into a false sense of security.
-Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality.
-A EM-MILG combats this concern by presenting all information clearly to the analyst.
+The analyst simply chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality.
+Our aim with EM-MILG is to combat this concern by presenting all information clearly to the analyst.
 In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
@@
 For example, when testing the first item for DIF, all of the other items are used as anchors.
-This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}).
+This is done using a LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}).
 Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
@@
 
-Edelen et al.
-(\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can they be sure that its those 10 items and not the other 11 items with DIF?
-We argue that this method cannot tell us with certainty that it is those 10 items that have DIF while the other 11 do not.
-Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
-Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method (Strobl et al.
-\protect\hyperlink{ref-strobl2018anchor}{2018}).
-In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
+Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, counter to the underlying rationale for the undertaking.
+On a practical level, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
 
@@
 The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF}, which is a significant problem in applications of AI methods as discussed below.
 
@@
 One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
-For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively --- with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model --- until no more items tested positively.
+For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively---with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model---until no more items tested positively.
 Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse) referring to beginning with the assumption that all items are DIF-free.
@@
 
-AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF?
+We argue that AOAA-AS, while a potential improvement, doesn't solve the fundamental problem of AOAA: What does one do when all items test positive for DIF?
 With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
@@
 Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
+Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method: Use AOAA and select the four items that exhibit the least amount of DIF as anchor items.
 It's unclear if one should proceed if those four items also test positive for DIF.
@@
 We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
-AOAA-OAT is inspired by Hagquist and Andrich (\protect\hyperlink{ref-hagquist2017recent}{2017}), who, in general, assert that \enquote{items showing DIF initially should not be resolved simultaneously but sequentially} {[}p.~6{]}.
+AOAA-OAT is inspired by Hagquist and Andrich (\protect\hyperlink{ref-hagquist2017recent}{2017}), who assert that \enquote{items showing DIF initially should not be resolved simultaneously but sequentially} {[}p.~6{]}.
 Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the \(\chi^2\) test statistic, is removed from the anchor set.
@@
 Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF.
+The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF (i.e., it is the least homogeneous in the sense that the group difference in performance is most different from the average group difference in performance on the other two items).
 In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
@@
 
-Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster.} They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
-For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\), where \(x\) and \(y\) both index the items.
-To be sure, \(\mathbf{R}^{\text{ref}}\) is a square with the number of rows and columns equal to the number of items.
+Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster}.
+They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
+For each group, the relative easinesses for each pair of items can be stored in the (square) matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\), where \(x\) and \(y\) both index the items.
+A similar matrix, \(\mathbf{R}^{\text{foc}}\), can be formed.
 The ultimate matrix of interest is \(\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}\) which is the \enquote{differences between groups in the pairwise differences in (easiness) between items} {[}p.~323{]}.
@@
 The general idea of identifying clusters of items is intriguing.
-However, their approach is needlessly complicated, and they did not describe a process for moving from \(\Delta \mathbf{R}\) to an anchor cluster.
+However, their approach is complicated, and they did not describe a process for moving from \(\Delta \mathbf{R}\) to an anchor cluster.
 Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing one such process.
 \(\Delta \mathbf{R}\) is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
-Accordingly, they use k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC.
+Accordingly, they recommend k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC.
 They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.
+They conducted a simulation study where some items contained DIF, and, unfortunately, found that BIC identifies only a single cluster, so all items were anchors.
 
@@
 Instead of choosing k with BIC, we use the gap statistic method recommended by Hastie, Tibshirani, and Walther (\protect\hyperlink{ref-hastie2001estimating}{2001}).
-The largest cluster is chosen as the anchor cluster.
+Based on the assumption that, at most, a minority of items contain DIF, we choose the largest cluster as the anchor cluster.
 If there is a tie for the largest cluster, the cluster with the lowest standard deviation of \(\tilde{\mathbf{d}}\) is selected.
-This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF.
 For our demonstration data, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
@@
 \(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (thus the abbreviation MAXGI).
-The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF.
+The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF and a small subset of items thus having large amounts of DIF.
 Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
@@
 
-For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values.
-The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is extraordinarily close to the data-generating value of -1.
+For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini index at a variety of possible \(\mu^\text{foc}\) values.
+The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is quite close to the data-generating value of -1.
 Moreover, there is a local maximum at \(\mu^{\text{foc}} \approx 1.5\), which corresponds to the items with DIF.
-This illustrates that --- as Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}) point out --- the search path is informative.
+This illustrates that---as Strobl et al.
+(\protect\hyperlink{ref-strobl2018anchor}{2018}) point out---the search path is informative.
 
@@
 To compare each of the methods in Table \ref{table:allmethods}, we conducted a simulated study.
-Our goal was to make the data generating process as realistic to the scenario described by Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) in which some items on a math test also depend on a student's verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
+Our goal was to use a relatively realistic data generating process motivated by the case study in Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) wherein some items on a math test also depend on a student's verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
 As described in \protect\hyperlink{intro}{the introduction}, nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
@@
 where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(a_{2j}\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
+
+Computing was done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the mirt R package (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the suite of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}).
+Code is available at \emph{(TODO)}.
 
@@
 
-Figure \ref{fig:simemmlg} shows the EM-MILG --- generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the \protect\hyperlink{anchoritems}{the anchor items section} --- for one run using the same item parameters that generated Figure \ref{fig:difmap}.
+Figure \ref{fig:simemmlg} shows the EM-MILG---generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the \protect\hyperlink{anchoritems}{the anchor items section}---for one run using the same item parameters that generated Figure \ref{fig:difmap}.
 As expected, \(\tilde{d_j}\) is about \(\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5\) for the first six items which are DIF free.
@@
 
-\hypertarget{individual-abilities}{%
-\subsubsection{Individual abilities}\label{individual-abilities}}
-
-Assessments are frequently used to make decisions about individual student abilities.
-We measured a method's ability to do so by calculating the rank correlation between the vector of estimated abilities and the vector of true target abilities.
-We used rank correlation as opposed to the more typically used RMSE in agreement with Lord's (\protect\hyperlink{ref-lord1986maximum}{1986}) argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
-\emph{(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary)}
-
 \hypertarget{anchor-items.}{%
@@
 Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual.
-AOAA-OAT is the clear winner.
+AOAA-OAT was the clear winner.
 It performed nearly perfectly for two, three, or four DIF items.
 Even when six of the 12 items on the test contained DIF, AOAA-OAT underestimated \(\mu_\theta^\text{foc}\) by only 0.05 standard deviations on its worst run.
-As expected, artificial DIF caused AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
+As expected, artificial DIF caused AOAA and AOAA-AS to begin to exhibit problematic performance as the number of DIF items increased.
+
+(START BEN DIDNT LOOK AT)
 
@@
 
+(STOP BEN DIDNT LOOK AT)
+
 \hypertarget{discussion-and-summary}{%
@@
 
-Measurement instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
-Inspired by Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}), we have resurrected the DIF-as-student-property view and shown its connection to the typical way in which AI methods and DIF simulation studies have been conceptualized.
-
-More importantly, we reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe to be more realistic than the typical DIF simulation study.
+Validity hinges on measurement instruments being relatively free from DIF.
+Such instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
+We reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe to be more realistic than the typical DIF simulation study.
 In particular, we simulated student ability as drawn from a two-dimensional distribution representing a student's target and nuisance ability, and then generated data using a noncompensatory item response model.
 Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
-On the other hand, AOAA-OAT, EMC, MINGI, and MAXBC all performed reasonably.
+This is concerning given both their widespread use and the fact that other method---AOAA-OAT, EMC, MINGI, and MAXBC---demonstrate superior performance.
 
-The fact that the AOAA-OAT exhibited superior performance is an important finding given that it is, to our knowledge, not used currently.
-We recommend its use whenever possible.
-One reason AOAA-OAT might not always be possible is that it can be computationally expensive.
+AOAA-OAT exhibiting superior performance is an important finding given that it is, to our knowledge, not used currently.
+We advocate for its widespread use.
+One drawback of the AOAA-OAT method is that it is computationally expensive.
 For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows.
@@
 
+\leavevmode\hypertarget{ref-bock1981marginal}{}%
+Bock, R Darrell, and Murray Aitkin. 1981. ``Marginal Maximum Likelihood Estimation of Item Parameters: Application of an Em Algorithm.'' \emph{Psychometrika} 46 (4): 443--59.
+
 \leavevmode\hypertarget{ref-camilli1992conceptual}{}%
@@
 
-\leavevmode\hypertarget{ref-edelen2006identification}{}%
-Edelen, Maria Orlando, David Thissen, Jeanne A Teresi, Marjorie Kleinman, and Katja Ocepek-Welikson.
-2006.
-``Identification of Differential Item Functioning Using Item Response Theory and the Likelihood-Based Model Comparison Approach: Application to the Mini-Mental State Examination.'' \emph{Medical Care}, S134--S142.
-
 \leavevmode\hypertarget{ref-gini1912variabilita}{}%
@@
 Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.
-
-\leavevmode\hypertarget{ref-lord1986maximum}{}%
-Lord, Frederic M. 1986. ``Maximum Likelihood and Bayesian Parameter Estimation in Item Response Theory.'' \emph{Journal of Educational Measurement} 23 (2): 157--62.
 

### 2020-03-21 paper/paper_1_intro.Rmd
after knitting revisions from ben d

@@
-Following @camilli1992conceptual, we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed.
+Following @camilli1992conceptual, we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when one imposes an item response model with too few ability dimensions.
 From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item.
@@
 For example, @kopf2015framework simulate students as belonging to either a reference or focal group.
-They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study.
+They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values that they obtained from a previous study.
 They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is
+They then simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
+Finally, they generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is
 \begin{align}
-P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
+P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j) \label{zopf}
 \end{align}
@@
 For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
-For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students.
+For example, to translate the @kopf2015framework simulation from the DIF-as-item-property to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students.
 Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for all students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for all students in the focal group.
-We now need to use a compensatory 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise.
+The target ability is the same as unidimensional ability above where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
+The nuisance ability takes one of two values: ${\eta_i}^{\text{ref}} = 0$ for students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for students in the focal group.
+We then reply on a 2PL compensatory 2PL model; slopes on the target ability are $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability is $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise.
 Again, 0.6 is the magnitude of DIF in logits.
-According to the two-dimensional compensatory 2PL model [@thissen1986taxonomy], the probability student $i$ responds correctly to item $j$ is
+According to the two-dimensional compensatory 2PL model [@thissen1986taxonomy], the probability that student $i$ responds correctly to item $j$ is
 \begin{align}
@@
 Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
+This approach depends upon a crucial identifying assumption; this assumption is the target of our investigation.

### 2020-03-21 paper/paper_2a_anchor_items.Rmd
after knitting revisions from ben d

@@
-Psychometricians have long been in search of the perfect DIF detection method.
-It has been well demonstrated that the IRT-based likelihood ratio test (LRT) is effective at detecting DIF [@meade2012solving].
-As a result, we avoid methods like the Mantel-Haenszel procedure [@holland1986differential], which muddies the waters by moving away from the IRT framework, and has been shown to perform no better — and in some cases worse — than IRT-based methods [@swaminathan1990detecting].
+Psychometricians have long been in search of the perfect method for detecting differential functioning of *individual* items.
+The IRT-based likelihood ratio test (LRT) is effective at detecting DIF [@meade2012solving].
+Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure [@holland1986differential] (which, for example, has been shown to perform no better—and in some cases worse—than IRT-based methods [@swaminathan1990detecting]).
 
-The unsolved problem that we address here is how to link groups in the typical setting in which no a prior assumptions — about either relative group ability or which items might have DIF — can be made.
-That is, we're interested in agnostic identification methods (hereafter referred to as "AI methods") so that LRTs can be used to test items for DIF.
-For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection.
-And, in search of a coherent framework, we sometimes edit names of existing methods.
-We recognize that others have done the same [e.g., @kopf2015framework], and that we risk contributing to a proliferation of names.
+However, use of LRT requires an identifying assumption.
+The assumption is needed so as to link groups.
+In the typical setting, no a prior assumptions—about either relative group ability or which items might have DIF—can be made.
+Thus, we're interested in *agnostic identification* methods (hereafter referred to as "AI methods").
+For simplicity, we focus on the Rasch model to isolate the fundamental issues in "the AI problem".
+Further, so as to offer a coherent framework, we sometimes edit names of existing methods.
+^[We recognize that others have done the same [e.g., @kopf2015framework], and that we risk contributing to a proliferation of names.]
 
 In this section, we summarize existing AI methods and propose a variety of extensions.
-We use simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test.
-Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
+We use a (simulated) demonstration dataset to illustrate key features of the different methods.
+^[This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.] The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+Our simulation is similar to our multidimensional alternative to the @kopf2015framework approach discussed above.
+Target ability is simulated as ${\theta_i}^{\text{ref}} \sim N(\mu^\text{ref} = 0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(\mu^\text{foc} = -1,1)$ for students in the focal group.
 Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{foc}} = -1$.
@@
 
-In general, we denote the mean ability of the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$, respectively.
-We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
-The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of $\hat\mu^\text{foc}$ from the estimation of $\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$, the difference in item easiness.
-The most common AI approach is to select a group of anchor items that are assumed to be DIF-free.
+The fundamental challenge is to use an AI method to disentangle estimation of $\hat\mu^\text{foc} - \mu^\text{ref}$, the difference in mean abilities, from the estimation of $\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$, the difference in item easiness.
+The most common AI approaches identify a group of anchor items that are assumed to be DIF-free.
 These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.
 
-Computing is done in R [@rcore], model fitting in the mirt R package [@chalmers2012mirt], and data wrangling/visualization in the suite of R packages known as the tidyverse [@tidy].
-Code is available at *(TODO)*.
+Throughout this paper, we estimate models using marginal maximum likelihood estimation [@bock1981marginal].
+In addition, we follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
 
@@
 
-AI methods typically work like a black box.
-The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting the final model.
-These types of AI methods have their role — and the main focus of this paper is on them — but wouldn't it be nice if there was a way for the analyst to visualize the possibility of DIF in their data?
+AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.
 
-We think so and, in response, propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
+While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position for understanding the scale of the problem in their data.
+We propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
 EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting $\mu^\text{foc} = 0$.
-Recall that $\mu^\text{ref} = 0$, so this implies that the groups have equal mean ability.
-The result is that all differences in performance — either from group ability differences or DIF — manifest in the item easiness difference parameter $\tilde{d_j}$.
-The tilde (e.g., $\tilde{d_j}$) is used to denote parameters estimated with $\mu^\text{foc}$ set to $0$ — as compared to the hat (e.g., $\hat{d_j}$), which is used for parameter estimates from the final model.
+Given that we also assume $\mu^\text{ref} = 0$, this is equivalent to the assumption that the groups have equal mean ability.
+As a result, all differences in performance—either from group ability differences or DIF—manifest in the item easiness difference parameter $\tilde{d_j}$.
+The tilde (e.g., $\tilde{d_j}$) is used to denote parameters estimated with $\mu^\text{foc}$ set to $0$—as compared to the hat (e.g., $\hat{d_j}$), which is used for parameter estimates from the properly identified model.
 
@@
 Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data.
-We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance.
-The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased.
-These unbiased items are known as anchor items.
+We emphasize that the EM-MILG contains all possible information about the difference in group performance.
+The analyst might assume that—because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits—items 1-5 are unbiased.
+Indeed, all AI methods rest on the logic that its the minority of items that contain DIF.
+These unbiased items are known as anchor items; setting anchor items is the most common identification assumption.
 After anchor items have been selected, the model is refit where the assumption that $\mu^\text{foc} = 0$ is dropped and replaced with the assumption that the anchor items are DIF-free (i.e., $d_j = 0$).
+Of course, the demonstration data is designed such that there are obvious groups; this will not typically be the case.
 
@@
 For example, when testing the first item for DIF, all of the other items are used as anchors.
-This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection].
+This is done using a LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection].
 Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
@@
 
-@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF.
-How can they be sure that its those 10 items and not the other 11 items with DIF?
-We argue that this method cannot tell us with certainty that it is those 10 items that have DIF while the other 11 do not.
-Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible.
-More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
-Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor].
-In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
+Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, counter to the underlying rationale for the undertaking.
+On a practical level, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
 
@@
 The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF."
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF", which is a significant problem in applications of AI methods as discussed below.
 
@@
 One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
-For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively.
+For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively—with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model—until no more items tested positively.
 @kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse) referring to beginning with the assumption that all items are DIF-free.
@@
 @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-@woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF.
+@woods2009empirical suggested a more straightforward, one-step method: Use AOAA and select the four items that exhibit the least amount of DIF as anchor items.
 It's unclear if one should proceed if those four items also test positive for DIF.
@@
 We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
-AOAA-OAT is inspired by @hagquist2017recent, who, in general, assert that "items showing DIF initially should not be resolved simultaneously but sequentially" [p.
+AOAA-OAT is inspired by @hagquist2017recent, who assert that "items showing DIF initially should not be resolved simultaneously but sequentially" [p.
 6].
@@
 
-@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
-For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$, where $x$ and $y$ both index the items.
-To be sure, $\mathbf{R}^{\text{ref}}$ is a square with the number of rows and columns equal to the number of items.
+@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster".
+They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
+For each group, the relative easinesses for each pair of items can be stored in the (square) matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$, where $x$ and $y$ both index the items.
+A similar matrix, $\mathbf{R}^{\text{foc}}$, can be formed.
 The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p.
@@
 The general idea of identifying clusters of items is intriguing.
-However, their approach is needlessly complicated, and they did not describe a process for moving from $\Delta \mathbf{R}$ to an anchor cluster.
+However, their approach is complicated, and they did not describe a process for moving from $\Delta \mathbf{R}$ to an anchor cluster.
 @pohl2017cluster extended their work by proposing one such process.
 $\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
-Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
+Accordingly, they recommend k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
 They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.
+They conducted a simulation study where some items contained DIF, and, unfortunately, found that BIC identifies only a single cluster, so all items were anchors.
 
@@
 Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating.
-The largest cluster is chosen as the anchor cluster.
+Based on the assumption that, at most, a minority of items contain DIF, we choose the largest cluster as the anchor cluster.
 If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected.
-This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF.
 For our demonstration data, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.

### 2020-03-21 paper/paper_2b_anchor_points.Rmd
after knitting revisions from ben d

@@
 $\mu^{\star\text{foc}}$ is selected by maximizing the Gini index (thus the abbreviation MAXGI).
-The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF.
+The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF and a small subset of items thus having large amounts of DIF.
 Denoting a function that calculates the Gini index from a vector of non-negative elements as  $G(\mathbf{x})$, MGI sets <!-- its weird to arg max over a parameter we already have a definition for - maybe switch to t or something -->
@@
 
-For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible $\mu^\text{foc}$ values.
-The result of MAXGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$, which is extraordinarily close to the data-generating value of -1.
+For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini index at a variety of possible $\mu^\text{foc}$ values.
+The result of MAXGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$, which is quite close to the data-generating value of -1.
 Moreover, there is a local maximum at $\mu^{\text{foc}} \approx 1.5$, which corresponds to the items with DIF.
-This illustrates that — as @strobl2018anchor point out — the search path is informative.
+This illustrates that—as @strobl2018anchor point out—the search path is informative.
 

### 2020-03-21 paper/paper_3_sim.Rmd
after knitting revisions from ben d

@@
 To compare each of the methods in Table \ref{table:allmethods}, we conducted a simulated study.
-Our goal was to make the data generating process as realistic to the scenario described by @ackerman1992didactic in which some items on a math test also depend on a student’s verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
+Our goal was to use a relatively realistic data generating process motivated by the case study in @ackerman1992didactic wherein some items on a math test also depend on a student’s verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
 As described in [the introduction](#intro), nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
@@
 where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $a_{2j}$ is the item's loading on nuisance ability [@demars2016partially].
+
+Computing was done in R [@rcore], model fitting in the mirt R package [@chalmers2012mirt], and data wrangling/visualization in the suite of R packages known as the tidyverse [@tidy].
+Code is available at *(TODO)*.
 
@@
 
-Figure \ref{fig:simemmlg} shows the EM-MILG — generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the [the anchor items section](#anchoritems) — for one run using the same item parameters that generated Figure \ref{fig:difmap}.
+Figure \ref{fig:simemmlg} shows the EM-MILG—generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the [the anchor items section](#anchoritems)—for one run using the same item parameters that generated Figure \ref{fig:difmap}.
 As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first six items which are DIF free.
@@
 
-<!-- Ben didn't totally foll -->
-
 An effective AI method should lead to a final model that accurately estimates the difference between the reference group's mean target ability and the focal group's mean target ability.
@@
 
-### Individual abilities
-
-Assessments are frequently used to make decisions about individual student abilities.
-We measured a method’s ability to do so by calculating the rank correlation between the vector of estimated abilities and the vector of true target abilities.
-We used rank correlation as opposed to the more typically used RMSE in agreement with Lord's -@lord1986maximum argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely.
-*(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary)*
+<!-- remember i removed rank order of abilities. in response, ben said "rather than abilities, i would argue in favor of proper estimation of measurement error as a function of theta?"  -->
 
@@
 Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual.
-AOAA-OAT is the clear winner.
+AOAA-OAT was the clear winner.
 It performed nearly perfectly for two, three, or four DIF items.
 Even when six of the 12 items on the test contained DIF, AOAA-OAT underestimated $\mu_\theta^\text{foc}$ by only 0.05 standard deviations on its worst run.
-As expected, artificial DIF caused AOAA and AOAA-AS to go off the rails as the number of DIF items increases.
+As expected, artificial DIF caused AOAA and AOAA-AS to begin to exhibit problematic performance as the number of DIF items increased.
+
+(START BEN DIDNT LOOK AT)
 
@@
 
+(STOP BEN DIDNT LOOK AT)
+
 # Discussion and summary
 
-Measurement instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
-Inspired by @ackerman1992didactic, we have resurrected the DIF-as-student-property view and shown its connection to the typical way in which AI methods and DIF simulation studies have been conceptualized.
-
-More importantly, we reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe to be more realistic than the typical DIF simulation study.
+Validity hinges on measurement instruments being relatively free from DIF.
+Such instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
+We reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe to be more realistic than the typical DIF simulation study.
 In particular, we simulated student ability as drawn from a two-dimensional distribution representing a student's target and nuisance ability, and then generated data using a noncompensatory item response model.
 Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
-On the other hand, AOAA-OAT, EMC, MINGI, and MAXBC all performed reasonably.
+This is concerning given both their widespread use and the fact that other method—AOAA-OAT, EMC, MINGI, and MAXBC—demonstrate superior performance.
 
-The fact that the AOAA-OAT exhibited superior performance is an important finding given that it is, to our knowledge, not used currently.
-We recommend its use whenever possible.
-One reason AOAA-OAT might not always be possible is that it can be computationally expensive.
+AOAA-OAT exhibiting superior performance is an important finding given that it is, to our knowledge, not used currently.
+We advocate for its widespread use.
+One drawback of the AOAA-OAT method is that it is computationally expensive.
 For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows.
@@
 Furthermore, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.
+
+<!-- This feels like it could go somewhere, but not sure where: Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor].
+In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.
+-->

### 2020-03-23 paper/paper.Rmd
finished paper print 6 and shipped back to ben

--- 2020-03-19 paper/paper.Rmd
@@
   It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
-However, to use an LRT, a model identifying assumption needs to be made so that differences in group ability can be disentangled from potential DIF.
+However, to use a LRT (or any other method), we require an identifying assumption to disentangle differences in group ability from potential DIF.
 We use the term "agnostic identification" (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
-We summarize existing AI methods and propose a variety of new methods.
-We conduct a simulation study — which we argue is more realistic than most DIF simulation studies in the literature — and find that one of the proposed new AI methods, All-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
-We also offer a new method, the equal means, multiple imputation logit graph (EM-MILG), that clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
+We first summarize existing AI methods and propose a variety of new methods.
+We then conduct a simulation study—which we argue is more realistic than most DIF simulation studies in the literature—and find that one of the proposed new AI methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
+We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
 \clearpage <!-- https://tinyurl.com/ybremelq -->
@@
     echo=FALSE,
-    warning=TRUE,
+    warning=FALSE,
+    message=FALSE
     fig.pos='H',
@@
 
+## Summary of AI methods
+
+```{r child = "paper_2c_summary_of_methods.Rmd"}
+```
+
 # Simulation study {#simstudy}

### 2020-03-23 paper/paper.tex
finished paper print 6 and shipped back to ben

@@
 It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
-However, to use a LRT, we require an identifying assumption to disentangle differences in group ability from potential DIF.
+However, to use a LRT (or any other method), we require an identifying assumption to disentangle differences in group ability from potential DIF.
 We use the term ``agnostic identification'' (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
@@
 
-\hypertarget{ben-d-notes}{%
-\section{Ben D Notes}\label{ben-d-notes}}
-
-\begin{itemize}
-\item
-  Switch 1 and 2 to theta and eta
-\item
-  Ben worries that in anchor items in particular having the demonstration data and the the thought experiment is too many threads
-\item
-  Be clear upfront that there are two ways to identify: anchor items and anchor points (and be sure to define both)
-\item
-  Explain AOAA-OAT inuition better. Perhaps using an EM-MILG or something? (in reference to the \enquote{least homogeneous in the sense of} part)
-\item
-  Compare performance across all methods at once, not one at a time
-\item
-  Show a paneled MILG for every method all at once. How cool.
-\item
-  About the simulation difmap figure: \enquote{It seems to me like more could be set about this.
-you might want to note, for example, that it need not be the case in general that we have 6 items \enquote{clustering} around the same location of the scale.
-you could have items 1-6 spread out in same way as items 7-12.
-much harder in that case.
-in fact, it seems like there is a geometry problem here we could attempt to solve.
-there are some item-easiness configureations (in sense of MILG) that we really are probably in bad shape with.
-if the item easiness params are uniformly distributed between -1 and 1 for example, we're stuck!
-maybe we could try to think about how to characterize such cases and write about what to do there?}
-\item
-  Fix that equation to take into account sample sizes
-\end{itemize}
-
 \hypertarget{intro}{%
@@
 For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies \enquote{DIF}.
+In this case, math ability is the \enquote{target ability}, and verbal ability is the \enquote{nuisance ability}.
+Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies \enquote{DIF}.
 
@@
 \begin{align}
-P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j) \label{zopf}
+\text{Pr}(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j) \label{zopf}
 \end{align}
@@
 For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students.
-Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is the same as unidimensional ability above where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-The nuisance ability takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group.
-We then reply on a 2PL compensatory 2PL model; slopes on the target ability are \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability is \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise.
-Again, 0.6 is the magnitude of DIF in logits.
 According to the two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}), the probability that student \(i\) responds correctly to item \(j\) is
 \begin{align}
-\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
+\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\theta j}, a_{\eta j}, b_j) = \sigma(a_{\theta j}\theta_i + a_{\eta j}\eta_i + b_j).
 \end{align}
-For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
+where \(\theta\) is target ability, \(\eta\) is nuisance ability, \(a_{\theta j}\) is an item's slope on target ability, and \(a_{\eta j}\) is an item's slope on nuisance ability.
+
+For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view, \(b_j\) is set to \({b_j}^{\text{ref}}\), \(\theta\) stays the same, \(a_{\theta j} = 1\) for all items (consistent with the Rasch model), and \(\eta\) takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group.
+For items with DIF, \(a_{\eta j} = 0.6\), and for items without DIF, \(a_{\eta j} = 0\) Again, 0.6 is the magnitude of DIF in logits.\footnote{The reader may notice that this is just one possible specification.
+Any specification that meets the following criteria has the same effect: First, \(a_{\eta j}\eta_i\) must be 0.6 lower for the focal group on items with DIF.
+Second, \(a_{\eta j}\eta_i\) must be equal across groups for items without DIF.} For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
 This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
@@
 The IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
-Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}) (which, for example, has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})).
+Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})).
+Within the IRT-framework, we---as is common---estimate models using marginal maximum likelihood estimation (Bock and Aitkin \protect\hyperlink{ref-bock1981marginal}{1981}) and identify the scale by setting \(\mu^\text{ref} = 0\).
 
-However, use of LRT requires an identifying assumption.
+Use of LRT requires an identifying assumption.
 The assumption is needed so as to link groups.
@@
 Thus, we're interested in \emph{agnostic identification} methods (hereafter referred to as \enquote{AI methods}).
-For simplicity, we focus on the Rasch model to isolate the fundamental issues in \enquote{the AI problem}.
-Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.}
+The fundamental challenge is to disentangle estimation of \(\hat\mu^\text{foc} - \mu^\text{ref} = \hat\mu^\text{foc} - 0 = \hat\mu^\text{foc}\), the difference in mean abilities, from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness.
+There are two types of AI methods: The most common, the use of anchor items, selects a group of items that are assumed to be DIF-free.
+These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.
+The other type, the use of an anchor point (Strobl et al.
+\protect\hyperlink{ref-strobl2018anchor}{2018}), directly sets the difference in mean abilities to some value, \(\mu^{\star\text{foc}}\).
+This anchor point identifies the model, and all items can be tested for DIF using an LRT.
 
 In this section, we summarize existing AI methods and propose a variety of extensions.
-We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+For simplicity, we focus on the Rasch model to isolate the fundamental issues in \enquote{the AI problem}.
+Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.} We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Our simulation is similar to our multidimensional alternative to the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) approach discussed above.
@@
 
-The fundamental challenge is to use an AI method to disentangle estimation of \(\hat\mu^\text{foc} - \mu^\text{ref}\), the difference in mean abilities, from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness.
-The most common AI approaches identify a group of anchor items that are assumed to be DIF-free.
-These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.
-
-Throughout this paper, we estimate models using marginal maximum likelihood estimation (Bock and Aitkin \protect\hyperlink{ref-bock1981marginal}{1981}).
-In addition, we follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\).
+The AI methods described in this paper are summarized in Table \ref{table:allmethods}.
+In addition the results for each method on the demonstration data are shown in a MILG in Figure \ref{fig:allmethodsfig}.
 
@@
 
-While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position for understanding the scale of the problem in their data.
+While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position to understand the scale of the problem in their data.
 We propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
@@
 As a result, all differences in performance---either from group ability differences or DIF---manifest in the item easiness difference parameter \(\tilde{d_j}\).
-The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\)---as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from the properly identified model.
+The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\)---as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from a properly identified model.
 
@@
 The analyst might assume that---because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits---items 1-5 are unbiased.
-Indeed, all AI methods rest on the logic that its the minority of items that contain DIF.
+This might seem to be a strong assumption, but all AI methods assume---in one way or another---that, if there is DIF, its a minority of items that are responsible.
 These unbiased items are known as anchor items; setting anchor items is the most common identification assumption.
-After anchor items have been selected, the model is refit where the assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the assumption that the anchor items are DIF-free (i.e., \(d_j = 0\)).
-Of course, the demonstration data is designed such that there are obvious groups; this will not typically be the case.
+After anchor items have been selected, the model is refit where the identifying assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the identifying assumption that the anchor items are DIF-free (i.e., \(d_j = 0\) for anchor items).
+Of course, the demonstration data is designed such that there are obvious groups; this will not necessarily typically be the case.
 
@@
 
-The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with the final model.
+The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with a properly identified model.
 Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
@@
 
-In general, one of our key concerns with traditional AI methods is that they can lull the analyst into a false sense of security.
-The analyst simply chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality.
+In general, one of our main concerns with traditional AI methods is that they can lull the analyst into a false sense of security: The analyst chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality.
 Our aim with EM-MILG is to combat this concern by presenting all information clearly to the analyst.
@@
 And, of course, the MILG can be used to visualize DIF regardless of how the model is identified.
-We now move on to more traditional AI methods, which don't require judgment from the analyst.
+We now move on to more traditional AI methods, which don't require judgment from the analyst, but make the majority-of-items-don't-have-DIF assumption more subtly.
 
@@
 The items for which the flexible model outperforms the baseline model (typically based on a \(\chi^2\) test) are identified as having DIF, and the rest of the items become anchor items.
-AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
+AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the baseline to the flexible model).
 
@@
 On a practical level, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
-
 Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
@@
 The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF}, which is a significant problem in applications of AI methods as discussed below.
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF}, which is a significant problem in applications of AI methods.
 
@@
 We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending \enquote{all-significant} indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
+We append \enquote{all-significant} to indicate that all items that test positive for DIF are removed from the anchor set in each iteration.
 AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.
@@
 Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method: Use AOAA and select the four items that exhibit the least amount of DIF as anchor items.
-It's unclear if one should proceed if those four items also test positive for DIF.
 
@@
 AOAA-OAT is inspired by Hagquist and Andrich (\protect\hyperlink{ref-hagquist2017recent}{2017}), who assert that \enquote{items showing DIF initially should not be resolved simultaneously but sequentially} {[}p.~6{]}.
-Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the \(\chi^2\) test statistic, is removed from the anchor set.
+Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF---based on the \(\chi^2\) test statistic---is removed from the anchor set.
 The process continues iteratively until no new items display DIF.
@@
 As a result, we believe that AOAA-OAT is less likely to be \enquote{fooled} by artifical DIF.
-AOAA-OAT is not currently implemented in the R package mirt.
+AOAA-OAT is not currently implemented in the mirt R package.
 
 Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF (i.e., it is the least homogeneous in the sense that the group difference in performance is most different from the average group difference in performance on the other two items).
+The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF (i.e., because the group difference in performance is most different from the average group difference in performance on the other two items).
 In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-To work, AOAA-OAT has two requirements: First, that at least two items do not have DIF, and second, that the set of items without DIF are more homogeneous than other sets of items.
-
-\hypertarget{summary-and-performance}{%
-\paragraph{Summary and performance}\label{summary-and-performance}}
-
-It's useful to remember that AOAA is not an iterative procedure.
-In constrast, the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method.
-All of the methods described in this paper are summarized in Table \ref{table:allmethods}.
-For our demonstration data, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
-On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.
 
@@
 Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster}.
-They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
+They pointed out that one way around the lack of model identifiability is to consider only relative item parameters.
 For each group, the relative easinesses for each pair of items can be stored in the (square) matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\), where \(x\) and \(y\) both index the items.
@@
 Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing one such process.
-\(\Delta \mathbf{R}\) is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
+They point out that \(\Delta \mathbf{R}\) is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
 Accordingly, they recommend k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC.
@@
 We propose a new cluster-based approach, which we call \enquote{equal means clustering} (EMC).
-Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\), which, as described in the \protect\hyperlink{emmilg}{EM-MILG section}, comes from setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), and thus the name \enquote{equal means clustering}).
-
+Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\), which, as described in the \protect\hyperlink{emmilg}{EM-MILG section}, comes from setting \(\mu^\text{foc} = 0\).
+Recall that \(\mu^\text{ref}\) is to \(0\) to identify the scale, and thus the name \enquote{equal means clustering}.
 Instead of choosing k with BIC, we use the gap statistic method recommended by Hastie, Tibshirani, and Walther (\protect\hyperlink{ref-hastie2001estimating}{2001}).
@@
 If there is a tie for the largest cluster, the cluster with the lowest standard deviation of \(\tilde{\mathbf{d}}\) is selected.
-For our demonstration data, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
 
@@
 
-The previously discussed AI methods select a set of anchor items, whether it is an algorithm or the analyst that makes that selection.
-The anchor items are used to estimate \(\hat \mu^\text{foc}\).
-An alternative strategy is to directly set the anchor point (Strobl et al.
+The previously discussed AI methods identify the model by selecting a set of anchor items that are assumed to be DIF-free.
+We now discuss the second type of AI methods: The use of an anchor point (Strobl et al.
 \protect\hyperlink{ref-strobl2018anchor}{2018}).
+An anchor point directly sets the difference in mean abilities to some value, \(\mu^{\star\text{foc}}\).
 Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF.
@@
 
-\(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (thus the abbreviation MAXGI).
+Strobl et al.
+(\protect\hyperlink{ref-strobl2018anchor}{2018}) select \(\mu^{\star\text{foc}}\) by maximizing the Gini index (thus our abbreviation MAXGI).
 The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF and a small subset of items thus having large amounts of DIF.
@@
 
-For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini index at a variety of possible \(\mu^\text{foc}\) values.
+For the demonstration data, Figure \ref{fig:ginipath} shows the gini index at a variety of possible \(\mu^\text{foc}\) values.
 The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is quite close to the data-generating value of -1.
@@
 \caption{Maximizing the Gini index (MAXGI) to select the anchor point.}\label{fig:ginipath}
-\end{figure}
-
-The model is then refit with the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\), and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginimilg-1}
-
-}
-
-\caption{The MILG with $\mu^{\text{foc}}$ set to $\mu^{\star\text{foc}} = -0.99$.}\label{fig:ginimilg}
 \end{figure}
@@
 If \(\mu^\text{foc} = 3\), wouldn't we also want to prioritize high \(\theta\) values?
-Accordingly, we set \(g(\theta)\) to be the average of the reference and focal group ability probability density functions:
+Accordingly, we set \(g(\theta)\) to be the weighted average of the reference and focal group ability probability density functions:
 \begin{align}
-g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
+g(\theta) = p^\text{ref} \cdot N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + (1 - p^\text{ref}) \cdot N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})
 \end{align}
+where \(p^\text{ref}\) is the proportion of all students that are in the reference group. In this paper, we always have \(p^\text{ref} = 0.5\).
 
-For our demonstration data simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\).
+For the demonstration data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\).
 In this case, MINBC works perfectly and the anchor point is found to be \(\mu^{\star\text{foc}} = -1\).
@@
 Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point.
-MILG and EM-MILG are somewhat of outliers in that they are methods for visualizing DIF or potential DIF.
 Table \ref{table:allmethods} summarizes all of these methods.
+Figure \ref{fig:allmethodsfig} shows the results for each method on the demonstration data in a MILG.
 
@@
 \centering
-\begin{tabular}{|p{4cm}|p{6cm}|p{4cm}|}
+\begin{tabular}{|p{3cm}|p{4cm}|p{3cm}|p{3cm}|}
 \toprule
 
-Method & Description & Literature \\
+Method & Description & Literature & Demonstration data \\
 
@@
 
-equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
-Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples by \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
-
-multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
-
 all-others-as-anchors (AOAA) & Test if each item has DIF by using all of the other items as anchors (not iterative).
-& Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
+& Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} & Did not find any anchor items \\\hline
 
@@
 Continue iterating until no new items test positive for DIF.
-& Proposed by \cite{drasgow1987study} \\\hline
+& Proposed by \cite{drasgow1987study} & Did not find any anchor items \\\hline
 
@@
 Continue iterating until no new items test positive for DIF.
-& To our knowledge, not proposed or used previously \\\hline
+& To our knowledge, not proposed or used previously & Correctly chose the first five items as anchors \\\hline
 
 equal means clustering (EMC) & Cluster items based on differences in performance across groups and choose one of the clusters to be the anchor cluster.
-& Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} \\\hline
+& Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} & Worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF \\\hline
 
-maximizing Gini index (MAXGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
+maximizing Gini index (MAXGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} & Set $\mu^{\star\text{foc}} = -0.99$, which is quite close to the data generating value of -1 \\\hline
 
-minimizing the area between curves (MINBC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
+minimizing the area between curves (MINBC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} & Set $\mu^{\star\text{foc}} = -1$, which is exactly the data generating value\\
 
@@
 
+\begin{verbatim}
+## Picking joint bandwidth of 0.00587
+\end{verbatim}
+
+\begin{verbatim}
+## Picking joint bandwidth of 0.00591
+\end{verbatim}
+
+\begin{verbatim}
+## Picking joint bandwidth of 0.00546
+\end{verbatim}
+
+\begin{verbatim}
+## Picking joint bandwidth of 0.0055
+\end{verbatim}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/allmethodsfig-1}
+
+}
+
+\caption{MILG for each of the AI methods applied to the demonstration data.
+AOAA-OAT and EMC are identified by anchor items (represented by points), MAXGI and MINBC are identified by setting $\mu^{\text{foc}}$ set to the anchor point $\mu^{\star\text{foc}}$, AOAA and AOAA-AS are not included as they did not find an identifying assumption.}\label{fig:allmethodsfig}
+\end{figure}
+
 \hypertarget{simstudy}{%
@@
 Our goal was to use a relatively realistic data generating process motivated by the case study in Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) wherein some items on a math test also depend on a student's verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
-As described in \protect\hyperlink{intro}{the introduction}, nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
-This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
+Nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
+As described in \protect\hyperlink{intro}{the introduction}, this setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
 One exception is Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}) who draw each student's target ability and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
@@
 \begin{align}
-\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{2j}) = \sigma(\theta_i) \cdot \sigma(a_{2j}\eta_i)
+\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\eta j}) = \sigma(\theta_i) \cdot \sigma(a_{\eta j}\eta_i)
 \end{align}
-where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(a_{2j}\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
+where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(a_{\eta j}\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
 
@@
 In each run, we simulated 10,000 students with half coming from each of the reference and focal groups.
-For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean {[}\(\mu_\theta^\text{ref} = 0\), \(\mu_\eta^\text{ref} = 0\){]} and covariance matrix \(\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}\).
-Abilities for students from the focal group are drawn using the same covariance matrix, but with means {[}\(\mu_\theta^\text{foc} = -0.5\), \(\mu_\eta^\text{foc} = -1\){]}.
+For students from the reference group, target ability and nuisance ability were drawn from the two-dimensional normal distribution with mean {[}\(\mu_\theta^\text{ref} = 0\), \(\mu_\eta^\text{ref} = 0\){]} and covariance matrix \(\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}\).
+Abilities for students from the focal group were drawn using the same covariance matrix, but with means {[}\(\mu_\theta^\text{foc} = -0.5\), \(\mu_\eta^\text{foc} = -1\){]}.
 
 The test always has 12 items, but we varied the number of items with DIF from two to six.
-For items without DIF, \(a_{2j} = \infty\) so that the model reduces to \(\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)\).
-For items with DIF, \(a_{2j}\) is calculated based on Ackerman's (\protect\hyperlink{ref-ackerman1994using}{1994}) angle equation as described in Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}):
+For items without DIF, \(a_{\eta j} = \infty\) so that the model reduces to \(\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)\).
+For items with DIF, \(a_{\eta j}\) is calculated based on Ackerman's (\protect\hyperlink{ref-ackerman1994using}{1994}) angle equation as described in Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}):
 \begin{align}
-\angle_j = \arccos \dfrac{a_{1j}^2}{a_{1j}^2 + a_{2j}^2}.
+\angle_j = \arccos \dfrac{a_{\theta j}^2}{a_{\theta j}^2 + a_{\eta j}^2}.
 \end{align}
@@
 An angle of 45\(^\circ\) indicates that the item loads equally on the target ability and nuisance ability.
-Our simple noncompensatory model has \(a_{1j} = 1\) for all items so the angle equation reduces to
+Our simple noncompensatory model has \(a_{\theta j} = 1\) for all items so the angle equation reduces to
 \begin{align}
-\angle_j = \arccos \dfrac{1}{1 + a_{2j}^2}.
+\angle_j = \arccos \dfrac{1}{1 + a_{\eta j}^2}.
 \end{align}
@@
 \end{align}
-For DIF items, we set \(a_{2j}\) based on angles with equal intervals between 20\(^\circ\) and 60\(^\circ\).
+For DIF items, we set \(a_{\eta j}\) based on angles with equal intervals between 20\(^\circ\) and 60\(^\circ\).
 For example, for a test with three DIF items the angles are 20\(^\circ\), 40\(^\circ\), and 60\(^\circ\).
@@
 
-Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between \(\theta_i\) and \(Pr(y_{ij} = 1)\) with \(\eta_i\) set to the group mean for a test with six DIF items.
+Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between \(\theta_i\) and \(\text{Pr}(y_{ij} = 1)\) with \(\eta_i\) set to the group mean for a test with six items with DIF.
 The items are ordered by the amount of DIF such that \(\angle_{j = 7} = 20^\circ\) up to \(\angle_{j = 12} = 60^\circ\).
@@
 
-\caption{For a 12 item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.}\label{fig:difmap}
+\caption{For a 12-item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.}\label{fig:difmap}
 \end{figure}
 
-Figure \ref{fig:simemmlg} shows the EM-MILG---generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the \protect\hyperlink{anchoritems}{the anchor items section}---for one run using the same item parameters that generated Figure \ref{fig:difmap}.
-As expected, \(\tilde{d_j}\) is about \(\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5\) for the first six items which are DIF free.
+Figure \ref{fig:simemmlg} shows the EM-MILG---generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the \protect\hyperlink{anchoritems}{anchor items section}---for one run using the same item parameters that generated Figure \ref{fig:difmap}.
+As expected, \(\tilde{d_j}\) is about \(\mu_\theta^\text{ref} - \mu_\theta^\text{foc}= 0 - (-0.5) = 0.5\) for the first six items which are DIF free.
 For the last six items, \(\tilde{d_j}\) increases as \(\angle_j\) increases.
@@
 
-\caption{For a 12 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.}\label{fig:simemmlg}
+\caption{The EM-MILG for one run using the same item parameters as generated Figure ef{fig:difmap}}\label{fig:simemmlg}
 \end{figure}
@@
 
-(START BEN DIDNT LOOK AT)
-
 EMC performed better than AOAA and AOAA-AS, but worse than the other methods.
 MINBC and MAXGI performed similarly well with MINBC estimating the achievement gap with more precision but more bias than MAXGI, especially for tests with more than four DIF items.
-We hypothesize that MINBC's susceptibility to bias results from considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.
+We hypothesize that MINBC's susceptibility to bias results from considering every item.
+AOAA-OAT, for example, completely disregards an item once it's removed from the anchor set.
 
@@
 
-(STOP BEN DIDNT LOOK AT)
-
 \hypertarget{discussion-and-summary}{%
@@
 
-Validity hinges on measurement instruments being relatively free from DIF.
+Validity hinges on measurement instruments being relatively free of DIF.
 Such instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
@@
 Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
-This is concerning given both their widespread use and the fact that other method---AOAA-OAT, EMC, MINGI, and MAXBC---demonstrate superior performance.
+This is concerning given both their widespread use and the fact that other methods---AOAA-OAT, EMC, MINGI, and MAXBC---demonstrate superior performance.
 
@@
 
-Future work should test these methods' performance under a greater variety of data-generating conditions.
+We emphasize that the EM-MILG gives a clear view of the possibility of DIF on a test.
+For example, in Figure \ref{fig:simemmlg} it's clear that the gap between focal and reference group performance across items on the test varies greatly.
+By assuming that the majority of items do not contain DIF, AI methods essentially make an educated guess about the relative responsibility of group ability differences vs.~DIF.
+Our simulation focused on a scenario wherein half of the items are DIF-free and the other half have escalating amounts of DIF.
+Different patterns of DIF (as visualized in an EM-MILG) will pose different challenges.
+
+For example, imagine a test in which the focal group outperforms the reference group by one logit on half of the items, and the reference group outperforms the focal group by one logit on the other half of the items.
+What should an AI method conclude in this case?
+AOAA and AOAA-AS will likely (perhaps rightfully) fail to find any anchor items; MINBC will choose an anchor point close to zero (and will find DIF in all items); MAXGI will arbitrarily choose an anchor point within one of the groups of items (and will find DIF in the other group of items); and AOAA-OAT and EMC will arbitrarily choose one of the groups of items to be anchor items (and, like MAXGI, will find DIF in the other group of items).
+This example illustrates that some patterns of DIF are such that the analyst should probably remain agnostic.
+In these cases, we suggest that the analyst report the nature of potential DIF using a EM-MILG.
+Future work might produce methods for quantifying what is visualized in a EM-MILG, essentially making a metric for the amount of potential DIF on a test.
+
+Moreover, future work should test these methods' performance under a greater variety of data-generating conditions.
 For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
-Furthermore, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.
+Finally, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting other types of item response models.
 

### 2020-03-23 paper/paper_1_intro.Rmd
finished paper print 6 and shipped back to ben

@@
 For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies "DIF".
+In this case, math ability is the "target ability", and verbal ability is the "nuisance ability".
+Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies "DIF".
 
@@
 For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
-For example, to translate the @kopf2015framework simulation from the DIF-as-item-property to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students.
-Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension.
-The target ability is the same as unidimensional ability above where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group.
-The nuisance ability takes one of two values: ${\eta_i}^{\text{ref}} = 0$ for students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for students in the focal group.
-We then reply on a 2PL compensatory 2PL model; slopes on the target ability are $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability is $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise.
-Again, 0.6 is the magnitude of DIF in logits.
 According to the two-dimensional compensatory 2PL model [@thissen1986taxonomy], the probability that student $i$ responds correctly to item $j$ is
 \begin{align}
-\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
+\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\theta j}, a_{\eta j}, b_j) = \sigma(a_{\theta j}\theta_i + a_{\eta j}\eta_i + b_j).
 \end{align}
-For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
+where $\theta$ is target ability, $\eta$ is nuisance ability, $a_{\theta j}$ is an item's slope on target ability, and $a_{\eta j}$ is an item's slope on nuisance ability.
+
+For example, to translate the @kopf2015framework simulation from the DIF-as-item-property to the DIF-as-student-property view, $b_j$ is set to ${b_j}^{\text{ref}}$, $\theta$ stays the same, $a_{\theta j} = 1$ for all items (consistent with the Rasch model), and $\eta$ takes one of two values: ${\eta_i}^{\text{ref}} = 0$ for students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for students in the focal group.
+For items with DIF, $a_{\eta j} = 0.6$, and for items without DIF, $a_{\eta j} = 0$.
+Again, 0.6 is the magnitude of DIF in logits.
+^[The reader may notice that this is just one possible specification.
+Any specification that meets the following criteria has the same effect: First, $a_{\eta j}\eta_i$ must be 0.6 lower for the focal group on items with DIF.
+Second, $a_{\eta j}\eta_i$ must be equal across groups for items without DIF.] For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
 This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.

### 2020-03-23 paper/paper_2a_anchor_items.Rmd
finished paper print 6 and shipped back to ben

@@
 The IRT-based likelihood ratio test (LRT) is effective at detecting DIF [@meade2012solving].
-Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure [@holland1986differential] (which, for example, has been shown to perform no better—and in some cases worse—than IRT-based methods [@swaminathan1990detecting]).
+Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure [@holland1986differential], which has been shown to perform no better—and in some cases worse—than IRT-based methods [@swaminathan1990detecting]).
+Within the IRT-framework, we—as is common—estimate models using marginal maximum likelihood estimation [@bock1981marginal] and identify the scale by setting $\mu^\text{ref} = 0$.
 
-However, use of LRT requires an identifying assumption.
+Use of LRT requires an identifying assumption.
 The assumption is needed so as to link groups.
@@
 Thus, we're interested in *agnostic identification* methods (hereafter referred to as "AI methods").
+The fundamental challenge is to disentangle estimation of $\hat\mu^\text{foc} - \mu^\text{ref} = \hat\mu^\text{foc} - 0 = \hat\mu^\text{foc}$, the difference in mean abilities, from the estimation of $\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$, the difference in item easiness.
+There are two types of AI methods: The most common, the use of anchor items, selects a group of items that are assumed to be DIF-free.
+These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.
+The other type, the use of an anchor point [@strobl2018anchor], directly sets the difference in mean abilities to some value, $\mu^{\star\text{foc}}$.
+This anchor point identifies the model, and all items can be tested for DIF using an LRT.
+
+In this section, we summarize existing AI methods and propose a variety of extensions.
 For simplicity, we focus on the Rasch model to isolate the fundamental issues in "the AI problem".
 Further, so as to offer a coherent framework, we sometimes edit names of existing methods.
-^[We recognize that others have done the same [e.g., @kopf2015framework], and that we risk contributing to a proliferation of names.]
-
-In this section, we summarize existing AI methods and propose a variety of extensions.
-We use a (simulated) demonstration dataset to illustrate key features of the different methods.
+^[We recognize that others have done the same [e.g., @kopf2015framework], and that we risk contributing to a proliferation of names.] We use a (simulated) demonstration dataset to illustrate key features of the different methods.
 ^[This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.] The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
@@
 
-The fundamental challenge is to use an AI method to disentangle estimation of $\hat\mu^\text{foc} - \mu^\text{ref}$, the difference in mean abilities, from the estimation of $\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$, the difference in item easiness.
-The most common AI approaches identify a group of anchor items that are assumed to be DIF-free.
-These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.
-
-Throughout this paper, we estimate models using marginal maximum likelihood estimation [@bock1981marginal].
-In addition, we follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$.
+The AI methods described in this paper are summarized in Table \ref{table:allmethods}.
+In addition the results for each method on the demonstration data are shown in a MILG in Figure \ref{fig:allmethodsfig}.
 
@@
 
-While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position for understanding the scale of the problem in their data.
+While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position to understand the scale of the problem in their data.
 We propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
@@
 As a result, all differences in performance—either from group ability differences or DIF—manifest in the item easiness difference parameter $\tilde{d_j}$.
-The tilde (e.g., $\tilde{d_j}$) is used to denote parameters estimated with $\mu^\text{foc}$ set to $0$—as compared to the hat (e.g., $\hat{d_j}$), which is used for parameter estimates from the properly identified model.
+The tilde (e.g., $\tilde{d_j}$) is used to denote parameters estimated with $\mu^\text{foc}$ set to $0$—as compared to the hat (e.g., $\hat{d_j}$), which is used for parameter estimates from a properly identified model.
 
@@
 
-Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data.
+Figure \ref{fig:emmilg} shows the EM-MILG for the demonstration data.
 We emphasize that the EM-MILG contains all possible information about the difference in group performance.
 The analyst might assume that—because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits—items 1-5 are unbiased.
-Indeed, all AI methods rest on the logic that its the minority of items that contain DIF.
+This might seem to be a strong assumption, but all AI methods assume—in one way or another—that, if there is DIF, its a minority of items that are responsible.
 These unbiased items are known as anchor items; setting anchor items is the most common identification assumption.
-After anchor items have been selected, the model is refit where the assumption that $\mu^\text{foc} = 0$ is dropped and replaced with the assumption that the anchor items are DIF-free (i.e., $d_j = 0$).
-Of course, the demonstration data is designed such that there are obvious groups; this will not typically be the case.
+After anchor items have been selected, the model is refit where the identifying assumption that $\mu^\text{foc} = 0$ is dropped and replaced with the identifying assumption that the anchor items are DIF-free (i.e., $d_j = 0$ for anchor items).
+Of course, the demonstration data is designed such that there are obvious groups; this will not necessarily typically be the case.
 
@@
 
-The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the final model.
+The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with a properly identified model.
 Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
@@
     mutate(difference_in_easy = difference_in_easy - difference_in_easy[anchor][1]) %>% 
+    mutate(item = as_factor(paste0("Item ", parse_number(as.character(item))))) %>%
     ggplot(aes(x = difference_in_easy, y = item)) +
@@
             gather(var, val) %>% 
+            mutate(var = as_factor(paste0("Item ", parse_number(as.character(var))))) %>%
             filter(val != 0),
@@
     ) +
-    geom_point() +
+    geom_point(
+        data =
+            out_3_biased_items$AOAA_OAT_final_dif[[1]] %>% 
+            mutate(difference_in_easy = difference_in_easy - difference_in_easy[anchor][1]) %>%
+            mutate(item = as_factor(paste0("Item ", parse_number(as.character(item))))) %>% 
+            filter(a_ref_easy == b_foc_easy)
+    ) +
     labs(
@@
 The items for which the flexible model outperforms the baseline model (typically based on a $\chi^2$ test) are identified as having DIF, and the rest of the items become anchor items.
-AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).
+AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the baseline to the flexible model).
 
@@
 On a practical level, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
-
 Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies [@stark2006detecting].
@@
 The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF", which is a significant problem in applications of AI methods as discussed below.
+This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF", which is a significant problem in applications of AI methods.
 
@@
 We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-Appending "all-significant" indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set.
+We append "all-significant" to indicate that all items that test positive for DIF are removed from the anchor set in each iteration.
 AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.
@@
 @woods2009empirical suggested a more straightforward, one-step method: Use AOAA and select the four items that exhibit the least amount of DIF as anchor items.
-It's unclear if one should proceed if those four items also test positive for DIF.
 
@@
 6].
-Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the $\chi^2$ test statistic, is removed from the anchor set.
+Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF—based on the $\chi^2$ test statistic—is removed from the anchor set.
 The process continues iteratively until no new items display DIF.
@@
 As a result, we believe that AOAA-OAT is less likely to be "fooled" by artifical DIF.
-AOAA-OAT is not currently implemented in the R package mirt.
+AOAA-OAT is not currently implemented in the mirt R package.
 
 <!-- improvement would be to explain what more homogeneous means in more detail -->
+<!-- the two kopf 2015 papers love the iterative forward SA. I bet AOAA-OAT is as good or better while fitting fewer models -->
 
 Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF (i.e., it is the least homogeneous in the sense that the group difference in performance is most different from the average group difference in performance on the other two items).
+The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF (i.e., because the group difference in performance is most different from the average group difference in performance on the other two items).
 In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-To work, AOAA-OAT has two requirements: First, that at least two items do not have DIF, and second, that the set of items without DIF are more homogeneous than other sets of items.
-
-#### Summary and performance
-
-<!-- the two kopf 2015 papers love the iterative forward SA. I bet AOAA-OAT is as good or better while fitting fewer models -->
-
-It's useful to remember that AOAA is not an iterative procedure.
-In constrast, the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method.
-All of the methods described in this paper are summarized in Table \ref{table:allmethods}.
-For our demonstration data, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified.
-On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.
 
@@
 @bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster".
-They pointed out that one way around the unidentifiability issue is to consider only relative item parameters.
+They pointed out that one way around the lack of model identifiability is to consider only relative item parameters.
 For each group, the relative easinesses for each pair of items can be stored in the (square) matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$, where $x$ and $y$ both index the items.
@@
 @pohl2017cluster extended their work by proposing one such process.
-$\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
+They point out that $\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
 Accordingly, they recommend k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC.
@@
 We propose a new cluster-based approach, which we call "equal means clustering" (EMC).
-Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$, which, as described in the [EM-MILG section](#emmilg), comes from setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, and thus the name "equal means clustering").
-
+Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$, which, as described in the [EM-MILG section](#emmilg), comes from setting $\mu^\text{foc} = 0$.
+Recall that $\mu^\text{ref}$ is to $0$ to identify the scale, and thus the name "equal means clustering".
 Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating.
@@
 If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected.
-For our demonstration data, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.

### 2020-03-23 paper/paper_3_sim.Rmd
finished paper print 6 and shipped back to ben

@@
 Our goal was to use a relatively realistic data generating process motivated by the case study in @ackerman1992didactic wherein some items on a math test also depend on a student’s verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
-As described in [the introduction](#intro), nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
-This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
+Nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
+As described in [the introduction](#intro), this setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
 One exception is @walker2017using who draw each student's target ability and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
@@
 \begin{align}
-\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{2j}) = \sigma(\theta_i) \cdot \sigma(a_{2j}\eta_i)
+\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\eta j}) = \sigma(\theta_i) \cdot \sigma(a_{\eta j}\eta_i)
 \end{align}
-where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $a_{2j}$ is the item's loading on nuisance ability [@demars2016partially].
+where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $a_{\eta j}$ is the item's loading on nuisance ability [@demars2016partially].
 
@@
 In each run, we simulated 10,000 students with half coming from each of the reference and focal groups.
-For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean [$\mu_\theta^\text{ref} = 0$, $\mu_\eta^\text{ref} = 0$] and covariance matrix $\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}$.
-Abilities for students from the focal group are drawn using the same covariance matrix, but with means [$\mu_\theta^\text{foc} = -0.5$, $\mu_\eta^\text{foc} = -1$].
+For students from the reference group, target ability and nuisance ability were drawn from the two-dimensional normal distribution with mean [$\mu_\theta^\text{ref} = 0$, $\mu_\eta^\text{ref} = 0$] and covariance matrix $\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}$.
+Abilities for students from the focal group were drawn using the same covariance matrix, but with means [$\mu_\theta^\text{foc} = -0.5$, $\mu_\eta^\text{foc} = -1$].
 
 The test always has 12 items, but we varied the number of items with DIF from two to six.
-For items without DIF, $a_{2j} = \infty$ so that the model reduces to $\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)$.
-For items with DIF, $a_{2j}$ is calculated based on Ackerman's -@ackerman1994using angle equation as described in @walker2017using:
+For items without DIF, $a_{\eta j} = \infty$ so that the model reduces to $\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)$.
+For items with DIF, $a_{\eta j}$ is calculated based on Ackerman's -@ackerman1994using angle equation as described in @walker2017using:
 \begin{align}
-\angle_j = \arccos \dfrac{a_{1j}^2}{a_{1j}^2 + a_{2j}^2}.
+\angle_j = \arccos \dfrac{a_{\theta j}^2}{a_{\theta j}^2 + a_{\eta j}^2}.
 \end{align}
@@
 An angle of 45$^\circ$ indicates that the item loads equally on the target ability and nuisance ability.
-Our simple noncompensatory model has $a_{1j} = 1$ for all items so the angle equation reduces to
+Our simple noncompensatory model has $a_{\theta j} = 1$ for all items so the angle equation reduces to
 \begin{align}
-\angle_j = \arccos \dfrac{1}{1 + a_{2j}^2}.
+\angle_j = \arccos \dfrac{1}{1 + a_{\eta j}^2}.
 \end{align}
@@
 \end{align}
-For DIF items, we set $a_{2j}$ based on angles with equal intervals between 20$^\circ$ and 60$^\circ$.
+For DIF items, we set $a_{\eta j}$ based on angles with equal intervals between 20$^\circ$ and 60$^\circ$.
 For example, for a test with three DIF items the angles are 20$^\circ$, 40$^\circ$, and 60$^\circ$.
@@
 
-Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean for a test with six DIF items.
+Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $\text{Pr}(y_{ij} = 1)$ with $\eta_i$ set to the group mean for a test with six items with DIF.
 The items are ordered by the amount of DIF such that $\angle_{j = 7} = 20^\circ$ up to $\angle_{j = 12} = 60^\circ$.
 
-```{r difmap, fig.cap = 'For a 12 item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.', out.width="70%", warning = FALSE, message = FALSE}
+```{r difmap, fig.cap = 'For a 12-item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.', out.width="70%", warning = FALSE, message = FALSE}
 out$bigsim[[5]]$pars[[1]] %>% 
@@
 
-Figure \ref{fig:simemmlg} shows the EM-MILG—generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the [the anchor items section](#anchoritems)—for one run using the same item parameters that generated Figure \ref{fig:difmap}.
-As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first six items which are DIF free.
+Figure \ref{fig:simemmlg} shows the EM-MILG—generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the [anchor items section](#anchoritems)—for one run using the same item parameters that generated Figure \ref{fig:difmap}.
+As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{ref} - \mu_\theta^\text{foc}=  0 - (-0.5) = 0.5$ for the first six items which are DIF free.
 For the last six items, $\tilde{d_j}$ increases as $\angle_j$ increases.
 
-```{r simemmlg, fig.cap = 'For a 12 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
+```{r simemmlg, fig.cap = 'The EM-MILG for one run using the same item parameters as generated Figure \ref{fig:difmap}', out.width="70%", warning = FALSE, message = FALSE}
 out$bigsim[[5]]$intuitive_mod[[1]] %>%
@@
     ) +
-    scale_y_continuous(labels = scales::percent, limits = c(0, 0.5))
+    scale_y_continuous(labels = scales::percent, limits = c(0, 0.5), breaks = c(0, 0.25, 0.5))
 ```
@@
 
-(STOP BEN DIDNT LOOK AT)
-
 # Discussion and summary
 
-Validity hinges on measurement instruments being relatively free from DIF.
+Validity hinges on measurement instruments being relatively free of DIF.
 Such instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
@@
 Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
-This is concerning given both their widespread use and the fact that other method—AOAA-OAT, EMC, MINGI, and MAXBC—demonstrate superior performance.
+This is concerning given both their widespread use and the fact that other methods—AOAA-OAT, EMC, MINGI, and MAXBC—demonstrate superior performance.
 
@@
 
-Future work should test these methods' performance under a greater variety of data-generating conditions.
+We emphasize that the EM-MILG gives a clear view of the possibility of DIF on a test.
+For example, in Figure \ref{fig:simemmlg} it's clear that the gap between focal and reference group performance across items on the test varies greatly.
+By assuming that the majority of items do not contain DIF, AI methods essentially make an educated guess about the relative responsibility of group ability differences vs.
+DIF.
+Our simulation focused on a scenario wherein half of the items are DIF-free and the other half have escalating amounts of DIF.
+Different patterns of DIF (as visualized in an EM-MILG) will pose different challenges.
+
+For example, imagine a test in which the focal group outperforms the reference group by one logit on half of the items, and the reference group outperforms the focal group by one logit on the other half of the items.
+What should an AI method conclude in this case?
+AOAA and AOAA-AS will likely (perhaps rightfully) fail to find any anchor items; MINBC will choose an anchor point close to zero (and will find DIF in all items); MAXGI will arbitrarily choose an anchor point within one of the groups of items (and will find DIF in the other group of items); and AOAA-OAT and EMC will arbitrarily choose one of the groups of items to be anchor items (and, like MAXGI, will find DIF in the other group of items).
+This example illustrates that some patterns of DIF are such that the analyst should probably remain agnostic.
+In these cases, we suggest that the analyst report the nature of potential DIF using a EM-MILG.
+Future work might produce methods for quantifying what is visualized in a EM-MILG, essentially making a metric for the amount of potential DIF on a test.
+
+Moreover, future work should test these methods' performance under a greater variety of data generating conditions.
 For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
-Furthermore, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.
+Finally, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting other types of item response models.
 

### 2020-03-23 R/dif_intuitive.R

@@ -24,6 +24,7 @@ draws_df_to_logit_plot <- function(draws_df){
     draws_df %>%
         select(-run) %>%
         gather(var, val, factor_key = TRUE) %>%
+        mutate(var = as_factor(paste0("Item ", parse_number(as.character(var))))) %>%
         ggplot(aes(x = val, y = var)) +
         ggridges::geom_density_ridges() +
         labs(x = "", y = "")

### 2020-10-14 paper/paper.tex
apparently i did a bunch of stuff and never committed it

@@
 \makeatother
-\subtitle{Ben Stenhaug, Ben Domingue, and Mike Frank\\
+\subtitle{Ben Stenhaug, Mike Frank, and Benjamin W. Domingue\\
 Stanford University}
@@
 We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
+Given the central importance of AI techniques to subsequent examinations of item-level bias, we believe these constitute important advances for the field.
 \clearpage
@@
 For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values that they obtained from a previous study.
+They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values that they obtained from a previous study (Wang, Shih, and Sun \protect\hyperlink{ref-wang2012dif}{2012}).
 They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
@@
 For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
-According to the two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}), the probability that student \(i\) responds correctly to item \(j\) is
+The two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}) asserts that the probability that student \(i\) responds correctly to item \(j\) is
 \begin{align}
@@
 
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view, \(b_j\) is set to \({b_j}^{\text{ref}}\), \(\theta\) stays the same, \(a_{\theta j} = 1\) for all items (consistent with the Rasch model), and \(\eta\) takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group.
-For items with DIF, \(a_{\eta j} = 0.6\), and for items without DIF, \(a_{\eta j} = 0\) Again, 0.6 is the magnitude of DIF in logits.\footnote{The reader may notice that this is just one possible specification.
+We can use this model to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view; \(b_j\) is set to \({b_j}^{\text{ref}}\), \(\theta\) stays the same, \(a_{\theta j} = 1\) for all items (consistent with the Rasch model), and \(\eta\) takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group.
+For items with DIF, \(a_{\eta j} = 0.6\), and for items without DIF, \(a_{\eta j} = 0\).
+Again, 0.6 is the magnitude of DIF in logits.\footnote{The reader may notice that this is just one possible specification.
 Any specification that meets the following criteria has the same effect: First, \(a_{\eta j}\eta_i\) must be 0.6 lower for the focal group on items with DIF.
 Second, \(a_{\eta j}\eta_i\) must be equal across groups for items without DIF.} For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
+
+However, this translation between views reveals an important point.
+As a side effect of this approach, simulation studies have tended to focus on the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
+Much as students vary in the target ability---a belief in such variation is central to the measurement enterprise---we suspect they also vary on additional ability dimensions.
+Indeed, previous work has shown that the best fitting model for item response data is often multidimensional where each dimension has significant variation (e.g., Furlow, Raiford Ross, and Gagné \protect\hyperlink{ref-furlow2009impact}{2009}).
+These dimensions are typically positively correlated, and the greater that correlation, the better unidimensional models tend to perform (Kirisci, Hsu, and Yu \protect\hyperlink{ref-kirisci2001robustness}{2001}).
+Explicitly accounting for this variation---which we do in our simulation study---should yield results more indicative of how DIF methods will perform in practice.
 
@@
 Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
-This approach depends upon a crucial identifying assumption; this assumption is the target of our investigation.
+This approach depends upon a crucial identifying assumption, which we discuss at length below, is the target of our investigation.
 
@@
 The IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
-Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})).
+Given both the efficacy of the LRT approach and our preference for methods that exist within the IRT-framework, we do not discuss alternative methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})).
 Within the IRT-framework, we---as is common---estimate models using marginal maximum likelihood estimation (Bock and Aitkin \protect\hyperlink{ref-bock1981marginal}{1981}) and identify the scale by setting \(\mu^\text{ref} = 0\).
+Both of these choices are---for our purposes---completely inconsequential.
 
-Use of LRT requires an identifying assumption.
+The use of LRT requires an identifying assumption.
 The assumption is needed so as to link groups.
-In the typical setting, no a prior assumptions---about either relative group ability or which items might have DIF---can be made.
+There are two possibilities.
+One possibility is to utilize information about the distribution of the focal group; in particular, knowledge that the focal group and reference group have the same mean ability would be valuable.
+Alternatively, we can use information from a set of unbiased items to derive such information.
+In the equating literature, the former is a \enquote{non-common item random groups} design and the latter is a \enquote{common-item nonequivalent groups} design (Cook and Paterson \protect\hyperlink{ref-cook1987problems}{1987}; Topczewski et al.
+\protect\hyperlink{ref-topczewski2013comparison}{2013}).
+
+However, in the typical setting, no a prior assumptions---about either relative group ability or which items might have DIF---can be made.
 Thus, we're interested in \emph{agnostic identification} methods (hereafter referred to as \enquote{AI methods}).
@@
 For simplicity, we focus on the Rasch model to isolate the fundamental issues in \enquote{the AI problem}.
-Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.} We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.} We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we conduct an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Our simulation is similar to our multidimensional alternative to the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) approach discussed above.
@@
 We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability.
-Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.
+From that perspective, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.
 To be sure, this data is used only for demonstrative purposes.
 We compare the performance of the AI methods in a more realistic scenario in the \protect\hyperlink{simstudy}{the simulation study}.
-
-The AI methods described in this paper are summarized in Table \ref{table:allmethods}.
-In addition the results for each method on the demonstration data are shown in a MILG in Figure \ref{fig:allmethodsfig}.
 
@@
 
-AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.
+Before we begin to discuss different AI methods, we begin by offering a suggestion for examining the identification problem in a given dataset.
+This approach is motivated by our observation that AI methods typically work like a black box.
+The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.
+From our perspective, tools that help an analyst to better understand their data are of great value.
 
-While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position to understand the scale of the problem in their data.
-We propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
-EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting \(\mu^\text{foc} = 0\).
+To that end, we propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
+EM-MILG begins by fitting a multigroup unidimensional Rasch model to the data that is identified by arbitrarily setting \(\mu^\text{foc} = 0\) (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}).
 Given that we also assume \(\mu^\text{ref} = 0\), this is equivalent to the assumption that the groups have equal mean ability.
@@
 
-Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data.
+Figure \ref{fig:emmilg} shows the EM-MILG for the demonstration data.
 We emphasize that the EM-MILG contains all possible information about the difference in group performance.
 The analyst might assume that---because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits---items 1-5 are unbiased.
-This might seem to be a strong assumption, but all AI methods assume---in one way or another---that, if there is DIF, its a minority of items that are responsible.
+This might seem to be a strong assumption, bit it is one we emphasize: All AI methods assume---in one way or another---that, if there is DIF, its a minority of items that are responsible (in this case, items 6-8).
 These unbiased items are known as anchor items; setting anchor items is the most common identification assumption.
 After anchor items have been selected, the model is refit where the identifying assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the identifying assumption that the anchor items are DIF-free (i.e., \(d_j = 0\) for anchor items).
-Of course, the demonstration data is designed such that there are obvious groups; this will not necessarily typically be the case.
+Of course, the demonstration data is designed such that there are obvious groups; this need not be the case in practice.
 
@@
 
-We described a variety of AI methods and their corresponding acronymns.
+We described a variety of AI methods and their corresponding acronyms.
 Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point.
@@
 
-\begin{verbatim}
-## Picking joint bandwidth of 0.00587
-\end{verbatim}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.00591
-\end{verbatim}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.00546
-\end{verbatim}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.0055
-\end{verbatim}
-
 \begin{figure}[H]
@@
 \caption{MILG for each of the AI methods applied to the demonstration data.
-AOAA-OAT and EMC are identified by anchor items (represented by points), MAXGI and MINBC are identified by setting $\mu^{\text{foc}}$ set to the anchor point $\mu^{\star\text{foc}}$, AOAA and AOAA-AS are not included as they did not find an identifying assumption.}\label{fig:allmethodsfig}
+AOAA-OAT and EMC are identified by anchor items (represented by points).
+MAXGI and MINBC are identified by setting $\mu^{\text{foc}}$ set to the anchor point $\mu^{\star\text{foc}}$.
+AOAA and AOAA-AS are not included as they did not find an identifying assumption.}\label{fig:allmethodsfig}
 \end{figure}
@@
 For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows.
+However, we believe the computational costs are worthwhile given the performance of this method relative to more widely implemented alternatives.
 To increase AOAA-OAT's use, we recommend its implementation (perhaps as the default) in popular IRT software such as the mirt R package.
@@
 
-Moreover, future work should test these methods' performance under a greater variety of data-generating conditions.
+Moreover, future work should test these methods' performance under a greater variety of data generating conditions.
 For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
@@
 
+\leavevmode\hypertarget{ref-cook1987problems}{}%
+Cook, Linda L, and Nancy S Paterson.
+1987.
+``Problems Related to the Use of Conventional and Item Response Theory Equating Methods in Less Than Optimal Circumstances.'' \emph{Applied Psychological Measurement} 11 (3): 225--44.
+
 \leavevmode\hypertarget{ref-demars2016partially}{}%
@@
 Drasgow, Fritz. 1987. ``Study of the Measurement Bias of Two Standardized Psychological Tests.'' \emph{Journal of Applied Psychology} 72 (1): 19.
+
+\leavevmode\hypertarget{ref-furlow2009impact}{}%
+Furlow, Carolyn F, Terris Raiford Ross, and Phill Gagné.
+2009.
+``The Impact of Multidimensionality on the Detection of Differential Bundle Functioning Using Simultaneous Item Bias Test.'' \emph{Applied Psychological Measurement} 33 (6): 441--64.
 
@@
 Holland, Paul W, and Dorothy T Thayer. 1986. ``Differential Item Functioning and the Mantel-Haenszel Procedure.'' \emph{ETS Research Report Series} 1986 (2): i--24.
+
+\leavevmode\hypertarget{ref-kirisci2001robustness}{}%
+Kirisci, Levent, Tse-chi Hsu, and Lifa Yu.
+2001.
+``Robustness of Item Parameter Estimation Programs to Assumptions of Unidimensionality and Normality.'' \emph{Applied Psychological Measurement} 25 (2): 146--62.
 
@@
 
+\leavevmode\hypertarget{ref-topczewski2013comparison}{}%
+Topczewski, Anna, Zhongmin Cui, David Woodruff, Hanwei Chen, and Yu Fang.
+2013.
+``A Comparison of Four Linear Equating Methods for the Common-Item Nonequivalent Groups Design Using Simulation Methods.
+ACT Research Report Series, 2013 (2).'' \emph{ACT, Inc.}
+
 \leavevmode\hypertarget{ref-walker2017using}{}%
@@
 ``Using a Multidimensional Irt Framework to Better Understand Differential Item Functioning (Dif): A Tale of Three Dif Detection Procedures.'' \emph{Educational and Psychological Measurement} 77 (6): 945--70.
+
+\leavevmode\hypertarget{ref-wang2012dif}{}%
+Wang, Wen-Chung, Ching-Lin Shih, and Guo-Wei Sun.
+2012.
+``The Dif-Free-Then-Dif Strategy for the Assessment of Differential Item Functioning.'' \emph{Educational and Psychological Measurement} 72 (4): 687--708.
 

### 2020-10-14 paper/paper_1_intro.Rmd
apparently i did a bunch of stuff and never committed it

@@
 For example, @kopf2015framework simulate students as belonging to either a reference or focal group.
-They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values that they obtained from a previous study.
+They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values that they obtained from a previous study [@wang2012dif].
 They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits.
@@
 For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
-According to the two-dimensional compensatory 2PL model [@thissen1986taxonomy], the probability that student $i$ responds correctly to item $j$ is
+The two-dimensional compensatory 2PL model [@thissen1986taxonomy] asserts that the probability that student $i$ responds correctly to item $j$ is
 \begin{align}
@@
 
-For example, to translate the @kopf2015framework simulation from the DIF-as-item-property to the DIF-as-student-property view, $b_j$ is set to ${b_j}^{\text{ref}}$, $\theta$ stays the same, $a_{\theta j} = 1$ for all items (consistent with the Rasch model), and $\eta$ takes one of two values: ${\eta_i}^{\text{ref}} = 0$ for students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for students in the focal group.
+We can use this model to translate the @kopf2015framework simulation from the DIF-as-item-property to the DIF-as-student-property view; $b_j$ is set to ${b_j}^{\text{ref}}$, $\theta$ stays the same, $a_{\theta j} = 1$ for all items (consistent with the Rasch model), and $\eta$ takes one of two values: ${\eta_i}^{\text{ref}} = 0$ for students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for students in the focal group.
 For items with DIF, $a_{\eta j} = 0.6$, and for items without DIF, $a_{\eta j} = 0$.
@@
 Second, $a_{\eta j}\eta_i$ must be equal across groups for items without DIF.] For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
+
+However, this translation between views reveals an important point.
+As a side effect of this approach, simulation studies have tended to focus on the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
+Much as students vary in the target ability—a belief in such variation is central to the measurement enterprise—we suspect they also vary on additional ability dimensions.
+Indeed, previous work has shown that the best fitting model for item response data is often multidimensional where each dimension has significant variation [e.g., @furlow2009impact].
+These dimensions are typically positively correlated, and the greater that correlation, the better unidimensional models tend to perform [@kirisci2001robustness].
+Explicitly accounting for this variation—which we do in our simulation study—should yield results more indicative of how DIF methods will perform in practice.
 
@@
 Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
-This approach depends upon a crucial identifying assumption; this assumption is the target of our investigation.
+This approach depends upon a crucial identifying assumption, which we discuss at length below, is the target of our investigation.

### 2020-10-14 paper/paper_2a_anchor_items.Rmd
apparently i did a bunch of stuff and never committed it

@@
 The IRT-based likelihood ratio test (LRT) is effective at detecting DIF [@meade2012solving].
-Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure [@holland1986differential], which has been shown to perform no better—and in some cases worse—than IRT-based methods [@swaminathan1990detecting]).
+Given both the efficacy of the LRT approach and our preference for methods that exist within the IRT-framework, we do not discuss alternative methods like the Mantel-Haenszel procedure [@holland1986differential], which has been shown to perform no better—and in some cases worse—than IRT-based methods [@swaminathan1990detecting].
 Within the IRT-framework, we—as is common—estimate models using marginal maximum likelihood estimation [@bock1981marginal] and identify the scale by setting $\mu^\text{ref} = 0$.
+Both of these choices are—for our purposes—completely inconsequential.
 
-Use of LRT requires an identifying assumption.
+The use of LRT requires an identifying assumption.
 The assumption is needed so as to link groups.
-In the typical setting, no a prior assumptions—about either relative group ability or which items might have DIF—can be made.
+There are two possibilities.
+One possibility is to utilize information about the distribution of the focal group; in particular, knowledge that the focal group and reference group have the same mean ability would be valuable.
+Alternatively, we can use information from a set of unbiased items to derive such information.
+In the equating literature, the former is a "non-common item random groups" design and the latter is a "common-item nonequivalent groups" design [@cook1987problems; @topczewski2013comparison].
+
+However, in the typical setting, no a prior assumptions—about either relative group ability or which items might have DIF—can be made.
 Thus, we're interested in *agnostic identification* methods (hereafter referred to as "AI methods").
@@
 ^[We recognize that others have done the same [e.g., @kopf2015framework], and that we risk contributing to a proliferation of names.] We use a (simulated) demonstration dataset to illustrate key features of the different methods.
-^[This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.] The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
+^[This data is used solely for illustrative purposes; we conduct an extensive simulation study later in the paper.] The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
 Our simulation is similar to our multidimensional alternative to the @kopf2015framework approach discussed above.
@@
 We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability.
-Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.
+From that perspective, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.
 To be sure, this data is used only for demonstrative purposes.
 We compare the performance of the AI methods in a more realistic scenario in the [the simulation study](#simstudy).
-
-The AI methods described in this paper are summarized in Table \ref{table:allmethods}.
-In addition the results for each method on the demonstration data are shown in a MILG in Figure \ref{fig:allmethodsfig}.
 
@@
 
-AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.
+Before we begin to discuss different AI methods, we begin by offering a suggestion for examining the identification problem in a given dataset.
+This approach is motivated by our observation that AI methods typically work like a black box.
+The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.
+From our perspective, tools that help an analyst to better understand their data are of great value.
 
-While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position to understand the scale of the problem in their data.
-We propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
-EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting $\mu^\text{foc} = 0$.
+To that end, we propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst.
+EM-MILG begins by fitting a multigroup unidimensional Rasch model to the data that is identified by arbitrarily setting $\mu^\text{foc} = 0$ [@chalmers2012mirt].
 Given that we also assume $\mu^\text{ref} = 0$, this is equivalent to the assumption that the groups have equal mean ability.
@@
 The analyst might assume that—because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits—items 1-5 are unbiased.
-This might seem to be a strong assumption, but all AI methods assume—in one way or another—that, if there is DIF, its a minority of items that are responsible.
+This might seem to be a strong assumption, bit it is one we emphasize: All AI methods assume—in one way or another—that, if there is DIF, its a minority of items that are responsible (in this case, items 6-8).
 These unbiased items are known as anchor items; setting anchor items is the most common identification assumption.
 After anchor items have been selected, the model is refit where the identifying assumption that $\mu^\text{foc} = 0$ is dropped and replaced with the identifying assumption that the anchor items are DIF-free (i.e., $d_j = 0$ for anchor items).
-Of course, the demonstration data is designed such that there are obvious groups; this will not necessarily typically be the case.
+Of course, the demonstration data is designed such that there are obvious groups; this need not be the case in practice.
 

### 2020-10-14 R/dif_AOAA.R

@@ -70,9 +70,11 @@ dif_AOAA_AS <- function(data, groups, dif_aoaa){
     status <-
         dif_aoaa %>%
         mutate(
-            status = ifelse(p < 0.05, "flex", "unknown")
+            status = ifelse(p < 0.05, "flex", "unknown"),
+            order = ifelse(p < 0.05, 1, NA)
         )
 
+    track_order <- 2
     while(TRUE){
         # for each unknown item, we use all other unknown items to test it
         new_status <- get_next_status(status, data, groups)
@@ -81,14 +83,14 @@ dif_AOAA_AS <- function(data, groups, dif_aoaa){
         status <-
             status %>%
             select(-p) %>%
-            left_join(new_status, by = "item") %>%
-            select(item, p, status)
+            left_join(new_status %>% select(-order), by = "item")
 
         # if no new significant tests then we're done
         if(all(status$p > 0.05, na.rm = TRUE)){break}
 
         # remove ALL SIGNIFICANT items from the anchor set
         status$status[which(status$p < 0.05)] <- "flex"
+        status$order[which(status$p < 0.05)] <- track_order
     }
 
     # unknowns become anchors

### 2020-10-14 VERBAL-INTUITIVE.R

@@ -1,0 +1,215 @@
+library(tidyverse)
+library(mirt)
+R.utils::sourceDirectory(here::here("R"))
+out <- read_rds("VERBAL2.rds")
+theme_set(theme_bw(base_size = 14))
+
+verbal <- read_rds("verbal.rds")
+
+data <- verbal %>% select(-gender)
+groups <- ifelse(verbal$gender == "female", "a_ref", "b_foc")
+
+vars <-
+    verbal %>%
+    select(-gender) %>%
+    names() %>%
+    str_replace("_", " - ") %>%
+    str_replace("s1", "Bus doesn't stop - ") %>%
+    str_replace("s2", "Miss a train - ") %>%
+    str_replace("s3", "Grocery store closed - ") %>%
+    str_replace("s4", "Operator disconnects me - ")
+
+intuitive_mod <- fit_mod_intuitive(data, groups)
+
+int_mod_df <-
+    intuitive_mod %>%
+    coef_1f() %>%
+    .$items %>%
+    mutate(title = paste0(item, " - ", vars)) %>%
+    select(item, title, a_ref_easy, b_foc_easy) %>%
+    mutate(male_more = b_foc_easy - a_ref_easy) %>%
+    mutate(title = fct_reorder(title, male_more))
+
+vars_df <-
+    tibble(
+        i = 1:24 %>% as.character(),
+        var = paste0("item", 1:24),
+        title = int_mod_df$title
+    )
+
+intuitive_mod %>%
+    mod_intuitive_to_draws_df() %>%
+    select(-run) %>%
+    gather(var, val) %>%
+    left_join(vars_df) %>%
+    mutate(val = -val) %>%
+    mutate(title = fct_reorder(title, val)) %>%
+    ggplot(aes(x = val, y = title)) +
+    ggridges::geom_density_ridges() +
+    labs(
+        x = latex2exp::TeX("$\\tilde{b_j}^{male} - \\tilde{b_j}^{female}$")
+    )
+
+intuitive_mod %>%
+    coef_1f() %>%
+    .$items %>%
+    mutate(title = paste0(item, " - ", vars)) %>%
+    select(item, title, a_ref_easy, b_foc_easy) %>%
+    mutate(male_more = b_foc_easy - a_ref_easy) %>%
+    mutate(title = fct_reorder(title, male_more)) %>%
+    ggplot(aes(x = male_more, y = title)) +
+    geom_point()
+
+p_df <-
+    verbal %>%
[…]

### 2021-01-13 paper/paper.Rmd
get all the paper stuff back on github

@@
 subtitle: |
-    Ben Stenhaug, Ben Domingue, and Mike Frank  
+    Ben Stenhaug, Mike Frank, and Benjamin W. Domingue  
     Stanford University
@@
 We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
+Given the central importance of AI techniques to subsequent examinations of item-level bias, we believe these constitute important advances for the field.
 \clearpage <!-- https://tinyurl.com/ybremelq -->
@@
     warning=FALSE,
-    message=FALSE
+    message=FALSE,
     fig.pos='H',

### 2021-01-13 paper/paper_2b_anchor_points.Rmd
get all the paper stuff back on github

@@
-The previously discussed AI methods select a set of anchor items, whether it is an algorithm or the analyst that makes that selection.
-The anchor items are used to estimate $\hat \mu^\text{foc}$.
-An alternative strategy is to directly set the anchor point [@strobl2018anchor].
+The previously discussed AI methods identify the model by selecting a set of anchor items that are assumed to be DIF-free.
+We now discuss the second type of AI methods: The use of an anchor point [@strobl2018anchor].
+An anchor point directly sets the difference in mean abilities to some value, $\mu^{\star\text{foc}}$.
 Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF.
@@
 
-$\mu^{\star\text{foc}}$ is selected by maximizing the Gini index (thus the abbreviation MAXGI).
+@strobl2018anchor select $\mu^{\star\text{foc}}$ by maximizing the Gini index (thus our abbreviation MAXGI).
 The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF and a small subset of items thus having large amounts of DIF.
@@
 
-For our demonstration data simulation, Figure \ref{fig:ginipath} shows the gini index at a variety of possible $\mu^\text{foc}$ values.
+For the demonstration data, Figure \ref{fig:ginipath} shows the gini index at a variety of possible $\mu^\text{foc}$ values.
 The result of MAXGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 2)`$, which is quite close to the data-generating value of -1.
@@
         linetype = "dashed"
-    )
-```
-
-The model is then refit with the identifying assumption that $\mu^{\text{foc}} = \mu^{\star\text{foc}}$, and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.
-
-```{r ginimilg, fig.cap = "The MILG with $\\mu^{\\text{foc}}$ set to $\\mu^{\\star\\text{foc}} = -0.99$.", out.width="70%", warning = FALSE, message = FALSE}
-# notice hard coding in this caption
-data <- out_3_biased_items$sim[[1]]$data
-groups <- out_3_biased_items$sim[[1]]$groups
-
-mod <- anchptfocmean_to_final_mod(gini_results$gini_anchor_points, data, groups)
-
-mod %>% 
-    mod_to_draws_df(n = 10000) %>% 
-    draws_df_to_logit_plot() +
-    labs(
-        x = latex2exp::TeX("$d_j = b_j^{ref} - b_j^{foc}$"),
-        y = ""
     )
@@
 If $\mu^\text{foc} = 3$, wouldn't we also want to prioritize high $\theta$ values?
-Accordingly, we set $g(\theta)$ to be the average of the reference and focal group ability probability density functions:
+Accordingly, we set $g(\theta)$ to be the weighted average of the reference and focal group ability probability density functions:
 \begin{align}
-g(\theta) = \dfrac{N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})}{2}.
+g(\theta) = p^\text{ref} \cdot N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + (1 - p^\text{ref}) \cdot N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})
 \end{align}
+where $p^\text{ref}$ is the proportion of all students that are in the reference group. In this paper, we always have $p^\text{ref} = 0.5$.
 
-For our demonstration data simulation, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$.
+For the demonstration data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for $\mu^\text{foc}$.
 In this case, MINBC works perfectly and the anchor point is found to be $\mu^{\star\text{foc}} = `r round(min_bc$between_curves_anchor_points, 2)`$.
@@
 ```
-
-## Summary of AI methods
-
-We described a variety of AI methods and their corresponding acronymns.
-Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point.
-MILG and EM-MILG are somewhat of outliers in that they are methods for visualizing DIF or potential DIF.
-Table \ref{table:allmethods} summarizes all of these methods.
-
-\begin{table}[H]
-\caption{Summary of agnostic identification (AI) methods}
-\centering
-\begin{tabular}{|p{4cm}|p{6cm}|p{4cm}|}
-\toprule
-
-Method & Description & Literature \\
-
-\midrule
-
-equal means, multiple imputation logit graph (EM-MILG) & Arbitrarily set both group means to 0, which pushes all group performance differences to the item paramaters, measure variability using multiple imputations, and graph the result.
-Can be used by an analyst to hand select anchor items & Inspired by pedagolical examples by \cite{pohl2017cluster} and \cite{talbot2013taking} \\\hline
-
-multiple imputation logit graph (MILG) & Similar to EM-MILG but used to visualize potential DIF once the model is already identified &  \\\hline
-
-all-others-as-anchors (AOAA) & Test if each item has DIF by using all of the other items as anchors (not iterative).
-& Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline
-
-all-others-as-anchors-all-significant (AOAA-AS) & The first iteration is AOAA.
-All items that test positive for DIF are removed from the anchor set.
-Continue iterating until no new items test positive for DIF.
-& Proposed by \cite{drasgow1987study} \\\hline
-
-all-others-as-anchors-one-at-a-time (AOAA-OAT) & The first iteration is AOAA.
-Only the item that shows the most extreme DIF is removed from the anchor set.
-Continue iterating until no new items test positive for DIF.
-& To our knowledge, not proposed or used previously \\\hline
-
-equal means clustering (EMC) & Cluster items based on differences in performance across groups and choose one of the clusters to be the anchor cluster.
-& Proposed by \cite{bechger2015statistical} and refined by \cite{pohl2017cluster} \\\hline
-
-maximizing Gini index (MAXGI) & Arbitrarily set both group means to 0 and then choose an anchor point by maximizing the Gini index & Adapted from work by \cite{strobl2018anchor} \\\hline
-
-minimizing the area between curves (MINBC) & Of the infinite number of model that maximizes the likelihood of the data, choose the one with the minimum total area between the two groups' item characteristic curves & Built on and inspired by work by \cite{raju1988area} and more recently, \cite{chalmers2016might} \\
-
-\bottomrule
-\end{tabular}
-\label{table:allmethods}
-\end{table}

### 2021-01-13 paper/paper_2c_summary_of_methods.Rmd
get all the paper stuff back on github

--- 2020-03-23 paper/paper.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-  11pt,
-]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Promise and peril: Agnostic identification methods for detecting differential item functioning},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-\usepackage{amsmath}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-\usepackage{float}
-\usepackage[autostyle, english = american]{csquotes}
-\setlength{\parindent}{4em}
-\setlength{\parskip}{2em}
-\usepackage{setspace}\doublespacing
-\AtBeginEnvironment{tabular}{\singlespacing}
-\usepackage{natbib}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\title{Promise and peril: Agnostic identification methods\\
-for detecting differential item functioning}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Ben Stenhaug, Ben Domingue, and Mike Frank\\
-Stanford University}
-\author{}
-\date{\vspace{-2.5em}}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models.
-However, to use a LRT (or any other method), we require an identifying assumption to disentangle differences in group ability from potential DIF.
-We use the term ``agnostic identification'' (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free.
-We first summarize existing AI methods and propose a variety of new methods.
-We then conduct a simulation study---which we argue is more realistic than most DIF simulation studies in the literature---and find that one of the proposed new AI methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods.
-We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst.
-\clearpage
-\end{abstract}
-
-{
-\setcounter{tocdepth}{5}
-\tableofcontents
-}
-\clearpage
-
-\hypertarget{intro}{%
-\section{Introduction}\label{intro}}
-
-Following Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when one imposes an item response model with too few ability dimensions.
-From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item.
-For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability.
-In this case, math ability is the \enquote{target ability}, and verbal ability is the \enquote{nuisance ability}.
-Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies \enquote{DIF}.
-
-Despite this, the usual setup of DIF simulation studies frames DIF as a property of the item.
-For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group.
-They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values that they obtained from a previous study.
-They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits.
-They then simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group.
-Finally, they generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
-\begin{align}
-\text{Pr}(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j) \label{zopf}
-\end{align}
-where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.
-
-For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional.
-According to the two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}), the probability that student \(i\) responds correctly to item \(j\) is
-\begin{align}
-\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\theta j}, a_{\eta j}, b_j) = \sigma(a_{\theta j}\theta_i + a_{\eta j}\eta_i + b_j).
-\end{align}
-where \(\theta\) is target ability, \(\eta\) is nuisance ability, \(a_{\theta j}\) is an item's slope on target ability, and \(a_{\eta j}\) is an item's slope on nuisance ability.
-
-For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view, \(b_j\) is set to \({b_j}^{\text{ref}}\), \(\theta\) stays the same, \(a_{\theta j} = 1\) for all items (consistent with the Rasch model), and \(\eta\) takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group.
-For items with DIF, \(a_{\eta j} = 0.6\), and for items without DIF, \(a_{\eta j} = 0\) Again, 0.6 is the magnitude of DIF in logits.\footnote{The reader may notice that this is just one possible specification.
-Any specification that meets the following criteria has the same effect: First, \(a_{\eta j}\eta_i\) must be 0.6 lower for the focal group on items with DIF.
-Second, \(a_{\eta j}\eta_i\) must be equal across groups for items without DIF.} For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.
-This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.
-
-If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model?
-Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities} {[}p.~144{]}.
-He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary {[}p.~144{]}.
-It's hard to imagine how one would have such knowledge.
-Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.
-This approach depends upon a crucial identifying assumption; this assumption is the target of our investigation.
-
-\hypertarget{agnostic-identification-methods}{%
-\section{Agnostic identification methods}\label{agnostic-identification-methods}}
-
-Psychometricians have long been in search of the perfect method for detecting differential functioning of \emph{individual} items.
-The IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}).
-Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})).
-Within the IRT-framework, we---as is common---estimate models using marginal maximum likelihood estimation (Bock and Aitkin \protect\hyperlink{ref-bock1981marginal}{1981}) and identify the scale by setting \(\mu^\text{ref} = 0\).
-
-Use of LRT requires an identifying assumption.
-The assumption is needed so as to link groups.
-In the typical setting, no a prior assumptions---about either relative group ability or which items might have DIF---can be made.
-Thus, we're interested in \emph{agnostic identification} methods (hereafter referred to as \enquote{AI methods}).
-The fundamental challenge is to disentangle estimation of \(\hat\mu^\text{foc} - \mu^\text{ref} = \hat\mu^\text{foc} - 0 = \hat\mu^\text{foc}\), the difference in mean abilities, from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness.
-There are two types of AI methods: The most common, the use of anchor items, selects a group of items that are assumed to be DIF-free.
-These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.
-The other type, the use of an anchor point (Strobl et al.
-\protect\hyperlink{ref-strobl2018anchor}{2018}), directly sets the difference in mean abilities to some value, \(\mu^{\star\text{foc}}\).
-This anchor point identifies the model, and all items can be tested for DIF using an LRT.
-
-In this section, we summarize existing AI methods and propose a variety of extensions.
-For simplicity, we focus on the Rasch model to isolate the fundamental issues in \enquote{the AI problem}.
-Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.} We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test.
-Our simulation is similar to our multidimensional alternative to the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) approach discussed above.
-Target ability is simulated as \({\theta_i}^{\text{ref}} \sim N(\mu^\text{ref} = 0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(\mu^\text{foc} = -1,1)\) for students in the focal group.
-Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{foc}} = -1\).
-The slope on target ability is set to \(a_{1j} = 1\) for all items.
-The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise.
-We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability.
-Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.
-To be sure, this data is used only for demonstrative purposes.
-We compare the performance of the AI methods in a more realistic scenario in the \protect\hyperlink{simstudy}{the simulation study}.
-
-The AI methods described in this paper are summarized in Table \ref{table:allmethods}.
-In addition the results for each method on the demonstration data are shown in a MILG in Figure \ref{fig:allmethodsfig}.
-
-\hypertarget{emmilg}{%
-\subsection{The equal means, multiple imputation logit graph (EM-MILG)}\label{emmilg}}
-
-AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.
-
-While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position to understand the scale of the problem in their data.
-We propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst.
-EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting \(\mu^\text{foc} = 0\).
-Given that we also assume \(\mu^\text{ref} = 0\), this is equivalent to the assumption that the groups have equal mean ability.
-As a result, all differences in performance---either from group ability differences or DIF---manifest in the item easiness difference parameter \(\tilde{d_j}\).
-The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\)---as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from a properly identified model.
-
-To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}).
-Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are drawn to estimate the distribution of \(\tilde{d_j}\) for each item.
-These are the distributions displayed in a EM-MILG.
-The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.
-
-Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data.
-We emphasize that the EM-MILG contains all possible information about the difference in group performance.
-The analyst might assume that---because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits---items 1-5 are unbiased.
-This might seem to be a strong assumption, but all AI methods assume---in one way or another---that, if there is DIF, its a minority of items that are responsible.
-These unbiased items are known as anchor items; setting anchor items is the most common identification assumption.
-After anchor items have been selected, the model is refit where the identifying assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the identifying assumption that the anchor items are DIF-free (i.e., \(d_j = 0\) for anchor items).
-Of course, the demonstration data is designed such that there are obvious groups; this will not necessarily typically be the case.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emmilg-1}
-
-}
-
-\caption{A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
-\end{figure}
-
-The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with a properly identified model.
-Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG).
-As expected, in our demonstration data, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/milg-1}
-
-}
-
-\caption{A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed by setting $d_j = 0$.}\label{fig:milg}
-\end{figure}
-
-In general, one of our main concerns with traditional AI methods is that they can lull the analyst into a false sense of security: The analyst chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality.
-Our aim with EM-MILG is to combat this concern by presenting all information clearly to the analyst.
-In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased.
-Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data.
-And, of course, the MILG can be used to visualize DIF regardless of how the model is identified.
-We now move on to more traditional AI methods, which don't require judgment from the analyst, but make the majority-of-items-don't-have-DIF assumption more subtly.
-
-\hypertarget{anchoritems}{%
-\subsection{Anchor items}\label{anchoritems}}
-
-\hypertarget{all-others-as-anchors-aoaa}{%
-\subsubsection{All-others-as-anchors (AOAA)}\label{all-others-as-anchors-aoaa}}
-
-Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method.
-AOAA tests each item for DIF one at a time using all of the other items as anchors.
-For example, when testing the first item for DIF, all of the other items are used as anchors.
-This is done using a LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}).
-Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on.
-The items for which the flexible model outperforms the baseline model (typically based on a \(\chi^2\) test) are identified as having DIF, and the rest of the items become anchor items.
-AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the baseline to the flexible model).
-
-Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, counter to the underlying rationale for the undertaking.
-On a practical level, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group.
-Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}).
-A simple thought experiment illustrates how AOAA fails: Imagine a three item test with a sufficiently large number of students where the first item has DIF, and the other two do not.
-Using AOAA, all items test positive for DIF.
-The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated.
-This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF}, which is a significant problem in applications of AI methods.
-
-\hypertarget{all-others-as-anchors-all-significant-aoaa-as}{%
-\paragraph{All-others-as-anchors-all-significant (AOAA-AS)}\label{all-others-as-anchors-all-significant-aoaa-as}}
-
-One way to attempt to counter artificial DIF is with purification through iterative anchor selection.
-For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively---with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model---until no more items tested positively.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse) referring to beginning with the assumption that all items are DIF-free.
-We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS).
-We append \enquote{all-significant} to indicate that all items that test positive for DIF are removed from the anchor set in each iteration.
-AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.
-
-We argue that AOAA-AS, while a potential improvement, doesn't solve the fundamental problem of AOAA: What does one do when all items test positive for DIF?
-With a sufficient sample size and at least one item with DIF, this will necessarily be the case.
-In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items.
-Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly.
-Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method: Use AOAA and select the four items that exhibit the least amount of DIF as anchor items.
-
-\hypertarget{all-other-as-anchors-one-at-a-time-aoaa-oat}{%
-\paragraph{All-other-as-anchors-one-at-a-time (AOAA-OAT)}\label{all-other-as-anchors-one-at-a-time-aoaa-oat}}
-
-We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed.
-AOAA-OAT is inspired by Hagquist and Andrich (\protect\hyperlink{ref-hagquist2017recent}{2017}), who assert that \enquote{items showing DIF initially should not be resolved simultaneously but sequentially} {[}p.~6{]}.
-Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF---based on the \(\chi^2\) test statistic---is removed from the anchor set.
-The process continues iteratively until no new items display DIF.
-AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only \emph{one} item in each iteration as opposed to \emph{all} items that test positive for DIF.
-As a result, we believe that AOAA-OAT is less likely to be \enquote{fooled} by artifical DIF.
-AOAA-OAT is not currently implemented in the mirt R package.
-
-Applying AOAA-OAT to our thought experiment demonstrates its effectiveness.
-The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF (i.e., because the group difference in performance is most different from the average group difference in performance on the other two items).
-In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion.
-
-\hypertarget{equal-means-clustering-emc}{%
-\subsubsection{Equal means clustering (EMC)}\label{equal-means-clustering-emc}}
-
-Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster}.
-They pointed out that one way around the lack of model identifiability is to consider only relative item parameters.
-For each group, the relative easinesses for each pair of items can be stored in the (square) matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\), where \(x\) and \(y\) both index the items.
-A similar matrix, \(\mathbf{R}^{\text{foc}}\), can be formed.
-The ultimate matrix of interest is \(\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}\) which is the \enquote{differences between groups in the pairwise differences in (easiness) between items} {[}p.~323{]}.
-
-The general idea of identifying clusters of items is intriguing.
-However, their approach is complicated, and they did not describe a process for moving from \(\Delta \mathbf{R}\) to an anchor cluster.
-Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing one such process.
-They point out that \(\Delta \mathbf{R}\) is skew-symmetric and of rank 1, which means that all information is contained in a single row or column.
-Accordingly, they recommend k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC.
-They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster.
-They conducted a simulation study where some items contained DIF, and, unfortunately, found that BIC identifies only a single cluster, so all items were anchors.
-
-We propose a new cluster-based approach, which we call \enquote{equal means clustering} (EMC).
-Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\), which, as described in the \protect\hyperlink{emmilg}{EM-MILG section}, comes from setting \(\mu^\text{foc} = 0\).
-Recall that \(\mu^\text{ref}\) is to \(0\) to identify the scale, and thus the name \enquote{equal means clustering}.
-Instead of choosing k with BIC, we use the gap statistic method recommended by Hastie, Tibshirani, and Walther (\protect\hyperlink{ref-hastie2001estimating}{2001}).
-Based on the assumption that, at most, a minority of items contain DIF, we choose the largest cluster as the anchor cluster.
-If there is a tie for the largest cluster, the cluster with the lowest standard deviation of \(\tilde{\mathbf{d}}\) is selected.
-
-\hypertarget{anchor-points}{%
-\subsection{Anchor points}\label{anchor-points}}
-
-The previously discussed AI methods identify the model by selecting a set of anchor items that are assumed to be DIF-free.
-We now discuss the second type of AI methods: The use of an anchor point (Strobl et al.
-\protect\hyperlink{ref-strobl2018anchor}{2018}).
-An anchor point directly sets the difference in mean abilities to some value, \(\mu^{\star\text{foc}}\).
-Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF.
-The question then becomes the following: How is the anchor point selected?
-
-\hypertarget{maximizing-the-gini-index-maxgi}{%
-\subsubsection{Maximizing the Gini index (MAXGI)}\label{maximizing-the-gini-index-maxgi}}
-
-Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (Gini \protect\hyperlink{ref-gini1912variabilita}{1912}) to select the anchor point.
-The Gini index is typically used to measure the inequality of wealth distribution in a country.
-For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution (Chitiga, Sekyere, and Tsoanamatsie \protect\hyperlink{ref-chitiga2015income}{2015}).
-In general, the Gini index \enquote{takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little} (Strobl et al.
-\protect\hyperlink{ref-strobl2018anchor}{2018}, 7).
-
-Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}) select \(\mu^{\star\text{foc}}\) by maximizing the Gini index (thus our abbreviation MAXGI).
-The intuition and assumption is that the anchor point should prioritize the majority of items having little to no DIF and a small subset of items thus having large amounts of DIF.
-Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
-\begin{align}
-\mu^{\star\text{foc}} = \mathop\mathrm{arg\,max}_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
-\end{align}
-where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\tilde{\mathbf{d}}\).
-
-For the demonstration data, Figure \ref{fig:ginipath} shows the gini index at a variety of possible \(\mu^\text{foc}\) values.
-The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is quite close to the data-generating value of -1.
-Moreover, there is a local maximum at \(\mu^{\text{foc}} \approx 1.5\), which corresponds to the items with DIF.
-This illustrates that---as Strobl et al.
-(\protect\hyperlink{ref-strobl2018anchor}{2018}) point out---the search path is informative.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginipath-1}
-
-}
-
-\caption{Maximizing the Gini index (MAXGI) to select the anchor point.}\label{fig:ginipath}
-\end{figure}
-
-\hypertarget{minimizing-between-curves-minbc}{%
-\subsubsection{Minimizing between curves (MINBC)}\label{minimizing-between-curves-minbc}}
-
-Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
-\begin{align}
-\text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
-\end{align}
-Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods (Magis et al.
-\protect\hyperlink{ref-magis2011generalized}{2011}).
-However, Raju's area method is not an AI method because the item characteristic curves still need to be linked by anchor items or an anchor point.
-An additional weakness is that the area is unweighted, so all values of \(\theta\) matter equally, despite some being much more realistic than others.
-
-To adapt Raju's area method into an AI method, we propose a new method, which we call \enquote{minimizing the area between curves} (MINBC).
-To understand MINBC, imagine a scenario in which the data-generating process is \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\), so that the groups have equal ability and there is no DIF.
-The fundamental identification problem is that there are an infinite number of models with the same likelihood from which to choose.
-For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\mu^{\star\text{foc}} = 0\).
-The model would then estimate \(\hat d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF.
-Alternatively, we could assume that the focal group has \(\mu^{\star\text{foc}} = 3\).
-The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability, but every item contains DIF against them.
-Both of these models have the same likelihood, so how should one choose which model to believe?
-MINBC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves.
-As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences (as opposed to DIF).
-
-Denote a function that takes \(\mu^\text{foc}\) as input and estimates \(\hat b_j^\text{foc}\) by fitting a unidimensional Rasch model as \(m_j(\mu^\text{foc})\).
-The amount of DIF on each item is calculated as
-\begin{align}
-\text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
-\end{align}
-where \(g(\theta)\) is a weighting function such that \(\int g(\theta)d\theta = 1\). The total DIF on the test, then, is
-\begin{align}
-\text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc})
-\end{align}
-In this way, \(\text{Total DIF}(\mu^\text{foc})\) is a function where the input is \(\mu^\text{foc}\) and the output is the total amount of DIF on the test. MINBC sets
-\begin{align}
-\mu^{\star\text{foc}} = \mathop\mathrm{arg\,min}_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
-\end{align}
-
-MINBC is inspired in part by Chalmers, Counsell, and Flora (\protect\hyperlink{ref-chalmers2016might}{2016}), who use the difference between test characteristic curves weighted by \(g(\theta)\) as a measure of differential test functioning (DTF).
-The selection of \(g(\theta)\) results in the relative weighting of \(\theta\) values.
-Chalmers, Counsell, and Flora do not discuss how to choose \(g(\theta)\) and in their empirical examples use \(g(\theta)\) uniform for \(-6 \le \theta \le 6\), which may be suboptimal in some cases.
-It might seem intuitive to choose \(g(\theta) \sim N(0, 1)\) because \(\mu^\text{ref} = 0\), but this choice doesn't take into account the ability distribution of the focal group.
-If \(\mu^\text{foc} = 3\), wouldn't we also want to prioritize high \(\theta\) values?
-Accordingly, we set \(g(\theta)\) to be the weighted average of the reference and focal group ability probability density functions:
-\begin{align}
-g(\theta) = p^\text{ref} \cdot N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + (1 - p^\text{ref}) \cdot N(\mu^{\text{foc}}, \sigma^{\text{foc}^2})
-\end{align}
-where \(p^\text{ref}\) is the proportion of all students that are in the reference group. In this paper, we always have \(p^\text{ref} = 0.5\).
-
-For the demonstration data, Figure \ref{fig:mabc} shows Total DIF at a variety of possible values for \(\mu^\text{foc}\).
-In this case, MINBC works perfectly and the anchor point is found to be \(\mu^{\star\text{foc}} = -1\).
-As with MAXGI, the model should then be refit using the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/mabc-1}
-
-}
-
-\caption{Minimizing the area between curves (MINBC) to select the anchor point.}\label{fig:mabc}
-\end{figure}
-
-\hypertarget{summary-of-ai-methods}{%
-\subsection{Summary of AI methods}\label{summary-of-ai-methods}}
-
-We described a variety of AI methods and their corresponding acronymns.
+We described a variety of AI methods and their corresponding acronyms.
 Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point.
@@
 
-\begin{verbatim}
-## Picking joint bandwidth of 0.00587
-\end{verbatim}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.00591
-\end{verbatim}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.00546
-\end{verbatim}
-
-\begin{verbatim}
-## Picking joint bandwidth of 0.0055
-\end{verbatim}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/allmethodsfig-1}
-
+```{r allmethodsfig, fig.cap = 'MILG for each of the AI methods applied to the demonstration data.
+AOAA-OAT and EMC are identified by anchor items (represented by points).
+MAXGI and MINBC are identified by setting $\\mu^{\\text{foc}}$ set to the anchor point $\\mu^{\\star\\text{foc}}$.
+AOAA and AOAA-AS are not included as they did not find an identifying assumption.', out.width="70%"}
+longify <- function(df, the_method){
+    df %>%
+        select(-run) %>%
+        gather(var, val) %>%
+        mutate(var = as_factor(paste0("Item ", parse_number(as.character(var))))) %>%
+        filter(val != 0) %>%
+        mutate(method = the_method)
 }
 
-\caption{MILG for each of the AI methods applied to the demonstration data.
-AOAA-OAT and EMC are identified by anchor items (represented by points), MAXGI and MINBC are identified by setting $\mu^{\text{foc}}$ set to the anchor point $\mu^{\star\text{foc}}$, AOAA and AOAA-AS are not included as they did not find an identifying assumption.}\label{fig:allmethodsfig}
-\end{figure}
+points <-
+    bind_rows(
+        out_3_biased_items$AOAA_OAT_final_dif[[1]] %>%
+            mutate(
+                item = as_factor(paste0("Item ", parse_number(as.character(item)))),
+                difference_in_easy = difference_in_easy - difference_in_easy[anchor][1],
+                method = "AOAA-OAT"
+            ),
 
-\hypertarget{simstudy}{%
-\section{Simulation study}\label{simstudy}}
+        out_3_biased_items$cluster_final_dif[[1]] %>%
+            mutate(
+                item = as_factor(paste0("Item ", parse_number(as.character(item)))),
+                difference_in_easy = difference_in_easy - difference_in_easy[anchor][1],
+                method = "EMC"
+            ),
+    ) %>% 
+    select(method, item, anchor) %>% 
+    filter(anchor) %>% 
+    mutate(val = 0)
 
-To compare each of the methods in Table \ref{table:allmethods}, we conducted a simulated study.
-Our goal was to use a relatively realistic data generating process motivated by the case study in Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) wherein some items on a math test also depend on a student's verbal ability (the target ability is math ability, and the nuisance ability is verbal ability).
-Nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group.
-As described in \protect\hyperlink{intro}{the introduction}, this setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group.
-One exception is Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}) who draw each student's target ability and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.
+data <- out_3_biased_items$sim[[1]]$data
+groups <- out_3_biased_items$sim[[1]]$groups
 
-In our simulation study, it was critical that student ability was drawn in a realistic way similar to Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}).
-However, we don't believe that a compensatory model is realistic in describing a math test where some items depend on verbal ability.
-For example, it's hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability.
-Accordingly, we generated item responses using a simplified version of Sympson's (\protect\hyperlink{ref-sympson1978model}{1978}) noncompensatory item response model in which
-\begin{align}
-\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\eta j}) = \sigma(\theta_i) \cdot \sigma(a_{\eta j}\eta_i)
-\end{align}
-where, as before, \(\theta_i\) is target ability, \(\eta_i\) is nuisance ability, and \(a_{\eta j}\) is the item's loading on nuisance ability (DeMars \protect\hyperlink{ref-demars2016partially}{2016}).
+mod_gini <- anchptfocmean_to_final_mod(gini_results$gini_anchor_points, data, groups)
+mod_minbc <- anchptfocmean_to_final_mod(min_bc$between_curves_anchor_points, data, groups)
 
-Computing was done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the mirt R package (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the suite of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}).
-Code is available at \emph{(TODO)}.
+data_for_graph <- 
+    bind_rows(
+        mod_gini %>% mod_to_draws_df(n = 10000) %>% longify("MAXGI"),
+        mod_minbc %>% mod_to_draws_df(n = 10000) %>% longify("MINBC"),
+        out_3_biased_items$AOAA_OAT_mod[[1]] %>% mod_to_draws_df(n = 10000) %>% longify("AOAA-OAT"),
+        out_3_biased_items$cluster_mod[[1]] %>% mod_to_draws_df(n = 10000) %>% longify("EMC")
+    )
 
-\hypertarget{drawing-parameters}{%
-\subsection{Drawing parameters}\label{drawing-parameters}}
-
-In each run, we simulated 10,000 students with half coming from each of the reference and focal groups.
-For students from the reference group, target ability and nuisance ability were drawn from the two-dimensional normal distribution with mean {[}\(\mu_\theta^\text{ref} = 0\), \(\mu_\eta^\text{ref} = 0\){]} and covariance matrix \(\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}\).
-Abilities for students from the focal group were drawn using the same covariance matrix, but with means {[}\(\mu_\theta^\text{foc} = -0.5\), \(\mu_\eta^\text{foc} = -1\){]}.
-
-The test always has 12 items, but we varied the number of items with DIF from two to six.
-For items without DIF, \(a_{\eta j} = \infty\) so that the model reduces to \(\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)\).
-For items with DIF, \(a_{\eta j}\) is calculated based on Ackerman's (\protect\hyperlink{ref-ackerman1994using}{1994}) angle equation as described in Walker and Gocer Sahin (\protect\hyperlink{ref-walker2017using}{2017}):
-\begin{align}
-\angle_j = \arccos \dfrac{a_{\theta j}^2}{a_{\theta j}^2 + a_{\eta j}^2}.
-\end{align}
-An item's angle measures the relative loading of the item on the two dimensions.
-The lower the angle, the more the item loads on target ability as compared to nuisance ability, which corresponds to less DIF.
-An angle of 45\(^\circ\) indicates that the item loads equally on the target ability and nuisance ability.
-Our simple noncompensatory model has \(a_{\theta j} = 1\) for all items so the angle equation reduces to
-\begin{align}
-\angle_j = \arccos \dfrac{1}{1 + a_{\eta j}^2}.
-\end{align}
-We are interested in specifying the angle of an item, so the relevant equation becomes
-\begin{align}
-a_{2j} = \sqrt{\dfrac{1 - \cos(\angle_j)^2}{\cos(\angle_j)^2}}.
-\end{align}
-For DIF items, we set \(a_{\eta j}\) based on angles with equal intervals between 20\(^\circ\) and 60\(^\circ\).
-For example, for a test with three DIF items the angles are 20\(^\circ\), 40\(^\circ\), and 60\(^\circ\).
-
-\hypertarget{visualizing-a-run}{%
-\subsection{Visualizing a run}\label{visualizing-a-run}}
-
-Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between \(\theta_i\) and \(\text{Pr}(y_{ij} = 1)\) with \(\eta_i\) set to the group mean for a test with six items with DIF.
-The items are ordered by the amount of DIF such that \(\angle_{j = 7} = 20^\circ\) up to \(\angle_{j = 12} = 60^\circ\).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/difmap-1}
-
-}
-
-\caption{For a 12-item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.}\label{fig:difmap}
-\end{figure}
-
-Figure \ref{fig:simemmlg} shows the EM-MILG---generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the \protect\hyperlink{anchoritems}{anchor items section}---for one run using the same item parameters that generated Figure \ref{fig:difmap}.
-As expected, \(\tilde{d_j}\) is about \(\mu_\theta^\text{ref} - \mu_\theta^\text{foc}= 0 - (-0.5) = 0.5\) for the first six items which are DIF free.
-For the last six items, \(\tilde{d_j}\) increases as \(\angle_j\) increases.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/simemmlg-1}
-
-}
-
-\caption{The EM-MILG for one run using the same item parameters as generated Figure ef{fig:difmap}}\label{fig:simemmlg}
-\end{figure}
-
-\hypertarget{outcomes}{%
-\subsection{Outcomes}\label{outcomes}}
-
-For each run, we applied each AI method to find the method's identifying assumption.
-The method's identifying assumption was then used to fit a final model.
-We compared the performance of those final models according to the following outcomes.
-
-\hypertarget{achievement-gap-residual}{%
-\subsubsection{Achievement gap residual}\label{achievement-gap-residual}}
-
-An effective AI method should lead to a final model that accurately estimates the difference between the reference group's mean target ability and the focal group's mean target ability.
-We refer to this quantity as the achievement gap.
-Recall that all models set \(\mu_\theta^\text{ref} = 0\), so the achievement gap reduces to \(\mu_\theta^\text{foc}\).
-The data-generating value of \(\mu_\theta^\text{foc}\) is 0.5, but each run, of course, includes sampling variability.
-To get at the heart of how well a method is doing, we calculated the achievement gap residual as the method's estimated achievement gap, \(\hat\mu_\theta^\text{foc}\), minus the achievement gap estimated when using only the DIF-free items as anchors.
-In summary, this outcome measures a method's ability to disentangle differences in target ability from nuisance ability at the group level.
-
-\hypertarget{anchor-items.}{%
-\subsubsection{Anchor items.}\label{anchor-items.}}
-
-For the methods that choose a set of anchor items, we looked directly at which anchor items were selected.
-An effective method should use most of the non-DIF items as anchors (the anchor hit rate) while avoiding using items with DIF as anchors (the false anchor rate).
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-In total, we executed 100 runs for each of two, three, four, five, and six DIF items.
-Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual.
-AOAA-OAT was the clear winner.
-It performed nearly perfectly for two, three, or four DIF items.
-Even when six of the 12 items on the test contained DIF, AOAA-OAT underestimated \(\mu_\theta^\text{foc}\) by only 0.05 standard deviations on its worst run.
-As expected, artificial DIF caused AOAA and AOAA-AS to begin to exhibit problematic performance as the number of DIF items increased.
-
-EMC performed better than AOAA and AOAA-AS, but worse than the other methods.
-MINBC and MAXGI performed similarly well with MINBC estimating the achievement gap with more precision but more bias than MAXGI, especially for tests with more than four DIF items.
-We hypothesize that MINBC's susceptibility to bias results from considering every item.
-AOAA-OAT, for example, completely disregards an item once it's removed from the anchor set.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/achievegap-1}
-
-}
-
-\caption{Achievement gap residual distributions across 100 runs for each AI method and number of DIF items.}\label{fig:achievegap}
-\end{figure}
-
-Figure \ref{fig:anchorfalse} shows the mean false anchor rates for each method and number of items with DIF.
-For example, when there were two items with DIF on the test, those two items had \(\angle_{11} = 20^\circ\) and \(\angle_{12} = 60^\circ\).
-AOAA-OAT never included item 12 in the anchor set, but incorrectly included item 11 in 60 out of the 100 runs.
-Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method was
-\begin{align}
-\dfrac{\text{Total number of DIF items in the anchor set}}{\text{Number of DIF items on each test} \cdot \text{Number of runs}} = \dfrac{60}{2 \cdot 100} = 30\%.
-\end{align}
-The fact that the item with 20\(^\circ\) of DIF is most commonly incorrectly included in the anchor set is what drove the counterintuitive result that the mean false anchor rate decreases with more DIF items.
-
-Similarly, Figure \ref{fig:anchorhit} shows the mean anchor hit rates.
-Remarkably, AOAA-OAT included an average of over \(90\%\) of DIF-free items in the anchor set regardless of the number of DIF items on the test.
-Interestingly, EMC had a better anchor hit rate on tests with more DIF items.
-This result appears to be driven by the clustering algorithm sometimes splitting all of the DIF-free items into two separate clusters, especially when most of the items are DIF-free.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/anchorfalse-1}
-
-}
-
-\caption{Mean false anchor rates across 100 runs for each AI method and number of DIF items.}\label{fig:anchorfalse}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/anchorhit-1}
-
-}
-
-\caption{Mean anchor hit rates across 100 runs for each AI method and number of DIF items.}\label{fig:anchorhit}
-\end{figure}
-
-\hypertarget{discussion-and-summary}{%
-\section{Discussion and summary}\label{discussion-and-summary}}
-
-Validity hinges on measurement instruments being relatively free of DIF.
-Such instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student.
-We reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe to be more realistic than the typical DIF simulation study.
-In particular, we simulated student ability as drawn from a two-dimensional distribution representing a student's target and nuisance ability, and then generated data using a noncompensatory item response model.
-Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows.
-This is concerning given both their widespread use and the fact that other methods---AOAA-OAT, EMC, MINGI, and MAXBC---demonstrate superior performance.
-
-AOAA-OAT exhibiting superior performance is an important finding given that it is, to our knowledge, not used currently.
-We advocate for its widespread use.
-One drawback of the AOAA-OAT method is that it is computationally expensive.
-For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows.
-To increase AOAA-OAT's use, we recommend its implementation (perhaps as the default) in popular IRT software such as the mirt R package.
-
-In addition to exploring and testing algorithmic AI methods, we introduced a method, the EM-MILG, that an analyst can use to visualize the amount of potential DIF in their data.
-This method can be used either to build their intuition or as a way in which they can select anchor items by hand.
-The EM-MILG's sibling method, the MILG, is, we believe, the best way to visualize the results of a DIF analysis after anchor items have been selected.
-
-We emphasize that the EM-MILG gives a clear view of the possibility of DIF on a test.
-For example, in Figure \ref{fig:simemmlg} it's clear that the gap between focal and reference group performance across items on the test varies greatly.
-By assuming that the majority of items do not contain DIF, AI methods essentially make an educated guess about the relative responsibility of group ability differences vs.~DIF.
-Our simulation focused on a scenario wherein half of the items are DIF-free and the other half have escalating amounts of DIF.
-Different patterns of DIF (as visualized in an EM-MILG) will pose different challenges.
-
-For example, imagine a test in which the focal group outperforms the reference group by one logit on half of the items, and the reference group outperforms the focal group by one logit on the other half of the items.
-What should an AI method conclude in this case?
-AOAA and AOAA-AS will likely (perhaps rightfully) fail to find any anchor items; MINBC will choose an anchor point close to zero (and will find DIF in all items); MAXGI will arbitrarily choose an anchor point within one of the groups of items (and will find DIF in the other group of items); and AOAA-OAT and EMC will arbitrarily choose one of the groups of items to be anchor items (and, like MAXGI, will find DIF in the other group of items).
-This example illustrates that some patterns of DIF are such that the analyst should probably remain agnostic.
-In these cases, we suggest that the analyst report the nature of potential DIF using a EM-MILG.
-Future work might produce methods for quantifying what is visualized in a EM-MILG, essentially making a metric for the amount of potential DIF on a test.
-
-Moreover, future work should test these methods' performance under a greater variety of data-generating conditions.
-For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions.
-Finally, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting other types of item response models.
-
-\clearpage
-
-\bibliographystyle{apa}
-\bibliography{paper.bib}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ackerman1992didactic}{}%
-Ackerman, Terry A. 1992. ``A Didactic Explanation of Item Bias, Item Impact, and Item Validity from a Multidimensional Perspective.'' \emph{Journal of Educational Measurement} 29 (1): 67--91.
-
-\leavevmode\hypertarget{ref-ackerman1994using}{}%
----------. 1994. ``Using Multidimensional Item Response Theory to Understand What Items and Tests Are Measuring.'' \emph{Applied Measurement in Education} 7 (4): 255--78.
-
-\leavevmode\hypertarget{ref-andrich2012real}{}%
-Andrich, David, and Curt Hagquist. 2012. ``Real and Artificial Differential Item Functioning.'' \emph{Journal of Educational and Behavioral Statistics} 37 (3): 387--416.
-
-\leavevmode\hypertarget{ref-bechger2015statistical}{}%
-Bechger, Timo M, and Gunter Maris. 2015. ``A Statistical Test for Differential Item Pair Functioning.'' \emph{Psychometrika} 80 (2): 317--40.
-
-\leavevmode\hypertarget{ref-bock1981marginal}{}%
-Bock, R Darrell, and Murray Aitkin. 1981. ``Marginal Maximum Likelihood Estimation of Item Parameters: Application of an Em Algorithm.'' \emph{Psychometrika} 46 (4): 443--59.
-
-\leavevmode\hypertarget{ref-camilli1992conceptual}{}%
-Camilli, Gregory. 1992. ``A Conceptual Analysis of Differential Item Functioning in Terms of a Multidimensional Item Response Model.'' \emph{Applied Psychological Measurement} 16 (2): 129--47.
-
-\leavevmode\hypertarget{ref-chalmers2012mirt}{}%
-Chalmers, R Philip. 2012. ``Mirt: A Multidimensional Item Response Theory Package for the R Environment.'' \emph{Journal of Statistical Software} 48 (6): 1--29.
-
-\leavevmode\hypertarget{ref-chalmers2018numerical}{}%
----------. 2018. ``Numerical Approximation of the Observed Information Matrix with Oakes' Identity.'' \emph{British Journal of Mathematical and Statistical Psychology} 71 (3): 415--36.
-
-\leavevmode\hypertarget{ref-chalmers2016might}{}%
-Chalmers, R Philip, Alyssa Counsell, and David B Flora.
-2016.
-``It Might Not Make a Big Dif: Improved Differential Test Functioning Statistics That Account for Sampling Variability.'' \emph{Educational and Psychological Measurement} 76 (1): 114--40.
-
-\leavevmode\hypertarget{ref-chitiga2015income}{}%
-Chitiga, Margaret, E Sekyere, and N Tsoanamatsie.
-2015.
-``Income Inequality and Limitations of the Gini Index: The Case of South Africa.'' \emph{Human Sciences Research Council (HSRC), Available at: Http://Www.
-Hsrc.
-Ac.
-Za/En/Review/Hsrc-Review-November-2014/Limitations-of-Gini-Index, Site Accessed} 2.
-
-\leavevmode\hypertarget{ref-demars2016partially}{}%
-DeMars, Christine E. 2016. ``Partially Compensatory Multidimensional Item Response Theory Models: Two Alternate Model Forms.'' \emph{Educational and Psychological Measurement} 76 (2): 231--57.
-
-\leavevmode\hypertarget{ref-drasgow1987study}{}%
-Drasgow, Fritz. 1987. ``Study of the Measurement Bias of Two Standardized Psychological Tests.'' \emph{Journal of Applied Psychology} 72 (1): 19.
-
-\leavevmode\hypertarget{ref-gini1912variabilita}{}%
-Gini, Corrado. 1912. ``Variabilità E Mutabilità (Variability and Mutability).'' \emph{Tipografia Di Paolo Cuppini, Bologna, Italy}, 156.
-
-\leavevmode\hypertarget{ref-hagquist2017recent}{}%
-Hagquist, Curt, and David Andrich.
-2017.
-``Recent Advances in Analysis of Differential Item Functioning in Health Research Using the Rasch Model.'' \emph{Health and Quality of Life Outcomes} 15 (1): 181.
-
-\leavevmode\hypertarget{ref-hastie2001estimating}{}%
-Hastie, Trevor, Robert Tibshirani, and Guenther Walther. 2001. ``Estimating the Number of Data Clusters via the Gap Statistic.'' \emph{J Roy Stat Soc B} 63: 411--23.
-
-\leavevmode\hypertarget{ref-holland1986differential}{}%
-Holland, Paul W, and Dorothy T Thayer. 1986. ``Differential Item Functioning and the Mantel-Haenszel Procedure.'' \emph{ETS Research Report Series} 1986 (2): i--24.
-
-\leavevmode\hypertarget{ref-kopf2015framework}{}%
-Kopf, Julia, Achim Zeileis, and Carolin Strobl. 2015. ``A Framework for Anchor Methods and an Iterative Forward Approach for Dif Detection.'' \emph{Applied Psychological Measurement} 39 (2): 83--103.
-
-\leavevmode\hypertarget{ref-magis2011generalized}{}%
-Magis, David, Gilles Raı̂che, Sébastien Béland, and Paul Gérard.
-2011.
-``A Generalized Logistic Regression Procedure to Detect Differential Item Functioning Among Multiple Groups.'' \emph{International Journal of Testing} 11 (4): 365--86.
-
-\leavevmode\hypertarget{ref-meade2012solving}{}%
-Meade, Adam W, and Natalie A Wright. 2012. ``Solving the Measurement Invariance Anchor Item Problem in Item Response Theory.'' \emph{Journal of Applied Psychology} 97 (5): 1016.
-
-\leavevmode\hypertarget{ref-pohl2017cluster}{}%
-Pohl, Steffi, Eric Stets, and Claus H Carstensen. 2017. ``Cluster-Based Anchor Item Identification and Selection.''
-
-\leavevmode\hypertarget{ref-raju1988area}{}%
-Raju, Nambury S. 1988. ``The Area Between Two Item Characteristic Curves.'' \emph{Psychometrika} 53 (4): 495--502.
-
-\leavevmode\hypertarget{ref-rcore}{}%
-R Core Team. 2019. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
-
-\leavevmode\hypertarget{ref-stark2006detecting}{}%
-Stark, Stephen, Oleksandr S Chernyshenko, and Fritz Drasgow.
-2006.
-``Detecting Differential Item Functioning with Confirmatory Factor Analysis and Item Response Theory: Toward a Unified Strategy.'' \emph{Journal of Applied Psychology} 91 (6): 1292.
-
-\leavevmode\hypertarget{ref-strobl2018anchor}{}%
-Strobl, Carolin, Julia Kopf, Raphael Hartmann, and Achim Zeileis. 2018. ``Anchor Point Selection: An Approach for Anchoring Without Anchor Items.'' Working Papers in Economics; Statistics.
-
-\leavevmode\hypertarget{ref-swaminathan1990detecting}{}%
-Swaminathan, Hariharan, and H Jane Rogers. 1990. ``Detecting Differential Item Functioning Using Logistic Regression Procedures.'' \emph{Journal of Educational Measurement} 27 (4): 361--70.
-
-\leavevmode\hypertarget{ref-sympson1978model}{}%
-Sympson, James B. 1978. ``A Model for Testing with Multidimensional Items.'' In \emph{Proceedings of the 1977 Computerized Adaptive Testing Conference}. 00014.
-
-\leavevmode\hypertarget{ref-talbot2013taking}{}%
-Talbot III, Robert M.
-2013.
-``Taking an Item-Level Approach to Measuring Change with the Force and Motion Conceptual Evaluation: An Application of Item Response Theory.'' \emph{School Science and Mathematics} 113 (7): 356--65.
-
-\leavevmode\hypertarget{ref-thissen1986taxonomy}{}%
-Thissen, David, and Lynne Steinberg. 1986. ``A Taxonomy of Item Response Models.'' \emph{Psychometrika} 51 (4): 567--77.
-
-\leavevmode\hypertarget{ref-thissen1993detection}{}%
-Thissen, David, Lynne Steinberg, and Howard Wainer. 1993. ``Detection of Differential Item Functioning Using the Parameters of Item Response Models.''
-
-\leavevmode\hypertarget{ref-walker2017using}{}%
-Walker, Cindy M, and Sakine Gocer Sahin.
-2017.
-``Using a Multidimensional Irt Framework to Better Understand Differential Item Functioning (Dif): A Tale of Three Dif Detection Procedures.'' \emph{Educational and Psychological Measurement} 77 (6): 945--70.
-
-\leavevmode\hypertarget{ref-tidy}{}%
-Wickham, Hadley. 2017. \emph{Tidyverse: Easily Install and Load the 'Tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}.
-
-\leavevmode\hypertarget{ref-woods2009empirical}{}%
-Woods, Carol M. 2009. ``Empirical Selection of Anchors for Tests of Differential Item Functioning.'' \emph{Applied Psychological Measurement} 33 (1): 42--57.
-
-\leavevmode\hypertarget{ref-yang2012characterizing}{}%
-Yang, Ji Seung, Mark Hansen, and Li Cai. 2012. ``Characterizing Sources of Uncertainty in Item Response Theory Scale Scores.'' \emph{Educational and Psychological Measurement} 72 (2): 264--90.
-
-\end{document}
+data_for_graph %>%
+    ggplot(aes(x = val, y = var)) +
+    ggridges::geom_density_ridges() +
+    geom_point(
+        data = points,
+        aes(x = val, y = item)
+    ) +
+    facet_wrap(~ method) +
+    labs(
+        x = latex2exp::TeX("$\\hat{d_j} = \\hat{b_j}^{ref} - \\hat{b_j}^{foc}$"),
+        y = ""
+    )
+```

### 2021-01-13 paper/paper_3_sim.Rmd
get all the paper stuff back on github

@@
   coord_flip() +
-  geom_hline(yintercept = -0.5, linetype = "dashed") +
+  geom_vline(xintercept = 0, linetype = "dashed", color = "chartreuse4", size = 0.5) +
   labs(

### 2021-01-13 paper/paper_2_sim.R

@@ -1,0 +1,80 @@
+out_3_biased_items <-
+    tibble(
+        pars = rerun(1, create_pars_compensatory_3_biased_items())
+    ) %>%
+    # sim and fit intuitive
+    mutate(
+        sim = pars %>% map(~ sim_students_compensatory(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
+        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
+    ) %>%
+    # dif: AOAA
+    mutate(
+        AOAA_status = sim %>% map(~ dif_AOAA(.$data, .$groups)),
+        AOAA_mod =
+            map2(
+                sim,
+                AOAA_status,
+                ~ mod_flexible(
+                    .x$data,
+                    .x$groups,
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        AOAA_final_dif = AOAA_mod %>% map(mod_to_final_dif)
+    ) %>%
+    # dif: AOAA AS
+    mutate(
+        AOAA_AS_status = map2(sim, AOAA_status, ~ dif_AOAA_AS(.x$data, .x$groups, .y)),
+        AOAA_AS_mod =
+            map2(
+                sim,
+                AOAA_AS_status,
+                ~ mod_flexible(
+                    .x$data,
+                    .x$groups,
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        AOAA_AS_final_dif = AOAA_AS_mod %>% map(mod_to_final_dif)
+    ) %>%
+    # dif: AOAA OAT
+    mutate(
+        AOAA_OAT_status = map(sim, ~ dif_AOAA_OAT(.$data, .$groups)),
+        AOAA_OAT_mod =
+            map2(
+                sim,
+                AOAA_OAT_status,
+                ~ mod_flexible(
+                    .x$data,
+                    .x$groups,
+                    which(.y$status == "flex"),
+                    se = TRUE
+                )
+            ),
+        AOAA_OAT_final_dif = AOAA_OAT_mod %>% map(mod_to_final_dif)
+    ) %>%
+    # dif: tibs cluster
+    mutate(
+        cluster =
+            intuitive_mod %>%
+            map(mod_to_final_dif) %>%
+            map(~ mutate(., cluster = irt_cluster(difference_in_easy)))
+    ) %>%
+    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>%
[…]

### 2021-01-13 paper/paper_3_sim.R

@@ -1,0 +1,22 @@
+# NEED TO ADD MORE DIF
+# DO AN EXTREME CASE THEN WORK BACK FROM THERE
+
+# (higher values of b_target are associated with less dif)
+
+# out_noncomp$intuitive_mod[[1]] %>%
+#     mod_intuitive_to_draws_df() %>%
+#     draws_df_to_logit_plot()
+
+library(furrr)
+options(future.fork.enable = TRUE)
+plan(multiprocess)
+
+out <-
+    tibble(n_dif_items = 2:6) %>%
+    mutate(
+        bigsim = n_dif_items %>% future_map(~ bigsim(runs = 20, n_dif_items = .))
+    )
+
+Sys.time()
+
+out %>% write_rds("paper/paper_3_out1.rds")
